### 2023-11-27 analysis/blots.qmd
adjust plot and bar themes

---
title: "Protein Expression"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/blots.qmd")

tar_config_set(store = here("_targets"))
```

```{r}
#| label: test
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_dual_col1a1) + 
  theme(legend.position = "bottom")
```

### 2023-11-27 R/blots.R

@@ -41,7 +41,8 @@ clean_blots <- function(df) {
 filter_blots <- function(df) {
   df |>
     dplyr::filter(!(.data$experiment == "ipf" & .data$batch == 2)) |>
-    dplyr::filter(!(.data$experiment == "ipf-lf" & .data$batch %in% 1:2))
+    dplyr::filter(!(.data$experiment == "ipf-lf" & .data$batch %in% 1:2)) |>
+    dplyr::filter(!(.data$experiment == "dual" & protein == "Col1a1" & batch %in% 1:4))
 }
 
 index_blots <- function(df) {

### 2023-11-27 R/stats.R

@@ -13,7 +13,7 @@ annot_stats <- function(stats) {
   stats |>
     dplyr::mutate(
       y = Inf,
-      vjust = 1.2,
+      vjust = 1.5,
       label = annot_p(.data[[col]])
     )
 }

### 2023-11-27 analysis/blots.qmd
add blot graphs to analysis qmd

@@
     warning: false
+    toc: true
+    toc-depth: 3
+    toc-location: left
 ---
@@
 
+## Experiment
+
+### Drugs in NHLF
+
 ```{r}
-#| label: test
+#| label: exp_ar_sma
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_ar_sma)
+```
+
+```{r}
+#| label: exp_diclofenac_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_diclofenac_sma)
+```
+
+```{r}
+#| label: exp_dual_akt
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_akt)
+```
+
+```{r}
+#| label: exp_dual_pakt
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_pakt)
+```
+
+```{r}
+#| label: exp_dual_ampk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_ampk)
+```
+
+```{r}
+#| label: exp_dual_pampk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_pampk)
+```
+
+```{r}
+#| label: exp_dual_cnn1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_cnn1)
+```
+
+```{r}
+#| label: exp_dual_col1a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_col1a1)
+```
+
+```{r}
+#| label: exp_dual_col3a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_col3a1)
+```
+
+```{r}
+#| label: exp_dual_erk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_erk)
+```
+
+```{r}
+#| label: exp_dual_perk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_perk)
+```
+
+```{r}
+#| label: exp_dual_fn1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_fn1)
+```
+
+```{r}
+#| label: exp_dual_kla
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_kla)
+```
+
+```{r}
+#| label: exp_dual_mct1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_mct1)
+```
+
+```{r}
+#| label: exp_dual_mct4
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_mct4)
+```
+
+```{r}
+#| label: exp_dual_smad3
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_smad3)
+```
+
+```{r}
+#| label: exp_dual_psmad3
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_psmad3)
+```
+
+```{r}
+#| label: exp_dual_sma
+#| fig-width: 5
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_sma)
+```
+
+```{r}
+#| label: exp_six_hif1a
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_six_hif1a)
+```
+
+### siRNA in NHLF
+
+```{r}
+#| label: exp_sirna_mct1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_sirna_mct1)
+```
+
+```{r}
+#| label: exp_sirna_mct4
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_sirna_mct4)
+```
+
+```{r}
+#| label: exp_sirna_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_sirna_sma)
+```
+
+### TGFβ
+
+```{r}
+#| label: exp_untreated_col1a1
 #| fig-width: 3
@@
 
-tar_read(blot_plot_dual_col1a1) + 
-  theme(legend.position = "bottom")
+tar_read(blot_plot_untreated_col1a1)
 ```
+
+```{r}
+#| label: exp_untreated_fn1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_fn1)
+```
+
+```{r}
+#| label: exp_untreated_mct1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_mct1)
+```
+
+```{r}
+#| label: exp_untreated_mct4
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_mct4)
+```
+
+```{r}
+#| label: exp_untreated_sma
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_sma)
+```
+
+### Lactate Treatment
+
+```{r}
+#| label: exp_lactate_col1a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_col1a1)
+```
+
+```{r}
+#| label: exp_lactate_fn1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_fn1)
+```
+
+```{r}
+#| label: exp_lactate_mct1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_mct1)
+```
+
+```{r}
+#| label: exp_lactate_mct4
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_mct4)
+```
+
+```{r}
+#| label: exp_lactate_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_sma)
+```
+
+### Drugs in Mice
+
+```{r}
+#| label: exp_bleo_mct1
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_bleo_mct1)
+```
+
+```{r}
+#| label: exp_bleo_mct4
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_bleo_mct4)
+```
+
+```{r}
+#| label: exp_bleo_sma
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_bleo_sma)
+```
+
+### IPF Lung Fibroblasts
+
+```{r}
+#| label: exp_comp_mct1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_comp_mct1)
+```
+
+```{r}
+#| label: exp_comp_mct4
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_comp_mct4)
+```
+
+```{r}
+#| label: exp_comp_sma
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_comp_sma)
+```
+
+```{r}
+#| label: exp_ipf_lf_col1a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_lf_col1a1)
+```
+
+```{r}
+#| label: exp_ipf_lf_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_lf_sma)
+```
+
+### IPF Lungs
+
+```{r}
+#| label: exp_ipf_mct1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_mct1)
+```
+
+```{r}
+#| label: exp_ipf_mct4
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_mct4)
+```
+
+```{r}
+#| label: exp_ipf_sma
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_sma)
+```
+
+## Protein
+
+### AKT
+
+```{r}
+#| label: prot_dual_akt
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_akt)
+```
+
+```{r}
+#| label: prot_dual_pakt
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_pakt)
+```
+
+### AMPK
+
+```{r}
+#| label: prot_dual_ampk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_ampk)
+```
+
+```{r}
+#| label: prot_dual_pampk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_pampk)
+```
+
+### CNN1
+
+```{r}
+#| label: prot_dual_cnn1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_cnn1)
+```
+
+### Col1a1
+
+```{r}
+#| label: prot_dual_col1a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_col1a1)
+```
+
+```{r}
+#| label: prot_ipf_lf_col1a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_lf_col1a1)
+```
+
+```{r}
+#| label: prot_lactate_col1a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_col1a1)
+```
+
+```{r}
+#| label: prot_untreated_col1a1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_col1a1)
+```
+
+### Col3a1
+
+```{r}
+#| label: prot_dual_col3a1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_col3a1)
+```
+
+### ERK
+
+```{r}
+#| label: prot_dual_erk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_erk)
+```
+
+```{r}
+#| label: prot_dual_perk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_perk)
+```
+
+### FN1
+
+```{r}
+#| label: prot_dual_fn1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_fn1)
+```
+
+```{r}
+#| label: prot_lactate_fn1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_fn1)
+```
+
+```{r}
+#| label: prot_untreated_fn1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_fn1)
+```
+
+### HIF-1α
+
+```{r}
+#| label: prot_six_hif1a
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_six_hif1a)
+```
+
+### Kla
+
+```{r}
+#| label: prot_dual_kla
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_kla)
+```
+
+### MCT1
+
+```{r}
+#| label: prot_bleo_mct1
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_bleo_mct1)
+```
+
+```{r}
+#| label: prot_comp_mct1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_comp_mct1)
+```
+
+```{r}
+#| label: prot_dual_mct1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_mct1)
+```
+
+```{r}
+#| label: prot_ipf_mct1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_mct1)
+```
+
+```{r}
+#| label: prot_lactate_mct1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_mct1)
+```
+
+```{r}
+#| label: prot_sirna_mct1
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_sirna_mct1)
+```
+
+```{r}
+#| label: prot_untreated_mct1
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_mct1)
+```
+
+### MCT4
+
+```{r}
+#| label: prot_bleo_mct4
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_bleo_mct4)
+```
+
+```{r}
+#| label: prot_comp_mct4
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_comp_mct4)
+```
+
+```{r}
+#| label: prot_dual_mct4
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_mct4)
+```
+
+```{r}
+#| label: prot_ipf_mct4
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_mct4)
+```
+
+```{r}
+#| label: prot_lactate_mct4
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_mct4)
+```
+
+```{r}
+#| label: prot_sirna_mct4
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_sirna_mct4)
+```
+
+```{r}
+#| label: prot_untreated_mct4
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_mct4)
+```
+
+### Smad3
+
+```{r}
+#| label: prot_dual_smad3
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_smad3)
+```
+
+```{r}
+#| label: prot_dual_psmad3
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_psmad3)
+```
+
+### α-SMA
+
+```{r}
+#| label: prot_ar_sma
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_ar_sma)
+```
+
+```{r}
+#| label: prot_bleo_sma
+#| fig-width: 3.5
+#| fig-height: 2.5
+
+tar_read(blot_plot_bleo_sma)
+```
+
+```{r}
+#| label: prot_comp_sma
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_comp_sma)
+```
+
+```{r}
+#| label: prot_diclofenac_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_diclofenac_sma)
+```
+
+```{r}
+#| label: prot_dual_sma
+#| fig-width: 5
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_sma)
+```
+
+```{r}
+#| label: prot_ipf_lf_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_lf_sma)
+```
+
+```{r}
+#| label: prot_ipf_sma
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_ipf_sma)
+```
+
+```{r}
+#| label: prot_lactate_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_lactate_sma)
+```
+
+```{r}
+#| label: prot_sirna_sma
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_sirna_sma)
+```
+
+```{r}
+#| label: prot_untreated_sma
+#| fig-width: 3
+#| fig-height: 2.5
+
+tar_read(blot_plot_untreated_sma)
+```

### 2023-11-27 R/plots.R

@@ -192,7 +192,9 @@ nice_limits <- function(x, n = 5) {
 
 clrs <- c(
   "Ctl"           = "#A0CBE8",
+  "Ctl-LF"        = "#A0CBE8",
   "IPF"           = "#4E79A7",
+  "IPF-LF"        = "#4E79A7",
   "Bleo"          = "#4E79A7",
   "TGFβ"          = "#4E79A7",
   "Veh"           = "#4E79A7",
@@ -309,13 +311,17 @@ plot_two_factor <- function(
       stat = "summary",
       fun = "mean",
       ggplot2::aes(fill = .data[["condition"]]),
-      position = ggplot2::position_dodge(width = dodge_width),
+      position = ggplot2::position_dodge(
+        width = dodge_width
+      ),
       width = 0.6,
       show.legend = TRUE
     ) +
     ggplot2::geom_errorbar(
       ggplot2::aes(group = .data[["condition"]]),
-      position = ggplot2::position_dodge(width = dodge_width),
+      position = ggplot2::position_dodge(
+        width = dodge_width
+      ),
       stat = "summary",
       fun.data = ggplot2::mean_se,
       width = 0.25,

### 2023-11-27 _targets.R

@@ -79,9 +79,9 @@ list(
       "dual",        "α-SMA",      "dual_sma",         FALSE,    rlang::sym("comps_dual"),
       "ipf-lf",      "Col1a1",     "ipf_lf_col1a1",    FALSE,    rlang::sym("comps_azd_2"),
       "ipf-lf",      "α-SMA",      "ipf_lf_sma",       FALSE,    rlang::sym("comps_azd_2"),
-      "ipf",         "MCT1",       "ipf_mct1",         FALSE,     NULL,
-      "ipf",         "MCT4",       "ipf_mct4",         FALSE,     NULL,
-      "ipf",         "α-SMA",      "ipf_sma",          FALSE,     NULL,
+      "ipf",         "MCT1",       "ipf_mct1",         FALSE,    NULL,
+      "ipf",         "MCT4",       "ipf_mct4",         FALSE,    NULL,
+      "ipf",         "α-SMA",      "ipf_sma",          FALSE,    NULL,
       "lactate",     "Col1a1",     "lactate_col1a1",   FALSE,    rlang::sym("comps_lactate"),
       "lactate",     "FN1",        "lactate_fn1",      FALSE,    rlang::sym("comps_lactate"),
       "lactate",     "MCT1",       "lactate_mct1",     FALSE,    rlang::sym("comps_lactate"),
@@ -117,6 +117,14 @@ list(
     #   format = "file"
     # ),
     NULL
+  ),
+
+
+  # analysis ----------------------------------------------------------------
+
+  tar_quarto(
+    blots_analysis,
+    path = "analysis/blots.qmd"
   ),
 
   NULL

### 2023-11-27 analysis/blots.qmd
analyze blot ratios

@@
 ```{r}
+#| label: exp_dual_p_akt
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_akt)
+```
+
+```{r}
 #| label: exp_dual_ampk
@@
 tar_read(blot_plot_dual_pampk)
+```
+
+```{r}
+#| label: exp_dual_p_ampk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_ampk)
 ```
@@
 tar_read(blot_plot_dual_perk)
+```
+
+```{r}
+#| label: exp_dual_p_erk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_erk)
 ```
@@
 tar_read(blot_plot_dual_psmad3)
+```
+
+```{r}
+#| label: exp_dual_p_smad3
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_smad3)
 ```
@@
 
+```{r}
+#| label: prot_dual_p_akt
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_akt)
+```
+
 ### AMPK
@@
 tar_read(blot_plot_dual_pampk)
+```
+
+```{r}
+#| label: prot_dual_p_ampk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_ampk)
 ```
@@
 tar_read(blot_plot_dual_perk)
+```
+
+```{r}
+#| label: prot_dual_p_erk
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_erk)
 ```
@@
 
+```{r}
+#| label: prot_dual_p_smad3
+#| fig-width: 4
+#| fig-height: 2.5
+
+tar_read(blot_plot_dual_p_smad3)
+```
+
 ### α-SMA

### 2023-11-27 _targets.R

@@ -111,14 +111,36 @@ list(
       plot_blot(blot_norm, blot_stats, title = protein),
       format = "rds"
     ),
-    # tar_target(
-    #   blot_img,
-    #   write_plot(blot_plot, name, path = "analysis/figures/blots"),
-    #   format = "file"
-    # ),
     NULL
   ),
-
+  tar_target(
+    blots_phospho,
+    phospho_ratio(blots_filtered)
+  ),
+  tar_map(
+    values = tibble::tribble(
+      ~experiment,   ~protein,     ~name,              ~paired, ~comp,
+      "dual",        "AKT",        "dual_p_akt",       FALSE,    rlang::sym("comps_azd_2"),
+      "dual",        "AMPK",       "dual_p_ampk",      FALSE,    rlang::sym("comps_azd_2"),
+      "dual",        "ERK",        "dual_p_erk",       FALSE,    rlang::sym("comps_azd_2"),
+      "dual",        "Smad3",      "dual_p_smad3",     FALSE,    rlang::sym("comps_azd_2")
+    ),
+    names = name,
+    tar_target(
+      blot_norm,
+      norm_blot(blots_phospho, experiment, protein)
+    ),
+    tar_target(
+      blot_stats,
+      analyze_blot(blot_norm, paired, comp)
+    ),
+    tar_target(
+      blot_plot,
+      plot_blot(blot_norm, blot_stats, title = paste0("p", protein, " / ", protein)),
+      format = "rds"
+    ),
+    NULL
+  ),
 
   # analysis ----------------------------------------------------------------
 

### 2023-11-28 _targets.R
add extracellular metabolite measurements

@@ -223,9 +223,153 @@ list(
     ),
     tar_target(
       plates_conc,
-      clean_data(plates_interp, cf)
+      clean_plates_interp(plates_interp, cf)
     ),
     NULL
+  ),
+
+  # extracellular -----------------------------------------------------------
+
+  tar_map(
+    values = list(
+      df = rlang::syms(c("plates_conc_lactate", "plates_conc_glucose")),
+      cf = rlang::syms(c("plates_conc_picogreen", "plates_conc_picogreen")),
+      names = c("lactate", "glucose")
+    ),
+    names = names,
+    tar_target(
+      plates_norm,
+      norm_plates(df, cf)
+    )
+  ),
+  tar_map(
+    values = list(
+      names = c(
+        "lactate_sirna",
+        "lactate_azd",
+        "lactate_ar",
+        "lactate_nsaid",
+        "glucose_azd"
+      ),
+      norm = rlang::syms(c(
+        "plates_norm_lactate",
+        "plates_norm_lactate",
+        "plates_norm_lactate",
+        "plates_norm_lactate",
+        "plates_norm_glucose"
+      )),
+      exp = list(
+        "sirna",
+        "dual",
+        c("dual", "ar"),
+        "diclofenac",
+        "dual"
+      ),
+      treat = list(
+        NULL,
+        c("Veh", "AZD", "VB", "AZD/VB"),
+        c("Veh", "AR", "VB", "AR/VB"),
+        NULL,
+        c("Veh", "AZD", "VB", "AZD/VB")
+      ),
+      comps = rlang::syms(c(
+        "comps_sirna",
+        "comps_azd_2",
+        "comps_ar_2",
+        "comps_nsaid",
+        "comps_azd_2"
+      )),
+      ytitle = c(
+        "Lactate (pmol/cell)",
+        "Lactate (pmol/cell)",
+        "Lactate (pmol/cell)",
+        "Lactate (pmol/cell)",
+        "Glucose (pmol/cell)"
+      )
+    ),
+    names = names,
+    tar_target(
+      plate_filter,
+      filter_plates(norm, exp = exp, treat = treat)
+    ),
+    tar_target(
+      plate_stats,
+      twofactor(plate_filter, "norm", comps = comps)
+    ),
+    tar_target(
+      plate_plot,
+      plot_two_factor(
+        plate_filter,
+        plate_stats,
+        "treatment",
+        "norm",
+        ytitle = ytitle
+      ),
+      format = "rds"
+    )
+  ),
+
[…]

### 2023-11-28 R/nad.R
add NAD analysis

@@ -1,0 +1,51 @@
+# nad.R
+
+clean_nad <- function(protein, nad) {
+  protein <-
+    protein |>
+    dplyr::mutate(
+      protein = 300 * .data$conc  # μg protein
+    ) |>
+    dplyr::select(-"conc")
+
+  nad <-
+    nad |>
+    dplyr::rename(measurement = "nucleotide") |>
+    dplyr::mutate(
+      conc = 900 * .data$conc  # pmol nucleotide
+    )
+
+  dplyr::left_join(
+    protein,
+    nad,
+    by = c("experiment", "sample", "condition", "treatment")
+  ) |>
+    tidyr::pivot_wider(
+      names_from = "measurement",
+      values_from = "conc"
+    ) |>
+    dplyr::mutate(
+      NAD_norm = .data$NAD / .data$protein,  # pmol / μg
+      NADH_norm = .data$NADH / .data$protein,
+      Ratio = .data$NADH / .data$NAD
+    ) |>
+    refactor() |>
+    tidyr::pivot_longer(
+      "NAD":"Ratio",
+      names_to = "measurement",
+      values_to = "value"
+    ) |>
+    dplyr::rename(group = "experiment") |>
+    dplyr::ungroup()
+}
+
+norm_nad <- function(df) {
+  df |>
+    dplyr::mutate(grand_mean = mean(.data$value)) |>
+    dplyr::group_by(.data$group) |>
+    dplyr::mutate(
+      exp_mean = mean(.data$value),
+      cf = .data$grand_mean - .data$exp_mean,
+      value_corr = .data$value + .data$cf
+    )
+}

### 2023-11-28 R/plates.R

@@ -27,7 +27,6 @@ clean_plates_interp <- function(df, cf) {
   dplyr::mutate(df, conc = .data$conc * cf)
 }
 
-
 norm_plates <- function(df, cf) {
   dplyr::left_join(
     df,

### 2023-11-28 R/stats.R

@@ -85,12 +85,16 @@ onefactor <- function(
 }
 
 
-twofactor <- function(df, y, mixed = TRUE, comps) {
+twofactor <- function(df, y, mixed = TRUE, comps, log = FALSE) {
   x <-
     df |>
     dplyr::mutate(
       dplyr::across(c("condition", "treatment"), forcats::fct_drop)
     )
+
+  if (log) {
+    y <- paste0("log(", y, ")")
+  }
 
   withCallingHandlers(
     message = \(cnd) rlang::warn(cnd$message),
@@ -122,4 +126,19 @@ twofactor <- function(df, y, mixed = TRUE, comps) {
     ) |>
     tidyr::complete(.data$condition, .data$treatment) |>
     annot_stats()
+}
+
+group_twofactor <- function(df, y, comps, ...) {
+  cols <- c("protein", "name", "rate", "stage", "stats", "measurement")
+  df |>
+    tidyr::nest() |>
+    dplyr::mutate(stats = purrr::map(
+      .data$data,
+      twofactor,
+      y = y,
+      comps = comps,
+      ...
+    )) |>
+    dplyr::select(tidyselect::any_of(cols)) |>
+    tidyr::unnest(c("stats"))
 }

### 2023-11-28 _targets.R

@@ -372,16 +372,76 @@ list(
     )
   ),
 
-  # analysis ----------------------------------------------------------------
+  # nad ---------------------------------------------------------------------
 
-  tar_quarto(
-    blots_analysis,
-    path = "analysis/blots.qmd"
+  tar_target(
+    nad_clean,
+    clean_nad(plates_conc_nad_protein, plates_conc_nad_nad)
   ),
-  tar_quarto(
-    extracellular_analysis,
-    path = "analysis/extracellular.qmd"
+  tar_map(
+    values = list(
+      names = c(
+        "nad",
+        "nad_norm",
+        "nadh",
+        "nadh_norm",
+        "ratio"
+      ),
+      measure = c(
+        "NAD",
+        "NAD_norm",
+        "NADH",
+        "NADH_norm",
+        "Ratio"
+      ),
+      ytitle = c(
+        "NAD (pmol)",
+        "NAD (pmol / μg protein)",
+        "NADH (pmol)",
+        "NADH (pmol / μg protein)",
+        "NADH / NAD"
+      )
+    ),
+    names = names,
+    tar_target(
+      nad_filter,
+      dplyr::filter(nad_clean, measurement == measure)
+    ),
+    tar_target(
+      nad_stats,
+      twofactor(nad_filter, y = "value", comps = comps_azd_2, log = TRUE)
+    ),
+    tar_target(
+      nad_norm,
+      norm_nad(nad_filter)
+    ),
+    tar_target(
+      nad_plot,
+      plot_two_factor(
+        nad_norm,
+        nad_stats,
+        "treatment",
+        "value_corr",
+        ytitle = ytitle
+        ),
+      format = "rds"
+    )
   ),
 
-  NULL
+# analysis ----------------------------------------------------------------
+
+tar_quarto(
+  blots_analysis,
+  path = "analysis/blots.qmd"
+),
+tar_quarto(
+  extracellular_analysis,
+  path = "analysis/extracellular.qmd"
+),
+tar_quarto(
+  redox_analysis,
+  path = "analysis/redox.qmd"
+),
+
+NULL
 )

### 2024-01-18 manuscript/manuscript.qmd
add manuscript infrastructure

---
title: "Lactate transport inhibition therapeutically reprograms fibroblast 
    metabolism in experimental pulmonary fibrosis"

authors:
- name: David R. Ziehr
  affiliations: 
  - ref: mgh
  - ref: bwh
  - ref: hms
- name: Nathan M. Krah
  affiliations:
  - ref: utah_gen
- name: Kevin Leahy 
  affiliations:
  - ref: bwh
- name: K. Mark Parnell
  affiliations:
  - ref: vettore
- name: Jack Varon
  affiliations: 
  - ref: bwh
  - ref: hms
- name: Rebecca M. Baron
  affiliations:
  - ref: bwh
  - ref: hms
- name: Nancy J. Philp
  affiliations:
  - ref: tj
- name: Lida P. Hariri
  affiliations: 
  - ref: hms
  - ref: mgh_path
- name: Edy Y. Kim
  affiliations:
  - ref: bwh
  - ref: hms
- name: Rachel S. Knipe
  affiliations:
  - ref: mgh
  - ref: hms
- name: Jared Rutter
  affiliations:
  - ref: utah_biochem
  - ref: utah_hhmi
- name: William M. Oldham
  affiliations:
  - ref: bwh
  - ref: hms
  email: woldham@bwh.harvard.edu
  corresponding: true

affiliations:
- id: mgh
  name: Department of Medicine, Massachusetts General Hospital, Boston, MA
- id: bwh
  name: Department of Medicine, Brigham and Women's Hospital and Harvard Medical School, Boston, MA
- id: hms
  name: Department of Medicine, Harvard Medical School, Boston, MA
- id: utah_gen
  name: Department of Human Genetics, University of Utah, Salt Lake City, UT
- id: vettore
  name: Vettore Biosciences, San Francisco, CA
- id: tj
  name: Department of Pathology, Anatomy & Cell Biology, Thomas Jefferson University, Philadelphia, PA
- id: mgh_path
  name: Department of Pathology, Massachusetts General Hospital, Boston, MA
- id: utah_biochem
  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
- id: utah_hhmi
  name: Howard Hughes Medical Institute, University of Utah School of Medicine, Salt Lake City, UT

format:
  docx:
    reference-doc: template.docx
    page-width: 7.5
    output-file: ziehr_2023_mcti_manuscript.docx
    fig-align: center

bibliography: 
- library.bib
- pkgs.bib
csl: jci.csl

crossref: 
  fig-title: ""
  title-delim: ""

filters:
- authors-block
---

### 2024-01-18 manuscript/supplement.qmd
add manuscript infrastructure

--- 2024-01-18 manuscript/manuscript.qmd
@@
 ---
-title: "Lactate transport inhibition therapeutically reprograms fibroblast 
+title: "Supplementary Material"
+subtitle: "Lactate transport inhibition therapeutically reprograms fibroblast 
     metabolism in experimental pulmonary fibrosis"
@@
     page-width: 7.5
-    output-file: ziehr_2023_mcti_manuscript.docx
+    output-file: ziehr_2023_mcti_supplement.docx
     fig-align: center

### 2024-01-18 R/manuscript.R

@@ -1,0 +1,15 @@
+# manuscript.R
+
+write_pkg_cites <- function() {
+  desc::desc_get_deps() |>
+    dplyr::pull(.data$package) |>
+    knitr::write_bib(file = manuscript_path("pkgs.bib")) |>
+    suppressWarnings()
+}
+
+test_quarto <- function(file = "manuscript.qmd") {
+  quarto::quarto_render(
+    manuscript_path(file),
+    execute_dir = getwd()
+  )
+}

### 2024-01-20 analysis/network.qmd
add cosmos analysis

---
title: "Network Analysis"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
    toc: true
    toc-depth: 3
    toc-location: left
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/network.qmd")

tar_config_set(store = here("_targets"))
```

## Build Networks

```{r}
#| label: universe

tt_metab <- tar_read(metab_mcti_tar_tt_tgfb)
all_metab <- tt_metab[["hmdb"]]
sig_metab <- all_metab[tt_metab$adj.P.Val < 0.1]

tt_genes <- tar_read(deg_tgfb) 
all_genes <- tt_genes[["symbol"]]
sig_genes <- all_genes[tt_genes$padj < 0.1]

universe <- 
  hmdb::hmdb_proteins |> 
  filter(hmdb %in% all_metab & protein %in% all_genes)

enriched <- 
  universe |> 
  filter(hmdb %in% sig_metab & protein %in% sig_genes)

edges <- nrow(enriched)
```

## Compare to Random Networks

```{r}
#| label: random-networks

sample_network <- function() {
  metab <- sample(all_metab, length(sig_metab), replace = FALSE)
  gene <- sample(all_genes, length(sig_genes), replace = FALSE)
  hmdb::hmdb_proteins |> 
    filter(hmdb %in% metab & protein %in% gene) |> 
    nrow()
}
edge_list <- replicate(1000, sample_network())

tibble(x = edge_list) |> 
ggplot() +
  geom_histogram(
    aes(x = x), 
    binwidth = 50
  ) + 
  geom_segment(
    x = edges, 
    xend = edges, 
    y = 10, 
    yend = 0, 
    arrow = arrow(length = unit(10, "points")), 
    color = "darkblue", 
    linewidth = 1, 
    lineend = "butt", 
    linejoin = "bevel"
  ) +
  labs(
    title = "Simulated networks", 
    x = NULL, 
    y = "Edge number"
  ) +
  scale_y_continuous(
    expand = expansion(c(0, 0)), 
    limits = nice_limits
  ) +
  theme_plot()
```

## Cluster Network

```{r}
#| label: graph

tidygraph::tbl_graph(edges = select(enriched, hmdb, protein)) |> 
  ggraph::ggraph(layout = "fr") +
  ggraph::geom_edge_link()
```

```{r}
#| label: cluster

clusters <- 
  enriched |> 
  select(red = protein, blue = name) |> 
  as.data.frame() |>
  netZooR::createCondorObject() |> 
  netZooR::condorCluster()
```

### 2024-01-20 R/cosmos.R

@@ -1,0 +1,119 @@
+# cosmos.R
+
+set_carnival_options <- function(){
+  path <- "analysis/carnival"
+  carnival_options <- cosmosR::default_CARNIVAL_options(solver = "cplex")
+  carnival_options$solverPath <- "/Applications/CPLEX_Studio2211/cplex/bin/x86-64_osx/cplex"
+  carnival_options$outputFolder <- path
+  carnival_options$workdir <- path
+  carnival_options$threads <- 4
+  carnival_options$clonelog <- -1
+  carnival_options$poolIntensity <- 2
+  carnival_options$mipGap <- 0.2
+  carnival_options
+}
+
+get_cosmos_data <- function() {
+  env <- rlang::new_environment()
+  x <- data(package = "cosmosR")[["results"]][, "Item"]
+  data(list = x, package = "cosmosR", envir = env)
+  mget(x, envir = env)
+}
+
+preprocess <- function(
+    route = c("forward", "reverse"),
+    network,
+    signaling,
+    metabolites,
+    transcripts,
+    diff_exp_threshold = 1,
+    maximum_network_depth = 6,
+    carnival_options
+){
+  carnival_options$timelimit <- 3600
+
+  network_members <- unique(c(network$source, network$target))
+  signaling <- signaling[names(signaling) %in% network_members]
+  metabolites <- metabolites[names(metabolites) %in% network_members]
+
+  route <- match.arg(route)
+  if (route == "forward") {
+    fun <- cosmosR::preprocess_COSMOS_signaling_to_metabolism
+    remove_nodes <- TRUE
+  } else if (route == "reverse") {
+    fun <- cosmosR::preprocess_COSMOS_metabolism_to_signaling
+    remove_nodes <- TRUE
+  }
+
+  args <-
+    list(
+      meta_network = network,
+      signaling_data = signaling,
+      metabolic_data = metabolites,
+      diff_expression_data = transcripts,
+      diff_exp_threshold = diff_exp_threshold,
+      maximum_network_depth = maximum_network_depth,
+      remove_unexpressed_nodes = remove_nodes,
+      CARNIVAL_options = carnival_options
+    )
+
+  do.call(fun, args)
+}
+
+run_cosmos <- function(
+    route = c("forward", "reverse"),
+    data,
+    carnival_options,
+    timelimit = 14400
+){
[…]

### 2024-01-20 manuscript/manuscript.qmd
start working on manuscript results

@@
 ---
+
+```{r}
+#| label: setup
+#| include: false
+
+# ensure refs are up-to-date
+targets::tar_load(refs)
+targets::tar_load(pkg_citations)
+
+# load data for paper
+
+# ensure figures are up-to-date
+# targets::tar_load(
+#   c(
+#     fig01, 
+#     fig02, 
+#     fig03,
+#     fig04,
+#     fig05, 
+#     fig06, 
+#     fig10, 
+#     fig11,
+#     resources_table,
+#     NULL
+#   )
+# )
+```
+
+\newpage
+
+# Abstract
+
+# Keywords
+
+\newpage
+
+# Introduction
+
+Idiopathic pulmonary fibrosis (IPF) is a chronic and progressive lung disease with high mortality and limited therapeutic options.
+IPF affects approximately 150,000 patients in the U.S.
+with a median survival of 3-5 years [@esposito2015; @raghu2016; @martinez2017].
+Currently approved pharmacotherapies for IPF are limited to the antifibrotics pirfenidone and nintedanib that slow, but do not stop, disease progression [@king2014; @richeldi2014], leaving lung transplantation as the only option available to eligible patients with progressive disease.
+The limited efficacy of antifibrotic therapies emphasizes the need for novel therapeutic approaches targeting different features of IPF pathobiology.
+
+Accumulating evidence suggests that metabolic reprogramming may be one such therapeutic vulnerability in IPF [@bueno2020; @selvarajah2021].
+Lung fibrosis is driven by the excessive deposition of extracellular matrix by myofibroblasts [@martinez2017].
+Fundamental changes in myofibroblast metabolism support myofibroblast differentiation and extracellular matrix production [@bernard2015; @xie2015; @goodwin2018; @hamanaka2019; @selvarajah2019].
+In particular, increased glycolysis and lactate production have been observed in IPF myofibroblasts *ex vivo* and following TGF-β1 (TGFβ)-induced myofibroblast differentiation *in vitro* [@kottmann2012; @xie2015; @kottmann2015; @bernard2015].
+These metabolic changes are critical for fibrogenesis, as small molecule inhibitors of glucose uptake, glycolysis, and lactate fermentation prevent myofibroblast differentiation *in vitro* and attenuate pulmonary fibrosis in animal models [@bernard2015; @kottmann2012; @xie2015; @goodwin2018; @judge2018; @cho2017].
+Unfortunately, low target affinities, poor specificity, narrow therapeutic indices, and common genetic resistance have all hampered the translation of these investigational compounds for clinical use [@rodriguez-enriquez2009; @michelakis2017; @pelicano2006].
+Moreover, the molecular mechanisms by which these metabolic inhibitors attenuate the myofibroblast differentiation transcriptional program remain unclear.
+In order to leverage metabolic therapies for IPF, more targeted and better characterized drugs must be developed.
+
+Toward this end, we sought to characterize the effects of a novel metabolic approach - inhibition of lactate transport - on myofibroblast differentiation and experimental pulmonary fibrosis.
+Lactate must be secreted to sustain glycolysis in myofibroblasts.
+A family of monocarboxylate transporters (MCT1-4) maintains lactate homeostasis and small molecule inhibitors of these transporters have been under active clinical development for oncologic applications where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
+Importantly, compared to previously investigated inhibitors of glycolysis, MCT inhibitors exhibit favorable pharmacologic profiles and have successfully translated to human clinic trials [@halford2023].
+Before the promise of this therapeutic approach in IPF may be realized, however, the preclinical efficacy and molecular mechanisms-of-action of lactate transport inhibitors must be demonstrated experimentally.
+
+In this work, we evaluated the contribution of lactate transporters to experimental pulmonary fibrosis.
+We found increased expression of the lactate transporters MCT1 and MCT4 IPF patient lungs, consistent with increased glycolysis previously observed.
+Inhibition of these transporters attenuated bleomycin-induced lung fibrosis *in vivo* and TGFβ-induced myofibroblast differentiation *in vitro*, where MCT4 inhibition demonstrated increased therapeutic efficacy.
+*Using stable isotope tracing, high-resolution spatial metabolomics, we find...* Based on these findings, we report a novel MCT4 inhibitor suitable for human clinical studies.
+Together, these data illuminate a novel approach to disrupt the metabolic program fueling the exuberant fibrosis at the core of IPF and provide proof-of-concept to target lactate transporters in the treatment of this deadly disease.
+
+# Results
+
+## MCT expression increases in human pulmonary fibrosis and experimental models
+
+Of the four lactate transporters, MCT1 and MCT4 are the most highly expressed in the lung [@frangogiannis2020].
+Based on this observation, we characterized MCT1 and MCT4 expression in human IPF lung explants obtained at the time of transplantation.
+Consistent with a pathologic role for MCT1 and MCT4 in IPF, we found significantly increased expression of these proteins in human IPF lung compared to non-fibrotic control samples (**Figure 1A-B**).

### 2024-01-20 R/blots.R

@@ -97,7 +97,8 @@ analyze_blot <- function(df, paired, comp = NULL) {
   } else {
     out <- twofactor(df, "norm", mixed = paired, comp)
   }
-  out
+  out |>
+    dplyr::mutate(protein = unique(df$protein), .before = 1)
 }
 
 plot_blot <- function(df, stats = NULL, title = NULL) {
@@ -131,4 +132,19 @@ phospho_ratio <- function(df) {
     ) |>
     dplyr::mutate(ratio = .data$phospho / .data$total) |>
     dplyr::filter(!is.na(.data$ratio))
+}
+
+combine_plots <- function(dfs, stats) {
+  df <- dplyr::bind_rows(dfs)
+  stat <- dplyr::bind_rows(stats)
+  prot <- unique(df$protein)
+  df$protein <- factor(df$protein, levels = prot)
+  stat$protein <- factor(stat$protein, levels = prot)
+
+  plot_blot(df, stat) +
+    ggplot2::facet_wrap(
+      ggplot2::vars(protein),
+      nrow = 1,
+      scales = "free_y"
+    )
 }

### 2024-01-20 R/manuscript.R

@@ -13,3 +13,18 @@ test_quarto <- function(file = "manuscript.qmd") {
     execute_dir = getwd()
   )
 }
+
+plot_image <- function(img, scale = 1, hjust = 0, vjust = 0) {
+  cowplot::ggdraw() +
+    cowplot::draw_image(
+      magick::image_read(img),
+      scale = scale,
+      hjust = hjust,
+      vjust = vjust
+    )  +
+    theme_plot() +
+    ggplot2::theme(
+      panel.border = ggplot2::element_blank(),
+      axis.line = ggplot2::element_blank()
+    )
+}

### 2024-01-22 manuscript/manuscript.qmd
add inhibitor effects on myofibroblast differentiation

@@
 # ensure figures are up-to-date
-# targets::tar_load(
-#   c(
-#     fig01, 
+targets::tar_load(
+  c(
+    fig01,
     # fig02,
@@
     # resources_table,
-#     NULL
-#   )
-# )
+    NULL
+  )
+)
 ```
@@
 Consistent with a pathologic role for MCT1 and MCT4 in IPF, we found significantly increased expression of these proteins in human IPF lung compared to non-fibrotic control samples (**Figure 1A-B**).
+These findings were recapitulated in experimental pulmonary fibrosis where intratracheal bleomycin increased both MCT1 and MCT4 expression (**Figure 1C-D**).
+
+A defining feature of IPF is the activation of tissue myofibroblasts, which are characterized by the *de novo* expression of smooth muscle α-actin (α-SMA); stress fiber formation; and increased migration, contraction, and extracellular matrix production [@selvarajah2021; @xie2015; @tomasek2002].
+TGFβ is the most potent inducer of myofibroblast activation *in vitro* and *in vivo*.
+TGFβ-dependent α-SMA expression is a well-accepted and widely utilized model of myofibroblast activation relevant to pulmonary fibrosis [@goodwin2018; @cui2019; @rangarajan2018; @locy2020].
+Consistent with our findings in human IPF lungs, we observed increased expression of MCT1 and MCT4 in normal human lung fibroblasts following TGFβ treatment (**Figure 1E-F**).
+These findings are consistent with increased expression of other glycolytic enzymes and the associated metabolic changes that have been previously observed in these cells [@xie2015].
+Together, these data demonstrate that pulmonary fibrosis is associated with increased expression of lactate transporters in the lung generally and in myofibroblasts specifically.
+
+## Lactate transport is essential for myofibroblast differentiation *in vitro*
+
+Using RNA interference and pharmacologic approaches, we next tested whether MCT expression and activity were required for myofibroblast differentiation *in vitro*.
+Lung fibroblasts were transfected with siRNA targeting MCT1 and MCT4 separately and together.
+After 24 h, the cells were treated with TGFβ for 48 h to induce myofibroblast differentiation.
+The siRNAs reduced lactate transporter protein levels > 80% (**Figure 2A-B**).
+Decreasing MCT1 or MCT4 expression caused a marked reduction in TGFβ-stimulated α-SMA expression.
+Notably, siMCT1 also decreased MCT4 expression.
+Interestingly, the mRNA sequence homology of these transporters is low and the MCT1 siRNA sequences do not match the MCT4 mRNA sequence, perhaps suggesting a post-translational explanation for this phenomenon.
+No adverse effects of lactate transporter knockdown on cell viability were observed, and siMCT1 significantly increased cell count in both control and TGFβ-treated cells (**Supplemental Figure 2A**).
+
+Given the reduction of MCT4 expression following siMCT1 treatment, we next evaluated the effects of pharmacologic MCT inhibitors on myofibroblast differentiation.
+AZD3965 is a high-affinity (*K~i~* 1.6 nM) inhibitor of MCT1 [@curtis2017].
+VB124 is a recently developed high-affinity (IC~50~ 19 nM) inhibitor of MCT4 [@cluntun2021].
+IPF lung fibroblasts were differentiated with TGFβ in the presence of these MCT inhibitors (**Figure 2C-D**).
+MCT4 inhibition by VB124 alone, or in combination with AZD3965, decreased Col1a1 and α-SMA expression.
+These effects were also observed in normal lung fibroblasts where AZD3965 as well as VB124, separately or together, decreased α-SMA expression (**Figure 2E-F**).
+AR-C155858, which is a high-affinity(*K~i~* 2 nM) inhibitor of both MCT1 and MCT2 [@ovens2010], also decreased α-SMA expression alone and in combination with VB124 (**Supplemental Figure 2B-C**).
+Pharmacologic MCT inhibition did not affect cell counts over 48 h of treatment (**Supplemental Figure 2D-E**).
+As anticipated, decreased Col1a1 and α-SMA expression was associated with decreased myofibroblast contractility as assessed by gel contraction assay (**Figure 2G-H**).
+Together, these data indicate that MCT expression and activity are required for myofibroblast differentiation *in vitro*.
+
+\newpage
+
+# Methods
+
+\newpage
+
+# Acknowledgements
+
+# Conflicts of Interest
+
+\newpage
+
+# References
+
+::: {#refs}
+:::
+
+\newpage
+
+# Figures
+
+![**Figure 1 - Lactate transporter expression increases in human IPF and experimental models.** (**A**) Immunoblot of whole lung homogenates from explanted IPF lungs and controls (Ctl).
+(**B**) Quantification of band densities from (A).
+(**C**) Immunoblot of whole lung homogenates from bleomycin (bleo)- and vehicle (Ctl)-treated mice.
+(**D**) Quantification of band densities from (C).
+(**E**) Representative immunoblot of cell lysates from normal human lung fibroblasts treated with TGFβ to induce myofibroblast differentiation.
+(**F**) Quantification of band densities from (E).
+Individual data points are biological replicates.
+Summary data are mean ± SEM (\* p < 0.05).](figs/Figure 01.png)

### 2024-01-22 R/blots.R

@@ -134,17 +134,26 @@ phospho_ratio <- function(df) {
     dplyr::filter(!is.na(.data$ratio))
 }
 
-combine_plots <- function(dfs, stats) {
+combine_plots <- function(dfs, stats, orient = "h") {
   df <- dplyr::bind_rows(dfs)
   stat <- dplyr::bind_rows(stats)
   prot <- unique(df$protein)
   df$protein <- factor(df$protein, levels = prot)
   stat$protein <- factor(stat$protein, levels = prot)
 
+  if (orient == "h") {
+    nrow <- 1
+    ncol <- NULL
+  } else if (orient == "v") {
+    nrow <- NULL
+    ncol <- 1
+  }
+
   plot_blot(df, stat) +
     ggplot2::facet_wrap(
       ggplot2::vars(protein),
-      nrow = 1,
+      nrow = nrow,
+      ncol = ncol,
       scales = "free_y"
     )
 }

### 2024-01-22 R/figures.R

@@ -62,3 +62,37 @@ make_fig01 <- function(p1, p2, p3, p4, p5, p6) {
       heights = ggplot2::unit(1, "in")
     )
 }
+
+make_fig02 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
+  design <- "
+  abcd
+  abcd
+  efgh
+  ef##
+  "
+
+  p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(c(1.5, 1.75, 1.5, 1.75), "in"),
+      heights = ggplot2::unit(c(0.75, 0.75, 1, 0.5), "in"),
+      guides = "collect"
+    ) &
+    ggplot2::theme(
+      legend.position = "bottom"
+    )
+}
+
+make_fig02s <- function(p1, p2, p3, p4, p5) {
+  design <- "a# \n bc \n de"
+  p1 + p2 + p3 + p4 + p5 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(c(1.5, 1.5), "in"),
+      heights = ggplot2::unit(1, "in"),
+      guides = "collect"
+    ) &
+    ggplot2::theme(
+      legend.position = "bottom"
+    )
+}

### 2024-01-22 R/plots.R

@@ -110,7 +110,7 @@ theme_plot <- function(
       half_line
     ),
     legend.key = ggplot2::element_rect(fill = "white", colour = NA),
-    legend.key.size = ggplot2::unit(0.8, "lines"),
+    legend.key.size = ggplot2::unit(0.6, "lines"),
     legend.key.height = NULL,
     legend.key.width = NULL,
     legend.text = ggplot2::element_text(size = ggplot2::rel(0.8)),
@@ -129,7 +129,7 @@ theme_plot <- function(
     panel.border = ggplot2::element_rect(fill = NA, color = "grey30"),
     panel.grid = ggplot2::element_blank(),
     panel.grid.minor = ggplot2::element_line(linewidth = ggplot2::rel(0.5)),
-    panel.spacing = ggplot2::unit(half_line, "pt"),
+    panel.spacing = ggplot2::unit(base_size, "pt"),
     panel.spacing.x = NULL,
     panel.spacing.y = NULL,
     panel.ontop    = FALSE,

### 2024-01-22 _targets.R

@@ -145,67 +145,103 @@ list(
   ),
   tar_map(
     values = list(
-      names = c(
-        "bleo_mct1",
-        "bleo_mct4",
-        "bleo_sma"
+      experiment = list(
+        "ipf",
+        "bleo",
+        "untreated",
+        "sirna",
+        "dual",
+        c("dual", "ar"),
+        "ipf-lf"
       ),
-      data = rlang::syms(c(
-        "blot_norm_bleo_mct1",
-        "blot_norm_bleo_mct4",
-        "blot_norm_bleo_sma"
-      ))
-    ),
-    names = names,
-    tar_target(
-      blot_norm_filter,
-      dplyr::filter(data, treatment == "Veh")
-    )
-  ),
-  tar_map(
-    values = list(
-      names = c(
-        "bleo_mct1",
-        "bleo_mct4",
-        "bleo_sma"
+      protein = list(
+        c("α-SMA", "MCT1", "MCT4"),
+        c("α-SMA", "MCT1", "MCT4"),
+        c("α-SMA", "MCT1", "MCT4"),
+        c("α-SMA", "MCT1", "MCT4"),
+        c("α-SMA", "Col1a1"),
+        "α-SMA",
+        c("α-SMA", "Col1a1")
       ),
-      data = rlang::syms(c(
-        "blot_stats_bleo_mct1",
-        "blot_stats_bleo_mct4",
-        "blot_stats_bleo_sma"
-      ))
-    ),
-    names = names,
-    tar_target(
-      blot_stats_filter,
-      dplyr::filter(data, treatment == "Veh" & is.na(condition)) |>
-        dplyr::mutate(x = 1.5)
-    )
-  ),
-  tar_map(
-    values = list(
-      names = c(
-        "mct_ipf",
-        "mct_bleo",
-        "mct_tgfb"
+      treatment = list(
+        NULL,
+        "Veh",
+        NULL,
+        NULL,
+        c("Veh", "AZD", "VB", "AZD/VB"),
+        c("Veh", "AR", "VB", "AR/VB"),
+        c("Veh", "AZD", "VB", "AZD/VB")
       ),
-      data = list(
-        rlang::syms(c("blot_norm_ipf_mct1", "blot_norm_ipf_mct4", "blot_norm_ipf_sma")),
-        rlang::syms(c("blot_norm_filter_bleo_mct1", "blot_norm_filter_bleo_mct4", "blot_norm_filter_bleo_sma")),
[…]

### 2024-01-23 manuscript/manuscript.qmd
revise figures 3

@@
     fig01,
-    # fig02,
+    fig02,
     # fig03,
@@
 
+## MCT inhibition reprograms myofibroblast metabolism
+
+MCTs are critical regulators of cellular lactate homeostasis.
+MCT1 is considered to be the main lactate importer in cells that use lactate to fuel oxidative phosphorylation or gluconeogenesis and is ubiquitously expressed.
+MCT1 also mediates lactate export in glycolytic cells [@halestrap2013; @brooks2018].
+MCT4 is considered to be the main lactate exporter in glycolytic cells and is up-regulated by hypoxia-inducible transcription factors [@ullah2006; @halestrap2013].
+MCT4 can also function as a lactate importer with a K~M~ ~1 mM [@contreras-baeza2019].
+
+To evaluate the metabolic consequences of lactate transporter inhibition, we measured extracellular lactate in the conditioned medium from cells treated with MCT siRNA or pharmacologic inhibitors (**Figure 3A-B**).
+Silencing MCT1 and MCT4, separately or together, decreased TGFβ-stimulated lactate efflux in lung fibroblasts (**Figure 3A**).
+Silencing both transporters was required to block TGFβ-stimulated lactate efflux completely.
+By contrast, pharmacologic inhibition of either MCT1 or MCT4 did not decrease TGFβ-stimulated lactate efflux (**Figure 3B**) and inhibition of both transporters was required to decrease extracellular lactate.
+Similarly, inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to attenuate lactate efflux fully (**Supplemental Figure 3A**).
+Compared to lactate production, the effect of MCT inhibitors on extracellular glucose was less pronounced, although we did observe a trend toward decreased glucose consumption by cells treated with both AZD3965 and VB124 (**Supplemental Figure 3B**).
+These findings are consistent with prior studies suggesting compensatory roles for MCT1 and MCT4 in lactate export [@benjamin2018].
+Moreover, these data also suggest that inhibition of glycolysis is not the mechanism by which lactate transport inhibition attenuates myofibroblast differentiation.
+
+To further interrogate the metabolic consequences of lactate transport inhibition, we next measured proton efflux (PER) and oxygen consumption (OCR) rates of lung fibroblasts treated with TGFβ in combination with MCT inhibitors (**Figure 3C, Supplemental Figure 3C**).
+Since MCTs co-transport protons with lactate, PER correlates with lactate efflux.
+Consistent with prior reports [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER and OCR following TGFβ stimulation for 48 h, consistent with increases in both glycolysis and oxidative phosphorylation.
+Similar to our direct measurements of extracellular lactate and glucose, simultaneous treatment with AZD3965 and VB124 was required to reduce PER.
+As anticipated, this decrease in PER was associated with a commensurate increase in OCR as myofibroblasts shifted their metabolism from glycolysis to oxidative phosphorylation.
+Unexpectedly, an increase in OCR was observed with AZD3965 and VB124 when administered individually.
+Indeed, the primary consequence of MCT1 or MCT4 inhibition alone was an increase in cellular ATP production rates driven by an increase in oxidative phosphorylation (**Figure 3D-E**).
+Consistent with a greater reliance on oxidative phosphorylation for ATP production, MCT inhibition decreased spare respiratory capacity, indicating that treated cells were respiring closer to their maximal oxidative capacity (**Supplemental Figure 3D**).
+No significant differences in glycolytic capacity or electron transport chain coupling efficiency were observed (**Supplemental Figure 3D**).
+Together, these data suggest that the primary metabolic consequence of MCT inhibition is the activation of oxidative phosphorylation rather than inhibition of glycolysis.
+
+To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based metabolomic profiling of cells treated with AZD3965 and VB124 to determine the impact of lactate transporter inhibition on intracellular metabolic pathways (**Figure 4**).
+Principal components (PC) analysis demonstrated excellent clustering of treatment groups with the drug effect corresponding to PC1 and the TGFβ treatment effect corresponding to PC2.
+Similar to the extracellular flux results described above, the magnitude of drug-induced perturbations were increased from MCT1 to MCT4 to combined inhibition (**Figure 4A**), as demonstrated by intracellular lactate levels (**Figure 4B**).
+As expected from measures of extracellular lactate following MCT inhibition, intracellular lactate accumulation increases modestly with MCT4 inhibition and markedly with MCT1/4 inhibition.
+TGFβ treatment was associated with diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Supplemental Figure 4A-B**).
+The effect of AZD3965 alone on intracellular metabolite levels was modest (**Supplemental Figure 4C-D**).
+By contrast, MCT4 inhibition by VB124 alone (**Supplemental Figure 4E-F**) or combined with AZD3965 (**Figure 4C-D**) caused substantial perturbations to intracellular metabolism.
+Specifically, we observed enrichment of the glycolysis and tricarboxylic acid (TCA) cycle metabolite sets with MCT4 inhibition.
+Together, these data indicate that inhibition of lactate export leads to accumulation of upstream glycolytic intermediates that are rerouted to mitochondrial oxidative metabolic pathways.
+
+To test this hypothesis, we next labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM)
+in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Figure 4E-F**).
+TGFβ increases ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), consistent with increased flux from glucose into the TCA cycle.
+While AZD3965 had little impact on these labeling patterns, MCT4 inhibition significantly increased the fractions of these metabolites labeled by glucose, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the bioenergetic and steady-state metabolomics experiments described above.
+
+Increased metabolite labeling by glucose must by offset by decreased labeling from other substrates.
+Since recent data suggest that lactate is a major oxidative fuel source in lung [@hui2017; @hui2020], we reasoned that MCT inhibition would decrease oxidation of exogenous lactate.
+To test this, lung fibroblasts were cultured with [U-^13^C~3~]-lactate (2 mM) in medium containing naturally labeled glucose, glutamine, and pyruvate (**Figure 4F**).
+Notably, extracellular [U-^13^C~3~]-lactate labeled ~ 50% of intracellular pyruvate and lactate at baseline with significant downstream incorporation into TCA metabolites.
+This labeling was decreased following TGFβ treatment, mirroring increased fractional labeling from glucose.
+MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Supplemental Figure 4G**).
+Interestingly, MCT4 inhibition alone or combined with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, emphasizing the importance of bi-directional lactate transport mediated by MCTs.
+Similar to our measurements of extracellular lactate, the effects of MCT inhibition were more pronounced when both inhibitors were used simultaneously, again demonstrating some functional redundancy of MCT1 and MCT4 in lung fibroblasts.
+
 \newpage
@@
 Summary data are mean ± SEM (\* p < 0.05).](figs/Figure 01.png)
+
+\newpage
+
+![**Figure 2 - Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Representative immunoblot of normal lung fibroblast cell lysates after treatment with siRNA targeting MCT1 or MCT4.
+(**B**) Quantification of protein expression levels following MCT1 and MCT4 knockdown.
+(**C**) Immunoblot of IPF lung fibroblast lysates from three separate donors after treatment with TGFβ and MCT inhibitors AZD3965 (AZD) or VB124 (VB).
+(**D**) Quantification of band densities from (C).
+(**E**) Representative immunoblot of normal lung fibroblasts treated with TGFβ and MCT inhibitors.
+(**F**) Quantification of α-SMA expression following MCT inhibition.
+(**G**) Representative image of a gel contraction assay after 24 h treatment with TGFβ and MCT inhibitors.
+(**H**) Quantification of contracted areas.
+Individual data points are biological replicates.
+Summary data are mean ± SEM (\* p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 02.png)

### 2024-01-23 R/seahorse.R

@@ -110,7 +110,13 @@ seahorse_group_onefactor <- function(x) {
 }
 
 plot_seahorse_summary <- function(x, stats) {
-  plot_one_factor(x, stats, x = "group", y = "value_corr", ytitle = "fmol/min/cell") +
+  plot_one_factor(
+    x,
+    stats,
+    x = "group",
+    y = "value_corr",
+    ytitle = "fmol/min/cell"
+  ) +
     ggplot2::facet_grid(
       cols = ggplot2::vars(.data$stage),
       rows = ggplot2::vars(.data$name),
@@ -200,8 +206,8 @@ plot_seahorse_atp_bars <- function(x, stats) {
         ymin = .data$ymin,
         ymax = .data$ymax
       ),
-      width = 0.25,
-      linewidth = 0.4
+      width = 0.2,
+      linewidth = 0.3,
     ) +
     ggbeeswarm::geom_quasirandom(
       data = pts,
@@ -212,7 +218,7 @@ plot_seahorse_atp_bars <- function(x, stats) {
       color = "black",
       width = 0.05,
       size = 0.75,
-      stroke = 0.3,
+      stroke = 0.2,
       show.legend = FALSE
     ) +
     ggplot2::geom_text(
@@ -284,12 +290,17 @@ plot_seahorse_pheno <- function(x) {
       color = "white",
       show.legend = FALSE
     ) +
-    ggplot2::geom_text(
-      ggplot2::aes(label = .data$group),
-      size = 1.75,
+    ggrepel::geom_text_repel(
+      ggplot2::aes(
+        label = .data$group
+      ),
+      point.padding = 6,
+      force_pull = 1,
+      lineheight = 1,
       nudge_x = 0.5,
-      nudge_y = -0.6,
+      nudge_y = -0.5,
       color = "black",
+      size = 1.75,
       show.legend = FALSE
     ) +
     ggplot2::scale_fill_manual(

### 2024-01-23 R/stats.R

@@ -13,7 +13,7 @@ annot_stats <- function(stats) {
   stats |>
     dplyr::mutate(
       y = Inf,
-      vjust = 1.5,
+      vjust = 1.2,
       label = annot_p(.data[[col]])
     )
 }

### 2024-01-23 _targets.R

@@ -182,9 +182,9 @@ list(
         NULL
       ),
       comp = list(
-        TRUE,
-        TRUE,
-        TRUE,
+        FALSE,
+        FALSE,
+        FALSE,
         rlang::sym("comps_sirna"),
         rlang::sym("comps_azd_2"),
         rlang::sym("comps_ar_2"),
@@ -1390,7 +1390,33 @@ list(
       pg_plot_azd,
       pg_plot_ar
     ) |>
-      write_figures("Figure 02_supplement"),
+      write_figures("Figure 02.supplement"),
+    format = "file"
+  ),
+
+  # figure 3 ----------------------------------------------------------------
+
+  tar_target(
+    fig03,
+    make_fig03(
+      plate_plot_lactate_sirna,
+      plate_plot_lactate_azd,
+      seahorse_mcti_timeline_plot,
+      seahorse_mcti_atp_bars,
+      seahorse_mcti_atp_pheno
+    ) |>
+      write_figures("Figure 03"),
+    format = "file"
+  ),
+  tar_target(
+    fig03s,
+    make_fig03s(
+      plate_plot_lactate_ar,
+      plate_plot_glucose_azd,
+      seahorse_mcti_summary_plot,
+      seahorse_mcti_stress_plot
+    ) |>
+      write_figures("Figure 03.supplement"),
     format = "file"
   ),
 

### 2024-01-24 manuscript/manuscript.qmd
start working on figure 4

@@
 Silencing both transporters was required to block TGFβ-stimulated lactate efflux completely.
-By contrast, pharmacologic inhibition of either MCT1 or MCT4 did not decrease TGFβ-stimulated lactate efflux (**Figure 3B**) and inhibition of both transporters was required to decrease extracellular lactate.
-Similarly, inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to attenuate lactate efflux fully (**Supplemental Figure 3A**).
+By contrast, pharmacologic inhibition of either MCT1 or MCT4 did not decrease TGFβ-stimulated lactate efflux (**Figure 3B**), and inhibition of both transporters was required to decrease extracellular lactate.
+Similarly, inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to prevent increases in lactate efflux (**Supplemental Figure 3A**).
 Compared to lactate production, the effect of MCT inhibitors on extracellular glucose was less pronounced, although we did observe a trend toward decreased glucose consumption by cells treated with both AZD3965 and VB124 (**Supplemental Figure 3B**).
@@
 Since MCTs co-transport protons with lactate, PER correlates with lactate efflux.
-Consistent with prior reports [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER and OCR following TGFβ stimulation for 48 h, consistent with increases in both glycolysis and oxidative phosphorylation.
+Consistent with prior reports [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER and OCR following TGFβ stimulation for 48 h, indicating increases in both glycolysis and oxidative phosphorylation.
 Similar to our direct measurements of extracellular lactate and glucose, simultaneous treatment with AZD3965 and VB124 was required to reduce PER.
@@
 Indeed, the primary consequence of MCT1 or MCT4 inhibition alone was an increase in cellular ATP production rates driven by an increase in oxidative phosphorylation (**Figure 3D-E**).
-Consistent with a greater reliance on oxidative phosphorylation for ATP production, MCT inhibition decreased spare respiratory capacity, indicating that treated cells were respiring closer to their maximal oxidative capacity (**Supplemental Figure 3D**).
+MCT inhibition decreased spare respiratory capacity, indicating that the basal respiratory rate of treated cells was closer to their maximal oxidative capacity (**Supplemental Figure 3D**).
 No significant differences in glycolytic capacity or electron transport chain coupling efficiency were observed (**Supplemental Figure 3D**).
@@
 
-To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based metabolomic profiling of cells treated with AZD3965 and VB124 to determine the impact of lactate transporter inhibition on intracellular metabolic pathways (**Figure 4**).
+### YOU ARE HERE ###
+
+To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of intracellular and extracellular metabolites from cells treated with AZD3965 and VB124 to determine the impact of lactate transporter inhibition on metabolic pathways.
+
 Principal components (PC) analysis demonstrated excellent clustering of treatment groups with the drug effect corresponding to PC1 and the TGFβ treatment effect corresponding to PC2.

### 2024-01-24 R/figures.R

@@ -127,3 +127,25 @@ make_fig03s <- function(p1, p2, p3, p4) {
       heights = ggplot2::unit(c(1.5, 2, 1.5), "in")
     )
 }
+
+make_fig04 <- function(p1, p2, p3, p4) {
+  design <- "abcd"
+  p1 + p2 + p3 + p4 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(2, "in"),
+      heights = ggplot2::unit(c(2, 2.5), "in")
+    ) &
+    ggplot2::theme(legend.position = "bottom")
+}
+
+make_fig04s <- function(p1, p2, p3, p4) {
+  design <- "abcd"
+  p1 + p2 + p3 + p4 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(2, "in"),
+      heights = ggplot2::unit(c(2, 2.5), "in")
+    ) &
+    ggplot2::theme(legend.position = "bottom")
+}

### 2024-01-24 R/metab.R

@@ -433,8 +433,8 @@ plot_mois <- function(se, tt, moi){
       position = ggplot2::position_dodge(width = 0.55),
       stat = "summary",
       fun.data = ggplot2::mean_se,
-      width = 0.25,
-      linewidth = 0.25,
+      width = 0.2,
+      linewidth = 0.3,
       show.legend = FALSE
     ) +
     ggbeeswarm::geom_quasirandom(
@@ -443,8 +443,8 @@ plot_mois <- function(se, tt, moi){
       pch = 1,
       color = "black",
       width = 0.05,
-      size = 0.5,
-      stroke = 0.25,
+      size = 0.75,
+      stroke = 0.2,
       show.legend = FALSE
     ) +
     ggplot2::geom_text(

### 2024-01-24 R/mids.R

@@ -271,89 +271,89 @@ format_mid_mcti <- function(files, correction_factors){
     dplyr::group_by(dplyr::across("run_date":"metabolite"))
 }
 
-# select_mid_mois <- function(df, trace) {
-#   metabs <- c("PYR", "LAC", "CIT", "2OG", "SUC", "MAL")
-#
-#   df |>
-#     dplyr::filter(.data$tracer == trace) |>
-#     dplyr::filter(.data$metabolite %in% metabs) |>
-#     dplyr::filter(.data$isotope == "M0") |>
-#     dplyr::group_by(.data$metabolite) |>
-#     dplyr::mutate(grand_mean = mean(.data$mid_corr)) |>
-#     dplyr::group_by(.data$metabolite, .data$replicate) |>
-#     dplyr::mutate(
-#       exp_mean = mean(.data$mid_corr),
-#       cf = grand_mean - exp_mean,
-#       # labeled = 1 - .data$mid_corr + .data$cf,
-#       labeled = 1 - .data$mid_corr,
-#       metabolite = factor(metabolite, levels = metabs)
-#     ) |>
-#     dplyr::select(
-#       "sample",
-#       group = "replicate",
-#       "condition":"metabolite",
-#       "area",
-#       "mid_corr",
-#       "labeled"
-#     ) |>
-#     dplyr::group_by(.data$metabolite)
-# }
-#
-#
-# format_bleo_mids <- function(files, mice) {
-#   nms <- stringr::str_extract(basename(files), "lung|plasma")
-#   files <- rlang::set_names(files, nms)
-#
-#   sample <- as.character(1:17)
-#   start_date <- lubridate::as_date(rep("2022-08-05", 17))
-#   mouse_no <- c(1:6, 8, 10:11, 13:18, 23, 24)
-#
-#   samples <-
-#     tibble::tibble(
-#       sample = sample,
-#       start_date = start_date,
-#       mouse_no = mouse_no
-#     ) |>
-#     dplyr::left_join(mice, by = c("start_date", "mouse_no")) |>
-#     dplyr::mutate(
-#       group = factor(
-#         group,
-#         levels = c("Veh", "Bleo", "AZD", "VB"),
-#         labels = c("Ctl", "Bleo", "AZD", "VB")
-#       )
-#     )
-#
-#   purrr::map_dfr(
-#     files,
-#     readxl::read_excel,
-#     sheet = 2,
-#     .id = "tissue"
-#   ) |>
[…]

### 2024-01-24 _targets.R

@@ -15,7 +15,8 @@ tar_option_set(
     "tidyverse",
     "patchwork",
     "scales",
-    "SummarizedExperiment"
+    "SummarizedExperiment",
+    "extrafont"
   ),
   format = "qs",
   controller = crew::crew_controller_local(workers = 4)
@@ -728,7 +729,7 @@ list(
     ),
     tar_target(
       metab_mcti_tar_pca_filter,
-      plot_pca_metab(metab_mcti_tar_treated),
+      plot_pca_metab(metab_mcti_tar_treated, show_reps = FALSE),
       format = "rds"
     ),
     tar_target(
@@ -983,59 +984,59 @@ list(
     mid_mcti_mids_corr,
     correct_mids(mid_mcti_mids, correction_matrices)
   ),
-  # tar_map(
-  #   values = list(
-  #     label = c("glc2", "gln5", "lac3", "glc6")
-  #   ),
-  #   names = label,
-  #   tar_target(
-  #     mid_mcti_plot,
-  #     plot_mids_all(mid_mcti_mids_corr, label = label)
-  #   ),
-  #   tar_target(
-  #     mid_mcti_img,
-  #     write_plot(mid_mcti_plot, label, path = "analysis/figures/mids"),
-  #     format = "file"
-  #   ),
-  #   NULL
-  # ),
-  # tar_map(
-  #   values = list(
-  #     label = c("gln5", "lac3", "glc6"),
-  #     title = list(
-  #       expression(paste("[U-"^13, "C"[5], "]-glutamine")),
-  #       expression(paste("[U-"^13, "C"[3], "]-lactate")),
-  #       expression(paste("[U-"^13, "C"[6], "]-glucose"))
-  #     )
-  #   ),
-  #   names = label,
-  #   tar_target(
-  #     mid_mcti_moi_data,
-  #     select_mid_mois(mid_mcti_mids_corr, label)
-  #   ),
-  #   tar_target(
-  #     mid_mcti_moi_stats,
-  #     group_twofactor(mid_mcti_moi_data, "labeled", comps_azd_2)
-  #   ),
-  #   tar_target(
-  #     mid_mcti_moi_plot,
-  #     plot_two_factor(
-  #       mid_mcti_moi_data,
-  #       mid_mcti_moi_stats,
-  #       x = "treatment",
-  #       y = "labeled",
-  #       ytitle = "Labeled fraction",
-  #       wrap = "metabolite",
-  #       title = title
-  #     ) +
-  #       ggplot2::scale_y_continuous(
[…]

### 2024-01-27 manuscript/manuscript.qmd
adjust titles on Figure 4

@@
 Since MCTs co-transport protons with lactate, PER correlates with lactate efflux.
-Consistent with prior reports [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER and OCR following TGFβ stimulation for 48 h, indicating increases in both glycolysis and oxidative phosphorylation.
+Consistent with prior reports [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER CR following TGFβ stimulation for 48 h, indicating increases in both glycolysis and oxidative phosphorylation.
 Similar to our direct measurements of extracellular lactate and glucose, simultaneous treatment with AZD3965 and VB124 was required to reduce PER.
@@
 
-### YOU ARE HERE ###
+To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of extracellular and intracellular metabolites from cells treated with AZD3965 and VB124.
+As suggested by the Seahorse analysis, inhibiting a single lactate transporter had modest effects on extracellular metabolite levels (**Figure 4A-D, Supplemental Figure 4**).
+This analysis confirmed the results of extracellular lactate measurements by enzyme assay, which showed dual inhibition was required to decrease lactate efflux (**Figure 4B**).
+In addition to lactate, dual lactate transporter inhibition also had broader impacts on amino acid fluxes (**Figure 4C-D**).
+Several metabolites were differentially regulated by TGFβ treatment (**Supplemental Figure 4A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, ketoleucine.
 
-To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of intracellular and extracellular metabolites from cells treated with AZD3965 and VB124 to determine the impact of lactate transporter inhibition on metabolic pathways.
+- intracellular metabolism correlated with extracellular measurements by MSEA
 

### 2024-01-27 R/figures.R

@@ -136,7 +136,9 @@ make_fig04 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
       title = "Extracellular",
       theme = ggplot2::theme(
         plot.title = ggplot2::element_text(
-          hjust = 0.5
+          hjust = 0.5,
+          face = "bold",
+          size = ggplot2::rel(1)
         )
       )
     ) +
@@ -153,7 +155,9 @@ make_fig04 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
       title = "Intracellular",
       theme = ggplot2::theme(
         plot.title = ggplot2::element_text(
-          hjust = 0.5
+          hjust = 0.5,
+          face = "bold",
+          size = ggplot2::rel(1)
         )
       )
     ) +
@@ -181,7 +185,13 @@ make_fig04s <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12) {
     p1 + p2 + p5 + p6 + p9 + p10 +
     patchwork::plot_annotation(
       title = "Extracellular",
-      theme = ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
     ) +
     theme_patchwork(
       design = design,
@@ -191,7 +201,13 @@ make_fig04s <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12) {
     p3 + p4 + p7 + p8 + p11 + p12 +
     patchwork::plot_annotation(
       title = "Intracellular",
-      theme = ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
     ) +
     theme_patchwork(
       design = design,

### 2024-01-27 manuscript/manuscript.qmd
add mid ratio calculations and plots

@@
 In addition to lactate, dual lactate transporter inhibition also had broader impacts on amino acid fluxes (**Figure 4C-D**).
-Several metabolites were differentially regulated by TGFβ treatment (**Supplemental Figure 4A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, ketoleucine.
+Several metabolites were differentially regulated by TGFβ treatment (**Supplemental Figure 4A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, and ketoleucine.
 
-- intracellular metabolism correlated with extracellular measurements by MSEA
-
+Similar effects on intracellular metabolites were observed.
 Principal components (PC) analysis demonstrated excellent clustering of treatment groups with the drug effect corresponding to PC1 and the TGFβ treatment effect corresponding to PC2.
-Similar to the extracellular flux results described above, the magnitude of drug-induced perturbations were increased from MCT1 to MCT4 to combined inhibition (**Figure 4A**), as demonstrated by intracellular lactate levels (**Figure 4B**).
+Similar to the extracellular flux results, the magnitude of drug-induced perturbations were increased from MCT1 to MCT4 to combined inhibition (**Figure 4E**), as demonstrated by intracellular lactate levels (**Figure 4F**).
 As expected from measures of extracellular lactate following MCT inhibition, intracellular lactate accumulation increases modestly with MCT4 inhibition and markedly with MCT1/4 inhibition.
-TGFβ treatment was associated with diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Supplemental Figure 4A-B**).
-The effect of AZD3965 alone on intracellular metabolite levels was modest (**Supplemental Figure 4C-D**).
-By contrast, MCT4 inhibition by VB124 alone (**Supplemental Figure 4E-F**) or combined with AZD3965 (**Figure 4C-D**) caused substantial perturbations to intracellular metabolism.
+TGFβ treatment was associated with diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Supplemental Figure 4C-D**).
+The effect of AZD3965 alone on intracellular metabolite levels was modest (**Supplemental Figure 4G-H**).
+By contrast, MCT4 inhibition by VB124 alone (**Supplemental Figure 4K-L**) or combined with AZD3965 (**Figure 4G-H**) caused substantial perturbations to intracellular metabolism.
 Specifically, we observed enrichment of the glycolysis and tricarboxylic acid (TCA) cycle metabolite sets with MCT4 inhibition.
@@
 
-To test this hypothesis, we next labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM)
-in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Figure 4E-F**).
+To test this hypothesis, we next labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM) in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Figure 4E-F**).
 TGFβ increases ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), consistent with increased flux from glucose into the TCA cycle.

### 2024-01-27 R/mids.R

@@ -279,12 +279,7 @@ select_mid_mois <- function(df, trace) {
     dplyr::filter(.data$metabolite %in% metabs) |>
     dplyr::filter(.data$isotope == "M0") |>
     dplyr::group_by(.data$metabolite) |>
-    dplyr::mutate(grand_mean = mean(.data$mid_corr)) |>
-    dplyr::group_by(.data$metabolite, .data$replicate) |>
     dplyr::mutate(
-      exp_mean = mean(.data$mid_corr),
-      cf = grand_mean - exp_mean,
-      # labeled = 1 - .data$mid_corr + .data$cf,
       labeled = 1 - .data$mid_corr,
       metabolite = factor(metabolite, levels = metabs)
     ) |>
@@ -356,4 +351,32 @@ format_bleo_mids <- function(files, mice) {
     dplyr::filter(!is.na(mid)) |>
     dplyr::arrange(metabolite, group, sample) |>
     dplyr::group_by(dplyr::across("tissue":"metabolite"))
+}
+
+calc_mid_ratio <- function(df, top, bottom) {
+  tibble::as_tibble(top) |>
+    dplyr::mutate(ratio = "top") |>
+    dplyr::bind_rows(
+      tibble::as_tibble(bottom) |>
+        dplyr::mutate(ratio = "bottom")
+    ) |>
+    dplyr::left_join(df, by = c("tracer", "metabolite", "isotope")) |>
+    tidyr::unite(metabolite, metabolite, isotope, sep = " ") |>
+    dplyr::group_by(tracer, sample, replicate, condition, treatment) |>
+    dplyr::summarise(
+      metabolite = stringr::str_c(
+        metabolite[ratio == "top"], " / ", metabolite[ratio == "bottom"]
+      ),
+      value = mid_corr[ratio == "top"] / mid_corr[ratio == "bottom"]
+    ) |>
+    dplyr::select(
+      tracer,
+      metabolite,
+      sample,
+      group = replicate,
+      condition,
+      treatment,
+      value
+    ) |>
+    dplyr::ungroup()
 }

### 2024-01-29 manuscript/manuscript.qmd
add MIDs to Fig 5 and start Fig 6

@@
     fig02,
-    # fig03,
-    # fig04,
-    # fig05,
+    fig03,
+    fig04,
+    fig05,
     # fig06,
@@
 )
+```
+
+```{r}
+#| label: calculations
+#| include: false
+
+targets::tar_load(
+  c(
+    deg_tgfb_dual
+  )
+)
+
+deg_tot <- nrow(deg_tgfb_dual)
+deg_diff <- nrow(dplyr::filter(deg_tgfb_dual, padj < 0.05)) 
+deg_frac <- sprintf("%1.0f", deg_diff / deg_tot * 100, "%")
+deg_tot <- scales::label_comma(accuracy = 1)(deg_tot)
 ```
@@
 
-## Lactate transport is essential for myofibroblast differentiation *in vitro*
+## Myofibroblast differentiation *in vitro* requires lactate transport
 
@@
 
-To test this hypothesis, we next labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM) in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Figure 4E-F**).
+To test this hypothesis, we next labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM) in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Figure 4I**).
 TGFβ increases ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), consistent with increased flux from glucose into the TCA cycle.
@@
 Increased metabolite labeling by glucose must by offset by decreased labeling from other substrates.
-Since recent data suggest that lactate is a major oxidative fuel source in lung [@hui2017; @hui2020], we reasoned that MCT inhibition would decrease oxidation of exogenous lactate.
-To test this, lung fibroblasts were cultured with [U-^13^C~3~]-lactate (2 mM) in medium containing naturally labeled glucose, glutamine, and pyruvate (**Figure 4F**).
-Notably, extracellular [U-^13^C~3~]-lactate labeled ~ 50% of intracellular pyruvate and lactate at baseline with significant downstream incorporation into TCA metabolites.
+Since recent data suggest that lactate is a major oxidative fuel source in the lung [@hui2017; @hui2020; @copeland2023], we reasoned that MCT inhibition would decrease oxidation of exogenous lactate.
+To test this, lung fibroblasts were cultured with [U-^13^C~3~]-lactate (2 mM) in medium containing naturally labeled glucose, glutamine, and pyruvate (**Figure 4J**).
+Extracellular [U-^13^C~3~]-lactate labeled ~ 50% of intracellular pyruvate and lactate at baseline with significant downstream incorporation into TCA metabolites.
 This labeling was decreased following TGFβ treatment, mirroring increased fractional labeling from glucose.
-MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Supplemental Figure 4G**).
-Interestingly, MCT4 inhibition alone or combined with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, emphasizing the importance of bi-directional lactate transport mediated by MCTs.
+MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Supplemental Figure 4M**).
+Interestingly, MCT4 inhibition alone or with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, emphasizing the importance of bi-directional lactate transport mediated by MCTs.
 Similar to our measurements of extracellular lactate, the effects of MCT inhibition were more pronounced when both inhibitors were used simultaneously, again demonstrating some functional redundancy of MCT1 and MCT4 in lung fibroblasts.
+
+## MCT inhibition attenuates pro-fibrotic transcriptional programs
+
+To identify anti-fibrotic transcriptional programs mediated by MCT inhibition, we performed RNA-seq on lung fibroblasts treated with TGFβ in combination with AZD3965 or VB124 (**Figure 5**).
+Principal components analysis (PCA) of the transcriptional changes mirrored the metabolomic findings.
+The first principal component corresponded to TGFβ treatment, accounting for 71% of the overall variance, while MCT4 inhibition, either alone or in combination with MCT1 inhibition aligned with the second principal component (**Figure 5A**).
+Samples treated with the MCT1 inhibitor AZD3965 were similar to vehicle-treated controls.
+Differential expression analysis of TGFβ-treated cells identified the expected increased expression of extracellular matrix proteins (**Supplemental Figure 5A**) and enrichment of the epithelial-to-mesenchymal (EMT) gene set, among others (**Figure 5B**).
+Consistent with the PCA results, only GRIK4 and BRI3 were differentially expressed following AZD3965 (**Supplemental Figure 5B**).
+By contrast, VB124, alone (**Supplemental Figure 5C**) or in combination with AZD3965 (**Figure 5C**), demonstrated more substantial reprogramming of myofibroblast transcription with `r deg_diff` of `r deg_tot` genes (`r deg_frac`%) differentially expressed at an FDR < 0.05.
+
+Co-administration of either AZD3965 or VB124 reversed TGFβ-dependent enrichment of the EMT gene set, which is the Hallmark gene set containing genes related to fibrosis (**Figure 5B**).
+Leading edge analysis of the EMT gene set identified seven genes shared among all three comparisons (*i.e.*, genes increased by TGFβ and decreased by both AZD3965 and VB124) (**Figure 5D**).
+These genes include: biglycan (BGN), COL6A3, Frizzled 8 (FZD8), matrix Gla protein (MGP), Prostate Transmembrane Protein, Androgen Induced 1 (PMEPA1), TIMP metallopeptidase inhibitor 1 (TIMP1), and tenascin-C (TNC).
+Many of these genes contribute to the pathobiology of pulmonary fibrosis or serve as biomarkers of disease or treatment response [@kwapiszewska2018; @williams2020; @baarsma2017; @todd2020; @bhattacharyya2016].
+Together, these data suggest lactate transport inhibitors attenuate the pro-fibrotic transcriptional program in TGFβ-treated LFs.
+
+## Lactate transport inhibition signals through histone lactylation
+
+We next sought to identify the signaling mechanism(s) linking lactate transport inhibition to anti-fibrotic transcriptional programs.
+Since increasing intracellular lactate was associated with more potent inhibition of α-SMA expression, we first treated cells with TGFβ in combination with extracellular lactate (10 mM).
+A prior studies demonstrated that extracellular lactate modestly increased α-SMA expression in the absence of TGFβ [@kottman2012].
+We found that extracellular lactate did not impact α-SMA expression in TGFβ-treated LFs (**Figure 6A-B**), suggesting that accumulation of intracellular lactate alone does not mediate the anti-fibrotic effects of MCT inhibition.
+
+Next, we examined TGFβ-dependent signaling.
+TGFβ activates both SMAD and non-SMAD signaling pathways.
+MCT inhibition did not decrease Smad3 or ERK phosphorylation after 48 h of TGFβ stimulation (**Figure 6C-D**).
+
+## Lactate transport inhibition / redox stress section
+
+We next examined whether MCT inhibition affects TGFβ-dependent signaling pathways.
+TGFβ activates both SMAD and non-SMAD signaling pathways.
+We found no impact of MCT inhibition on decreasing Smad3 or ERK phosphorylation in LFs 48 h following TGFβ stimulation (**Figure 5**).
+
+Hypoxia-inducible factor 1α (HIF-1α) is activated in lung fibroblasts following TGFβ and is required for myofibroblast differentiation and experimental pulmonary fibrosis partly through activation of glycolysis [@goodwin2018].
+We found maximal HIF-1α stabilization 6 h following TGFβ treatment.
+At this time point, MCT inhibition did not decrease HIF-1α (**Figure 5E-F**), rather, HIF-1α protein levels were increased with combined MCT1 and MCT4 inhibition.
+
+\newpage
+
+# Discussion
+
+Craig Thompson paper: Lactate activates ETC independently of its metabolism
+
+Craig Thompson paper: Isotope labeling in lung fibroblasts
 

### 2024-01-29 R/figures.R

@@ -128,7 +128,7 @@ make_fig03s <- function(p1, p2, p3, p4) {
     )
 }
 
-make_fig04 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
+make_fig04 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10) {
   design <- "abcd"
   extra <-
     p1 + p2 + p3 + p4 +
@@ -168,17 +168,27 @@ make_fig04 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
       heights = ggplot2::unit(2, "in")
     )
 
+  mids <-
+    p9 + p10 +
+    patchwork::plot_layout(guides = "collect") +
+    theme_patchwork(
+      design = "ab",
+      tags = list(c("I", "J"))
+    ) &
+    ggplot2::theme(legend.position = "bottom")
+
   patchwork::wrap_elements(full = extra) +
     patchwork::wrap_elements(full = intra) +
+    patchwork::wrap_elements(full = mids) +
     theme_patchwork(
-      design = "a \n b",
+      design = "a \n b \n c",
       widths = ggplot2::unit(9.5, "in"),
       heights = ggplot2::unit(3.5, "in"),
       tags = NULL
     )
 }
 
-make_fig04s <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12) {
+make_fig04s <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12, p13) {
   design <- "ab \n cd \n ef"
 
   extra <-
@@ -213,13 +223,53 @@ make_fig04s <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12) {
       design = design,
       tags = list(c("C", "D", "G", "H", "K", "L"))
     )
-
-  patchwork::wrap_elements(full = extra) + patchwork::wrap_elements(full = intra) +
+  mids <-
+    p13 +
     theme_patchwork(
-      design = "ab",
-      widths = ggplot2::unit(4, "in"),
-      heights = ggplot2::unit(6, "in"),
-      tags = NULL
+      tags = list("M")
     ) &
     ggplot2::theme(legend.position = "bottom")
+
+  patchwork::wrap_elements(full = extra) +
+    patchwork::wrap_elements(full = intra) +
+    patchwork::wrap_elements(full = mids) +
+    theme_patchwork(
[…]

### 2024-01-29 R/rnaseq.R

@@ -361,7 +361,7 @@ plot_edge <- function(x, dds) {
       color = NULL,
       fill = NULL
     ) +
-    ggplot2::guides(color = ggplot2::guide_legend(ncol = 1, byrow = TRUE)) +
+    ggplot2::guides(color = ggplot2::guide_legend(ncol = 2, byrow = TRUE)) +
     ggplot2::coord_cartesian(clip = "off") +
     theme_plot() +
     ggplot2::theme(

### 2024-01-29 _targets.R

@@ -1041,16 +1041,30 @@ list(
   ),
   tar_map(
     values = list(
-      names = c("lac_cit_pyr", "lac_lac_pyr"),
+      names = c(
+        "glc_cit_m2_pyr",
+        "glc_cit_m3_pyr",
+        "glc_cit_m4_pyr",
+        "glc_lac_pyr"
+      ),
       top = alist(
-        list(tracer = "lac3", metabolite = "CIT", isotope = "M4"),
-        list(tracer = "lac3", metabolite = "LAC", isotope = "M3")
+        list(tracer = "glc6", metabolite = "CIT", isotope = "M2"),
+        list(tracer = "glc6", metabolite = "CIT", isotope = "M3"),
+        list(tracer = "glc6", metabolite = "CIT", isotope = "M4"),
+        list(tracer = "glc6", metabolite = "LAC", isotope = "M3")
       ),
       bottom = alist(
-        list(tracer = "lac3", metabolite = "PYR", isotope = "M3"),
-        list(tracer = "lac3", metabolite = "PYR", isotope = "M3")
+        list(tracer = "glc6", metabolite = "PYR", isotope = "M3"),
+        list(tracer = "glc6", metabolite = "PYR", isotope = "M3"),
+        list(tracer = "glc6", metabolite = "PYR", isotope = "M3"),
+        list(tracer = "glc6", metabolite = "PYR", isotope = "M3")
       ),
-      ytitle = c("CIT M2 / PYR M3", "LAC M3 / PYR M3")
+      ytitle = c(
+        "CIT M2 / PYR M3",
+        "CIT M3 / PYR M3",
+        "CIT M4 / PYR M3",
+        "LAC M3 / PYR M3"
+      )
     ),
     names = names,
     tar_target(
@@ -1373,7 +1387,8 @@ list(
       "bleo-timeline.png",          1,      0,      0,      "timeline",
       "dual-azd-smad3-blot.png",    1,      0,      -0.05,  "smad3_azd",
       "dual-azd-erk-blot.png",      1,      0,      -0.05,  "erk_azd",
-      "dual-azd-hif1a-blot.png",    1.3,    0,      -0.05,  "hif_azd"
+      "dual-azd-hif1a-blot.png",    1.3,    0,      -0.05,  "hif_azd",
+      "lactate-sma-blot.png",       1,      0,      0,      "sma_lac"
     ),
     names = names,
     tar_target(
@@ -1477,7 +1492,9 @@ list(
[…]

### 2024-01-29 manuscript/manuscript.qmd
combine phospho protein ratio analysis

@@
     fig05,
-    # fig06,
+    fig06,
     # fig10,
@@
   c(
-    deg_tgfb_dual
+    deg_dual
   )
@@
 
-deg_tot <- nrow(deg_tgfb_dual)
-deg_diff <- nrow(dplyr::filter(deg_tgfb_dual, padj < 0.05)) 
+deg_tot <- nrow(deg_dual)
+deg_diff <- nrow(dplyr::filter(deg_dual, padj < 0.05)) 
 deg_frac <- sprintf("%1.0f", deg_diff / deg_tot * 100, "%")
@@
 Since increasing intracellular lactate was associated with more potent inhibition of α-SMA expression, we first treated cells with TGFβ in combination with extracellular lactate (10 mM).
-A prior studies demonstrated that extracellular lactate modestly increased α-SMA expression in the absence of TGFβ [@kottman2012].
+A prior study demonstrated that extracellular lactate modestly increased α-SMA expression in the absence of TGFβ [@kottmann2012].
 We found that extracellular lactate did not impact α-SMA expression in TGFβ-treated LFs (**Figure 6A-B**), suggesting that accumulation of intracellular lactate alone does not mediate the anti-fibrotic effects of MCT inhibition.
 
+Prior work has also linked inhibition of glycolysis and lactate production to hypoxia-inducible factor 1α activation (HIF-1α), which contributes to myofibroblast differentiation [@kottmann2012; @goodwin2018].
+We found maximal HIF-1α activation 6 h following TGFβ stimulation.
+At this time point, HIF-1α protein levels were similar across vehicle-, AZD3965-, and VB124-treated cells (**Figure 6C-D**).
+In LFs treated with dual inhibitors, HIF-1α protein was increased compared to vehicle, consistent with enrichment of the "Hypoxia" gene set in our RNA-seq data (**Figure 5B**).
+These findings, in light of prior studies, indicate that lactate transport inhibition acts downstream of HIF-1α-dependent transcriptional programs to inhibit myofibroblast differentiation.
+
 Next, we examined TGFβ-dependent signaling.
-TGFβ activates both SMAD and non-SMAD signaling pathways.
-MCT inhibition did not decrease Smad3 or ERK phosphorylation after 48 h of TGFβ stimulation (**Figure 6C-D**).
+TGFβ activates both SMAD and non-SMAD signaling pathways through a cascade of protein phosphorylation events.
+MCT inhibitors did not decrease Smad3 or ERK phosphorylation after 48 h of TGFβ stimulation (**Figure 6E-H**).
+
+Recently,
 

### 2024-01-29 R/blots.R

@@ -44,8 +44,10 @@ filter_blots <- function(df) {
     dplyr::filter(!(.data$experiment == "ipf" & .data$batch == 2)) |>
     dplyr::filter(!(.data$experiment == "ipf-lf" & .data$batch %in% 1:2)) |>
     dplyr::filter(!(.data$experiment == "dual" & .data$protein == "Col1a1" & .data$batch %in% 1:4)) |>
-    dplyr::filter(!(.data$experiment == "dual" & .data$protein %in% prots & .data$batch %in% 1:4)) |>
-    dplyr::filter(!(.data$experiment == "dual" & .data$protein %in% c("pAKT", "AKT") & .data$trial == 2))
+    # dplyr::filter(!(.data$experiment == "dual" & .data$protein %in% prots & .data$batch %in% 1:4)) |>
+    dplyr::filter(!(.data$experiment == "dual" & .data$protein %in% c("pAKT", "AKT") & .data$batch == 5 & .data$trial == 2)) |>
+    dplyr::filter(!(.data$experiment == "dual" & .data$protein %in% c("pERK", "ERK") & .data$batch == 3 & .data$trial %in% 3:4)) |>
+    dplyr::filter(!(.data$experiment == "six" & .data$protein == "HIF-1α" & .data$trial == 5))
 }
 
 index_blots <- function(df) {
@@ -81,7 +83,8 @@ norm_blot <- function(df, exp, prot, treat = NULL, cond = NULL) {
             .data$treatment == min(.data$treatment)
         ],
         na.rm = TRUE
-      )
+      ),
+      protein = factor(.data$protein, levels = prot)
     )
 }
 
@@ -119,10 +122,11 @@ phospho_ratio <- function(df) {
     dplyr::mutate(
       phospho = ifelse(
         stringr::str_detect(.data$protein, "^p\\D"),
-        "phospho",
-        "total"
+        "density",
+        "background"
       ),
       protein = stringr::str_replace(.data$protein, "^p", ""),
+      protein = stringr::str_c("p", .data$protein, "/", .data$protein),
       .after = "dose"
     ) |>
     dplyr::select("group", "experiment":"protein", "ratio") |>
@@ -130,30 +134,6 @@ phospho_ratio <- function(df) {
       names_from = "phospho",
       values_from = "ratio"
     ) |>
-    dplyr::mutate(ratio = .data$phospho / .data$total) |>
[…]

### 2024-01-29 R/figures.R

@@ -264,12 +264,14 @@ make_fig05s <- function(p1, p2, p3) {
     )
 }
 
-make_fig06 <- function(p1, p2) {
-  design <- "ab"
-  p1 + p2 +
+make_fig06 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
+  design <- "abcd \n efgh"
+  p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 +
     theme_patchwork(
       design = design,
       widths = ggplot2::unit(1.5, "in"),
-      heights = ggplot2::unit(1.5, "in")
-    )
+      heights = ggplot2::unit(c(1, 2), "in"),
+      guides = "collect"
+    ) &
+    ggplot2::theme(legend.position = "bottom")
 }

### 2024-01-29 _targets.R

@@ -50,60 +50,72 @@ list(
     filter_blots(blots_clean)
   ),
   tar_target(
+    blots_phospho,
+    phospho_ratio(blots_filtered)
+  ),
+  tar_target(
+    blots_all,
+    dplyr::bind_rows(blots_filtered, blots_phospho)
+  ),
+  tar_target(
     blots_index,
-    index_blots(blots_filtered)
+    index_blots(blots_all)
   ),
   tar_map(
     values = tibble::tribble(
-      ~experiment,   ~protein,     ~name,              ~paired, ~comp,
-      "ar",          "α-SMA",      "ar_sma",           FALSE,    rlang::sym("comps_ar"),
-      "bleo",        "MCT1",       "bleo_mct1",        FALSE,    rlang::sym("comps_bleo"),
-      "bleo",        "MCT4",       "bleo_mct4",        FALSE,    rlang::sym("comps_bleo"),
-      "bleo",        "α-SMA",      "bleo_sma",         FALSE,    rlang::sym("comps_bleo"),
-      "comp",        "MCT1",       "comp_mct1",        FALSE,    NULL,
-      "comp",        "MCT4",       "comp_mct4",        FALSE,    NULL,
-      "comp",        "α-SMA",      "comp_sma",         FALSE,    NULL,
-      "diclofenac",  "α-SMA",      "diclofenac_sma",   FALSE,    rlang::sym("comps_nsaid"),
-      "dual",        "AKT",        "dual_akt",         FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "AMPK",       "dual_ampk",        FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "CNN1",       "dual_cnn1",        FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "Col1a1",     "dual_col1a1",      FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "Col3a1",     "dual_col3a1",      FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "ERK",        "dual_erk",         FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "FN1",        "dual_fn1",         FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "Kla",        "dual_kla",         FALSE,    rlang::sym("comps_azd_2"),
[…]

### 2024-02-13 manuscript/manuscript.qmd
add NADP data

@@
 
-## Lactate transport inhibition signals through histone lactylation
+## Lactate transport inhibition may signal through histone lactylation
 
@@
 
-Recently,
+Post-translational protein modification of histone lysine residues by lactylation has been identified as a novel mechanism of metabolic regulation of gene expression [@zhang2019a].
+Given the increase in intracellular lactate following MCT4 inhibition, we measured H3 histone lactylation by immunoblot (**Figure 6I-J**).
+We found that inhibition of both MCT1 and MCT4 increased histone lactylation and may account for some of the transcriptional changes observed following these treatments.
 
-## Lactate transport inhibition / redox stress section
+## Lactate transport inhibition contributes to antioxidant defense
 
-We next examined whether MCT inhibition affects TGFβ-dependent signaling pathways.
-TGFβ activates both SMAD and non-SMAD signaling pathways.
-We found no impact of MCT inhibition on decreasing Smad3 or ERK phosphorylation in LFs 48 h following TGFβ stimulation (**Figure 5**).
-
-Hypoxia-inducible factor 1α (HIF-1α) is activated in lung fibroblasts following TGFβ and is required for myofibroblast differentiation and experimental pulmonary fibrosis partly through activation of glycolysis [@goodwin2018].
-We found maximal HIF-1α stabilization 6 h following TGFβ treatment.
-At this time point, MCT inhibition did not decrease HIF-1α (**Figure 5E-F**), rather, HIF-1α protein levels were increased with combined MCT1 and MCT4 inhibition.
+Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which couple lactate oxidation to NADH production.
+Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression, we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
+Consistent with an increase in intracellular lactate, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A-C**).
 

### 2024-02-13 R/blots.R

@@ -19,7 +19,8 @@ clean_blots <- function(df) {
       "cnn1" = "CNN1",
       "hif1a" = "HIF-1α",
       "akt" = "AKT",
-      "pakt" = "pAKT"
+      "pakt" = "pAKT",
+      "h3" = "H3"
     )
 
   df |>
@@ -112,26 +113,30 @@ plot_blot <- function(df, stats = NULL, title = NULL) {
   }
 }
 
-phospho_ratio <- function(df) {
-  prots <- c("AKT", "AMPK", "ERK", "Smad3")
+blot_ratios <- function(df) {
+  num <- c("pAKT", "pAMPK", "pERK", "pSmad3", "Kla")
+  denom <- c("AKT", "AMPK", "ERK", "Smad3", "H3")
+  ratios <- stringr::str_c(num, denom, sep = "/")
 
   df |>
     dplyr::filter(
-      stringr::str_detect(.data$protein, paste(prots, collapse = "|"))
+      stringr::str_detect(.data$protein, paste(c(num, denom), collapse = "|"))
     ) |>
     dplyr::mutate(
-      phospho = ifelse(
-        stringr::str_detect(.data$protein, "^p\\D"),
+      temp = ifelse(
+        .data$protein %in% num,
         "density",
         "background"
       ),
-      protein = stringr::str_replace(.data$protein, "^p", ""),
-      protein = stringr::str_c("p", .data$protein, "/", .data$protein),
+      protein = purrr::map_chr(
+        .data$protein,
+        \(x) ratios[stringr::str_detect(ratios, x)]
+      ),
       .after = "dose"
     ) |>
     dplyr::select("group", "experiment":"protein", "ratio") |>
     tidyr::pivot_wider(
-      names_from = "phospho",
+      names_from = "temp",
       values_from = "ratio"
     ) |>
     dplyr::mutate(ratio = .data$density / .data$background) |>

### 2024-02-13 R/figures.R

@@ -264,14 +264,26 @@ make_fig05s <- function(p1, p2, p3) {
     )
 }
 
-make_fig06 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
-  design <- "abcd \n efgh"
-  p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 +
+make_fig06 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10) {
+  design <- "abcd \n efgh \n ij##"
+  p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + p10 +
     theme_patchwork(
       design = design,
       widths = ggplot2::unit(1.5, "in"),
-      heights = ggplot2::unit(c(1, 2), "in"),
+      heights = ggplot2::unit(c(1, 2, 1), "in"),
       guides = "collect"
+    ) &
+    ggplot2::theme(legend.position = "bottom")
+}
+
+make_fig07 <- function(p1, p2, p3, p4, p5, p6) {
+  ((p1 + p2 + p3 + p4 + p5 + p6) +
+     patchwork::plot_layout(guides = "collect")) +
+    theme_patchwork(
+      design = "abc \n def",
+      widths = ggplot2::unit(1.5, "in"),
+      heights = ggplot2::unit(c(1.5), "in"),
+      tags = list(c("A", "", "", "B", "", ""))
     ) &
     ggplot2::theme(legend.position = "bottom")
 }

### 2024-02-13 R/nad.R

@@ -1,19 +1,19 @@
 # nad.R
 
 clean_nad <- function(protein, nad) {
+  metabs <- unique(nad$nucleotide)
+  top <- metabs[metabs %in% c("NADH", "NADPH")]
+  bottom <- metabs[metabs %in% c("NAD", "NADP")]
+
   protein <-
     protein |>
-    dplyr::mutate(
-      protein = 300 * .data$conc  # μg protein
-    ) |>
-    dplyr::select(-"conc")
+    dplyr::rename(protein = "conc") |>
+    # μg protein or divide cell by 1000
+    dplyr::mutate(protein = ifelse(top == "NADPH", protein / 1000, protein))
 
   nad <-
     nad |>
-    dplyr::rename(measurement = "nucleotide") |>
-    dplyr::mutate(
-      conc = 900 * .data$conc  # pmol nucleotide
-    )
+    dplyr::rename(measurement = "nucleotide") # pmol nucleotide
 
   dplyr::left_join(
     protein,
@@ -25,13 +25,15 @@ clean_nad <- function(protein, nad) {
       values_from = "conc"
     ) |>
     dplyr::mutate(
-      NAD_norm = .data$NAD / .data$protein,  # pmol / μg
-      NADH_norm = .data$NADH / .data$protein,
-      Ratio = .data$NADH / .data$NAD
+      # pmol / μg protein or fmol / cell
+      "{bottom}_norm" := .data[[bottom]] / .data$protein,
+      "{top}_norm" := .data[[top]] / .data$protein,
+
+      "{top}/{bottom}" := .data[[top]] / .data[[bottom]]
     ) |>
     refactor() |>
     tidyr::pivot_longer(
-      "NAD":"Ratio",
+      !("experiment":"protein"),
       names_to = "measurement",
       values_to = "value"
     ) |>

### 2024-02-13 _targets.R

@@ -51,7 +51,7 @@ list(
   ),
   tar_target(
     blots_phospho,
-    phospho_ratio(blots_filtered)
+    blot_ratios(blots_filtered)
   ),
   tar_target(
     blots_all,
@@ -110,7 +110,10 @@ list(
       "dual",        "pAKT/AKT",     "dual_p_akt",       FALSE,    rlang::sym("comps_azd_2"),
       "dual",        "pAMPK/AMPK",   "dual_p_ampk",      FALSE,    rlang::sym("comps_azd_2"),
       "dual",        "pERK/ERK",     "dual_p_erk",       FALSE,    rlang::sym("comps_azd_2"),
-      "dual",        "pSmad3/Smad3", "dual_p_smad3",     FALSE,    rlang::sym("comps_azd_2")
+      "dual",        "pSmad3/Smad3", "dual_p_smad3",     FALSE,    rlang::sym("comps_azd_2"),
+      "lactyl",      "Kla",          "lactyl_kla",       FALSE,    rlang::sym("comps_azd_2"),
+      "lactyl",      "H3",           "lactyl_h3",        FALSE,    rlang::sym("comps_azd_2"),
+      "lactyl",      "Kla/H3",       "lactyl_kla_h3",    FALSE,    rlang::sym("comps_azd_2")
     ),
     names = name,
     tar_target(
@@ -139,7 +142,8 @@ list(
         c("dual", "ar"),
         "ipf-lf",
         "dual",
-        "dual"
+        "dual",
+        "lactyl"
       ),
       protein = list(
         c("α-SMA", "MCT1", "MCT4"),
@@ -150,7 +154,8 @@ list(
         "α-SMA",
         c("α-SMA", "Col1a1"),
         c("pSmad3", "Smad3", "pSmad3/Smad3"),
-        c("pERK", "ERK", "pERK/ERK")
+        c("pERK", "ERK", "pERK/ERK"),
+        c("Kla", "H3", "Kla/H3")
       ),
       treatment = list(
         NULL,
@@ -161,9 +166,11 @@ list(
         c("Veh", "AR", "VB", "AR/VB"),
         c("Veh", "AZD", "VB", "AZD/VB"),
         c("Veh", "AZD", "VB", "AZD/VB"),
+        c("Veh", "AZD", "VB", "AZD/VB"),
         c("Veh", "AZD", "VB", "AZD/VB")
       ),
       condition = list(
+        NULL,
         NULL,
[…]

### 2024-02-14 manuscript/manuscript.qmd
add ROS data

@@
     fig06,
+    fig07,
     # fig10,
@@
 Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression, we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
-Consistent with an increase in intracellular lactate, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A-C**).
+Consistent with an increase in intracellular lactate with MCT4 inhibition, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A**), while the NADPH/NADP^+^ ratio was unchanged (**Figure 7B**).
+Treatment-dependent increases in NADH/NAD^+^ corresponded to reductions in total ROS as measured by CellROX (**Figure 7C**).
+Notably, we did not observe TGFβ-dependent increases in total ROS production nor an impact of TGFβ or inhibitors on mitochondrial superoxide production (**Supplemental Figure 7A**).
+
+Proline biosynthesis has been previousyl suggested as a mechanism by which NIH-3T3 fibroblasts prevent excessive ROS accumulation following TGFβ stimulation [@schworer2020].
 

### 2024-02-14 R/figures.R

@@ -276,14 +276,15 @@ make_fig06 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10) {
     ggplot2::theme(legend.position = "bottom")
 }
 
-make_fig07 <- function(p1, p2, p3, p4, p5, p6) {
+make_fig07 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
   ((p1 + p2 + p3 + p4 + p5 + p6) +
      patchwork::plot_layout(guides = "collect")) +
+    p7 +
+    ((p8) + patchwork::plot_layout(guides = "keep")) +
     theme_patchwork(
-      design = "abc \n def",
+      design = "abc \n def \n gh#",
       widths = ggplot2::unit(1.5, "in"),
       heights = ggplot2::unit(c(1.5), "in"),
-      tags = list(c("A", "", "", "B", "", ""))
-    ) &
-    ggplot2::theme(legend.position = "bottom")
+      tags = list(c("A", "", "", "B", "", "", "C", "D"))
+    )
 }

### 2024-02-14 R/ros.R

@@ -1,0 +1,32 @@
+# ros.R
+
+clean_ros <- function(df) {
+  out <-
+    df |>
+    dplyr::group_by(.data$experiment) |>
+    dplyr::mutate(
+      dplyr::across(
+        tidyselect::any_of(c("value", "MitoSOX", "MitoTracker")),
+        \(x) x - max(0, mean(x[.data$type == "blank"]), na.rm = TRUE)
+      )
+    ) |>
+    dplyr::filter(.data$type != "blank") |>
+    dplyr::select(-"type") |>
+    refactor() |>
+    dplyr::ungroup()
+
+  if ("value" %nin% names(out)) {
+    out |>
+      dplyr::mutate(
+        Ratio = .data$MitoSOX / .data$MitoTracker
+      ) |>
+      tidyr::pivot_longer(
+        c("MitoSOX", "MitoTracker", "Ratio"),
+        names_to = "measurement",
+        values_to = "value"
+      )
+  } else if ("value" %in% names(out)) {
+    out |>
+      dplyr::mutate(measurement = "CellROX", .before = "value")
+  }
+}

### 2024-02-14 _targets.R

@@ -572,6 +572,32 @@ list(
     )
   ),
 
+  # ros ---------------------------------------------------------------------
+
+  tar_map(
+    values = list(
+      names = c("cellrox", "mitosox")
+    ),
+    names = names,
+    tar_target(
+      ros_file,
+      raw_data_path(stringr::str_c(names, ".csv")),
+      format = "file_fast"
+    ),
+    tar_target(
+      ros_raw,
+      readr::read_csv(ros_file, show_col_types = FALSE)
+    ),
+    tar_target(
+      ros_clean,
+      clean_ros(ros_raw)
+    )
+  ),
+  tar_target(
+    ros_clean,
+    dplyr::bind_rows(ros_clean_cellrox, ros_clean_mitosox)
+  ),
+
   # seahorse ----------------------------------------------------------------
 
   tar_target(
@@ -851,6 +877,11 @@ list(
     tar_target(
       metab_mcti_tar_lactate,
       plot_mois(metab_mcti_tar_treated, metab_mcti_tar_tt, "lactate"),
+      format = "rds"
+    ),
+    tar_target(
+      metab_mcti_tar_proline,
+      plot_mois(metab_mcti_tar_treated, metab_mcti_tar_tt, "proline"),
       format = "rds"
     ),
     NULL
@@ -1615,7 +1646,9 @@ list(
       nad_plot_nadh_ratio,
       nad_plot_nadp_norm,
       nad_plot_nadph_norm,
-      nad_plot_nadph_ratio
+      nad_plot_nadph_ratio,
+      patchwork::plot_spacer(), # CellROX
+      metab_mcti_tar_proline_intra
     ) |>
       write_figures("Figure 07"),
     format = "file"

### 2024-02-15 manuscript/manuscript.qmd
finish up ROS data

--- 2024-02-13 manuscript/manuscript.qmd
@@
 
-Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which couple lactate oxidation to NADH production.
-Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression, we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
-Consistent with an increase in intracellular lactate, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A-C**).
+Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which transfer electrons from lactate to NADH.
+Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression [@jain2013, @schworer2020], we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
+Consistent with an increase in intracellular lactate with MCT4 inhibition, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A**), which was associated with a reduction in total ROS as measured by the CellROX fluorescent probe (**Figure 7B**).
+By contrast, the NADPH/NADP^+^ was unchanged by either TGFβ or MCT inhibitors (**Supplemental Figure 7A**).
+Notably, we did not observe TGFβ-dependent increases in total ROS production nor an impact of TGFβ or inhibitors on mitochondrial superoxide production (**Supplemental Figure 7B**).
+
+Proline biosynthesis has been previously suggested as a mechanism by which NIH-3T3 fibroblasts prevent excessive ROS accumulation following TGFβ stimulation [@schworer2020].
+In this model, proline synthesis from glutamine consumes reducing equivalents from NADPH and NADH, thereby ameliorating reductive stress and decreasing ROS production.
+Compared to this prior study, we did not find that TGFβ increased proline in our primary lung fibroblasts (**Supplemental Figure 7C**).
+Overall, the fractional labeling of proline from [U-^13^C~5~]-glutamine was modest (~10%) compared to the 40% previously reported (**Supplemental Figure 7D**).
+While we observed trends toward increased proline production from glutamine with MCT4 inhibition, this mechanism seems unlikely to account for the effects of MCT inhibitors on ROS generation.
 

### 2024-02-15 R/figures.R

@@ -276,15 +276,29 @@ make_fig06 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10) {
     ggplot2::theme(legend.position = "bottom")
 }
 
-make_fig07 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
-  ((p1 + p2 + p3 + p4 + p5 + p6) +
-     patchwork::plot_layout(guides = "collect")) +
-    p7 +
-    ((p8) + patchwork::plot_layout(guides = "keep")) +
+make_fig07 <- function(p1, p2, p3, p4) {
+  x <- p1 + p2 + p3 + patchwork::plot_layout(guides = "collect")
+  x - p4 +
+    theme_patchwork(
+      design = "aaab",
+      widths = ggplot2::unit(1.5, "in"),
+      heights = ggplot2::unit(1.5, "in"),
+      tags = list(c("A", "", "", "B"))
+    ) &
+    ggplot2::theme(
+      legend.position = "bottom",
+      legend.margin = ggplot2::margin(t = -10)
+    )
+}
+
+make_fig07s <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
+  p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 +
     theme_patchwork(
       design = "abc \n def \n gh#",
       widths = ggplot2::unit(1.5, "in"),
-      heights = ggplot2::unit(c(1.5), "in"),
+      heights = ggplot2::unit(1.5, "in"),
+      guides = "collect",
       tags = list(c("A", "", "", "B", "", "", "C", "D"))
-    )
+    ) &
+    ggplot2::theme(legend.position = "bottom")
 }

### 2024-02-15 R/mids.R

@@ -62,7 +62,28 @@ correct_mids <- function(mid, correction_matrices) {
     dplyr::mutate(data = purrr::map2(.data$matrix, .data$data, mmult)) |>
     tidyr::unnest(c("data")) |>
     dplyr::filter(.data$isotope %in% stringr::str_c("M", 0:6)) |>
-    dplyr::select(-"matrix")
+    dplyr::select(-"matrix") |>
+    dplyr::ungroup()
+}
+
+format_mids_all <- function(df) {
+  df |>
+    dplyr::group_by(tracer, condition, treatment, metabolite, isotope) |>
+    dplyr::summarize(
+      mean = mean(mid_corr),
+      se = sd(mid) / sqrt(dplyr::n())
+    ) |>
+    dplyr::group_by(tracer, condition, treatment, metabolite) |>
+    dplyr::mutate(
+      means = 1 - cumsum(mean) + mean,
+      ymin = means + se,
+      ymax = means - se,
+      group = stringr::str_c(condition, treatment, sep = "."),
+      group = factor(group, levels = c(
+        "Ctl.Veh", "Ctl.AZD", "Ctl.VB", "Ctl.AZD/VB",
+        "TGFβ.Veh", "TGFβ.AZD", "TGFβ.VB", "TGFβ.AZD/VB"
+      ))
+    )
 }
 
 plot_mids_all <- function(df, label) {
@@ -74,28 +95,55 @@ plot_mids_all <- function(df, label) {
       lac3 = expression(paste("[U-"^13, "C"[3], "]-lactate"))
     )
 
-  x <-
-    df |>
+  df |>
     dplyr::filter(tracer %in% label) |>
-    dplyr::group_by(condition, treatment, metabolite, isotope) |>
-    dplyr::summarize(
-      mean = mean(mid_corr),
-      se = sd(mid) / sqrt(dplyr::n())
-    ) |>
-    dplyr::group_by(condition, treatment, metabolite) |>
-    dplyr::mutate(
-      means = 1 - cumsum(mean) + mean,
-      ymin = means + se,
-      ymax = means - se,
-      group = stringr::str_c(condition, treatment, sep = "."),
-      group = factor(group, levels = c(
-        "Ctl.Veh", "Ctl.AZD", "Ctl.VB", "Ctl.AZD/VB",
-        "TGFβ.Veh", "TGFβ.AZD", "TGFβ.VB", "TGFβ.AZD/VB"
-      ))
-    )
-
-  ggplot2::ggplot(x) +
[…]

### 2024-02-15 _targets.R

@@ -1091,6 +1091,10 @@ list(
     mid_mcti_mids_corr,
     correct_mids(mid_mcti_mids, correction_matrices)
   ),
+  tar_target(
+    mid_mcti_stacked_data,
+    format_mids_all(mid_mcti_mids_corr)
+  ),
   tar_map(
     values = list(
       label = c("glc2", "gln5", "lac3", "glc6")
@@ -1098,7 +1102,7 @@ list(
     names = label,
     tar_target(
       mid_mcti_plot,
-      plot_mids_all(mid_mcti_mids_corr, label = label),
+      plot_mids_all(mid_mcti_stacked_data, label = label),
       format = "rds"
     ),
     tar_target(
@@ -1200,6 +1204,54 @@ list(
     ),
     NULL
   ),
+  tar_map(
+    values = list(
+      names = c("gln5_proline"),
+      tracer = c("gln5"),
+      metabolite = c("PRO"),
+      isotope = c("M5"),
+      ytitle = c("M5 Mole fraction"),
+      title = c("[U-^13^C~5~]-GLN → PRO")
+    ),
+    names = names,
+    tar_target(
+      mid_filter,
+      filter_mids(
+        mid_mcti_mids_corr,
+        tracer,
+        metabolite,
+        isotope
+      )
+    ),
+    tar_target(
+      mid_stats,
+      twofactor(
+        mid_filter,
+        "mid_corr",
+        mixed = FALSE,
+        comps = comps_azd_2
+      )
+    ),
+    tar_target(
+      mid_plot,
+      plot_two_factor(
+        mid_filter,
+        mid_stats,
+        x = "treatment",
+        y = "mid_corr",
+        ytitle = ytitle,
+        title = title
+      ) +
+        ggplot2::theme(
+          plot.title = ggtext::element_markdown(
+            margin = ggplot2::margin(b = -5)
+          )
+        ),
+      format = "rds"
+    ),
+    NULL
+  ),
+
 
   # rnaseq ------------------------------------------------------------------
 
@@ -1680,23 +1732,26 @@ list(
       nad_plot_nad_norm,
       nad_plot_nadh_norm,
       nad_plot_nadh_ratio,
-      nad_plot_nadp_norm,
-      nad_plot_nadph_norm,
-      nad_plot_nadph_ratio,
-      ros_plot_cellrox,
-      metab_mcti_tar_proline_intra
+      ros_plot_cellrox
     ) |>
[…]

### 2024-02-15 manuscript/manuscript.qmd
add mice phenotype figures

@@
     fig07,
+    fig08, 
     # fig10,
@@
 
+## MCT inhibition decreases experimental pulmonary fibrosis
+
+Based on our *in vitro* data suggesting an anti-fibrotic effect of lactate transporter inhibition, we next tested the efficacy of MCT inhibitors in the bleomycin-induced mouse model of pulmonary fibrosis.
+Mice received 1.2 U/kg bleomycin by intratracheal administration.
+After seven days, the animals began treatment with AZD3965 (100 mg/kg twice daily) or VB124 (30 mg/kg once daily) or vehicle by oral gavage (**Figure 8A**).
+We anticipated that dual inhibition would likely be toxic *in vivo* and did not assess the combination of inhibitors.
+Compared to vehicle, mice treated with VB124 had increased weight recovery 21 days after bleomycin administration (**Supplemental Figure 8**).
+Lung mechanics were improved by ~ 50% of baseline compared to vehicle controls following 14 days of MCT inhibitor treatment (**Figure 8B-C**).
+Consistent with improved elastance and compliance measures, histologic fibrosis severity assessed by Ashcroft scoring [@ashcroft1988] also improved.
+
 \newpage

### 2024-02-15 R/figures.R

@@ -302,3 +302,24 @@ make_fig07s <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
     ) &
     ggplot2::theme(legend.position = "bottom")
 }
+
+make_fig08 <- function(p1, p2, p3, p4, p5, p6) {
+  design <- "abc \n def"
+  p1 + p2 + p3 + p4 + p5 + p6 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(c(2, 1.5, 1.5), "in"),
+      heights = ggplot2::unit(1.5, "in")
+    )
+}
+
+make_fig08s <- function(p1) {
+  design <- "a"
+  p1 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(4.5, "in"),
+      heights = ggplot2::unit(1.25, "in"),
+      tags = NULL
+    )
+}

### 2024-02-15 R/mice.R

@@ -83,6 +83,15 @@ plot_wt_timeline <- function(df, stat) {
       linewidth = 0.1,
       linetype = 2
     ) +
+    ggbeeswarm::geom_quasirandom(
+      ggplot2::aes(group = .data$group),
+      pch = 1,
+      color = "black",
+      width = 0.05,
+      size = 0.5,
+      stroke = 0.25,
+      show.legend = FALSE
+    ) +
     ggplot2::geom_errorbar(
       ggplot2::aes(color = .data$group),
       stat = "summary",
@@ -107,15 +116,6 @@ plot_wt_timeline <- function(df, stat) {
       size = 1.5,
       color = "white",
       position = ggplot2::position_nudge(x = 1),
-      show.legend = FALSE
-    ) +
-    ggbeeswarm::geom_quasirandom(
-      ggplot2::aes(group = .data$group),
-      pch = 1,
-      color = "black",
-      width = 0.05,
-      size = 0.5,
-      stroke = 0.25,
       show.legend = FALSE
     ) +
     ggplot2::geom_text(

### 2024-02-15 _targets.R

@@ -1542,7 +1542,7 @@ list(
       "dual-azd-sma-blot.png",      1,      0,      -0.05,  "sma_azd",
       "dual-ar-sma-blot.png",       1.2,    0.1,    -0.05,  "sma_ar",
       "dual-azd-contract.png",      1,      0,      -0.05,  "contraction",
-      "bleo-trichrome.png",         1,      0,      0,      "trichrome",
+      "bleo-trichrome.png",         1.1,    0,      0,      "trichrome",
       "bleo-timeline.png",          1,      0,      0,      "timeline",
       "dual-azd-smad3-blot.png",    1,      0,      -0.05,  "smad3_azd",
       "dual-azd-erk-blot.png",      1,      0,      -0.05,  "erk_azd",
@@ -1750,6 +1750,30 @@ list(
       mid_plot_gln5_proline
     ) |>
       write_figures("Figure 07.supplement"),
+    format = "file"
+  ),
+
+  # figure 8 ----------------------------------------------------------------
+
+  tar_target(
+    fig08,
+    make_fig08(
+      fig_img_timeline,
+      mice_vent_mcti_plot_cst,
+      mice_vent_mcti_plot_prime8,
+      fig_img_trichrome,
+      mice_mcti_plot_ashcroft,
+      mice_mcti_plot_ohp
+    ) |>
+      write_figures("Figure 08"),
+    format = "file"
+  ),
+  tar_target(
+    fig08s,
+    make_fig08s(
+      mice_wt_mcti_plot
+    ) |>
+      write_figures("Figure 08.supplement"),
     format = "file"
   ),
 

### 2024-02-16 analysis/bleo-mid.qmd
working on isotope labeling data

---
title: "Bleo ^13^C Tracing"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
    toc: true
    toc-depth: 3
    toc-location: left
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/bleo-mid.qmd")

tar_config_set(store = here("_targets"))
```

## Overview

Mice were treated with bleomycin on Day 1. Treatment with AZD3965 or VB124 began on Day 7. Mice were euthanized on Day 21. One hour before organ harvest, mice were injected with an intraperitoneal bolus of [U-^13^C~6~]-glucose. Lungs and plasma were collected and isotope enrichment was analyzed by LC-MS. The goal is to characterize changes in isotope labeling patterns in disease and with treatment.

## Data

```{r}
#| label: data

mids <- tar_read(mid_bleo_mids_corr)
```

### Are there significant impacts of bleomycin?

This would be the first report of isotope labeling in the bleomycin model.

```{r}
#| label: overall-enrichment

mids |> 
  dplyr::filter(isotope == "M0") |> 
  group_by(tissue, metabolite) |> 
  nest() |> 
  mutate(
    model = purrr::map(data, \(x) lm(mid_corr ~ group, data = x)), 
    stats = purrr::map(
      model, 
      \(x) emmeans::emmeans(x, ~ group) |> 
        emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx") |> 
        broom::tidy()
    )
  ) |> 
  unnest(c(stats)) |> 
  select(tissue, group = contrast, metabolite, adj.p.value) |> 
  arrange(adj.p.value) |> 
  head()
```

Based on these data, there are no significantly different patterns of enrichment. This may be a function of relatively few mice (n = 3) and variability in tracer dose.

### Does labeling suggest differences in lactate fate?

For these analyses, can look at product/substrate ratios for specific isotopes. This analysis may have more power because they are ratios. For the lung, 3PG M3 is produced primarily by glycolysis. Can look at the ratio of downstream isotope enrichment to 3PG.

```{r}
#| label: 3PG-ratios

df <- 
  mids |> 
  filter(tissue == "lung") |> 
  filter(
    (metabolite == "3PG" & isotope == "M3") |
      (metabolite == "CIT" & isotope == "M2") |
      (metabolite == "LAC" & isotope == "M3")
  ) |> 
  group_by(id, group) |> 
  summarize(
    lac = mid_corr[metabolite == "LAC"] / mid_corr[metabolite == "3PG"], 
    cit = mid_corr[metabolite == "CIT"] / mid_corr[metabolite == "3PG"]
  ) |> 
  pivot_longer(c("lac", "cit"), names_to = "ratio", values_to = "value") |> 
  filter(!(id == "S17" & ratio == "cit"))

ggplot(df) +
  facet_wrap(vars(ratio)) +
  aes(
    x = group, 
    y = value, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  )
```

The data demonstrate that M2 citrate labeling is ~ 80% of M3 3PG labeling, suggesting that most of the intracellular citrate is coming from 3PG from glycolysis. In addition, the M3 LAC / M3 3PG labeling ratio is greater than 1. This suggests an alternative source of labeled lactate other than glycolysis, which is likely plasma M3 lactate given the plasma levels. This is also consistent with the observation that this ratio is higher in AZD-treated animals where the plasma enrichment tended to be higher.

```{r}
#| label: CIT-LAC

df <- 
  mids |> 
  filter(tissue == "lung") |> 
  filter(
    (metabolite == "LAC" & isotope == "M3") |
      (metabolite == "CIT" & isotope == "M2")
  ) |> 
  group_by(id, group) |> 
  summarize(
    cit = (1-mid_corr[metabolite == "CIT"]) / mid_corr[metabolite == "LAC"]
  ) |> 
  filter(!(id == "S17"))

ggplot(df) +
  aes(
    x = group, 
    y = cit, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  )
```

Looking at the ratio of M3 lactate to M2 citrate, there seems to be decreased oxidation in AZD-treated cells, possibly consistent with decreased uptake, while there was no difference with VB124.

### What about plasma normalization?

```{r}
#| label: plasma-GLC-3PG

df <- 
  mids |> 
  filter(
    (tissue == "plasma" & metabolite == "GLC" & isotope == "M6") |
      (tissue == "lung" & metabolite == "3PG" & isotope == "M3")
  ) |> 
  group_by(id, group) |> 
  summarize(
    cit = (mid_corr[tissue == "lung"]) / mid_corr[tissue == "plasma"]
  ) |> 
  filter(!(id == "S17"))

lm(cit ~ group, data = df) |> 
  emmeans::emmeans(~ group) |> 
  emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx")

ggplot(df) +
  aes(
    x = group, 
    y = cit, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  )
```

If we normalize intracellular M3 3PG by extracellular M6 GLC, we see a decrease with AZD3965 and an increase with VB124. This may suggest a trend toward decreased glycolysis following AZD, but an increase following VB124.

```{r}
#| label: plasma-lung-LAC

df <- 
  mids |> 
  filter(
    (tissue == "plasma" & metabolite == "LAC" & isotope == "M3") |
      (tissue == "lung" & metabolite == "LAC" & isotope == "M3")
  ) |> 
  group_by(id, group) |> 
  summarize(
    ratio = (mid_corr[tissue == "lung"]) / mid_corr[tissue == "plasma"]
  ) |> 
  filter(id != "S17")

lm(ratio ~ group, data = df) |> 
  emmeans::emmeans(~ group) |> 
  emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx")

ggplot(df) +
  aes(
    x = group, 
    y = ratio, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  )
```

In these data, we see that AZD and VB seem to decrease the fraction of intracellular lactate from extracellular lactate.

```{r}
#| label: normalize-plasma-GLC

df <- 
  mids |> 
  filter(
    (tissue == "lung" & metabolite == "3PG" & isotope == "M3") |
      (tissue == "lung" & metabolite == "LAC" & isotope == "M3") |
      (tissue == "plasma" & metabolite == "GLC" & isotope == "M6")
  ) |> 
  group_by(id, group) |> 
  summarize(
    lac = mid_corr[metabolite == "LAC"] / mid_corr[metabolite == "GLC"], 
    pg = mid_corr[metabolite == "3PG"] / mid_corr[metabolite == "GLC"]
  ) |> 
  pivot_longer(c("lac", "pg"), names_to = "ratio", values_to = "value") |> 
  filter(!(id == "S17" & ratio == "cit"))

ggplot(df) +
  facet_wrap(vars(ratio)) +
  aes(
    x = group, 
    y = value, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  )
```

Normalizing by plasma tracer labeling suggests increased intracellular lactate from glucose in VB124-treated cells, this is likely consistent with increased glycolysis and decreased lactate efflux. In AZD-treated cells, there was a decrease in 3PG labeling from plasma glucose, suggesting decreased glycolysis?

### Plasma Normalization by Mass

If using a different compound, we would likely normalize the tissue concentration by the plasma concentration. We will explore those possibilities here.

```{r}
#| label: plasma-tissue-LAC-mass

df <- 
  mids |> 
  dplyr::filter(metabolite == "LAC" & isotope == "M3") |> 
  dplyr::group_by(id, sample, group) |> 
  dplyr::summarise(ratio = area[tissue == "lung"] / area[tissue == "plasma"]) |> 
  dplyr::ungroup() |> 
  dplyr::mutate(ratio_norm = ratio / mean(ratio[group == "Bleo"], na.rm = TRUE)) |> 
  dplyr::group_by(group) |> 
  wmo::remove_nested_outliers("ratio_norm", remove = TRUE)

lm(ratio_norm ~ group, data = df) |> 
  emmeans::emmeans(~ group) |> 
  emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx")

ggplot2::ggplot(df) +
  aes(
    x = group, 
    y = ratio_norm, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  ) +
  labs(
    y = "Lung / Plasma M3 LAC"
  )
```

```{r}
#| label: plasma-tissue-GLC-LAC-mass

df <- 
  mids |> 
  dplyr::filter(
    (tissue == "lung" & metabolite == "LAC" & isotope == "M3") | 
      (tissue == "plasma" & metabolite == "GLC" & isotope == "M6")
    ) |> 
  dplyr::group_by(id, sample, group) |> 
  dplyr::summarise(ratio = area[tissue == "lung"] / area[tissue == "plasma"]) |> 
  dplyr::ungroup() |> 
  dplyr::mutate(ratio_norm = ratio / mean(ratio[group == "Bleo"], na.rm = TRUE)) |> 
  dplyr::group_by(group) |> 
  wmo::remove_nested_outliers("ratio_norm", remove = TRUE)

lm(ratio_norm ~ group, data = df) |> 
  emmeans::emmeans(~ group) |> 
  emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx")

ggplot2::ggplot(df) +
  aes(
    x = group, 
    y = ratio_norm, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  ) +
  labs(
    y = "Lung M3 LAC / Plasma M6 GLC"
  )
```

```{r}
#| label: plasma-tissue-LAC-CIT-mass

df <- 
  mids |> 
  dplyr::filter(
    (tissue == "lung" & metabolite == "CIT" & isotope == "M2") | 
      (tissue == "plasma" & metabolite == "LAC" & isotope == "M3")
    ) |> 
  dplyr::group_by(id, sample, group) |> 
  dplyr::summarise(ratio = area[tissue == "lung"] / area[tissue == "plasma"]) |> 
  dplyr::ungroup() |> 
  dplyr::mutate(ratio_norm = ratio / mean(ratio[group == "Bleo"], na.rm = TRUE)) |> 
  dplyr::group_by(group) |> 
  wmo::remove_nested_outliers("ratio_norm", remove = TRUE)

lm(ratio_norm ~ group, data = df) |> 
  emmeans::emmeans(~ group) |> 
  emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx")

ggplot2::ggplot(df) +
  aes(
    x = group, 
    y = ratio_norm, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  ) +
  labs(
    y = "Lung M2 CIT / Plasma M3 LAC"
  )
```

```{r}
#| label: plasma-tissue-GLC-CIT-mass

df <- 
  mids |> 
  dplyr::filter(
    (tissue == "lung" & metabolite == "CIT" & isotope == "M2") | 
      (tissue == "plasma" & metabolite == "GLC" & isotope == "M6")
    ) |> 
  dplyr::group_by(id, sample, group) |> 
  dplyr::summarise(ratio = area[tissue == "lung"] / area[tissue == "plasma"]) |> 
  dplyr::ungroup() |> 
  dplyr::mutate(ratio_norm = ratio / mean(ratio[group == "Bleo"], na.rm = TRUE)) |> 
  dplyr::group_by(group) |> 
  wmo::remove_nested_outliers("ratio_norm", remove = TRUE)

lm(ratio_norm ~ group, data = df) |> 
  emmeans::emmeans(~ group) |> 
  emmeans::contrast(method = comps_bleo_2, adjust = "dunnettx")

ggplot2::ggplot(df) +
  aes(
    x = group, 
    y = ratio_norm, 
    fill = group
  ) +
  geom_bar(
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(), 
    show.legend = FALSE
  ) +
  geom_point(
    pch = 21, 
    color = "white", 
    show.legend = FALSE
  ) +
  geom_errorbar(
    stat = "summary", 
    fun.data = "mean_se", 
    width = 0.2, 
    linewidth = 0.25
  ) +
  scale_fill_manual(
    values = clrs
  ) +
  labs(
    y = "Lung M2 CIT / Plasma M6 GLC"
  )
```

### 2024-02-16 analysis/mims.qmd
working on isotope labeling data

---
title: "MIMS Analysis"
date-modified: today
format: html
excecute:
  message: false
---

```{r}
#| label: setup
#| include: false

devtools::load_all()
suppressPackageStartupMessages({
  library(tidyverse)
  library(targets)
})

here::i_am("analysis/mims.qmd")

theme_set(ggthemes::theme_few())
```

```{r}
#| label: data

withr::with_dir(here::here(), {
  mims <- tar_read(mims_clean)
})

df <- 
  mims |> 
  dplyr::mutate(
    label = dplyr::case_when(
      ion == "16O" ~ "O", 
      ion == "12C14N" ~ "C^14^N", 
      ion == "12C15N" ~ "C^15^N", 
      ion == "12C15N/12C14N" ~ "C^15^N / C^14^N", 
      ion == "12C21H" ~ "C~2~^1^H", 
      ion == "12C22H" ~ "C~2~^2^H", 
      ion == "12C22H/12C21H" ~ "C~2~^2^H / C~2~^1^H", 
      ion == "31P" ~ "P", 
      ion == "32S" ~ "S", 
      ion == "31P/32S" ~ "P / S"
    ), 
    across(c(mean, stddev), \(x) ifelse(str_detect(ion, "/"), x / 10000, x)),
    across(c(mean, stddev), \(x) ifelse(str_detect(ion, "N/"), x / 0.00368, x)),
    across(c(mean, stddev), \(x) ifelse(str_detect(ion, "H/"), x / 0.000115, x)),
  )
```

## Overview

To further characterize the metabolic consequences of lactate transport inhibition *in vivo*, we performed multi-isotope imaging mass spectrometry of mouse lungs labeled with 6 doses of ^15^N-proline and ^2^H~7~-glucose over three days prior to euthanasia. We imaged one animal from Ctl, Bleo, AZD, and VB groups. For bleomcyin-treated animals, one fibrotic region and one region with more normal appearing alveolar structure were measured. Two planes of images were acquired. The 26 m/*z* ion was ^12^C^14^N in the first plane and ^12^C~2~^2^H in the second plane.

## Analysis

Image data were analyzed using the OpenMIMS Image J plugin. For the initial analysis, regions-of-interest (ROIs) were drawn on the compressed images using the P/S signal, which highlights nuclei. These ROIs were background, nuclei, or lamella. Nuclei were subcategorized as alveolar (cells entirely within the alveoli, likely airway macrophages), airway (nuclei directly adjacent to an airway), AT2 (a large airway cell with lamella), and fibrosis (nuclei not adjacent to an airway). Putative lamellar bodies were characterized by high P and O and low S. I assume an isotopic enrichment of ^2^H to be 0.0115% and of ^15^N to be 0.368%.

## Plots

```{r}
#| label: nuclei-by-region

df |> 
  filter(feature == "nuclei") |> 
  filter(ion %in% c("12C15N/12C14N", "12C22H/12C21H")) |> 
  ggplot() +
  facet_wrap(
    facets = vars(label), 
    scales = "free_y"
  ) +
  aes(
    x = group, 
    y = mean
  ) +
  geom_jitter(
    aes(color = region), 
    position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.3), 
    alpha = 0.2
  ) +
  geom_crossbar(
    aes(group = region), 
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(width = 0.8), 
    width = 0.35
  ) +
  guides(
    color = guide_legend(override.aes = list(alpha = 1))
  ) + 
  scale_color_brewer(
    palette = "Set1", 
    labels = list("Alveolar", "Fibrotic")
  ) +
  labs(
    x = NULL, 
    y = "Enrichment", 
    color = NULL, 
    title = "Nuclear isotope enrichment"
  ) +
  theme(
    strip.text = ggtext::element_markdown(size = 10), 
    legend.position = "bottom"
  )
```

Across all nuclei, there is increased ^15^N signal associated with bleomycin compared to control. This would be consistent with expectation. There may be a trend toward decreased ^15^N nuclear signal with AZD and VB treatment. This would also be consistent with our expectation. Interestingly, bleomycin treatment decreases ^2^H signal in the fibrotic region while AZD and VB treatment increase this signal. Could this be consistent with rerouting of glucose carbons into oxidative and catabolic pathways rather than elimination as lactate, which would fit with our hypothesis? Regardles of the biological meaning, it does suggest a homeostatic metabolic reprogramming as AZD and VB increase glucose enrichment back toward control.

```{r}
#| label: nuclei-by-type

df |> 
  filter(feature == "nuclei") |> 
  filter(ion %in% c("12C15N/12C14N", "12C22H/12C21H")) |> 
  ggplot() +
  facet_grid(
    rows = vars(label), 
    cols = vars(tag), 
    scales = "free_y"
  ) +
  aes(
    x = group, 
    y = mean
  ) +
  geom_jitter(
    aes(color = tag), 
    position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.3), 
    alpha = 0.2
  ) +
  geom_crossbar(
    aes(group = tag), 
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(width = 0.8), 
    width = 0.35
  ) +
  guides(
    color = guide_legend(override.aes = list(alpha = 1))
  ) + 
  scale_color_brewer(
    palette = "Set1"
  ) +
  labs(
    x = NULL, 
    y = "Enrichment", 
    color = NULL, 
    title = "Nuclear isotope enrichment"
  ) +
  theme(
    strip.text.y = ggtext::element_markdown(size = 10), 
    legend.position = "bottom"
  )
```

When separating nuclear ROIs by their environment, different patterns emerge. For ^15^N, nuclei adjacent to the airways, including likely AT2 cells, and alveolar cells demonstrate increased uptake that is attenuated by AZD and VB. This pattern was attenuated or possibly reversed in cells within fibrotic areas. ^2^H uptake in airway, alveolar, and AT2 cell nuclei seems to increase with VB, while nuclei in fibrotic areas have increased labeling with either AZD or VB.

```{r}
#| label: fibrosis-nuclei-by-type

df |> 
  filter(feature == "nuclei") |> 
  filter(region == "fib") |> 
  filter(ion %in% c("12C15N/12C14N", "12C22H/12C21H")) |> 
  ggplot() +
  facet_grid(
    rows = vars(label), 
    cols = vars(tag), 
    scales = "free_y"
  ) +
  aes(
    x = group, 
    y = mean
  ) +
  geom_jitter(
    aes(color = tag), 
    position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.3), 
    alpha = 0.2
  ) +
  geom_crossbar(
    aes(group = tag), 
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(width = 0.8), 
    width = 0.35
  ) +
  guides(
    color = guide_legend(override.aes = list(alpha = 1))
  ) + 
  scale_color_brewer(
    palette = "Set1"
  ) +
  labs(
    x = NULL, 
    y = "Enrichment", 
    color = NULL, 
    title = "Nuclear isotope enrichment"
  ) +
  theme(
    strip.text.y = ggtext::element_markdown(size = 10), 
    legend.position = "bottom"
  )
```

These overall patterns appear preserved when only the more fibrotic section is considered. Importantly, these data suggest this general classification scheme (airway *v.* fibrosis) identifies cells with distinct metabolic phenotypes.

```{r}
#| label: nuclear-enrichment-correlation

df |>
  filter(feature == "nuclei") |> 
  filter(ion %in% c("12C15N/12C14N", "12C22H/12C21H")) |> 
  select(group:ion, mean) |> 
  pivot_wider(
    names_from = "ion", 
    values_from = "mean"
  ) |> 
  ggplot() +
  facet_wrap(
    facets = vars(region), 
    ncol = 1
  ) +
  aes(
    x = `12C22H/12C21H`, 
    y = `12C15N/12C14N`
  ) +
  geom_point(
    aes(
      color = tag
    ), 
    alpha = 0.25
  ) +
  geom_smooth(
    aes(
      color = tag
    ), 
    method = "lm", 
    formula = "y ~ x", 
    show.legend = FALSE
  ) +
  guides(
    color = guide_legend(override.aes = list(alpha = 1))
  ) + 
  scale_color_brewer(
    palette = "Set1"
  ) +
  labs(
    x = "C~2~^2^H / C~2~^1^H", 
    y = "C^15^N / C^14^N", 
    color = NULL, 
    title = "Nuclear isotope enrichment"
  ) +
  coord_fixed() +
  theme(
    axis.title.x = ggtext::element_markdown(size = 10), 
    axis.title.y = ggtext::element_markdown(size = 10), 
    legend.position = "bottom"
  )
```

It appears that there is generally a correlation between ^2^H and ^15^N enrichment among all nuclei that is stronger in fibrotic (*fib*) than alveolar (*alv*) regions.

```{r}
#| label: nuclear-enrichment-correlation-disease

df |>
  filter(feature == "nuclei") |> 
  filter(ion %in% c("12C15N/12C14N", "12C22H/12C21H")) |> 
  select(group:ion, mean) |> 
  pivot_wider(
    names_from = "ion", 
    values_from = "mean"
  ) |> 
  ggplot() +
  # facet_wrap(
  #   facets = vars(region), 
  #   ncol = 1
  # ) +
  aes(
    x = `12C22H/12C21H`, 
    y = `12C15N/12C14N`
  ) +
  geom_point(
    aes(
      color = group
    ), 
    alpha = 0.25
  ) +
  geom_smooth(
    aes(
      color = group
    ), 
    method = "lm", 
    formula = "y ~ x", 
    show.legend = FALSE
  ) +
  guides(
    color = guide_legend(override.aes = list(alpha = 1))
  ) + 
  scale_color_brewer(
    palette = "Set1"
  ) +
  labs(
    x = "C~2~^2^H / C~2~^1^H", 
    y = "C^15^N / C^14^N", 
    color = NULL, 
    title = "Nuclear isotope enrichment"
  ) +
  coord_fixed() +
  theme(
    axis.title.x = ggtext::element_markdown(size = 10), 
    axis.title.y = ggtext::element_markdown(size = 10), 
    legend.position = "bottom"
  )
```

Across treatment groups, it seems that there is a decrease in the slope associated with AZD and VB treatment, trending toward the control, suggesting decreasing correlation between isotope enrichment associated with treatment.

```{r}
#| label: lamella

df |>
  filter(feature == "lamella") |> 
  ggplot() +
  facet_wrap(
    facets = vars(label),
    scales = "free_y"
  ) +
  aes(
    x = group, 
    y = mean
  ) +
  geom_jitter(
    aes(
      color = group, 
      shape = region
    ), 
    position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.3), 
    alpha = 0.2
  ) +
  geom_crossbar(
    aes(
      group = region
    ), 
    stat = "summary", 
    fun = "mean", 
    position = position_dodge(width = 0.8), 
    width = 0.35
  ) +
  guides(
    color = guide_legend(override.aes = list(alpha = 1))
  ) + 
  scale_color_brewer(
    palette = "Set1"
  ) +
  labs(
    x = NULL, 
    y = "Enrichment", 
    color = NULL, 
    shape = NULL, 
    title = "Lamellar signal"
  ) +
  theme(
    strip.text = ggtext::element_markdown(size = 10), 
    legend.position = "bottom"
  )

```

We noted many small circular features with intense O and P signal with decreased S signal that were adjacent to airways. We suspected these were lamellar bodies rich in phospholipids. As an identifiable feature, we included these ROIs and are noting some differences in signal intensities. These data raise a question about comparisons among absolute signal intensities across samples. For example, the fibrotic bleomycin signal is decreased across all samples, but the ratiometric values for isotope enrichment are similar. If we wanted to compare the P signal across samples, does it make sense to normalize this by the C^14^N signal?

## To Do

- [ ] Add ROIs for non-nuclear interstitium
- [ ] Add ROIs for low P nuclei (appear to be ^15^N hotspots)
- [ ] Calculate correlation between ^15^N and ^2^H enrichment
- [ ] Identify regions of high and low relative enrichment

### 2024-02-16 manuscript/manuscript.qmd
working on isotope labeling data

@@
 
+## MCT inhibition reprograms lung metabolism *in vivo*
+
+To investigate metabolic perturbations following lactate transporter inhibition *in vivo*, we next performed metabolomic profiling of lung and plasma from mice treated with bleomycin and MCT inhibitors (**Supplemental Figure 9.1**).
+While bleomycin treatment substantially altered the lung metabolic profile, treatment with AZD3965 or VB124 had little additional impact on total metabolite levels.
+Unlike bleomycin alone or following AZD3965 treatment, VB124 treatment was associated with significant changes in several circulating metabolites (**Supplemental Figure 9.1I**).
+Among these, VB124 significantly increased circulating lactate, while AZD3965 was associated with decreased circulating lactate (**Figure 9A**).
+This pattern was also reflected in the lung, where only modest changes were observed following bleomycin, contrary to prior reports [@kottmann2012].
+Given the impact of our treatments on circulating lactate, we calculated the ratio of lung-to-plasma lactate.
+This analysis demonstrated the expected increase in lactate in fibrotic lungs with levels returning to baseline following the addition of VB124 (**Figure 9A**).
+
+We were surprised to observe a relative paucity of metabolic changes in whole lung lysates following MCT inhibition compared to our *in vitro* studies.
+We hypothesized that this might be a consequence of the cellular heterogeneity of MCT expression in the fibrotic lung.
+To examine this hypothesis, we performed multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following treatment with ^15^N-proline and ^2^H-glucose.
+MIMS enables the quantification of stable isotope tracer uptake into tissue biomass with subcellular spatial resolution [@steinhauser2012; @narendra2020].
+For three days prior to tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a tracer for tissue fibrosis and 50 mg ^2^H-glucose as a metabolic tracer.
+Lung tissue sections were subsequently imaged by nanoscale seconary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B**).
+
 \newpage

### 2024-02-16 R/figures.R

@@ -323,3 +323,78 @@ make_fig08s <- function(p1) {
       tags = NULL
     )
 }
+
+make_fig09 <- function(p1) {
+  design <- "aaa"
+  p1 + # p2 + p3 + p4 + p5 + p6 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(c(1.5), "in"),
+      heights = ggplot2::unit(1.5, "in")
+    )
+}
+
+make_fig09s1 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12) {
+  # design <- "abc \n def \n ghi \n jkl"
+  # p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + p10 + p11 + p12 +
+  #   theme_patchwork(
+  #     design = design,
+  #     widths = ggplot2::unit(2, "in"),
+  #     heights = ggplot2::unit(c(2, 2.5, 2, 2.5), "in")
+  #   )
+
+  design <- "ab \n cd \n ef"
+
+  extra <-
+    p1 + p2 + p5 + p6 + p9 + p10 +
+    patchwork::plot_annotation(
+      title = "Plasma",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
+    ) +
+    theme_patchwork(
+      design = design,
+      tags = list(c("A", "B", "E", "F", "I", "J"))
+    )
+  intra <-
+    p3 + p4 + p7 + p8 + p11 + p12 +
+    patchwork::plot_annotation(
+      title = "Lung",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
+    ) +
+    theme_patchwork(
+      design = design,
+      tags = list(c("C", "D", "G", "H", "K", "L"))
+    )
+
+  patchwork::wrap_elements(full = extra) +
+    patchwork::wrap_elements(full = intra) +
+    theme_patchwork(
+      design = "ab",
+      widths = ggplot2::unit(4, "in"),
+      heights = ggplot2::unit(6, "in"),
+      tags = NULL
+    )
+}
+
+make_fig09s2 <- function(p1, p2) {
+  design <- "aab"
+  p1 + p2 +
[…]

### 2024-02-16 R/mims.R

@@ -1,0 +1,44 @@
+# mims.R
+
+clean_mims <- function(files) {
+  names(files) <- stringr::str_replace(basename(files), "(-interleaved)*_data.csv", "")
+
+  df_slice <-
+    purrr::map(
+      files[!stringr::str_detect(files, "interleaved")],
+      \(x) readr::read_csv(x, show_col_types = FALSE)
+    ) |>
+    purrr::list_rbind(names_to = "file")
+
+  purrr::map(
+    files[stringr::str_detect(files, "interleaved")],
+    \(x) readr::read_csv(x, show_col_types = FALSE)
+  ) |>
+    purrr::list_rbind(names_to = "file") |>
+    dplyr::left_join(
+      df_slice,
+      by = c("file", "Roi group", "Roi tags", "Roi name")
+    ) |>
+    dplyr::rename(
+      feature = "Roi group",
+      tag = "Roi tags",
+      roi_id = "Roi name"
+    ) |>
+    tidyr::pivot_longer(
+      cols = tidyselect::matches("\\d+\\w+"),
+      names_to = "measurement",
+      values_to = "value"
+    ) |>
+    tidyr::separate_wider_delim("measurement", names = c("ion", "measurement"), delim = " ") |>
+    tidyr::pivot_wider(
+      names_from = "measurement",
+      values_from = "value"
+    ) |>
+    tidyr::separate_wider_delim(
+      cols = file,
+      names = c("group", "region", "image", "date", "sample"),
+      delim = "_"
+    ) |>
+    dplyr::mutate(group = toupper(group)) |>
+    refactor()
+}

### 2024-02-20 manuscript/manuscript.qmd
start analysis of MIMS images

@@
 To examine this hypothesis, we performed multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following treatment with ^15^N-proline and ^2^H-glucose.
-MIMS enables the quantification of stable isotope tracer uptake into tissue biomass with subcellular spatial resolution [@steinhauser2012; @narendra2020].
+MIMS enables the quantification of stable isotope tracer uptake into tissue biomass with subcellular spatial resolution [@steinhauser2012; @gyngard2019].
 For three days prior to tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a tracer for tissue fibrosis and 50 mg ^2^H-glucose as a metabolic tracer.
-Lung tissue sections were subsequently imaged by nanoscale seconary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B**).
+Lung tissue sections were subsequently imaged by nanoscale secondary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B**).
 

### 2024-02-20 R/figures.R

@@ -324,13 +324,13 @@ make_fig08s <- function(p1) {
     )
 }
 
-make_fig09 <- function(p1) {
-  design <- "aaa"
-  p1 + # p2 + p3 + p4 + p5 + p6 +
+make_fig09 <- function(p1, p2) {
+  design <- "aaa \n bbb"
+  p1 + p2 + # p3 + p4 + p5 + p6 +
     theme_patchwork(
       design = design,
       widths = ggplot2::unit(c(1.5), "in"),
-      heights = ggplot2::unit(1.5, "in")
+      heights = ggplot2::unit(c(1.5, 3), "in")
     )
 }
 

### 2024-02-20 R/mims.R

@@ -42,3 +42,49 @@ clean_mims <- function(files) {
     dplyr::mutate(group = toupper(group)) |>
     refactor()
 }
+
+read_mims <- function(file) {
+  img <- mims::mims_read(file)
+
+  n_ratio <- mims::mims_ratio(img, "12C15N", "12C14N")
+  h_ratio <- mims::mims_ratio(img, "12C22H", "12C21H")
+
+  idx <- which(img@metadata$globalMetadata$symbols == "12C14N")
+  mims::mims_sum_frames(img, idx) |>
+    EBImage::abind(n_ratio[,,1], along = 3) |>
+    EBImage::abind(h_ratio[,,1], along = 3)
+}
+
+extract_ratios <- function(img, nm) {
+  histo <- EBImage::normalize(img[,,1])
+
+  threshold <- EBImage::otsu(histo)
+  mask <- histo > threshold
+
+  # EBImage::display(mask)
+
+  n_ratio <- img[,,2][mask]
+  h_ratio <- img[,,3][mask]
+  tibble::tibble(
+    name = nm,
+    n_ratio = n_ratio,
+    h_ratio = h_ratio
+  )
+}
+
+format_ratios <- function(x) {
+  dplyr::bind_rows(x) |>
+    tidyr::separate(name, c("group", "region", "replicate"), sep = "_") |>
+    dplyr::mutate(
+      n_ratio_norm = n_ratio / 0.0037,
+      h_ratio_norm = h_ratio / 0.0001,
+      group = toupper(group),
+      region = factor(
+        region,
+        levels = c("alv", "fib"),
+        labels = c("Alveolar", "Fibrotic")
+      )
+    ) |>
+    refactor() |>
+    dplyr::mutate(group = forcats::fct_drop(group))
+}

### 2024-02-20 _targets.R

@@ -25,6 +25,24 @@ tar_option_set(
 # source functions
 tar_source()
 
+# branching
+mims_values <-
+  tibble::tibble(
+    names = c(
+      "azd_alv_1",
+      "azd_fib_1",
+      "azd_fib_2",
+      "bleo_alv_1",
+      "bleo_fib_1",
+      "control_alv_1",
+      "vb_alv_1",
+      "vb_fib_1"
+    ),
+    files = purrr::map_chr(
+      names,
+      \(x) raw_data_path(stringr::str_c(x, ".+\\d\\.nrrd"))
+    )
+  )
 
 # targets -----------------------------------------------------------------
 
@@ -1628,6 +1646,33 @@ list(
     mims_clean,
     clean_mims(mims_files)
   ),
+  tar_map(
+    values = mims_values,
+    names = names,
+    tar_target(
+      mims_img,
+      read_mims(files)
+    ),
+    tar_target(
+      mims_ratios,
+      extract_ratios(mims_img, names)
+    )
+  ),
+  tar_target(
+    mims_ratios,
+    format_ratios(
+      list(
+        mims_ratios_azd_alv_1,
+        mims_ratios_azd_fib_1,
+        mims_ratios_azd_fib_2,
+        mims_ratios_bleo_alv_1,
+        mims_ratios_bleo_fib_1,
+        mims_ratios_control_alv_1,
+        mims_ratios_vb_alv_1,
+        mims_ratios_vb_fib_1
+      )
+    )
+  ),
 
   # analysis ----------------------------------------------------------------
 
@@ -1671,7 +1716,8 @@ list(
       "dual-azd-erk-blot.png",      1,      0,      -0.05,  "erk_azd",
       "dual-azd-hif1a-blot.png",    1,      0,      -0.05,  "hif_azd",
       "lactate-sma-blot.png",       1.2,    0,      -0.05,  "sma_lac",
-      "dual-azd-kla-h3.png",        1,      0,      0,      "azd_kla"
+      "dual-azd-kla-h3.png",        1,      0,      0,      "azd_kla",
+      "mims-panel.png",             1,      0,      0,      "mims"
     ),
     names = names,
     tar_target(
@@ -1905,7 +1951,8 @@ list(
   tar_target(
     fig09,
     make_fig09(
-      metab_bleo_lacate_plot
+      metab_bleo_lacate_plot,
+      fig_img_mims
     ) |>
       write_figures("Figure 09"),
     format = "file"

### 2024-02-23 manuscript/manuscript.qmd
add mims data and figure to paper

@@
 We were surprised to observe a relative paucity of metabolic changes in whole lung lysates following MCT inhibition compared to our *in vitro* studies.
-We hypothesized that this might be a consequence of the cellular heterogeneity of MCT expression in the fibrotic lung.
-To examine this hypothesis, we performed multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following treatment with ^15^N-proline and ^2^H-glucose.
-MIMS enables the quantification of stable isotope tracer uptake into tissue biomass with subcellular spatial resolution [@steinhauser2012; @gyngard2019].
-For three days prior to tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a tracer for tissue fibrosis and 50 mg ^2^H-glucose as a metabolic tracer.
-Lung tissue sections were subsequently imaged by nanoscale secondary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B**).
+We hypothesized that this might be a consequence of measuring steady-state metabolite levels that do not account for differences in metabolic flux.
+To examine this hypothesis, we performed  multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following administration of ^15^N-proline and ^2^H-glucose or ^13^C-glucose.
+MIMS enables the quantification of stable isotope tracer flux into tissue biomass with subcellular spatial resolution [@steinhauser2012; @gyngard2019].
+For three days prior to tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a fibrosis tracer and 50 mg ^2^H-glucose or ^13^C-glucose as a metabolic tracer.
+Lung tissue sections were subsequently imaged by nanoscale secondary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B-C**).
+This labeling strategy was quite effective, demonstrating ~1.5-fold ^15^N, ~4.6-fold for ^2^H, and ~2.4-fold for ^13^C enrichment above natural isotope abundance.
+Pulmonary fibrosis was associated with a significant increase in ^15^N labeling from proline, consistent with increased collagen synthesis and deposition during the labeling period.
+Similar to the lung function and histological analyses, less ^15^N incorporation per tissue area was observed in drug-treated animals.
+Glucose labeling followed a similar pattern where lactate transporter inhibition caused a decrease in glucose incorporation into tissue biomass.
+Importantly, these data findings are not merely a consequence of decreased tissue fibrosis as we chose more fibrotic areas to image (**Figure 9B**) and the enrichment values are averaged over the tissue area.
 
@@
 Craig Thompson paper: Isotope labeling in lung fibroblasts
+
+<!-- spatial metabolomics -->
 

### 2024-02-23 R/mice.R

@@ -252,8 +252,14 @@ format_mice_col <- function(df, column = "Ashcroft_avg") {
     )
 }
 
-stats_histo <- function(df, measure) {
-    stats::lm(value ~ group, data = df) |>
+stats_histo <- function(df, measure, mixed = FALSE) {
+  if (mixed) {
+    m <- lmerTest::lmer(value ~ group + (1 | animal), data = df)
+  } else {
+    m <- stats::lm(value ~ group, data = df)
+  }
+
+  m |>
     emmeans::emmeans(~ group) |>
     emmeans::contrast(
       method = list(

### 2024-02-23 R/mims.R

@@ -50,13 +50,16 @@ read_mims <- function(file) {
   n_ratio <- h_ratio <- c_ratio <- empty
 
   if ("12C15N" %in% symbols) {
-    n_ratio <- mims::mims_ratio(img, "12C15N", "12C14N")[,,1]
+    n_ratio <-
+      mims::mims_ratio(img, "12C15N", "12C14N")[,,1]
   }
   if ("12C22H" %in% symbols) {
-    h_ratio <- mims::mims_ratio(img, "12C22H", "12C21H")[,,1]
+    h_ratio <-
+      mims::mims_ratio(img, "12C22H", "12C21H")[,,1]
   }
   if ("12C13C" %in% symbols) {
-    c_ratio <- mims::mims_ratio(img, "12C13C", "12C2")[,,1]
+    c_ratio <-
+      mims::mims_ratio(img, "12C13C", "12C2")[,,1]
   }
 
   idx <- which(symbols == "12C14N")
@@ -76,19 +79,26 @@ extract_ratios <- function(img, nm) {
 
   n_ratio <- img[,,2][mask]
   h_ratio <- img[,,3][mask]
+  c_ratio <- img[,,4][mask]
   tibble::tibble(
     name = nm,
     n_ratio = n_ratio,
-    h_ratio = h_ratio
+    h_ratio = h_ratio,
+    c_ratio = c_ratio
   )
 }
 
 format_ratios <- function(x) {
-  dplyr::bind_rows(x) |>
-    tidyr::separate(name, c("group", "region", "replicate"), sep = "_") |>
+  x |>
+    tidyr::separate(
+      name,
+      c("group", "animal", "region", "replicate"),
+      sep = "_"
+    ) |>
     dplyr::mutate(
-      n_ratio_norm = n_ratio / 0.0037,
-      h_ratio_norm = h_ratio / 0.0001,
+      n_ratio_norm = n_ratio / (0.368/99.632),
+      h_ratio_norm = h_ratio / (0.0115/99.9885),
+      c_ratio_norm = c_ratio / (1.07/98.93),
       group = toupper(group),
       region = factor(
         region,
@@ -96,6 +106,69 @@ format_ratios <- function(x) {
         labels = c("Alveolar", "Fibrotic")
       )
     ) |>
+    tidyr::pivot_longer(
+      -c("group":"replicate"),
+      names_to = c("element", ".value"),
+      names_pattern = c("(\\w)_(.+)"),
+      cols_vary = "slowest"
+    ) |>
[…]

### 2024-02-23 _targets.R

@@ -25,58 +25,6 @@ tar_option_set(
 # source functions
 tar_source()
 
-# branching
-mims_values <-
-  tibble::tibble(
-    names = c(
-      "azd_1_alv_1",
-      "azd_1_fib_1",
-      "azd_1_fib_2",
-      "azd_2_fib_1",
-      "azd_3_fib_1",
-      "azd_3_fib_2",
-      "azd_3_fib_3",
-      "azd_3_fib_4",
-      "azd_3_fib_5",
-      "azd_3_fib_6",
-      "azd_3_fib_7",
-      "azd_3_fib_8",
-      "azd_3_fib_9",
-      "bleo_1_alv_1",
-      "bleo_1_fib_1",
-      "bleo_2_fib_1",
-      "bleo_2_fib_2",
-      "bleo_3_fib_1",
-      "bleo_3_fib_2",
-      "bleo_3_fib_3",
-      "bleo_3_fib_4",
-      "bleo_3_fib_5",
-      "bleo_3_fib_6",
-      "bleo_3_fib_7",
-      "control_1_alv_1",
-      "control_2_alv_1",
-      "control_2_alv_2",
-      "control_3_alv_1",
-      "control_3_alv_2",
-      "control_3_alv_3",
-      "vb_1_alv_1",
-      "vb_1_fib_1",
-      "vb_2_fib_1",
-      "vb_3_fib_1",
-      "vb_3_fib_2",
-      "vb_3_fib_3",
-      "vb_3_fib_4",
-      "vb_3_fib_5",
-      "vb_3_fib_6",
-      "vb_3_fib_7",
-      "vb_3_fib_8",
-      "vb_3_fib_9"
-    ),
-    files = purrr::map_chr(
-      names,
-      \(x) raw_data_path(stringr::str_c(x, ".+(\\d|r)\\.nrrd"))
-    )
-  )
 
 # targets -----------------------------------------------------------------
 
@@ -1681,33 +1629,144 @@ list(
     clean_mims(mims_files)
   ),
   tar_map(
-    values = mims_values,
+    values = list(
+      names = c(
+        "azd_1_alv_1",
+        "azd_1_fib_1",
+        "azd_1_fib_2",
+        "azd_2_fib_1",
+        "azd_3_fib_1",
+        "azd_3_fib_2",
+        "azd_3_fib_3",
+        "azd_3_fib_4",
+        "azd_3_fib_5",
+        "azd_3_fib_6",
[…]

### 2024-02-23 manuscript/manuscript.qmd
add VB253 raw data

@@
 Glucose labeling followed a similar pattern where lactate transporter inhibition caused a decrease in glucose incorporation into tissue biomass.
-Importantly, these data findings are not merely a consequence of decreased tissue fibrosis as we chose more fibrotic areas to image (**Figure 9B**) and the enrichment values are averaged over the tissue area.
+Importantly, these  findings are not merely a consequence of decreased tissue fibrosis as we chose more fibrotic areas to image (**Figure 9B**) and the enrichment values are normalized to tissue area.
+
+## Novel MCT4 inhibitor improves experimental pulmonary fibrosis
+
+Our preclinical data suggest that MCT4 inhibition, as a single drug therapy, is more potently antifibrotic than MCT1 inhibition, both in terms of inhibiting myofibroblast differentiation and decreasing bleomycin-induced fibrosis.
+VB124
 

### 2024-02-23 R/figures.R

@@ -398,3 +398,22 @@ make_fig09s2 <- function(p1, p2) {
     )
 }
 
+make_fig10 <- function(p1, p2) {
+  design <- "ab"
+  p1 + p2 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(1.5, "in"),
+      heights = ggplot2::unit(1.5, "in")
+    )
+}
+
+make_fig10s <- function(p1) {
+  design <- "a"
+  p1 +
+    theme_patchwork(
+      design = design,
+      widths = ggplot2::unit(1.5, "in"),
+      heights = ggplot2::unit(1.5, "in")
+    )
+}

### 2024-02-23 R/plots.R

@@ -220,6 +220,8 @@ clrs <- c(
   "TGFβ\nVB"      = "#B6992D",
   "Ctl\nAZD/VB"   = "#FABFD2",
   "TGFβ\nAZD/VB"  = "#D37295",
+  "VB253"         = "#E15759",
+  "Nintedanib"    = "#499894",
   NULL
 )
 

### 2024-02-23 R/vb.R

@@ -1,0 +1,97 @@
+# vb.R
+
+read_vb_invitro_sma <- function(file) {
+  readr::read_csv(file, show_col_types = FALSE) |>
+    tidyr::pivot_longer(
+      tidyselect::matches("_\\d{1}$"),
+      names_to = c("measurement", "replicate"),
+      names_pattern = "(.+)_(.)",
+      values_to = "value"
+    ) |>
+    tidyr::pivot_wider(
+      names_from = "measurement",
+      values_from = "value"
+    ) |>
+    dplyr::mutate(
+      type = ifelse(stringr::str_detect(donor, "IPF"), "IPF", "Ctl"),
+      .before = donor
+    ) |>
+    dplyr::group_by(drug, donor) |>
+    dplyr::mutate(
+      dplyr::across(
+        tidyselect::any_of(c("sma", "nuclei")),
+        \(x) x / mean(x[dose == min(dose)]),
+        .names = "{.col}_norm"
+      )
+    ) |>
+    dplyr::group_by(drug, dose, type, donor) |>
+    dplyr::summarise(dplyr::across(
+      tidyselect::any_of(c("sma", "sma_norm", "nuclei", "nuclei_norm")), mean
+    )) |>
+    dplyr::mutate(
+      drug = factor(drug, levels = c("VB253", "Nintedanib"))
+    )
+}
+
+plot_vb_invitro_sma <- function(df, source = "IPF", ytitle = "α-SMA fold change") {
+  df |>
+    dplyr::filter(type == source) |>
+    dplyr::group_by(dose) |>
+    dplyr::mutate(width = 0.025 * dplyr::n()) |>
+    ggplot2::ggplot() +
+    ggplot2::aes(
+      x = dose,
+      y = sma_norm,
+      color = drug
+    ) +
+    ggplot2::geom_smooth(
+      method = drc::drm,
+      method.args = list(
+        fct = drc::L.4()
+      ),
+      linewidth = 0.5,
+      se = FALSE,
+      show.legend = FALSE
+    ) +
+    ggplot2::geom_errorbar(
+      ggplot2::aes(
+        color = .data$drug,
+        width = .data$width
+      ),
+      stat = "summary",
+      fun.data = ggplot2::mean_se,
+      # width = 0.1,
+      linewidth = 0.25,
+      position = ggplot2::position_dodge(width = 0.1),
+      show.legend = FALSE
+    ) +
+    ggplot2::geom_point(
[…]

### 2024-02-26 R/figures.R
add VB253 data

@@ -398,19 +398,86 @@ make_fig09s2 <- function(p1, p2) {
     )
 }
 
-make_fig10 <- function(p1, p2) {
-  design <- "ab"
-  p1 + p2 +
+make_fig10 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
+  # design <- "abcc \n defg \n hijk"
+  # p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + p10 + p11 +
+  #   theme_patchwork(
+  #     design = design,
+  #     widths = ggplot2::unit(1.5, "in"),
+  #     heights = ggplot2::unit(1.5, "in")
+  #   )
+
+  top <-
+    ((p1 + p2) +
+       patchwork::plot_layout(guides = "collect") &
+       ggplot2::theme(
+         legend.position = "bottom",
+         legend.box.margin = ggplot2::margin(t = -10)
+       )) -
+    p3 +
     theme_patchwork(
-      design = design,
-      widths = ggplot2::unit(1.5, "in"),
-      heights = ggplot2::unit(1.5, "in")
+      # design = "abc",
+      widths = c(2, 2),
+      tags = list(c("A", "B", "C")),
+      guides = "collect"
+    ) &
+    ggplot2::theme(legend.position = "bottom")
+
+  mid <-
+    p4 + p5 + p6 + p7 +
+    patchwork::plot_annotation(
+      title = "Young mice",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
+    ) +
+    theme_patchwork(
+      design = "abcd",
+      tags = list(c("D", "E", "F", "G"))
+    )
+
+  bot <-
+    p8 + p9 + p10 + p11 +
+    patchwork::plot_annotation(
+      title = "Aged mice",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
+    ) +
+    theme_patchwork(
+      design = "abcd",
+      tags = list(c("H", "I", "J", "K"))
+    )
+
+  patchwork::wrap_elements(full = top) +
+    patchwork::wrap_elements(full = mid) +
[…]

### 2024-02-26 R/mice.R

@@ -252,7 +252,7 @@ format_mice_col <- function(df, column = "Ashcroft_avg") {
     )
 }
 
-stats_histo <- function(df, measure, mixed = FALSE) {
+stats_histo <- function(df, measure, mixed = FALSE, comps = comps_bleo_2) {
   if (mixed) {
     m <- lmerTest::lmer(value ~ group + (1 | animal), data = df)
   } else {
@@ -262,17 +262,14 @@ stats_histo <- function(df, measure, mixed = FALSE) {
   m |>
     emmeans::emmeans(~ group) |>
     emmeans::contrast(
-      method = list(
-        Ctl = c(-1, 1, 0, 0),
-        AZD = c(0, -1, 1, 0),
-        VB =  c(0, -1, 0, 1)
-      ),
+      method = comps,
       adjust = "dunnettx"
     ) |>
     tibble::as_tibble() |>
     dplyr::rename(group = .data$contrast) |>
     dplyr::mutate(measurement = measure) |>
-    annot_stats()
+    annot_stats() |>
+    refactor()
 }
 
 
@@ -287,7 +284,8 @@ plot_mice <- function(
     y = "value",
     ytitle = NULL
 ) {
-  df |>
+  p <-
+    df |>
     dplyr::filter(measurement == measure) |>
     ggplot2::ggplot() +
     ggplot2::aes(
@@ -323,15 +321,6 @@ plot_mice <- function(
       stroke = 0.25,
       show.legend = FALSE
     ) +
-    ggplot2::geom_text(
-      data = dplyr::filter(stats, measurement == measure),
-      ggplot2::aes(
-        y = y,
-        vjust = vjust,
-        label = label
-      ),
-      color = "black"
-    ) +
     ggplot2::labs(
       x = NULL,
       y = ytitle
@@ -348,4 +337,19 @@ plot_mice <- function(
     ggplot2::coord_cartesian(clip = "off") +
     theme_plot() +
     NULL
+
+  if (!is.null(stats)) {
+    p <-
+      p +
+      ggplot2::geom_text(
+        data = dplyr::filter(stats, measurement == measure),
+        ggplot2::aes(
+          y = y,
+          vjust = vjust,
+          label = label
+        ),
+        color = "black"
+      )
+  }
+  p
 }

### 2024-02-26 R/plots.R

@@ -221,7 +221,8 @@ clrs <- c(
   "Ctl\nAZD/VB"   = "#FABFD2",
   "TGFβ\nAZD/VB"  = "#D37295",
   "VB253"         = "#E15759",
-  "Nintedanib"    = "#499894",
+  "NIN"           = "#499894",
+  "PFD"           = "#B07AA1",
   NULL
 )
 

### 2024-02-26 R/utils.R

@@ -21,14 +21,24 @@ manuscript_path <- function(nm) {
   list_files(nm = nm, path = "manuscript")
 }
 
+# plot_image <- function(img, scale = 1, hjust = 0, vjust = 0) {
+#   cowplot::ggdraw() +
+#     cowplot::draw_image(
+#       magick::image_read(img),
+#       scale = scale,
+#       hjust = hjust,
+#       vjust = vjust
+#     )  +
+#     theme_plot() +
+#     ggplot2::theme(
+#       panel.border = ggplot2::element_blank(),
+#       axis.line = ggplot2::element_blank()
+#     )
+# }
+
 plot_image <- function(img, scale = 1, hjust = 0, vjust = 0) {
-  cowplot::ggdraw() +
-    cowplot::draw_image(
-      magick::image_read(img),
-      scale = scale,
-      hjust = hjust,
-      vjust = vjust
-    )  +
+  png::readPNG(img, native = TRUE) |>
+    patchwork::wrap_elements(plot = _) +
     theme_plot() +
     ggplot2::theme(
       panel.border = ggplot2::element_blank(),
@@ -137,6 +147,7 @@ refactor <- function(df) {
       "TGFβ\nAZD/VB" = "TGFβ\nAZD/VB",
       "Ctl" = "Ctl",
       "Ctl" = "CONTROL",
+      "Ctl" = "Veh",
       "TGFβ" = "TGFβ",
       "IPF" = "IPF",
       "Bleo" = "Bleo",
@@ -154,6 +165,9 @@ refactor <- function(df) {
       "Ctl\nVB" = "control\nVB124",
       "Ctl\nVB" = "control\nVB",
       "VB" = "VB",
+      "VB253" = "VB253",
+      "NIN" = "Nintedanib",
+      "PFD" = "Pirfenidone",
       "Ctl\nAZD/VB" = "control\nAZD/VB",
       "Ctl\nAZD/VB" = "control\nDual",
       "Ctl\nMed" = "Ctl\nMed",

### 2024-02-26 R/vb.R

@@ -29,11 +29,20 @@ read_vb_invitro_sma <- function(file) {
       tidyselect::any_of(c("sma", "sma_norm", "nuclei", "nuclei_norm")), mean
     )) |>
     dplyr::mutate(
-      drug = factor(drug, levels = c("VB253", "Nintedanib"))
+      drug = factor(
+        drug,
+        levels = c("VB253", "Nintedanib"),
+        labels = c("VB253", "NIN")
+      )
     )
 }
 
-plot_vb_invitro_sma <- function(df, source = "IPF", ytitle = "α-SMA fold change") {
+plot_vb_invitro_sma <- function(
+    df,
+    source = "IPF",
+    y = "sma_norm",
+    ytitle = "α-SMA fold change"
+) {
   df |>
     dplyr::filter(type == source) |>
     dplyr::group_by(dose) |>
@@ -41,11 +50,12 @@ plot_vb_invitro_sma <- function(df, source = "IPF", ytitle = "α-SMA fold change
     ggplot2::ggplot() +
     ggplot2::aes(
       x = dose,
-      y = sma_norm,
+      y = .data[[y]],
       color = drug
     ) +
     ggplot2::geom_smooth(
       method = drc::drm,
+      formula = y ~ x,
       method.args = list(
         fct = drc::L.4()
       ),
@@ -94,4 +104,15 @@ plot_vb_invitro_sma <- function(df, source = "IPF", ytitle = "α-SMA fold change
     ) +
     ggplot2::coord_cartesian(clip = "off") +
     theme_plot()
+}
+
+read_vb_mice <- function(file) {
+  readr::read_csv(file, show_col_types = FALSE) |>
+    refactor() |>
+    tidyr::pivot_longer(
+      c("ashcroft", "sma", "Penh", "lactate"),
+      names_to = "measurement",
+      values_to = "value"
+    ) |>
+    dplyr::filter(!is.na(value))
 }

### 2024-02-26 R/comps.R
add VB253 data

@@ -19,6 +19,19 @@ comps_bleo_2 <- list(
   "VB" =  c(0, -1, 0, 1)
 )
 
+comps_bleo_3 <- list(
+  # "Ctl" =           c(1, -1, 0, 0),
+  "VB253" =         c(0, -1, 1, 0),
+  "Pirfenidone" =   c(0, -1, 0, 1)
+)
+
+comps_bleo_4 <- list(
+  # "Ctl" =           c(1, -1, 0, 0, 0),
+  "VB253" =         c(0, -1, 1, 0, 0),
+  "Nintedanib" =    c(0, -1, 0, 1, 0),
+  "Pirfenidone" =   c(0, -1, 0, 0, 1)
+)
+
 comps_nsaid <- list(
   "DIC.Ctl" =  c(-1, 0, 1, 0, 0, 0, 0, 0),
   "DIC.TGFβ" = c(0, -1, 0, 1, 0, 0, 0, 0),

### 2024-02-26 R/mims.R

@@ -29,17 +29,24 @@ clean_mims <- function(files) {
       names_to = "measurement",
       values_to = "value"
     ) |>
-    tidyr::separate_wider_delim("measurement", names = c("ion", "measurement"), delim = " ") |>
+    tidyr::separate_wider_delim(
+      "measurement",
+      names = c("ion", "measurement"),
+      delim = " "
+    ) |>
     tidyr::pivot_wider(
       names_from = "measurement",
       values_from = "value"
     ) |>
     tidyr::separate_wider_delim(
       cols = file,
-      names = c("group", "region", "image", "date", "sample"),
+      names = c("group", "replicate", "region", "image", "date", "sample"),
       delim = "_"
     ) |>
-    dplyr::mutate(group = toupper(group)) |>
+    dplyr::mutate(
+      replicate = factor(replicate),
+      group = toupper(group)
+    ) |>
     refactor()
 }
 

### 2024-02-26 _targets.R

@@ -1601,7 +1601,7 @@ list(
     ),
     tar_target(
       mice_mcti_stats,
-      stats_histo(mice_mcti, measure = names)
+      stats_histo(mice_mcti, measure = names, comps = rlang::sym("comps_bleo_2"))
     ),
     tar_target(
       mice_mcti_plot,
@@ -1785,10 +1785,18 @@ list(
       names = c("ipf", "ctl")
     ),
     names = names,
-    tar_target(
-      vb_invitro_sma_plot,
-      plot_vb_invitro_sma(vb_invitro_sma_data, type),
-      format = "rds"
+    tar_map(
+      values = list(
+        y = c("sma_norm", "nuclei_norm"),
+        ytitle = c("α-SMA fold change", "Nuclei fold change"),
+        names = c("sma", "nuclei")
+      ),
+      names = names,
+      tar_target(
+        vb_invitro_plot,
+        plot_vb_invitro_sma(vb_invitro_sma_data, type, y, ytitle),
+        format = "rds"
+      )
     )
   ),
   tar_target(
@@ -1802,8 +1810,52 @@ list(
   ),
   tar_target(
     vb_invitro_smad_plot,
-    plot_vb_invitro_sma(vb_invitro_smad_data, "IPF", "Nuclear Smad3"),
+    plot_vb_invitro_sma(vb_invitro_smad_data, "IPF", ytitle = "Nuclear Smad3"),
     format = "rds"
+  ),
+  tar_target(
+    vb_mice_file,
+    raw_data_path("vb_mice.csv"),
+    format = "file"
+  ),
+  tar_target(
+    vb_mice_data,
+    read_vb_mice(vb_mice_file)
+  ),
+  tar_map(
+    values =
+      tibble::tribble(
+        ~ages,   ~measure,   ~ytitle,                       ~comps,
+        "young", "ashcroft", "Ashcroft score",              rlang::sym("comps_bleo_3"),
+        "young", "sma",      "α-SMA area",                  rlang::sym("comps_bleo_3"),
+        "young", "Penh",     "Penh",                        rlang::sym("comps_bleo_4"),
+        "old",   "ashcroft", "Ashcroft score",              rlang::sym("comps_bleo_4"),
+        "old",   "sma",      "α-SMA area",                  rlang::sym("comps_bleo_4"),
+        "old",   "lactate",  "Lactate (nmol / mg protein)", rlang::sym("comps_bleo_4")
+      ) |>
+      tidyr::unite(names, ages, measure, remove = FALSE),
[…]

### 2024-03-02 manuscript/manuscript.qmd
add aged mice histology examples

@@
     fig08, 
-    # fig10,
-    # fig11,
+    fig09, 
+    fig10,
     # resources_table,

### 2024-03-02 _targets.R

@@ -1902,7 +1902,9 @@ list(
       "lactate-sma-blot.png",       1.2,    0,      -0.05,  "sma_lac",
       "dual-azd-kla-h3.png",        1,      0,      0,      "azd_kla",
       "mims-panel.png",             1.05,   0.05,   0,      "mims",
-      "vb-bleo-timeline.png",       1,      0,      0,      "vb_timeline"
+      "vb-bleo-timeline.png",       1,      0,      0,      "vb_timeline",
+      "bleo-vb-young-sma.png",      1,      0,      0,      "vb_young_sma",
+      "bleo-vb-aged-sma.png",        1,      0,      0,     "vb_aged_sma"
     ),
     names = names,
     tar_target(
@@ -2181,8 +2183,8 @@ list(
       vb_invitro_plot_nuclei_ipf,
       fig_img_vb_timeline,
       vb_mice_plot_young_Penh,
-      patchwork::plot_spacer(),
       vb_mice_plot_young_ashcroft,
+      fig_img_vb_young_sma,
       vb_mice_plot_young_sma +
         ggplot2::scale_y_continuous(
           labels = scales::label_percent(),
@@ -2190,8 +2192,8 @@ list(
           expand = ggplot2::expansion(c(0, 0)),
           limits = nice_limits
         ),
-      patchwork::plot_spacer(),
       vb_mice_plot_old_ashcroft,
+      fig_img_vb_aged_sma,
       vb_mice_plot_old_sma +
         ggplot2::scale_y_continuous(
           labels = scales::label_percent(),

### 2024-03-03 manuscript/manuscript.qmd
finish results draft and remake all targets

@@
 
-## Novel MCT4 inhibitor improves experimental pulmonary fibrosis
+## Novel MCT4 inhibitor mitigates experimental pulmonary fibrosis
 
-Our preclinical data suggest that MCT4 inhibition, as a single drug therapy, is more potently antifibrotic than MCT1 inhibition, both in terms of inhibiting myofibroblast differentiation and decreasing bleomycin-induced fibrosis.
-VB124
+Our preclinical data suggest that MCT4 inhibition, as a single target therapy, is more potently antifibrotic than MCT1 inhibition, both in terms of inhibiting myofibroblast differentiation *in vitro* and decreasing bleomycin-induced fibrosis.
+VB253 is a novel inhibitor of MCT4 with similar selectivity, but ~10-fold increased potency, for MCT4 inhibition.
+Like VB124, VB253 dose-dependently decreases **TGFβ-stimulated?** α-SMA in human IPF lung fibroblasts *ex vivo* (**Figure 10A**).
+Given the robust effect of MCT4 inhibition on myofibroblast differentiation, we compared VB253 to nintedanib, an anti-fibrotic medication clinically indicated for pulmonary fibrosis (**Figure 10A-B**).
+Nintedanib inhibits TGFβ-mediated fibroblast to myofibroblast differentiation and decreases Col1a1 expression *in vitro* and *in vivo*  [@wollin2014].
+While both drugs potently attenuated α-SMA expression, nintedanib demonstrated a moderate cytotoxic effect that was not observed with VB253 (**Figure 10B**).
+Nintedanib prevents myofibroblast differentiation, in part, through inhibition of TGFβ receptor phosphorylation and activation of Smad-dependent signaling pathways [@rangarajan2016].
+Consistent with these findings, nintedanib dose-dependently attenuates TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 have no impact on this signaling pathway (**Figure 6**, **Supplemental Figure 10**).
+
+We next assessed the ability of VB253 to inhibit bleomycin-induced pulmonary fibrosis.
+In experiments conducted independently from those reported above, VB253 was administered to young mice (8-10 weeks) beginning on day 7 following bleomycin (**Figure 10C**).
+Mice treated with nintedanib or pirfenidone provided comparisons to currently available therapies for pulmonary fibrosis.
+Three weeks after bleomycin administration, whole-body plethysmography was performed to assess breathing patterns in unrestrained mice.
+Enhanced pause (Penh) is a dimensionless index describing airflow during tidal breathing found to increase significantly in bleomycin-treated mice [@vanoirbeek2010].
+VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterens in bleomycin-treated mice (**Figure 10D**).
+This was associated with decreased pulmonary fibrosis by histologic assessment (**Figure 10E**) and α-SMA expression (**Figure 10F-G**).
+Morevover, the effects of VB253 compared favorably to the antifibrotic effects of nintedanib and pirfenidone.
+
+Compared to young mice, bleomycin-induced fibrosis in aged mice is more severe and does not resolve spontaneously [@redente2011].
+Given that IPF incidence increases with age, some studies have suggested such aged mouse models may be more clinically relevant.
+Thus, we examined the antifibrotic effects of VB253 in aged (60+ weeks) mice.
+Similar to young mice, VB253 decreased fibrosis severity as quantified by Ashcroft score and α-SMA expression (**Figure 10H-J**).
+The magnitude of improvement was similar to that observed with nintedanib and pirfenidone.
+As anticipated, VB2253 decreased total lung lactate similar to VB124 (**Figure 9A**).
+Taken together, these data provide compelling preclinical data for the development of lactate transporter inhibition as a therapeutic strategy for fibrotic lung disease.
 

### 2024-03-03 R/figures.R

@@ -469,18 +469,13 @@ make_fig10 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
 }
 
 
-
-
-
-
-
-
-make_fig10s <- function(p1, p2) {
-  design <- "ab"
-  p1 + p2 +
+make_fig10s <- function(p1) {
+  design <- "a"
+  p1 +
     theme_patchwork(
       design = design,
       widths = ggplot2::unit(1.5, "in"),
-      heights = ggplot2::unit(1.5, "in")
+      heights = ggplot2::unit(1.5, "in"),
+      tags = NULL
     )
 }

### 2024-03-03 R/mice.R

@@ -266,7 +266,7 @@ stats_histo <- function(df, measure, mixed = FALSE, comps = comps_bleo_2) {
       adjust = "dunnettx"
     ) |>
     tibble::as_tibble() |>
-    dplyr::rename(group = .data$contrast) |>
+    dplyr::rename(group = "contrast") |>
     dplyr::mutate(measurement = measure) |>
     annot_stats() |>
     refactor()

### 2024-03-03 _targets.R
finish results draft and remake all targets

@@ -1542,7 +1542,7 @@ list(
     format = "rds"
   ),
 
-  # mice_vent ---------------------------------------------------------------
+  # mice vent ---------------------------------------------------------------
 
   tar_target(
     mice_vent,
@@ -1601,7 +1601,7 @@ list(
     ),
     tar_target(
       mice_mcti_stats,
-      stats_histo(mice_mcti, measure = names, comps = rlang::sym("comps_bleo_2"))
+      stats_histo(mice_mcti, measure = names, comps = comps_bleo_2)
     ),
     tar_target(
       mice_mcti_plot,
@@ -2206,14 +2206,14 @@ list(
       write_figures("Figure 10"),
     format = "file"
   ),
-  # tar_target(
-  #   fig10s,
-  #   make_fig10s(
-  #     vb_invitro_smad_plot
-  #   ) |>
-  #     write_figures("Figure 10.supplement"),
-  #   format = "file"
-  # ),
+  tar_target(
+    fig10s,
+    make_fig10s(
+      vb_invitro_smad_plot
+    ) |>
+      write_figures("Figure 10.supplement"),
+    format = "file"
+  ),
 
   # manuscript --------------------------------------------------------------
 

### 2024-03-11 R/mids.R
finish Discussion first draft

@@ -658,3 +658,87 @@ plot_plasma_ratio <- function(mids) {
     theme_plot() +
     NULL
 }
+
+plot_tgf_mids <- function(df) {
+  metab <- c(
+    "FBP",
+    "6PG",
+    "PYR",
+    "LAC",
+    "CIT",
+    "2OG",
+    "SUC",
+    "MAL",
+    "ASP",
+    "ALA",
+    "SER",
+    "GLY",
+    "GLN",
+    "GLU",
+    "PRO"
+  )
+
+  x <-
+    df |>
+    dplyr::filter(tracer != "glc2") |>
+    dplyr::filter(treatment == "Veh") |>
+    dplyr::filter(metabolite %in% metab) |>
+    dplyr::mutate(
+      metabolite = factor(metabolite, levels = metab),
+      tracer = factor(
+        tracer,
+        levels = c("glc6", "lac3", "gln5"),
+        labels = c(
+          "[U-^13^C~6~]-glucose",
+          "[U-^13^C~3~]-lactate",
+          "[U-^13^C~5~]-glutamine"
+        )
+      )
+    )
+
+  stats <-
+    x |>
+    dplyr::group_by(tracer, metabolite) |>
+    tidyr::nest() |>
+    dplyr::mutate(
+      stats = purrr::map(data, \(x) {
+        stats::lm(mid_corr ~ condition * isotope, data = x) |>
+          emmeans::emmeans(~ condition * isotope) |>
+          emmeans::contrast(
+            method = "pairwise",
+            by = "isotope",
+            adjust = "mvt"
+          ) |>
+          broom::tidy() |>
+          dplyr::select("isotope", "p.value")
+      })
+    ) |>
+    tidyr::unnest(c(stats)) |>
+    dplyr::mutate(
+      condition = NA
+    ) |>
+    annot_stats()
+
+  plot_two_factor(
+    df = x,
+    stats = stats,
+    x = "isotope",
+    y = "mid_corr",
+    ytitle = "Mole fraction"
+  ) +
+    ggplot2::facet_grid(
+      cols = ggplot2::vars(tracer),
+      rows = ggplot2::vars(metabolite)
+    ) +
+    ggplot2::scale_y_continuous(
+      breaks = seq(0, 1, by = 0.25),
+      expand = ggplot2::expansion(c(0, 0.15)),
+      limits = nice_limits
+    ) +
+    ggplot2::theme(
+      strip.text = ggtext::element_markdown(),
[…]

### 2024-03-11 _targets.R

@@ -1252,6 +1252,20 @@ list(
     ),
     NULL
   ),
+  tar_target(
+    mid_tgf_plot,
+    plot_tgf_mids(mid_mcti_mids_corr) +
+      theme_patchwork(
+        widths = ggplot2::unit(7.5, "in"),
+        heights = ggplot2::unit(11, "in"),
+        tags = NULL
+      ),
+    format = "rds"
+  ),
+  tar_target(
+    mid_tgf_plot_img,
+    write_figures(mid_tgf_plot, "Figure 04.supplement.2")
+  ),
 
   # mid bleo ----------------------------------------------------------------
 
@@ -2032,7 +2046,7 @@ list(
       msea_table_img_vb_intra,
       mid_mcti_moi_plot_gln5
     ) |>
-      write_figures("Figure 04.supplement"),
+      write_figures("Figure 04.supplement.1"),
     format = "file"
   ),
 

### 2024-03-11 manuscript/manuscript.qmd
finish Discussion first draft

@@
 A family of monocarboxylate transporters (MCT1-4) maintains lactate homeostasis and small molecule inhibitors of these transporters have been under active clinical development for oncologic applications where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
-Importantly, compared to previously investigated inhibitors of glycolysis, MCT inhibitors exhibit favorable pharmacologic profiles and have successfully translated to human clinic trials [@halford2023].
+Importantly, compared to previously investigated inhibitors of glycolysis, MCT inhibitors exhibit favorable pharmacologic profiles and have successfully translated to human clinical trials [@halford2023].
 Before the promise of this therapeutic approach in IPF may be realized, however, the preclinical efficacy and molecular mechanisms-of-action of lactate transport inhibitors must be demonstrated experimentally.
@@
 Inhibition of these transporters attenuated bleomycin-induced lung fibrosis *in vivo* and TGFβ-induced myofibroblast differentiation *in vitro*, where MCT4 inhibition demonstrated increased therapeutic efficacy.
-*Using stable isotope tracing, high-resolution spatial metabolomics, we find...* Based on these findings, we report a novel MCT4 inhibitor suitable for human clinical studies.
+Using metabolomics, stable isotope tracing, and high-resolution spatial metabolomic imaging, we find that lactate transport inhibition promotes glucose oxidation.
+These metabolic changes signal, in part, through increased protein lactylation and decreased intracellular reactive oxygen species (ROS) production.
+Based on these findings, we report a novel MCT4 inhibitor suitable for human clinical studies.
 Together, these data illuminate a novel approach to disrupt the metabolic program fueling the exuberant fibrosis at the core of IPF and provide proof-of-concept to target lactate transporters in the treatment of this deadly disease.
@@
 Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which transfer electrons from lactate to NADH.
-Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression [@jain2013, @schworer2020], we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
+Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression [@jain2013; @schworer2020], we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
 Consistent with an increase in intracellular lactate with MCT4 inhibition, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A**), which was associated with a reduction in total ROS as measured by the CellROX fluorescent probe (**Figure 7B**).
+AZD3965, but not VB124, also decreased mitochondrial superoxide production (**Supplemental Figure 7B**).
 By contrast, the NADPH/NADP^+^ was unchanged by either TGFβ or MCT inhibitors (**Supplemental Figure 7A**).
-Notably, we did not observe TGFβ-dependent increases in total ROS production nor an impact of TGFβ or inhibitors on mitochondrial superoxide production (**Supplemental Figure 7B**).
+Notably, we did not observe TGFβ-dependent increases in total ROS production in our model system.
 
@@
 In this model, proline synthesis from glutamine consumes reducing equivalents from NADPH and NADH, thereby ameliorating reductive stress and decreasing ROS production.
-Compared to this prior study, we did not find that TGFβ increased proline in our primary lung fibroblasts (**Supplemental Figure 7C**).
-Overall, the fractional labeling of proline from [U-^13^C~5~]-glutamine was modest (~10%) compared to the 40% previously reported (**Supplemental Figure 7D**).
+Compared to this prior study, we did not find that TGFβ increased proline in primary lung fibroblasts (**Supplemental Figure 7C**).
+Overall, the fractional labeling of proline from [U-^13^C~5~]-glutamine was modest at 10% compared to the 40% previously reported (**Supplemental Figure 7D**).
 While we observed trends toward increased proline production from glutamine with MCT4 inhibition, this mechanism seems unlikely to account for the effects of MCT inhibitors on ROS generation.
@@
 The magnitude of improvement was similar to that observed with nintedanib and pirfenidone.
-As anticipated, VB2253 decreased total lung lactate similar to VB124 (**Figure 9A**).
+As anticipated, VB253 decreased total lung lactate similar to VB124 (**Figure 9A**).
 Taken together, these data provide compelling preclinical data for the development of lactate transporter inhibition as a therapeutic strategy for fibrotic lung disease.
@@
 
-Craig Thompson paper: Lactate activates ETC independently of its metabolism
+Our findings demonstrate that lactate transport is an essential component of the metabolic reprogramming associated with myofibroblast differentiation *in vitro* and *in vivo*.
+We observed increased expression of the lactate transporters MCT1 and MCT4 in IPF lung explants and experimental model systems.
+Inhibition of these transporters decreased TGFβ-stimulated myofibroblast differentiation and the severity of bleomcyin-induced pulmonary fibrosis.
+Metabolically, MCT antagonists promoted glucose oxidation and decreased glucose carbon incorporation into fibrotic lung regions, which were associated with decreased oxidative stress.
+Beyond metabolic reprogramming, we observed changes in gene expression profiles consistent with antifibrotic metabolic signaling downstream of lactate transport inhibition, which may be mediated by changes in ROS production, protein lactylation, or other mechanisms.
+MCT4 inhibition consistently proved more potently antifibrotic than MCT1 inhibition, and we report a novel MCT4 inhibitor, VB253, with a more favorable pharmacologic profile than VB124 that is now in Phase 1 clinical trials [**REF?**].
+Collectively, these data establish lactate transport as a metabolic opportunity for therapeutic intervention in pulmonary fibrosis.
 
-Craig Thompson paper: Isotope labeling in lung fibroblasts
+Metabolic reprogramming is a hallmark of myofibroblast differentiation.
+Previous studies have identified alterations in carbohydrate, amino acid, and lipid metabolic pathways that support fibrogenesis [@selvarajah2021].
+Of these, increased lactate production was among the earliest recognized metabolic changes associated with pulmonary fibrosis [@kottmann2012].
+Subsequent work demonstrated increased lactate production was driven by activation of glycolysis in myofibroblasts [@xie2015] and that inhibition of glycolysis prevented myofibroblast differentiation [@xie2015; @kottmann2015; @bernard2015] and, in turn, experimental pulmonary fibrosis [@judge2018; @cho2017].
+Consistent with increased glycolysis in myofibroblasts, we have now demonstrated increased expression of the lactate transporters MCT1 and MCT4 support this metabolic phenotype.
 
-<!-- spatial metabolomics -->
+MCTs are proton-coupled monocarboxylate symporters with varying affinities for lactate, pyruvate, and other monocarboxylates [@halestrap2013].
+MCT1 is constitutively and ubiquitously expressed and primarily considered to be a lactate importer with a K~M~ for lactate of 3-6 mM.
+By contrast, MCT4 expression is dynamically regulated, including by HIF-1α, and has primarily been considered to be a lactate exporter with a low lactate affinity (K~M~ 30-40 mM).
+Recent data suggests, however, that MCT4 has a much higher affinity for lactate import than previously appreciated (K~M~ 1 mM) [@contreras-baeza2019].
+These data are most consistent with our findings that inhibition of both MCT1 and MCT4 was required to inhibit lactate export (**Figure 3**) and isotope incorporation from extracellular [U-^13^C~3~]-lactate import (**Figure 4J**).
+Previous studies of cancer cells have similarly recognized that MCT4-expressing cells are resistant to the cytotoxic effects of MCT1 inhibition [@curtis2017].
+Interestingly, inhibition of either MCT1 or MCT4 reduced myofibroblast differentiation without affecting lactate export.
+
+To clarify the metabolic consequences of lactate transporter inhibition, we performed a comprehensive metabolic analysis of myofibroblasts treated with MCT inhibitors, including bioenergetic measurements, metabolomic profiling, and stable isotope tracing experiments.
+The primary metabolic consequence of MCT inhibition is activation of oxidative phosphorylation.
+MCT inhibitors increased the fraction of mitochondrial ATP production, which was associated with increased levels of tricarboxylic acid cycle intermediates, and increased isotope incorporation from glucose substrates.
+Collectively, these data indicate that MCT inhibition reroutes carbon flow from glucose away from lactate fermentation and toward glucose oxidation.
+Several mechanisms may account for this phenomenon.
+Since the magnitude of these effects generally correlates with intracellular lactate levels, mass action may be an important driver of this metabolic phenomenon.
+Lactate accumulation is also closely coupled in NADH production through lactate dehydrogenase activity.
+Cytoplasmic NADH may be transported into the mitochondria through the malate-aspartate shuttle and oxidized by the electron transport chain.
+Recently, lactate itself was shown activate the mitochondrial electron transport chain independently of its metabolism [@cai2023], although the molecular mechanism for this effect is unknown.
+
+How these metabolic effects signal to antifibrotic transcriptional programs remains unclear.
+We investigated several hypotheses based on previous studies.
+First, we did not observe significant changes in canonical (*i.e.*, Smad3) or non-canonical (*i.e.*, ERK) TGFβ signaling pathways.
+Second, TGFβ-dependent activation of HIF-1α seems to be required for myofibroblast differentiation [@goodwin2018].
+MCT1 and MCT4 potentiated HIF-1α stabilization following TGFβ (**Figure 6D**), indicating that MCT inhibition acts downstream of HIF signaling pathways.
+Third, post-translational modification of histone lysines by lactate serves as an epigenetic modification that stimulates gene expression [@zhang2019a].
+We observed increased histone lactylation following dual MCT inhibition, which may account for some of the transcriptional changes we observed.
+In addition to histones, protein lactylation has been increasingly identified as a critical regulator of protein function, including cytoskeletal proteins [@yang2023].
+Fourth, increased ROS production has been observed following TGFβ treatment [@jain2013; @schworer2020].
+While we were unable to replicate TGFβ-dependent increases in ROS in our cells and model system, we did observe decreased CellROX oxidation with MCT4 inhibition and decreased MitoSOX oxidation with AZD3965.
+Collectively, these studies identify two putative mechanisms for antifibrotic lactate signaling following MCT inhibition, protein lactylation and antioxidant activity; however, novel lactate targets are being increasingly recognized as important mediators of metabolic signaling [@liu2023; @liu2023b].
+
+Through our detailed metabolic investigation of lactate transport inhibition, we have also generated valuable data on the metabolic and transcriptional consequences of human lung myofibroblast differentation.
+TGFβ stimulated marked changes in amino acid and nucleic acid metabolic pathways (**Supplemental Figure 4A-D**).
+Previous studies have identified important roles for glutamine, proline, and taurine [@selvarajah2021; @choudhury2023; @phan1995], while the contributions of branched chain amino acid metabolism, for example, remain unexplored.
+Integrating these multi-omics data sets may identify novel molecular targets for future drug development.
+
+In addition to metabolomic profiling, we have performed stable isotope tracing using glucose, lactate, and glutamine substrates in TGFβ-treated primary human lung fibroblasts.
+To our knowledge, these are the first comprehensive data on intracellular substrate metabolism in this commonly utilized model of myofibroblast differentiation.
+Overall, isotope labeling patterns changed little following TGFβ treatment (**Supplemental Figure 4.2**).
+Together with the results of our extracellular flux experiments, this finding suggests that TGFβ primarily increases the flow of metabolites through metabolic pathways without substantially altering the pathways themselves.
+Consistent with increased glucose uptake and accumulation of glycolytic metabolites, a greater fraction of pyruvate, lactate, alanine, and serine are labeled by [U-^13^C~6~]-glucose, which is offset by a decrease in the fractional labeling from [U-^13^C~3~]-lactate.
+Similar to our previous work [@copeland2023], nearly 50% of tricarboxylic acid metabolites (citrate, 2-oxoglutarate, succinate, and malate) are labeled by [U-^13^C~3~]-lactate, highlighting the importance of lactate as a respiratory fuel source in these cells.
+
+In addition to cellular bioenergetics, the contribution of fibroblast metabolic pathways to extracellular matrix production has been an intense area of investigation [@selvarajah2021].
+Serine and glycine may be synthesized from the glycolytic intermediate 3-phosphoglycerate, and proline may be synthesized from glutamine.
+These pathways have been implicated in myofibroblast differentation and pulmonary fibrosis [@bernard2018; @hamanaka2019; @selvarajah2019; @schworer2020; @choudhury2023].
+Interestingly, our data suggest little incorporation of glucose or lactate carbon into serine and little incorporation of glutamine carbon into proline.
+These differences may be secondary to the cells studied (IMR-90 fetal lung fibroblasts or NIH-3T3 spontaneously immortalized mouse embryonic fibroblasts) or the culture medium utilized (Eagle's minimum essential medium, which lacks serine, glycine, and proline).
+In this context, our study highlights that fibroblasts will preferentially use available amino acids rather than reroute substrates into biosynthetic pathways.
+Developing metabolic flux models using human physiologic medium [@cantor2017] may provide a more accurate and complete characterization of substrate flow into energetic and biosynthetic pathways that more closely reflects fibroblast metabolism *in vivo*.
+
+Even better than the use of *in vitro* model systems are approaches that enable the study of cell metabolism *in vivo*.
+Here, we performed metabolomic profiling of lung and plasma samples from mice treated with bleomycin and MCT inhibitors (**Supplemental Figure 9**).
+These studies did not reveal consistent changes in whole lung metabolite levels and demonstrate a potential limitation of using bulk metabolomics to monitor metabolic changes in heterogeneous cell populations *in vivo*.
+We sought to overcome this limitation with spatial metabolic imaging.
+Mice received stable isotope tracers of proline and glucose prior to euthanasia.
+Since these tracers are administered to live animals, multi-isotope imaging mass spectrometry identifies lung regions that are metabolically active over the labeling period.
+Since tissues are fixed and processed *ex vivo*, the isotope signals indicate substrate incorporation into fixable biomass [@zhang2020; @wertheim2023].
+Thus, using multi-isotope imaging mass spectrometry, we are directly quantifying isotope flux from glucose into fibrotic lung regions.
+We found that ^15^N-proline enrichment levels correlate with tissue fibrosis from histologic and biochemical assessments.
+This finding emphasizes the ability of cells to incorporate circulating proline for protein synthesis, and likely de-emphasizes the importance of *de novo* proline biosynthesis *in vivo*.
+The ^2^H-glucose signal provides some direct evidence for MCT-dependent metabolic reprogramming in mice, with less carbohydrate incorporation into matrix proteins.
+Owing to the financial and time costs of isotope tracing and imaging, only a few animals per group were studied.
+Nonetheless, metabolic imaging at high spatial resolution offers tremendous promise for correlating cell identities determined from spatial transcriptomic profiles with metabolic features.
+Stable isotopes may be safely administered to human patients prior lung biopsy or explant, an approach that would considerably advance our understanding of how cell metabolism contributes to pulmonary fibrosis.
+
+MCT1 and MCT4 are expressed by many cells in the lung, particularly macrophages and dendritic cells [@travaglini2020].
+Although our *ex vivo* and *in vitro* experiments suggest that inhibiting myofibroblast differentiation is a primary antifibrotic mechanism, inhibition of lactate transporters expressed by other cell types may also play a role in their therapeutic effect.
+For example, MCT4 expression is increased as part of a HIF-1α gene expression signature in transitional AT2 cells that accumulate in pulmonary fibrosis and contribute to aberrant repair [@choi2020; @zhu2022].
+We are in the process of generating fibroblast-specific conditional knockout mice to explore the role of these transporters in fibroblasts specifically.
+
+Pharmacologic interest in lactate transporters has been driven by the recognition of increased lactate transporter expression in a variety of cancers [@payen2020].
+AZD3965 was selected for this study as it has been investigated in human clinical trials for advanced solid organ malignancies [@halford2023].
+In this Phase 1 study, AZD3965 was generally well-tolerated with 7 of 40 patients experiencing dose-limiting toxicities including asymptomatic, reversible ocular changes; acidosis; and increased troponin.
+VB124 was the first selective MCT4 inhibitor developed [@cluntun2021] and here we report a second generation MCT4 inhibitor, VB253.
+A Phase 1 clinical trial of VB253 is currently enrolling [**REF?**].
+MCT4 global knockout mice are viable and breed normally [@bisetto2018; @bisetto2019], increasing optimism that VB253 will be well tolerated in humans.
+One of the barriers to translating metabolic therapies to humans has been the poor pharmacologic properties of the small molecule laboratory compounds that have been studied previously.
+The IC~50~ values for AZD3965 and VB253 are ~ 2 nM, compared to the next most potent inhibitor, lonidamine, with an IC~50~ of 7,000 nM [@yin2019].
+As a metabolic target downstream of glycolysis, lactate transport inhibitors may also be better tolerated by allowing glycolysis to continue to support of glucose oxidation, while upstream inhibitors of glycolysis have a more severe impact on cellular bioenergetics.
+
+In conclusion, we have demonstrated that lactate transporters promote myofibroblast differentiation and pulmonary fibrosis.
+Using a comprehensive metabolic phenotyping strategy, we have identified antifibrotic mechanisms for MCT inhibition and demonstrated evidence for metabolic reprogramming in animal models.
+We have demonstrated the antifibrotic efficacy of currently available lactate transport inhibitors with accepted preclinical disease models.
+Collectively, our data advance lactate transport inhibition as a viable therapeutic strategy for patients with pulmonary fibrosis.
 

### 2024-03-12 manuscript/manuscript.qmd
working on adding figures

@@
     fig10,
-    # resources_table,
+    resources_table,
     NULL
@@
 To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of extracellular and intracellular metabolites from cells treated with AZD3965 and VB124.
-As suggested by the Seahorse analysis, inhibiting a single lactate transporter had modest effects on extracellular metabolite levels (**Figure 4A-D, Supplemental Figure 4**).
+As suggested by the Seahorse analysis, inhibiting a single lactate transporter had modest effects on extracellular metabolite levels (**Figure 4A-D, Supplemental Figure 4.1**).
 This analysis confirmed the results of extracellular lactate measurements by enzyme assay, which showed dual inhibition was required to decrease lactate efflux (**Figure 4B**).
 In addition to lactate, dual lactate transporter inhibition also had broader impacts on amino acid fluxes (**Figure 4C-D**).
-Several metabolites were differentially regulated by TGFβ treatment (**Supplemental Figure 4A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, and ketoleucine.
+Several metabolites were differentially regulated by TGFβ treatment (**Supplemental Figure 4.1A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, and ketoleucine.
 
@@
 As expected from measures of extracellular lactate following MCT inhibition, intracellular lactate accumulation increases modestly with MCT4 inhibition and markedly with MCT1/4 inhibition.
-TGFβ treatment was associated with diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Supplemental Figure 4C-D**).
-The effect of AZD3965 alone on intracellular metabolite levels was modest (**Supplemental Figure 4G-H**).
-By contrast, MCT4 inhibition by VB124 alone (**Supplemental Figure 4K-L**) or combined with AZD3965 (**Figure 4G-H**) caused substantial perturbations to intracellular metabolism.
+TGFβ treatment was associated with diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Supplemental Figure 4.1C-D**).
+The effect of AZD3965 alone on intracellular metabolite levels was modest (**Supplemental Figure 4.1G-H**).
+By contrast, MCT4 inhibition by VB124 alone (**Supplemental Figure 4.1K-L**) or combined with AZD3965 (**Figure 4G-H**) caused substantial perturbations to intracellular metabolism.
 Specifically, we observed enrichment of the glycolysis and tricarboxylic acid (TCA) cycle metabolite sets with MCT4 inhibition.
@@
 This labeling was decreased following TGFβ treatment, mirroring increased fractional labeling from glucose.
-MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Supplemental Figure 4M**).
+MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Supplemental Figure 4.1M**).
 Interestingly, MCT4 inhibition alone or with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, emphasizing the importance of bi-directional lactate transport mediated by MCTs.
@@
 Through our detailed metabolic investigation of lactate transport inhibition, we have also generated valuable data on the metabolic and transcriptional consequences of human lung myofibroblast differentation.
-TGFβ stimulated marked changes in amino acid and nucleic acid metabolic pathways (**Supplemental Figure 4A-D**).
+TGFβ stimulated marked changes in amino acid and nucleic acid metabolic pathways (**Supplemental Figure 4.1A-D**).
 Previous studies have identified important roles for glutamine, proline, and taurine [@selvarajah2021; @choudhury2023; @phan1995], while the contributions of branched chain amino acid metabolism, for example, remain unexplored.
@@
 
+## Key Resources
+
+```{r}
+#| results: asis
+#| tbl-cap: "Key Resources Table"
+
+flextable::flextable_to_rmd(resources_table)
+```
+
+\newpage
+
+## Chemicals
+
+Recombinant human TGF-β1 was purchased commercially, dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquotting and storing at -80 °C.
+AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
+VB124 (10 μM) was a generous gift from Vettore, LLC.
+The drugs were prepared in DMSO and used at the indicated doses unless otherwise indicated.
+These doses were chosen based on prior reports and their target affinities.
+
+## Human lung tissue and fibroblasts
+
+Human IPF lung samples were obtained at the time of lung transplant.
+Control samples were prepared from non-fibrotic and unaffected border regions of lung nodule biopsy specimens or from failed donor controls.
+Samples were obtained through the BWH Biorepository for Understanding Inflammatory Lung Disease (BUILD) or the MGH ILD Translational Research Program and their collection was approved by the Mass General Brigham Institutional Review Board (2013P002332, 2016P001890, 2019P003592, 2020P002765).
+
+To isolate human lung fibroblasts, lung tissue was minced, digested with liberase and DNase, and passed through a sterile 70 μm filter.
+The filtrate was centrifuged at 300 ×*g* and washed once with RPMI medium (Lonza).
+Cells were plated in DMEM (Lonza) supplemented with 10% FBS, penicillin, and streptomycin.
+After two passages, medium was changed to FGM-2 (Lonza).
+
+## Bleomycin model
+
+All animal experiments were approved by the Brigham and Women's Hospital IACUC (2020N000199).
+C57Bl6/J mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
+Control animals received an equivalent volume of PBS.
+For drug treatments, AZD3965 (100 mg/kg twice daily) and VB124 (30 mg/kg daily) were prepared in 0.5% methylcellulose and 0.1% Tween-20 with sonication and administered by oral gavage every day beginning on day 7 following bleomycin.
+Control animals received vehicle alone twice daily.
+After 21 d, animals were euthanized.
+
+For lung function measurements, animals were anesthetized with pentobarbital (100 μg/g; Covetrus, Portland, ME) and tracheally cannulated for pulmonary function measurements using a flexiVent (SCIREQ).
+Animals were ventilated for 3-5 min with 150 breaths per minute at 10 mL/kg and 3 cmH~2~O positive end-expiratory pressure.
+After two total lung capacity maneuvers, respiratory elastance was measured using two Prime-8 maneuvers.
+A pressure-volume loop was then generated and static compliance was calculated using the Salazar-Knowles equation.
+
+Histologic fibrosis severity was assessed by Aschcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
+The average of twenty 10× fields is reported for each animal.
+
+Lung hydroxyproline content was determined as described previously [@tager2008].
+The four lobes of the R lung were homogenized in PBS with Halt protease inhibitors (Thermo) and hydrolyzed in 6 N HCl at 120 °C for 12 h, with an aliquot added to 1.4% chloramine-T, 10% 1-propanol, and 0.5 M sodium acetate, pH 6.0.
+After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
+Samples were incubated at 65 °C for 15 minutes.
+Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
+
+For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanisia on day 21 after bleomycin.
+
+## Myofibroblast differentiation *in vitro*
+
+Normal lung fibroblasts were purchased commercially (Lonza).
+All fibroblasts were cultured in FGM-2 medium (Lonza).
+Passages 3-8 were used for experiments.
+Myofibroblast differentiation was induced *in vitro* using recombinant human TGF-β1 (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
+
+## Isotope labeling
+
+Cells were seeded and serum-starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
+The concentrations of glucose and glutamine match the concentrations of these substrates determined in standard (FGM-2) growth medium.
+After 24 h, cells were washed with PBS and the medium was changed to "heavy" labeling medium containing [1,2-^13^C~2~]-glucose, [U-^13^C~6~]-glucose, [U-^13^C~5~]-glutamine, or [U-^13^C~2~]-lactate (Cambridge Isotope Labs) along with TGFβ and MCT inhibitors.
+Cell lysates were prepared for metabolomics as described below.
+
+## Immunoblotting
+
+Human and mouse lung samples were mechanically homogenized in RIPA lysis buffer containing Halt Protease Inhibitor Cocktail (Thermo).
+Cells were washed with one volume of PBS and collected by scraping in PBS.
+Cell suspensions were centrifuged at 5,000 ×*g* for 5 min at 4 °C.
+Pellets were lysed in buffer containing Tris 10 mM, pH 7.4, NaCl 150 mM, EDTA 1 mM, EGTA 1 mM, Triton X-100 1% v/v, NP-40 0.5% v/v, and protease inhibitors.
+Protein concentrations were determined by BCA Protein Assay (Thermo).
+Lysates were normalized for protein concentration and subjected to SDS-PAGE separation on stain-free tris-glycine gels (Bio-Rad), cross-linked and imaged with the Chemidoc system (Bio-Rad), transferred to PVDF membranes with the Trans-Blot Turbo transfer system (Bio-Rad), imaged, blocked in 5% blocking buffer (Bio-Rad), blotted in primary and secondary antibodies, and developed using WesternBright ECL (Advansta).
+Band signal intensity was normalized to total protein per lane as determined from the stain-free gel or membrane images using Image Lab software (Bio-Rad).
+
+## Histone purification
+
+Histones were extracted from lung fibroblast cell lysates using the EpiQuik Total Histone Extraction Kit (Epigentek).
+
+## RNA interference
+
+Lung fibroblasts were reverse transfected with ON-TARGETplus SMARTpool siRNA (Dharmacon) targeting MCT1 or MCT4 using Lipofectamine RNAiMAX (Thermo).
+A non-targeting siRNA pool was used a transfection control (Dharmacon).
+For combined siMCT1 and siMCT4 treatment, the total siRNA dose was kept constant.
+After 24 h, cells were serum starved for an additional 24 h prior to treatment with TGFβ.
+
+## Gel contraction assay
+
+Lung fibroblasts in FBM and neutralized collagen solution (TeloCol-3, Advanced Biomatrix) were combined 1:2 to yield 100,000 cells/mL and 1.8 mg/mL collagen.
+This solution was added to 24-well plates (0.5 mL) and allowed to harden for 10 min at 37 °C prior to the addition of 1 mL FBM.
+Cells were serum starved overnight prior to releasing the gels and treating them with TGFβ and MCT inhibitors in triplicate.
+Images were acquired after 24 and 48 h using a ChemiDoc imager (Bio-Rad).
+Gel areas were measured using Image Lab software (Bio-Rad).
+Data are expressed as the ratio of TGFβ-treated gel area to control gel area.
+
+## Cell count
+
+Cell counts were estimated from total DNA quantified using the Quant-iT PicoGreen dsDNA Assay Kit (Thermo) as described previously [@copeland2023].
+
+## Lactate assay
+
+Lactate was determined using an enzymatic cycling assay resulting in the reduction of resazurin to the fluorescent chemical resorufin.
+Medium samples were diluted 1:10 in PBS.
+Samples were combined with a solution of diaphorase (125 mU), lactate dehydrogenase, NAD (250 μM), and resazurin (25 μM) in Tris buffer (100 mM, pH 8.5) then incubated for 30 min at room temperature.
+Resorufin fluorescence was measured using a SpectraMax i3x microplate reader (Molecular Devices) and lactate concentration was interpolated from a standard curve.
+Data were normalized to estimated cell count.
+
+## Glucose assay
+
+Medium samples were diluted 10-fold in PBS.
+Glucose concentration was determined using the Glucose Colorimetric Assay Kit (Cayman) according to the manufacturer's protocol.
+Standards were prepared in PBS.
+
+## Seahorse assays
+
+Lung fibroblasts were seeded in Seahorse assay plates at a density of 20,000 cells per well.
+On the following day, the cells were serum starved for 24 h prior to treatment with TGFβ and MCT inhibitors for 48 h.
+On the day of the assay, the medium was exchanged to Seahorse DMEM without phenol red supplemented with 5 mM HEPES, 10 mM glucose, 2 mM glutamine, and 1 mM pyruvate.
+OCR and extracellular acidification rates (ECAR) were measured at baseline and following the sequential injections of oligomycin (1 μM final), FCCP (2 μM final), and rotenone plus antimycin A (0.5 μM final each) using an XFe24 bioanalyzer (Agilent).
+At the end of the assay, cells were stained for 30 min with the fluorescent dye Nuclear Green LCS1 (10 μM; Abcam).
+Stained cells were quantified using a SpectraMax MiniMax 300 Imaging Cytometer (Molecular Devices) and these cell counts were used to normalize the rate data.
+Proton efflux rate (PER) was calculated from the ECAR and buffer capacity of the medium.
+PER from glycolysis was determined by subtracting the contribution of oxidative phosphorylation to total PER using the empirically measured CO~2~ correction factor of 0.41.
+ATP from glycolysis was assumed to be equal to basal glycolytic PER.
+ATP from oxidative phosphorylation was assumed to be 2 × 2.75 × oligomycin-sensitive OCR [@mookerjee2017].
+
+## Metabolomics
+
+Metabolomics data acquisition and analysis were performed as described previously [@copeland2023].
+
+### Metabolite extraction
+
+Extracellular metabolites were obtained by mixing conditioned medium 1:4 with 100% MeOH precooled to -80 °C.
+Intracellular metabolites were obtained after washing cells with 2 volumes of ice-cold PBS and floating on liquid nitrogen.
+Plates were stored at -80 °C until extraction.
+Metabolites were extracted with 1 mL 80% MeOH pre-cooled to -80 °C.
+Insoluble material from extracellular and intracelelular metabolite samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
+The supernatant was evaporated to dryness at 42 °C using a SpeedVac concentrator (Thermo Savant).
+Samples were resuspended in 35 μL 20 mM ammonium phosphate in LC-MS-grade water prior to analysis.
+
+### Acquisition parameters
+
+LC-MS analysis was performed on a Vanquish ultra-high-performance liquid chromatography system coupled to a Q Exactive orbitrap mass spectrometer by a HESI-II electrospray ionization probe (Thermo).
+External mass calibration was performed weekly.
+Metabolite samples (2.5 μL) were separated using a ZIC-pHILIC stationary phase (2.1 × 150 mm, 5 μm) (Merck).
+The autosampler temperature was 4 °C and the column compartment was maintained at 25 °C.
+Mobile phase A was 20 mM ammonium carbonate, 0.1% ammonium hydroxide, and 5 μM medronic acid.
+Mobile phase B was acetonitrile.
+The flow rate was 0.1 mL/min.
+Solvent was introduced to the mass spectrometer *via* electrospray ionization with the following source parameters: sheath gas 40, auxiliary gas 15, sweep gas 1, spray voltage +3.0 kV for positive mode and -3.1 kV for negative mode, capillary temperature 275 °C, S-lens RF level 40, and probe temperature 350 °C.
+Data were acquired and peaks integrated using TraceFinder 4.1 (Thermo).
+
+### Metabolomic profiling
+
+For metabolomic profiling, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B.
+The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*.
+Resolution was set to 70,000 and the AGC target was 1×10^6^ ions.
+Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method.
+For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run.
+Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas.
+Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018].
+Missing values were imputed using random forest.
+Sample peak areas were normalized using probabilistic quotient normalization [@dieterle2006].
+Differentially regulated metabolites were identified using limma [@ritchie2015].
+Exploratory data analysis and review of PCA plots suggested orthogonal effects of TGFβ and MCT inhibitor treatment.
+For this reason, differentially regulated metabolites were determined based on the main effects of TGFβ or MCT inhibitor.
+Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021] with KEGG metabolite pathways [@kanehisa2000].
+
+### Stable isotope quantification
+
+Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B.
+The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite.
+The resolution was set at 70,000 and AGC target was 1×10^5^ ions.
+Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015].
+Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].
+
+\newpage
+
+## RNA-seq transcriptomics
+
+RNA was collected from LFs treated with TGFβ and MCT inhibitors as described above.
+Four biological replicates were analyzed.
+Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample.
+Sequences were deposited in the NIH SRA (PRJNA1011992).
+Sequences were mapped to the human GRCh38 primary assembly and counts summarized using salmon [@patro2017].
+Details of the processing pipeline and summarized data are available from the Oldham Lab GitHub repository (github.com/oldhamlab/rnaseq.lf.tgfb.mcti).
+Differentially expressed genes were identified using DESeq2 [@love2014].
+Gene set enrichment was performed using the fgsea R package [@korotkevich2021].
+
+## Pyridine dinucleotide measurements
+
+Cellular NAD^+^ and NADH were measured using an enzymatic fluorometric cycling assay as previously described [@oldham2015; @copeland2023].
+Samples for NADP^+^ and NADPH quantification were similarly prepared and analyzed using the luminescence-based NADP/NADPH-Glo assay (Promega).
+
+## Reactive oxygen species
+
+*Fei Li to write.*
+
 \newpage
@@
 
-![**Figure 1 - Lactate transporter expression increases in human IPF and experimental models.** (**A**) Immunoblot of whole lung homogenates from explanted IPF lungs and controls (Ctl).
+![**Figure 1 - Lactate transporter expression increases in human IPF lungs and experimental models.** (**A**) Immunoblot of whole lung homogenates from explanted IPF lungs and controls (Ctl).
 (**B**) Quantification of band densities from (A).
-(**C**) Immunoblot of whole lung homogenates from bleomycin (bleo)- and vehicle (Ctl)-treated mice.
+(**C**) Immunoblot of whole lung homogenates from bleomycin (Bleo)- and vehicle (Ctl)-treated mice.
 (**D**) Quantification of band densities from (C).
@@
 Individual data points are biological replicates.
-Summary data are mean ± SEM (\* p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 02.png)
+Summary data are mean ± SEM (\* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 02.png)
+
+\newpage
+
+![**Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined by enzymatic assay following TGFβ stimulation of cells treated with siRNA targeting MCT1 or MCT4, separately or together (N = 6 biological replicates, \* adjusted p < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares siMCT *v.* siCTL).
+(**B**) TGFβ-stimulated lactate secretion was measured following treatment with MCT1 inhibitor AZD3965 (AZD), MCT4 inhibitor VB124 (VB), or both (N = 4-13 biological replicates, \* adjusted p < 0.05, *black* compared TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+(**C**) Oxygen consumption (OCR) and proton efflux (PER) rates of lung fibroblasts treated with TGFβ and MCT inhibitors for 48 h prior to the assay.
+Measurements were performed at baseline and following injection of ATP synthase inhibitor oligomycin (Oligo), mitochondrial membrane uncoupler FCCP, and Complex I and III inhibitors rotenone and antimycin A (Rot/AMA) (N = 4 biological replicates, data are mean ± SEM).
+(**D**) Glycolytic (Glyco) and mitochondrial (Mito) ATP production rates were calculated from PER and OCR, respectively (\* adjusted p < 0.05 compared to TGFβ/Veh).
+(**E**) Energy phenogram derived from data presented in (D).
+TGFβ increases ATP production from both glycolysis and oxidative phosphorylation, favoring the former, while MCT inhibition increases mitochondrial ATP production.
+Summary data are mean ± SEM.](figs/Figure 03.png)
+
+\newpage
+
+![**Figure 4 - Lactate transporter inhibition promotes oxidative phosphorylation.** (**A**) Principal components analysis of metabolites in conditioned medium from lung fibroblasts treated with TGFβ (*bold colors*) or vehicle (*light colors*) and lactate transport inhibitors (N = 5).
+(**B**) Extracellular lactate determined by LC-MS.
+(**C**) Volcano plot of significantly altered extracellular metabolites following combined treatment with AZD3965 and VB124.
+Differentially regulated metabolites are colored (adjusted p < 0.1), the top 10 up- and down-regulated of which are labeled.
+(**D**) KEGG pathways significantly enriched (adjusted p < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
+Positive NES are relatively enriched in AZD/VB-treated cells while negative NES are relatively enriched in Vehicle-treated cells.
+(**E**) Principal components analysis of intracellular metabolites extracted from lung fibroblasts stimulated with TGFβ (*bold colors*) or vehicle (*light colors*) in the presence of lactate transport inhibitors (N = 5).
+(**F**) Intracellular lactate increases significantly with MCT4 inhibition alone (VB) or when combined with MCT1 inhibition (AZD/VB) (\* adjusted p < 0.05 compared to vehicle control for the main effect of the inhibitor).
+(**G**) Volcano plot of significantly altered metabolites following combined treatment with AZD3965 and VB124.
+Differentially regulated metabolites are colored (adjusted p < 0.1), the top 10 up- and down-regulated of which are labeled.
+(**H**) KEGG pathways significantly enriched (adjusted p < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
+Positive NES are relatively enriched in AZD/VB-treated cells while negative NES are relatively enriched in Vehicle-treated cells.
+(**I**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~6~\]-glucose.
+(**J**) Labeled fraction of intracellular metabolites following \[U-^13^C~3~\]-lactate treatment.
+(N = 4 biological replicates, \* adjusted p < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+Summary data are mean ± SEM.](figs/Figure 04.png)
+
+\newpage
+
+![**Figure 5 - Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A**) Principal components analysis of RNA-seq data from lung fibroblasts treated with TGFβ (*bold colors*) or vehicle (*light colors*) and lactate transport inhibitors (N = 4).
+(**B**) Dot plot of Hallmark gene set enrichment analysis.
+Significantly enriched pathways with adjusted p < 0.05 are included and points are colored by normalized enrichment score (NES).
+Positive NES indicated relative enrichment following TGFβ compared to control (TGFβ) or with MCT inhibitor compared to vehicle in TGFβ-treated cells.
+(**C**) Volcano plot of significantly differentially expressed genes in TGFβ-stimulated cells treated with AZD3965 and VB124 compared to vehicle control.
+Significantly differentially expressed geness are highlighted (adjusted p < 0.05), the top 15 up- and down-regulated of which are labeled.
+(**D**) Transcript counts from the leading edge of enrichment for the epithelial mesenchymal transition Hallmark gene set demonstrating down-regulation of pro-fibrotic genes with MCT inhibition.](figs/Figure 05.png)
+
+\newpage
+
+![**Figure 6 - Lactate transport inhibition does not alter TGFβ signaling pathways.** (**A**) Representative immunoblot of normal lung fibroblast cell lysates after treatment with extracellular lactate (10 mM).
+(**B**) Quantification of protein expression levels from (A).
+(**C**) Immunoblot of HIF-1α stabilization 6 h after TGFβ treatment in the presence of MCT inhibitors.
+(**D**) Quantification of band densities from (C).
+(**E**) Representative immunoblot of canonical TGFβ receptor target, Smad3.
+(**F**) Quantification of band densities from (E).
+(**G**) Representative immunoblot of non-canonical TGFβ receptor target, ERK (**H**) Quantification of band densities from (E).
+(**I**) Representative immunoblot of histones purified from lung fibroblasts treated with TGFβ and MCT inhibitors.
+(**J**) Quantification of band densities from (I).
+Summary data are mean ± SEM (\* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 06.png)
+
+\newpage
+
+![**Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADH/NAD^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**B**) Relative CellROX fluorescence, an indicator of cellular reactive oxygen species, normalized to TGFβ-treated cells (N = 3 biological replicates; \* adjusted p < 0.05 compared to TGFβ-treated cells).
+Summary data are mean ± SEM.](figs/Figure 07.png)
+
+\newpage
+
+![**Figure 8 - Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl6N mice were administered bleomycin (1.2 U/kg) on day 0.
+Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21.
+On day 21, mice were anesthetized with pentobarbital for lung function measurements.
+(**B**) Static compliance of the lung was measured by pressure-volume loop analysis.
+(**C**) Elastance was measured using the forced oscillation technique for 8 s (Prime-8) (N = 13-17).
+(**D**) Representative trichrome-stained histologic sections of mouse lungs.
+(**E**) Ashcroft scores from histologic sections from (D) determined by a lung pathologist blinded to treatment assignment (N = 6-7).
+(**F**) Right lung hydroxyproline content (N = 10-13).
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.png)
+
+# Supplemental Figures
+
+![**Supplemental Figure 2 - Lactate transport inhibition decreases myofibroblast differentation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
+(**B**) Representative immunoblot of normal lung fibroblasts treated with TGFβ and MCT inhibitors AR-C155858 (AR) or VB124 (VB).
+(**C**) Quantification of α-SMA expression following MCT inhibition.
+(**D-E**) Cell counts following pharmacologic inhibition of lactate transporters.
+Summary data are mean ± SEM (\* p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+](figs/Figure 02.supplement.png)
+
+\newpage
+
+![**Supplemental Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+(**B**) Extracellular glucose was determined following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
+(**C**) Summary data from Seahorse extracellular flux analysis.
+(**D**) Cell bioenergetic stress testing results from Seahorse analysis.
+Glycolytic capacity measures the increase in proton efflux rate following oligomycin treatment.
+Spare respiratory capacity measures the increase in oxygen consumption rate (OCR) following FCCP treatment compared to basal OCR.
+Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin).
+The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p < 0.05 compared to TGFβ/Veh).
+Summary data are mean ± SEM.](figs/Figure 03.supplement.png)
+
+\newpage
+
+![**Supplemental Figure 4.1 - Metabolomic profiling of lung fibroblasts treated with lactate transport inhibitors.** (**A-D**) Extracellular (A-B) and intracellular (C-D) metabolomic profile and metabolite set enrichment analyses of lung fibroblasts treated with TGFβ.
+(**E-H**) Extracellular (E-F) and intracellular (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment.
+(**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment.
+Differentially regulated metabolites are colored (adjusted p < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (adjusted p < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Positive NES indicate enrichment in treated cells while negative NES indicate relative enrichment in Vehicle-treated cells.
+(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+Summary data are mean ± SEM.](figs/Figure 04.supplement.1.png)
+
+\newpage
+
+![**Supplemental Figure 4.2 - Stable isotope tracing in lung fibroblasts treated with TGFβ.** Lung fibroblasts were cultured with stable isotopes of glucose, lactate, and glutamine during treatment with TGFβ (*columns*).
+Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*).
+Significant differences in isotopic enrichment are indicated (\* adjusted p < 0.05).
+Summary data are mean ± SEM (N = 4 biological replicates).](figs/Figure 04.supplement.2.png)
+
+\newpage
+
+![**Supplemental Figure 5 - Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A-C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh.
+Significantly differentially expressed geness are highlighted (adjusted p < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/Figure 05.supplement.png)
+
+\newpage
+
+![**Supplemental Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**B**) MitoSOX fluorescence, a marker of mitochondrial superoxide production, was measured and normalized to MitoTracker fluorescence, a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p < 0.05 compared to TGFβ-treated cells).
+(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p < 0.05 for the main treatment effect).
+(**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates).
+Summary data are mean ± SEM.](figs/Figure 07.supplement.png)
+
+\newpage
+
+![**Supplemental Figure 8 - Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly.
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.supplement.png)
+
+\newpage

### 2024-03-12 R/seahorse.R

@@ -226,6 +226,8 @@ plot_seahorse_atp_bars <- function(x, stats) {
         y = .data$y.stat,
         label = .data$label
       ),
+      size = 10 / ggplot2::.pt,
+      family = "Calibri",
       nudge_y = 2
     ) +
     ggplot2::labs(
@@ -300,6 +302,7 @@ plot_seahorse_pheno <- function(x) {
       nudge_x = 0.5,
       nudge_y = -0.5,
       color = "black",
+      family = "Calibri",
       size = 1.75,
       show.legend = FALSE
     ) +

### 2024-03-12 R/figures.R
working on adding figures

@@ -479,3 +479,12 @@ make_fig10s <- function(p1) {
       tags = NULL
     )
 }
+
+create_resources <- function(path) {
+  readr::read_csv(path, show_col_types = FALSE) |>
+    flextable::flextable() |>
+    flextable::bold(part = "header", bold = TRUE) |>
+    flextable::font(fontname = "Calibri", part = "all") |>
+    flextable::fontsize(size = 9, part = "all") |>
+    flextable::set_table_properties(layout = "autofit")
+}

### 2024-03-12 R/mice.R

@@ -125,7 +125,9 @@ plot_wt_timeline <- function(df, stat) {
         x = 14.5,
         y = y,
         vjust = vjust
-      )
+      ),
+      size = 10 / ggplot2::.pt,
+      family = "Calibri"
     ) +
     ggplot2::labs(
       x = "Time (d)",
@@ -348,7 +350,9 @@ plot_mice <- function(
           vjust = vjust,
           label = label
         ),
-        color = "black"
+        color = "black",
+        size = 10 / ggplot2::.pt,
+        family = "Calibri"
       )
   }
   p

### 2024-03-13 manuscript/manuscript.qmd
finish first complete draft and spell check

@@
   affiliations: 
+  - ref: bwh
   - ref: mgh
+  - ref: hms
+- name: Fei Li
+  affiliations:
   - ref: bwh
@@
   - ref: utah_gen
-- name: Kevin Leahy 
+- name: Kevin J. Leahy 
   affiliations:
@@
   - ref: vettore
+- name: Christelle Guillermier
+  affiliations:
+  - ref: bwh
+  - ref: hms
 - name: Jack Varon
@@
   - ref: hms
+- name: Bradley A. Maron
+  affiliations:
+  - ref: umd
 - name: Nancy J. Philp
@@
   - ref: tj
-- name: Lida P. Hariri
+- name: Lida p-value. Hariri
   affiliations: 
@@
   - ref: hms
+- name: Matthew L. Steinhauser
+  affiliations:
+  - ref: pitt
+  - ref: upmc
 - name: Rachel S. Knipe
@@
 affiliations:
+- id: bwh
+  name: Department of Medicine, Brigham and Women's Hospital, Boston, MA
 - id: mgh
   name: Department of Medicine, Massachusetts General Hospital, Boston, MA
-- id: bwh
-  name: Department of Medicine, Brigham and Women's Hospital and Harvard Medical School, Boston, MA
 - id: hms
@@
   name: Vettore Biosciences, San Francisco, CA
+- id: umd
+  name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MA
 - id: tj
@@
   name: Department of Pathology, Massachusetts General Hospital, Boston, MA
+- id: pitt
+  name: Aging Institute, University of Pittsburgh, Pittsburgh, PA
+- id: upmc
+  name: UPMC Heart and Vascular Institute, UPMC Presbyterian, Pittsburgh, PA
 - id: utah_biochem
@@
     fig02,
+    fig02s, 
     fig03,
+    fig03s, 
     fig04,
+    fig04s, 
+    mid_tgf_plot_img,
     fig05,
+    fig05s, 
     fig06,
     fig07,
+    fig07s, 
     fig08, 
+    fig08s, 
     fig09, 
+    fig09s1, 
     fig10,
+    fig10s, 
     resources_table,
@@
 
+# Conflicts of Interest
+
+*Email Will with COI.*
+
 \newpage
@@
 
+Myofibroblast differentiation requires metabolic reprogramming to support the extracellular matrix synthesis that drives pulmonary fibrosis.
+While prior studies demonstrate that targeting fibroblast metabolism mitigates experimental pulmonary fibrosis, this approach has been slow to translate to clinical practice.
+Here, we examined the contribution of lactate transporters MCT1 and MCT4 to myofibroblast differentiation and experimental pulmonary fibrosis using novel drugs suitable for human clinical trials, including a newly developed inhibitor of MCT4, VB253.
+We demonstrate that inhibition of MCT1 or MCT4 decreases TGFβ-stimulated myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
+Using bioenergetic analyses, stable isotope tracing, metabolomic profiling, and imaging mass spectrometry in cells and mice, we demonstrate lactate transport inhibition promotes oxidative phosphorylation and decreases glucose metabolite incorporation into fibrotic lung regions.
+We characterize the transcriptional consequences of these inhibitors and propose that metabolic signaling from lactate by protein lactylation and by reactive oxygen species detoxification account for changes in gene expression.
+Collectively, these data establish lactate transport as a requirement for myofibroblast differentiation, establish these transporters as a therapeutic vulnerability in experimental fibrosis, and support continued clinical study in patients with pulmonary fibrosis for whom limited therapeutic options exist.
+
 # Keywords
+
+pulmonary fibrosis, lactate transporters, myofibroblast, imaging mass spectrometry
 
@@
 These effects were also observed in normal lung fibroblasts where AZD3965 as well as VB124, separately or together, decreased α-SMA expression (**Figure 2E-F**).
-AR-C155858, which is a high-affinity(*K~i~* 2 nM) inhibitor of both MCT1 and MCT2 [@ovens2010], also decreased α-SMA expression alone and in combination with VB124 (**Supplemental Figure 2B-C**).
+AR-C155858, which is a high-affinity (*K~i~* 2 nM) inhibitor of both MCT1 and MCT2 [@ovens2010], also decreased α-SMA expression alone and in combination with VB124 (**Supplemental Figure 2B-C**).
 Pharmacologic MCT inhibition did not affect cell counts over 48 h of treatment (**Supplemental Figure 2D-E**).
@@
 MCT4 is considered to be the main lactate exporter in glycolytic cells and is up-regulated by hypoxia-inducible transcription factors [@ullah2006; @halestrap2013].
-MCT4 can also function as a lactate importer with a K~M~ ~1 mM [@contreras-baeza2019].
+MCT4 can also function as a lactate importer with a *K*~M~ of 1 mM [@contreras-baeza2019].
 
@@
 Principal components (PC) analysis demonstrated excellent clustering of treatment groups with the drug effect corresponding to PC1 and the TGFβ treatment effect corresponding to PC2.
-Similar to the extracellular flux results, the magnitude of drug-induced perturbations were increased from MCT1 to MCT4 to combined inhibition (**Figure 4E**), as demonstrated by intracellular lactate levels (**Figure 4F**).
+Similar to the extracellular flux results, the magnitude of drug-induced perturbations was increased from MCT1 to MCT4 to combined inhibition (**Figure 4E**), as demonstrated by intracellular lactate levels (**Figure 4F**).
 As expected from measures of extracellular lactate following MCT inhibition, intracellular lactate accumulation increases modestly with MCT4 inhibition and markedly with MCT1/4 inhibition.
@@
 Consistent with the PCA results, only GRIK4 and BRI3 were differentially expressed following AZD3965 (**Supplemental Figure 5B**).
-By contrast, VB124, alone (**Supplemental Figure 5C**) or in combination with AZD3965 (**Figure 5C**), demonstrated more substantial reprogramming of myofibroblast transcription with `r deg_diff` of `r deg_tot` genes (`r deg_frac`%) differentially expressed at an FDR < 0.05.
+By contrast, VB124, alone (**Supplemental Figure 5C**) or in combination with AZD3965 (**Figure 5C**), demonstrated more substantial reprogramming of myofibroblast transcription with `r deg_diff` of `r deg_tot` genes (`r deg_frac`%) differentially expressed at a FDR < 0.05.
 
@@
 Glucose labeling followed a similar pattern where lactate transporter inhibition caused a decrease in glucose incorporation into tissue biomass.
-Importantly, these  findings are not merely a consequence of decreased tissue fibrosis as we chose more fibrotic areas to image (**Figure 9B**) and the enrichment values are normalized to tissue area.
+Importantly, these  findings are not merely a consequence of decreased tissue fibrosis as we chose more fibrotic areas to image (**Figure 9B**), and the enrichment values are normalized to tissue area.
 
@@
 Enhanced pause (Penh) is a dimensionless index describing airflow during tidal breathing found to increase significantly in bleomycin-treated mice [@vanoirbeek2010].
-VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterens in bleomycin-treated mice (**Figure 10D**).
+VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterns in bleomycin-treated mice (**Figure 10D**).
 This was associated with decreased pulmonary fibrosis by histologic assessment (**Figure 10E**) and α-SMA expression (**Figure 10F-G**).
-Morevover, the effects of VB253 compared favorably to the antifibrotic effects of nintedanib and pirfenidone.
+Moreover, the effects of VB253 compared favorably to the antifibrotic effects of nintedanib and pirfenidone.
 
@@
 We observed increased expression of the lactate transporters MCT1 and MCT4 in IPF lung explants and experimental model systems.
-Inhibition of these transporters decreased TGFβ-stimulated myofibroblast differentiation and the severity of bleomcyin-induced pulmonary fibrosis.
+Inhibition of these transporters decreased TGFβ-stimulated myofibroblast differentiation and the severity of bleomycin-induced pulmonary fibrosis.
 Metabolically, MCT antagonists promoted glucose oxidation and decreased glucose carbon incorporation into fibrotic lung regions, which were associated with decreased oxidative stress.
@@
 MCTs are proton-coupled monocarboxylate symporters with varying affinities for lactate, pyruvate, and other monocarboxylates [@halestrap2013].
-MCT1 is constitutively and ubiquitously expressed and primarily considered to be a lactate importer with a K~M~ for lactate of 3-6 mM.
-By contrast, MCT4 expression is dynamically regulated, including by HIF-1α, and has primarily been considered to be a lactate exporter with a low lactate affinity (K~M~ 30-40 mM).
-Recent data suggests, however, that MCT4 has a much higher affinity for lactate import than previously appreciated (K~M~ 1 mM) [@contreras-baeza2019].
+MCT1 is constitutively and ubiquitously expressed and primarily considered to be a lactate importer with a *K*~M~ for lactate of 3-6 mM.
+By contrast, MCT4 expression is dynamically regulated, including by HIF-1α, and has primarily been considered to be a lactate exporter with a low lactate affinity (*K*~M~ 30-40 mM).
+Recent data suggests, however, that MCT4 has a much higher affinity for lactate import than previously appreciated (*K*~M~ 1 mM) [@contreras-baeza2019].
 These data are most consistent with our findings that inhibition of both MCT1 and MCT4 was required to inhibit lactate export (**Figure 3**) and isotope incorporation from extracellular [U-^13^C~3~]-lactate import (**Figure 4J**).
@@
 
-Through our detailed metabolic investigation of lactate transport inhibition, we have also generated valuable data on the metabolic and transcriptional consequences of human lung myofibroblast differentation.
+Through our detailed metabolic investigation of lactate transport inhibition, we have also generated valuable data on the metabolic and transcriptional consequences of human lung myofibroblast differentiation.
 TGFβ stimulated marked changes in amino acid and nucleic acid metabolic pathways (**Supplemental Figure 4.1A-D**).
@@
 Serine and glycine may be synthesized from the glycolytic intermediate 3-phosphoglycerate, and proline may be synthesized from glutamine.
-These pathways have been implicated in myofibroblast differentation and pulmonary fibrosis [@bernard2018; @hamanaka2019; @selvarajah2019; @schworer2020; @choudhury2023].
+These pathways have been implicated in myofibroblast differentiation and pulmonary fibrosis [@bernard2018; @hamanaka2019; @selvarajah2019; @schworer2020; @choudhury2023].
 Interestingly, our data suggest little incorporation of glucose or lactate carbon into serine and little incorporation of glutamine carbon into proline.
@@
 
-Recombinant human TGF-β1 was purchased commercially, dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquotting and storing at -80 °C.
+Recombinant human TGF-β1 was purchased commercially, dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
 AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
@@
 Control samples were prepared from non-fibrotic and unaffected border regions of lung nodule biopsy specimens or from failed donor controls.
-Samples were obtained through the BWH Biorepository for Understanding Inflammatory Lung Disease (BUILD) or the MGH ILD Translational Research Program and their collection was approved by the Mass General Brigham Institutional Review Board (2013P002332, 2016P001890, 2019P003592, 2020P002765).
-
 To isolate human lung fibroblasts, lung tissue was minced, digested with liberase and DNase, and passed through a sterile 70 μm filter.
@@
 
-All animal experiments were approved by the Brigham and Women's Hospital IACUC (2020N000199).
-C57Bl6/J mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
+C57Bl6/N mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
 Control animals received an equivalent volume of PBS.
@@
 After 21 d, animals were euthanized.
+Lung function measurements were performed.
+Plasma samples were obtained by right ventricular puncture and collected in EDTA-coated tubes.
+The lungs were dissected and perfused with PBS.
+The right lung was isolated while the left lung was inflated to 20 cmH~2~O prior to fixation.
+The right lungs were snap frozen.
 
@@
 After two total lung capacity maneuvers, respiratory elastance was measured using two Prime-8 maneuvers.
-A pressure-volume loop was then generated and static compliance was calculated using the Salazar-Knowles equation.
+A pressure-volume loop was then generated, and static compliance was calculated using the Salazar-Knowles equation.
 
-Histologic fibrosis severity was assessed by Aschcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
+Histologic fibrosis severity was assessed by Ashcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
 The average of twenty 10× fields is reported for each animal.
@@
 The four lobes of the R lung were homogenized in PBS with Halt protease inhibitors (Thermo) and hydrolyzed in 6 N HCl at 120 °C for 12 h, with an aliquot added to 1.4% chloramine-T, 10% 1-propanol, and 0.5 M sodium acetate, pH 6.0.
-After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
+After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-value-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
 Samples were incubated at 65 °C for 15 minutes.
 Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
-
-For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanisia on day 21 after bleomycin.
 
@@
 
-Cells were seeded and serum-starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
+Cells were seeded and serum starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
 The concentrations of glucose and glutamine match the concentrations of these substrates determined in standard (FGM-2) growth medium.
@@
 Metabolites were extracted with 1 mL 80% MeOH pre-cooled to -80 °C.
-Insoluble material from extracellular and intracelelular metabolite samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
+Mouse plasma and lung homogenates were mixed 1:4 with 100% MeOH precooled to -80 °C.
+Samples were extracted at -80 °C for 4 h.
+Insoluble material from these samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
 The supernatant was evaporated to dryness at 42 °C using a SpeedVac concentrator (Thermo Savant).
@@
 Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample.
-Sequences were deposited in the NIH SRA (PRJNA1011992).
 Sequences were mapped to the human GRCh38 primary assembly and counts summarized using salmon [@patro2017].
-Details of the processing pipeline and summarized data are available from the Oldham Lab GitHub repository (github.com/oldhamlab/rnaseq.lf.tgfb.mcti).
 Differentially expressed genes were identified using DESeq2 [@love2014].
@@
 
-*Fei Li to write.*
+*Fei Li to write method for CellROX and MitoSOX/MitoTracker.*
+
+## Multi-isotope imaging mass spectrometry
+
+For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on day 21 after bleomycin.
+
+*Christelle to write.*
+
+Isotopic enrichment for each animal/image was summarized as follows.
+The ratio of "heavy" to "light" ion intensities were calculated.
+An image mask was generated by Otsu thresholding to exclude background regions.
+Pixel ratios equal to zero were excluded as these are non-biologic.
+The average of the remaining pixel ratios was then calculated to provide a mean tissue isotope enrichment value for the image.
+
+## VB253
+
+VB253 was developed by Vettore, LLC as a more potent small molecule inhibitor of MCT4.
+Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC.
+These experiments were completed independently of the other studies presented in the manuscript and offer independently validation of the antifibrotic effects of MCT4 inhibition in cells and mice.
+
+*Nathan, Mark, and Jared.
+Add the names of the CROs that did the experiments?
+I think the aSMA/Smad3/nuclei experiments will need a description.
+The others are pretty similar to what is described above.*
+
+## Statistics
+
+Data analysis, statistical comparisons, and visualization were performed in R [@R-base].
+Experiments included technical and biological replicates as noted above.
+The number of biological replicates (N) is indicated in the figure legends.
+Summary data show the mean ± SEM.
+Outliers were identified using twice the median absolute deviation as a cutoff threshold.
+Comparisons were performed using linear mixed-effects models with condition (±TGFβ), treatment and their interaction as fixed effects and biological replicate as a random effect.
+Significant differences in estimated marginal means were identified by comparisons to the multivariate *t* distribution.
+Metabolomics and RNA-seq data were analyzed as described above.
+Probability values less than 0.05 were considered statistically significant.
+
+## Study approval
+
+Human samples were obtained through the BWH Biorepository for Understanding Inflammatory Lung Disease (BUILD) or the MGH ILD Translational Research Program and their collection was approved by the Mass General Brigham Institutional Review Board (2013P002332, 2016P001890, 2019P003592, 2020P002765).
+
+All animal experiments were approved by the Brigham and Women's Hospital IACUC (2020N000199).
+
+## Data availability
+
+The raw data and annotated analysis code necessary to reproduce this manuscript are contained in an R package research compendium available from the Oldham Lab GitHub repository (github.com/oldhamlab/Ziehr.2023.ipf.mcti).
+
+RNA sequences were deposited in the NIH SRA (PRJNA1011992).
+Details of the processing pipeline and summarized data are available from the Oldham Lab GitHub repository (github.com/oldhamlab/rnaseq.lf.tgfb.mcti).
 
@@
 
+# Author Contributions
+
+The authors confirm contribution to the paper as follows: designed research studies - D.R.Z., F.L., K.M.p-value., R.M.B., B.A.M., N.J.p-value., E.Y.K., M.L.S., R.S.K., J.R., W.M.O.; conducted experiments - D.R.Z., F.L., K.J.L., W.M.O.; acquired data - D.R.Z., F.L., N.M.K., K.J.L., K.M.p-value., C.G., J.V., L.p-value.H., R.S.K., W.M.O.; analyzed data - D.R.Z., F.L., N.M.K., K.M.p-value., C.G., M.L.S., R.S.K., J.R., W.M.O.; wrote the manuscript - W.M.O.
+All authors reviewed the results, revised the manuscript, and approve the final version.
+
 # Acknowledgements
 
-# Conflicts of Interest
+Research reported in this publication was supported by the National Heart Lung and Blood Institute of the National Institutes of Health under award number R01GM987654 (to W.M.O.), and by the Brigham and Women's Hospital Research Institute.
+The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health.
+VB124 was a gift from Vettore, LLC.
 
@@
 Individual data points are biological replicates.
-Summary data are mean ± SEM (\* p < 0.05).](figs/Figure 01.png)
+Summary data are mean ± SEM (\* p-value < 0.05).](figs/Figure 01.png)
 
@@
 Individual data points are biological replicates.
-Summary data are mean ± SEM (\* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 02.png)
+Summary data are mean ± SEM (\* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 02.png)
 
@@
 
-![**Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined by enzymatic assay following TGFβ stimulation of cells treated with siRNA targeting MCT1 or MCT4, separately or together (N = 6 biological replicates, \* adjusted p < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares siMCT *v.* siCTL).
-(**B**) TGFβ-stimulated lactate secretion was measured following treatment with MCT1 inhibitor AZD3965 (AZD), MCT4 inhibitor VB124 (VB), or both (N = 4-13 biological replicates, \* adjusted p < 0.05, *black* compared TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+![**Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined by enzymatic assay following TGFβ stimulation of cells treated with siRNA targeting MCT1 or MCT4, separately or together (N = 6 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares siMCT *v.* siCTL).
+(**B**) TGFβ-stimulated lactate secretion was measured following treatment with MCT1 inhibitor AZD3965 (AZD), MCT4 inhibitor VB124 (VB), or both (N = 4-13 biological replicates, \* adjusted p-value < 0.05, *black* compared TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
 (**C**) Oxygen consumption (OCR) and proton efflux (PER) rates of lung fibroblasts treated with TGFβ and MCT inhibitors for 48 h prior to the assay.
 Measurements were performed at baseline and following injection of ATP synthase inhibitor oligomycin (Oligo), mitochondrial membrane uncoupler FCCP, and Complex I and III inhibitors rotenone and antimycin A (Rot/AMA) (N = 4 biological replicates, data are mean ± SEM).
-(**D**) Glycolytic (Glyco) and mitochondrial (Mito) ATP production rates were calculated from PER and OCR, respectively (\* adjusted p < 0.05 compared to TGFβ/Veh).
+(**D**) Glycolytic (Glyco) and mitochondrial (Mito) ATP production rates were calculated from PER and OCR, respectively (\* adjusted p-value < 0.05 compared to TGFβ/Veh).
 (**E**) Energy phenogram derived from data presented in (D).
@@
 (**C**) Volcano plot of significantly altered extracellular metabolites following combined treatment with AZD3965 and VB124.
-Differentially regulated metabolites are colored (adjusted p < 0.1), the top 10 up- and down-regulated of which are labeled.
-(**D**) KEGG pathways significantly enriched (adjusted p < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+(**D**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
 Positive NES are relatively enriched in AZD/VB-treated cells while negative NES are relatively enriched in Vehicle-treated cells.
 (**E**) Principal components analysis of intracellular metabolites extracted from lung fibroblasts stimulated with TGFβ (*bold colors*) or vehicle (*light colors*) in the presence of lactate transport inhibitors (N = 5).
-(**F**) Intracellular lactate increases significantly with MCT4 inhibition alone (VB) or when combined with MCT1 inhibition (AZD/VB) (\* adjusted p < 0.05 compared to vehicle control for the main effect of the inhibitor).
+(**F**) Intracellular lactate increases significantly with MCT4 inhibition alone (VB) or when combined with MCT1 inhibition (AZD/VB) (\* adjusted p-value < 0.05 compared to vehicle control for the main effect of the inhibitor).
 (**G**) Volcano plot of significantly altered metabolites following combined treatment with AZD3965 and VB124.
-Differentially regulated metabolites are colored (adjusted p < 0.1), the top 10 up- and down-regulated of which are labeled.
-(**H**) KEGG pathways significantly enriched (adjusted p < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+(**H**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
 Positive NES are relatively enriched in AZD/VB-treated cells while negative NES are relatively enriched in Vehicle-treated cells.
@@
 (**J**) Labeled fraction of intracellular metabolites following \[U-^13^C~3~\]-lactate treatment.
-(N = 4 biological replicates, \* adjusted p < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+(N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
 Summary data are mean ± SEM.](figs/Figure 04.png)
@@
 (**B**) Dot plot of Hallmark gene set enrichment analysis.
-Significantly enriched pathways with adjusted p < 0.05 are included and points are colored by normalized enrichment score (NES).
+Significantly enriched pathways with adjusted p-value < 0.05 are included and points are colored by normalized enrichment score (NES).
 Positive NES indicated relative enrichment following TGFβ compared to control (TGFβ) or with MCT inhibitor compared to vehicle in TGFβ-treated cells.
 (**C**) Volcano plot of significantly differentially expressed genes in TGFβ-stimulated cells treated with AZD3965 and VB124 compared to vehicle control.
-Significantly differentially expressed geness are highlighted (adjusted p < 0.05), the top 15 up- and down-regulated of which are labeled.
+Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.
 (**D**) Transcript counts from the leading edge of enrichment for the epithelial mesenchymal transition Hallmark gene set demonstrating down-regulation of pro-fibrotic genes with MCT inhibition.](figs/Figure 05.png)
@@
 (**J**) Quantification of band densities from (I).
-Summary data are mean ± SEM (\* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 06.png)
+Summary data are mean ± SEM (\* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 06.png)
 
@@
 
-![**Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADH/NAD^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-(**B**) Relative CellROX fluorescence, an indicator of cellular reactive oxygen species, normalized to TGFβ-treated cells (N = 3 biological replicates; \* adjusted p < 0.05 compared to TGFβ-treated cells).
+![**Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADH/NAD^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**B**) Relative CellROX fluorescence, an indicator of cellular reactive oxygen species, normalized to TGFβ-treated cells (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
 Summary data are mean ± SEM.](figs/Figure 07.png)
@@
 
-![**Figure 8 - Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl6N mice were administered bleomycin (1.2 U/kg) on day 0.
+![**Figure 8 - Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
 Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21.
@@
 (**F**) Right lung hydroxyproline content (N = 10-13).
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.png)
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.png)
+
+\newpage
+
+![**Figure 9 - Lactate transporter inhibition reprograms lung metabolism *in vivo*.** (**A**) Plasma and lung lactate levels following 14-day treatment with AZD3965 or VB124 (N = 6-9; \* adjusted p-value < 0.05 compared to bleomycin control).
+(**B**) Representative images from multi-isotope imaging mass spectrometry (MIMS).
+The ^12^C^14^N ion illustrates tissue histology.
+The ^12^C^15^N/^12^C^14^N ion ratio image indicates enrichment from ^15^N-proline (colorbar 1-3-fold of natural abundance).
+The ^12^C~2~^2^H/^12^C~2~^1^H ratio indicates enrichment above natural abundance from ^2^H-glucose (colorbar 1-10-fold of natural abundance).
+(**C**) Quantification of isotope enrichment (N = 3 biological replicates, \* adjusted p-value-value < 0.05 compared to bleomycin-treated control).](figs/Figure 09.png)
+
+\newpage
+
+![**Figure 10 - VB253, a novel MCT4 inhibitor, is potently antifibrotic *in vitro* and *in vivo*.** (**A**) Dose-response of α-SMA protein expression in IPF lung fibroblasts treated with VB253 or nintedanib (NIN).
+(**B**) Number of nuclei, a marker of cell viability, following treatment with VB253 or NIN.
+Summary data are mean ± SEM of cell responses from N = 3 different donors.
+(**C**) Young (8-10 w) or aged (60+ w) mice were treated with bleomycin (1 U/kg) on day 0.
+Beginning on day 7, animals were treated with VB253 or the FDA-approved antifibrotics nintedanib (NIN) and pirfenidone (PFD) for 14 days prior to euthanasia on day 21.
+(**D**) Enhanced pause (Penh) is a non-invasive index of mouse tidal breathing that increases with bleomycin treatment.
+(**E-G**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (E) and decreased α-SMA protein expression (F-G).
+(**H-J**) Similar histologic improved was observed in aged mice with decreased histologic fibrosis (H) and decreased α-SMA protein expression (I-J).
+(**K**) VB253 decreased lung lactate, similar to established antifibrotics.
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 10.png)
+
+\newpage
 
@@
 
-![**Supplemental Figure 2 - Lactate transport inhibition decreases myofibroblast differentation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
+![**Supplemental Figure 2 - Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
 (**B**) Representative immunoblot of normal lung fibroblasts treated with TGFβ and MCT inhibitors AR-C155858 (AR) or VB124 (VB).
@@
 (**D-E**) Cell counts following pharmacologic inhibition of lactate transporters.
-Summary data are mean ± SEM (\* p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+Summary data are mean ± SEM (\* p-value-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
 ](figs/Figure 02.supplement.png)
@@
 
-![**Supplemental Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+![**Supplemental Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
 (**B**) Extracellular glucose was determined following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
@@
 Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin).
-The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p < 0.05 compared to TGFβ/Veh).
+The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p-value < 0.05 compared to TGFβ/Veh).
 Summary data are mean ± SEM.](figs/Figure 03.supplement.png)
@@
 (**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment.
-Differentially regulated metabolites are colored (adjusted p < 0.1), the top 10 up- and down-regulated of which are labeled.
-KEGG pathways significantly enriched (adjusted p < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
 Positive NES indicate enrichment in treated cells while negative NES indicate relative enrichment in Vehicle-treated cells.
-(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
 Summary data are mean ± SEM.](figs/Figure 04.supplement.1.png)
@@
 Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*).
-Significant differences in isotopic enrichment are indicated (\* adjusted p < 0.05).
+Significant differences in isotopic enrichment are indicated (\* adjusted p-value < 0.05).
 Summary data are mean ± SEM (N = 4 biological replicates).](figs/Figure 04.supplement.2.png)
@@
 ![**Supplemental Figure 5 - Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A-C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh.
-Significantly differentially expressed geness are highlighted (adjusted p < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/Figure 05.supplement.png)
+Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/Figure 05.supplement.png)
 
@@
 
-![**Supplemental Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-(**B**) MitoSOX fluorescence, a marker of mitochondrial superoxide production, was measured and normalized to MitoTracker fluorescence, a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p < 0.05 compared to TGFβ-treated cells).
-(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p < 0.05 for the main treatment effect).
+![**Supplemental Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**B**) MitoSOX fluorescence, a marker of mitochondrial superoxide production, was measured and normalized to MitoTracker fluorescence, a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
+(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p-value < 0.05 for the main treatment effect).
 (**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates).
@@
 ![**Supplemental Figure 8 - Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly.
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.supplement.png)
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.supplement.png)
 
 \newpage
+
+![**Supplemental Figure 9 - Metabolic reprogramming by lactate transport inhibition *in vivo*.** (**A-D**) Plasma (A-B) and lung (C-D) metabolomic profile and metabolite set enrichment analyses in bleomycin-treated mice.
+(**E-H**) Plasma (E-F) and lung (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment in bleomycin-treated mice.
+(**I-L**) Plasma (I-J) and lung (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment in bleomycin-treated mice.
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by bleomycin treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Positive NES indicate enrichment in treated mice while negative NES indicate relative enrichment in control or untreated mice.](figs/Figure 09.supplement.1.png)
+
+\newpage
+
+![**Supplemental Figure 10 - VB253 does not attenuate canonical TGFβ signaling.** Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
+Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/Figure 10.supplement.png)
+
+\newpage
+
+# Fin

### 2024-03-13 _targets.R

@@ -1256,8 +1256,8 @@ list(
     mid_tgf_plot,
     plot_tgf_mids(mid_mcti_mids_corr) +
       theme_patchwork(
-        widths = ggplot2::unit(6, "in"),
-        heights = ggplot2::unit(10, "in"),
+        widths = ggplot2::unit(5.5, "in"),
+        heights = ggplot2::unit(9, "in"),
         tags = NULL
       ),
     format = "rds"

### 2024-03-14 manuscript/manuscript.qmd
address Maron comments

@@
   - ref: umd
+  - ref: ihc
 - name: Nancy J. Philp
@@
   - ref: tj
-- name: Lida p-value. Hariri
+- name: Lida P. Hariri
   affiliations: 
@@
   name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MA
+- id: ihc
+  name: University of Maryland Institute for Health Computing, Bethesda, MD
 - id: tj
@@
 targets::tar_load(pkg_citations)
-
-# load data for paper
 
@@
 
+B.A.M. has received consulting fees from Actelion and Tenax and has performed investigator-initiated research with support from Deerfield, all outside the scope of this research.
+
 \newpage
@@
 While prior studies demonstrate that targeting fibroblast metabolism mitigates experimental pulmonary fibrosis, this approach has been slow to translate to clinical practice.
-Here, we examined the contribution of lactate transporters MCT1 and MCT4 to myofibroblast differentiation and experimental pulmonary fibrosis using novel drugs suitable for human clinical trials, including a newly developed inhibitor of MCT4, VB253.
+Here, we examined the contribution of lactate transporters MCT1 and MCT4 to myofibroblast differentiation and experimental pulmonary fibrosis using novel drugs suitable for human clinical trials, including a newly developed small molecule inhibitor of MCT4, VB253.
 We demonstrate that inhibition of MCT1 or MCT4 decreases TGFβ-stimulated myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
@@
 We characterize the transcriptional consequences of these inhibitors and propose that metabolic signaling from lactate by protein lactylation and by reactive oxygen species detoxification account for changes in gene expression.
-Collectively, these data establish lactate transport as a requirement for myofibroblast differentiation, establish these transporters as a therapeutic vulnerability in experimental fibrosis, and support continued clinical study in patients with pulmonary fibrosis for whom limited therapeutic options exist.
+Collectively, these data establish lactate transport as a requirement for myofibroblast differentiation, establish these transporters as a pharmacologic target in experimental fibrosis, and support continued clinical study in patients with pulmonary fibrosis for whom limited therapeutic options exist.
 
@@
 A family of monocarboxylate transporters (MCT1-4) maintains lactate homeostasis and small molecule inhibitors of these transporters have been under active clinical development for oncologic applications where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
-Importantly, compared to previously investigated inhibitors of glycolysis, MCT inhibitors exhibit favorable pharmacologic profiles and have successfully translated to human clinical trials [@halford2023].
+Importantly, compared to previously investigated inhibitors of glycolysis, MCT inhibitors exhibit favorable pharmacologic profiles and have successfully translated to human clinical trials for advanced solid tumors [@halford2023].
 Before the promise of this therapeutic approach in IPF may be realized, however, the preclinical efficacy and molecular mechanisms-of-action of lactate transport inhibitors must be demonstrated experimentally.
@@
 Notably, siMCT1 also decreased MCT4 expression.
-Interestingly, the mRNA sequence homology of these transporters is low and the MCT1 siRNA sequences do not match the MCT4 mRNA sequence, perhaps suggesting a post-translational explanation for this phenomenon.
+The mRNA sequence homology of these transporters is low and the MCT1 siRNA sequences do not target the MCT4 mRNA sequence, suggesting a post-translational mechanism accounting for decreased MCT1 expression decreasing MCT4 protein.
 No adverse effects of lactate transporter knockdown on cell viability were observed, and siMCT1 significantly increased cell count in both control and TGFβ-treated cells (**Supplemental Figure 2A**).
@@
 VB253 is a novel inhibitor of MCT4 with similar selectivity, but ~10-fold increased potency, for MCT4 inhibition.
-Like VB124, VB253 dose-dependently decreases **TGFβ-stimulated?** α-SMA in human IPF lung fibroblasts *ex vivo* (**Figure 10A**).
+Like VB124, VB253 dose-dependently decreases TGFβ-stimulated α-SMA in human IPF lung fibroblasts *ex vivo* (**Figure 10A**).
 Given the robust effect of MCT4 inhibition on myofibroblast differentiation, we compared VB253 to nintedanib, an anti-fibrotic medication clinically indicated for pulmonary fibrosis (**Figure 10A-B**).
@@
 
-The authors confirm contribution to the paper as follows: designed research studies - D.R.Z., F.L., K.M.p-value., R.M.B., B.A.M., N.J.p-value., E.Y.K., M.L.S., R.S.K., J.R., W.M.O.; conducted experiments - D.R.Z., F.L., K.J.L., W.M.O.; acquired data - D.R.Z., F.L., N.M.K., K.J.L., K.M.p-value., C.G., J.V., L.p-value.H., R.S.K., W.M.O.; analyzed data - D.R.Z., F.L., N.M.K., K.M.p-value., C.G., M.L.S., R.S.K., J.R., W.M.O.; wrote the manuscript - W.M.O.
+The authors confirm contribution to the paper as follows: designed research studies - D.R.Z., F.L., K.M.P., R.M.B., B.A.M., N.J.P., E.Y.K., M.L.S., R.S.K., J.R., W.M.O.; conducted experiments - D.R.Z., F.L., K.J.L., W.M.O.; acquired data - D.R.Z., F.L., N.M.K., K.J.L., K.M.P., C.G., J.V., L.P.H., R.S.K., W.M.O.; analyzed data - D.R.Z., F.L., N.M.K., K.M.P., C.G., M.L.S., R.S.K., J.R., W.M.O.; wrote the manuscript - W.M.O.
 All authors reviewed the results, revised the manuscript, and approve the final version.

### 2024-03-20 manuscript/manuscript.qmd
add EMS acknowledgement
update authors

@@
   - ref: hms
+- name: K. Mark Parnell
+  affiliations:
+  - ref: vettore
 - name: Nathan M. Krah
@@
   - ref: bwh
-- name: K. Mark Parnell
-  affiliations:
-  - ref: vettore
 - name: Christelle Guillermier
@@
   name: Department of Medicine, Harvard Medical School, Boston, MA
+- id: vettore
+  name: Vettore Biosciences, San Francisco, CA
 - id: utah_gen
   name: Department of Human Genetics, University of Utah, Salt Lake City, UT
-- id: vettore
-  name: Vettore Biosciences, San Francisco, CA
 - id: umd
@@
 Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].
-
-\newpage
-
 ## RNA-seq transcriptomics
@@
 
+The authors thank Louise Trakimas and the HMS Electron Microscopy Facility for MIMS sample preparation and consultation.
 Research reported in this publication was supported by the National Heart Lung and Blood Institute of the National Institutes of Health under award number R01GM987654 (to W.M.O.), and by the Brigham and Women's Hospital Research Institute.

### 2024-03-21 _targets.R
update R and packages

@@ -1518,7 +1518,7 @@ list(
 
   tar_target(
     mice_file,
-    raw_data_path("mice.csv"),
+    raw_data_path("^mice.csv"),
     format = "file"
   ),
   tar_target(

### 2024-03-21 manuscript/manuscript.qmd
update R and packages

@@
 While we were unable to replicate TGFβ-dependent increases in ROS in our cells and model system, we did observe decreased CellROX oxidation with MCT4 inhibition and decreased MitoSOX oxidation with AZD3965.
-Collectively, these studies identify two putative mechanisms for antifibrotic lactate signaling following MCT inhibition, protein lactylation and antioxidant activity; however, novel lactate targets are being increasingly recognized as important mediators of metabolic signaling [@liu2023; @liu2023b].
+Collectively, these studies identify two putative mechanisms for antifibrotic lactate signaling following MCT inhibition, protein lactylation and antioxidant activity; however, novel lactate targets are being increasingly recognized as important mediators of metabolic signaling [@liu2023a; @liu2023c].
 

### 2024-03-27 manuscript/manuscript.qmd
fix liu2023b reference

--- 2024-03-20 manuscript/manuscript.qmd
@@
 While we were unable to replicate TGFβ-dependent increases in ROS in our cells and model system, we did observe decreased CellROX oxidation with MCT4 inhibition and decreased MitoSOX oxidation with AZD3965.
-Collectively, these studies identify two putative mechanisms for antifibrotic lactate signaling following MCT inhibition, protein lactylation and antioxidant activity; however, novel lactate targets are being increasingly recognized as important mediators of metabolic signaling [@liu2023; @liu2023b].
+Collectively, these studies identify two putative mechanisms for antifibrotic lactate signaling following MCT inhibition, protein lactylation and antioxidant activity; however, novel lactate targets are being increasingly recognized as important mediators of metabolic signaling [@liu2023a; @liu2023b].
 

### 2024-04-12 R/figures.R
revise figures for submission

@@ -16,7 +16,7 @@ theme_patchwork <- function(
     ),
     patchwork::plot_annotation(
       tag_levels = tags,
-      theme = ggplot2::theme(plot.margin = ggplot2::margin(5, 5, 5, 5))
+      theme = ggplot2::theme(plot.margin = ggplot2::margin(rep(3, 4)))
     )
   )
 }
@@ -45,7 +45,8 @@ write_figures <- function(plot, filename, path = "manuscript/figs") {
     path = path,
     width = overall_width,
     height = overall_height,
-    units = "in"
+    units = "in",
+    res = 220
   )
 
   if (file.exists("Rplots.pdf")) unlink("Rplots.pdf")
@@ -54,46 +55,49 @@ write_figures <- function(plot, filename, path = "manuscript/figs") {
 }
 
 make_fig01 <- function(p1, p2, p3, p4, p5, p6) {
+  figs <- plot_scale(p1, p3, p5)
   design <- "ab \n cd \n ef"
-  p1 + p2 + p3 + p4 + p5 + p6 +
+  figs[[1]] + p2 + figs[[2]] + p4 + figs[[3]] + p6 +
     theme_patchwork(
       design = design,
-      widths = ggplot2::unit(c(1, 1.75), "in"),
-      heights = ggplot2::unit(1, "in")
+      widths = ggplot2::unit(c(1.5, 2), "in"),
+      heights = ggplot2::unit(1.5, "in"),
+      tags = list(c("A", "", "B", "", "C", ""))
     )
 }
 
-make_fig02 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
-  design <- "
-  abcd
-  abcd
-  efgh
-  ef##
-  "
+make_fig02 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11, p12) {
+  figs <- plot_scale(p1, p3, p5)
 
-  p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 +
+  col1 <-
+    figs[[1]] + p2 + figs[[2]] + p4 + figs[[3]] + p6 + p7 + p8 +
     theme_patchwork(
-      design = design,
-      widths = ggplot2::unit(c(1.5, 1.75, 1.5, 1.75), "in"),
-      heights = ggplot2::unit(c(0.75, 0.75, 1, 0.5), "in"),
-      guides = "collect"
+      design = "ab \n cd \n ef \n gh",
[…]

### 2024-04-12 R/metab.R

@@ -706,7 +706,7 @@ plot_bleo_lactate <- function(x) {
     ) +
     ggplot2::labs(
       x = NULL,
-      y = "Lactate"
+      y = "Lactate (normalized)"
     ) +
     ggplot2::scale_color_manual(
       values = clrs,

### 2024-04-12 R/rnaseq.R

@@ -92,7 +92,7 @@ plot_rnaseq_pca <- function(pca_data) {
     ggplot2::coord_fixed(clip = "off") +
     ggplot2::theme(
       legend.position = "bottom",
-      legend.key.size = ggplot2::unit(1, units = "lines"),
+      # legend.key.size = ggplot2::unit(1, units = "lines"),
       legend.margin = ggplot2::margin(t = -60),
       legend.text.align = 0
     )

### 2024-04-12 R/seahorse.R

@@ -303,7 +303,7 @@ plot_seahorse_pheno <- function(x) {
       nudge_y = -0.5,
       color = "black",
       family = "Calibri",
-      size = 1.75,
+      size = 2,
       show.legend = FALSE
     ) +
     ggplot2::scale_fill_manual(

### 2024-04-12 manuscript/manuscript.qmd
revise figures for submission

@@
     fig02,
-    fig02s, 
     fig03, 
-    fig03s, 
     fig04, 
-    fig04s, 
-    mid_tgf_plot_img,
     fig05, 
-    fig05s, 
     fig06, 
     fig07, 
-    fig07s, 
     fig08, 
-    fig08s, 
-    fig09, 
-    fig09s1, 
-    fig10,
-    fig10s, 
-    resources_table,
     NULL
@@
 \newpage
+
+# Summary
+
+Small molecule inhibitors of lactate transporters, including the novel MCT4 inhibitor VB253, reprogram fibroblast metabolism to prevent myofibroblast differentiation and to decrease bleomycin-induced pulmonary fibrosis.
 
@@
 
-flextable::flextable_to_rmd(resources_table)
+# flextable::flextable_to_rmd(resources_table)
 ```
@@
 
-![**Figure 1 - Lactate transporter expression increases in human IPF lungs and experimental models.** (**A**) Immunoblot of whole lung homogenates from explanted IPF lungs and controls (Ctl).
-(**B**) Quantification of band densities from (A).
-(**C**) Immunoblot of whole lung homogenates from bleomycin (Bleo)- and vehicle (Ctl)-treated mice.
-(**D**) Quantification of band densities from (C).
-(**E**) Representative immunoblot of cell lysates from normal human lung fibroblasts treated with TGFβ to induce myofibroblast differentiation.
-(**F**) Quantification of band densities from (E).
+![**Fig.
+1.
+Lactate transporter expression increases in human IPF lung and experimental models.** (**A**) MCT1, MCT4, and α-SMA protein expression in whole lung homogenates from explanted IPF lungs and controls (Ctl).
+(**B**) MCT1, MCT4, and α-SMA protein expression in whole lung homogenates from bleomycin (Bleo)- and vehicle (Ctl)-treated mice.
+(**C**) MCT1, MCT4, and α-SMA protein expression in cell lysates from normal human lung fibroblasts treated with TGFβ to induce myofibroblast differentiation.
 Individual data points are biological replicates.
-Summary data are mean ± SEM (\* p-value < 0.05).](figs/Figure 01.png)
+Summary data are mean ± SEM (\* p-value < 0.05).](figs/F_01.png)
 
@@
 
-![**Figure 2 - Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Representative immunoblot of normal lung fibroblast cell lysates after treatment with siRNA targeting MCT1 or MCT4.
-(**B**) Quantification of protein expression levels following MCT1 and MCT4 knockdown.
-(**C**) Immunoblot of IPF lung fibroblast lysates from three separate donors after treatment with TGFβ and MCT inhibitors AZD3965 (AZD) or VB124 (VB).
-(**D**) Quantification of band densities from (C).
-(**E**) Representative immunoblot of normal lung fibroblasts treated with TGFβ and MCT inhibitors.
-(**F**) Quantification of α-SMA expression following MCT inhibition.
-(**G**) Representative image of a gel contraction assay after 24 h treatment with TGFβ and MCT inhibitors.
-(**H**) Quantification of contracted areas.
+![**Fig.
+2.
+Lactate transport inhibition decreases myofibroblast differentiation and pro-fibrotic gene transcription *in vitro*.** (**A**) RNA interference targeting MCT1 or MCT4 decreases TGFβ-stimulated α-SMA expression in normal human lung fibroblasts.
+(**B**-**C**) Small molecule inhibitors of MCT1 (AZD3965, AZD) or MCT4 (VB124, VB) decrease TGFβ-stimulated α-SMA expression in IPF lung fibroblasts (B) and normal human lung fibroblasts (C).
+(**D**) Lactate transport inhibitors decrease gel contractility measured 24 h following TGFβ.
+(**E**) Principal components analysis of RNA-seq data from lung fibroblasts treated with TGFβ (*bold colors*) or vehicle (*light colors*) and lactate transport inhibitors (N = 4).
+(**F**) Dot plot of Hallmark gene set enrichment analysis.
+Significantly enriched pathways with adjusted p-value < 0.05 are included and points are colored by normalized enrichment score (NES).
+Positive NES indicated relative enrichment following TGFβ compared to control (TGFβ) or with MCT inhibitor compared to vehicle in TGFβ-treated cells.
+(**G**) Volcano plot of significantly differentially expressed genes in TGFβ-stimulated cells treated with AZD3965 and VB124 compared to vehicle control.
+Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.
+(**H**) Transcript counts from the leading edge of enrichment for the epithelial mesenchymal transition Hallmark gene set demonstrating down-regulation of pro-fibrotic genes with MCT inhibition.
 Individual data points are biological replicates.
-Summary data are mean ± SEM (\* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 02.png)
+Summary data are mean ± SEM (\* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/F_02.png)
 
@@
 
-![**Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined by enzymatic assay following TGFβ stimulation of cells treated with siRNA targeting MCT1 or MCT4, separately or together (N = 6 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares siMCT *v.* siCTL).
+![**Fig.
+3.
+Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined by enzymatic assay following TGFβ stimulation of cells treated with siRNA targeting MCT1 or MCT4, separately or together (N = 6 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares siMCT *v.* siCTL).
 (**B**) TGFβ-stimulated lactate secretion was measured following treatment with MCT1 inhibitor AZD3965 (AZD), MCT4 inhibitor VB124 (VB), or both (N = 4-13 biological replicates, \* adjusted p-value < 0.05, *black* compared TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
@@
 TGFβ increases ATP production from both glycolysis and oxidative phosphorylation, favoring the former, while MCT inhibition increases mitochondrial ATP production.
-Summary data are mean ± SEM.](figs/Figure 03.png)
+Summary data are mean ± SEM.](figs/F_03.png)
 
@@
 
-![**Figure 4 - Lactate transporter inhibition promotes oxidative phosphorylation.** (**A**) Principal components analysis of metabolites in conditioned medium from lung fibroblasts treated with TGFβ (*bold colors*) or vehicle (*light colors*) and lactate transport inhibitors (N = 5).
+![**Fig.
+4.
+Lactate transporter inhibition promotes oxidative phosphorylation.** (**A**) Principal components analysis of metabolites in conditioned medium from lung fibroblasts treated with TGFβ (*bold colors*) or vehicle (*light colors*) and lactate transport inhibitors (N = 5).
 (**B**) Extracellular lactate determined by LC-MS.
@@
 (**D**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
-Positive NES are relatively enriched in AZD/VB-treated cells while negative NES are relatively enriched in Vehicle-treated cells.
+Positive NES indicates enrichment in AZD/VB-treated cells while negative NES indicates enrichment in Vehicle-treated cells.
 (**E**) Principal components analysis of intracellular metabolites extracted from lung fibroblasts stimulated with TGFβ (*bold colors*) or vehicle (*light colors*) in the presence of lactate transport inhibitors (N = 5).
@@
 Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-(**H**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
-Positive NES are relatively enriched in AZD/VB-treated cells while negative NES are relatively enriched in Vehicle-treated cells.
-(**I**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~6~\]-glucose.
-(**J**) Labeled fraction of intracellular metabolites following \[U-^13^C~3~\]-lactate treatment.
-(N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
-Summary data are mean ± SEM.](figs/Figure 04.png)
+(**H**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by NES.
+Positive NES indicates enrichment in AZD/VB-treated cells while negative NES indicates enrichment in Vehicle-treated cells.
+(**I**) Schematic of ^13^C isotope labeling of carbon atoms from glucose (*blue*) or lactate (*pink*) into tricarboxylic acid cycle metabolites.
+Each circle represents a carbon atom.
+(**J**-**K**) Stable isotope incorporation from \[U-^13^C~6~\]-glucose (J) or \[U-^13^C~3~\]-lactate (K) into the intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+(**L**) Intracellular lactate oxidation is coupled to NADH/NAD^+^, which was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**M**) Normalized CellROX fluorescence, an indicator of cellular reactive oxygen species, normalized to TGFβ-treated cells (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
+(**N**) MitoSOX fluorescence, a marker of mitochondrial superoxide production, was measured and normalized to MitoTracker fluorescence, a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
+Summary data are mean ± SEM.](figs/F_04.png)
 
@@
 
-![**Figure 5 - Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A**) Principal components analysis of RNA-seq data from lung fibroblasts treated with TGFβ (*bold colors*) or vehicle (*light colors*) and lactate transport inhibitors (N = 4).
-(**B**) Dot plot of Hallmark gene set enrichment analysis.
-Significantly enriched pathways with adjusted p-value < 0.05 are included and points are colored by normalized enrichment score (NES).
-Positive NES indicated relative enrichment following TGFβ compared to control (TGFβ) or with MCT inhibitor compared to vehicle in TGFβ-treated cells.
-(**C**) Volcano plot of significantly differentially expressed genes in TGFβ-stimulated cells treated with AZD3965 and VB124 compared to vehicle control.
-Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.
-(**D**) Transcript counts from the leading edge of enrichment for the epithelial mesenchymal transition Hallmark gene set demonstrating down-regulation of pro-fibrotic genes with MCT inhibition.](figs/Figure 05.png)
+![**Fig.
+5.
+Lactate transport inhibition does not alter TGFβ signaling pathways.** (**A**) Phosphorylation of canonical TGFβ receptor target, Smad3, in lung fibroblasts treated with MCT inhibitors.
+(**B**) Phosphorylation of non-canonical TGFβ receptor target, ERK, in lung fibroblasts treated with MCT inhibitors.
+Summary data are mean ± SEM (\* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/F_05.png)
 
@@
 
-![**Figure 6 - Lactate transport inhibition does not alter TGFβ signaling pathways.** (**A**) Representative immunoblot of normal lung fibroblast cell lysates after treatment with extracellular lactate (10 mM).
-(**B**) Quantification of protein expression levels from (A).
-(**C**) Immunoblot of HIF-1α stabilization 6 h after TGFβ treatment in the presence of MCT inhibitors.
-(**D**) Quantification of band densities from (C).
-(**E**) Representative immunoblot of canonical TGFβ receptor target, Smad3.
-(**F**) Quantification of band densities from (E).
-(**G**) Representative immunoblot of non-canonical TGFβ receptor target, ERK (**H**) Quantification of band densities from (E).
-(**I**) Representative immunoblot of histones purified from lung fibroblasts treated with TGFβ and MCT inhibitors.
-(**J**) Quantification of band densities from (I).
-Summary data are mean ± SEM (\* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).](figs/Figure 06.png)
-
-\newpage
-
-![**Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADH/NAD^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-(**B**) Relative CellROX fluorescence, an indicator of cellular reactive oxygen species, normalized to TGFβ-treated cells (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
-Summary data are mean ± SEM.](figs/Figure 07.png)
-
-\newpage
-
-![**Figure 8 - Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
+![**Fig.
+6.
+Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
 Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21.
@@
 (**F**) Right lung hydroxyproline content (N = 10-13).
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.png)
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_06.png)
 
@@
 
-![**Figure 9 - Lactate transporter inhibition reprograms lung metabolism *in vivo*.** (**A**) Plasma and lung lactate levels following 14-day treatment with AZD3965 or VB124 (N = 6-9; \* adjusted p-value < 0.05 compared to bleomycin control).
+![**Fig.
+7.
+Lactate transporter inhibition reprograms lung metabolism *in vivo*.** (**A**) Plasma and lung lactate levels following 14-day treatment with AZD3965 or VB124 (N = 6-9; \* adjusted p-value < 0.05 compared to bleomycin control).
 (**B**) Representative images from multi-isotope imaging mass spectrometry (MIMS).
 The ^12^C^14^N ion illustrates tissue histology.
-The ^12^C^15^N/^12^C^14^N ion ratio image indicates enrichment from ^15^N-proline (colorbar 1-3-fold of natural abundance).
-The ^12^C~2~^2^H/^12^C~2~^1^H ratio indicates enrichment above natural abundance from ^2^H-glucose (colorbar 1-10-fold of natural abundance).
-(**C**) Quantification of isotope enrichment (N = 3 biological replicates, \* adjusted p-value-value < 0.05 compared to bleomycin-treated control).](figs/Figure 09.png)
+The ^12^C^15^N/^12^C^14^N ion ratio image shows enrichment from ^15^N-proline (colorbar 1-3-fold of natural abundance).
+The ^12^C~2~^2^H/^12^C~2~^1^H ratio shows enrichment above natural abundance from ^2^H-glucose (colorbar 1-10-fold of natural abundance).
+(**C**) Quantification of isotope enrichment (N = 3 biological replicates, \* adjusted p-value-value < 0.05 compared to bleomycin-treated control).](figs/F_07.png)
 
@@
 
-![**Figure 10 - VB253, a novel MCT4 inhibitor, is potently antifibrotic *in vitro* and *in vivo*.** (**A**) Dose-response of α-SMA protein expression in IPF lung fibroblasts treated with VB253 or nintedanib (NIN).
+![**Fig.
+8.
+VB253, a novel MCT4 inhibitor, is potently antifibrotic *in vitro* and *in vivo*.** (**A**) Dose-response of α-SMA protein expression in IPF lung fibroblasts treated with VB253 or nintedanib (NIN).
 (**B**) Number of nuclei, a marker of cell viability, following treatment with VB253 or NIN.
@@
 (**D**) Enhanced pause (Penh) is a non-invasive index of mouse tidal breathing that increases with bleomycin treatment.
-(**E-G**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (E) and decreased α-SMA protein expression (F-G).
-(**H-J**) Similar histologic improved was observed in aged mice with decreased histologic fibrosis (H) and decreased α-SMA protein expression (I-J).
+(**E**-**G**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (E) and decreased α-SMA protein expression (F-G).
+(**H**-**J**) Similar histologic improved was observed in aged mice with decreased histologic fibrosis (H) and decreased α-SMA protein expression (I-J).
 (**K**) VB253 decreased lung lactate, similar to established antifibrotics.
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 10.png)
-
-\newpage
-
-# Supplemental Figures
-
-![**Supplemental Figure 2 - Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
-(**B**) Representative immunoblot of normal lung fibroblasts treated with TGFβ and MCT inhibitors AR-C155858 (AR) or VB124 (VB).
-(**C**) Quantification of α-SMA expression following MCT inhibition.
-(**D-E**) Cell counts following pharmacologic inhibition of lactate transporters.
-Summary data are mean ± SEM (\* p-value-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-](figs/Figure 02.supplement.png)
-
-\newpage
-
-![**Supplemental Figure 3 - Extracellular fluxes following lactate transport inhibition.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
-(**B**) Extracellular glucose was determined following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
-(**C**) Summary data from Seahorse extracellular flux analysis.
-(**D**) Cell bioenergetic stress testing results from Seahorse analysis.
-Glycolytic capacity measures the increase in proton efflux rate following oligomycin treatment.
-Spare respiratory capacity measures the increase in oxygen consumption rate (OCR) following FCCP treatment compared to basal OCR.
-Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin).
-The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p-value < 0.05 compared to TGFβ/Veh).
-Summary data are mean ± SEM.](figs/Figure 03.supplement.png)
-
-\newpage
-
-![**Supplemental Figure 4.1 - Metabolomic profiling of lung fibroblasts treated with lactate transport inhibitors.** (**A-D**) Extracellular (A-B) and intracellular (C-D) metabolomic profile and metabolite set enrichment analyses of lung fibroblasts treated with TGFβ.
-(**E-H**) Extracellular (E-F) and intracellular (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment.
-(**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment.
-Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
-Positive NES indicate enrichment in treated cells while negative NES indicate relative enrichment in Vehicle-treated cells.
-(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
-Summary data are mean ± SEM.](figs/Figure 04.supplement.1.png)
-
-\newpage
-
-![**Supplemental Figure 4.2 - Stable isotope tracing in lung fibroblasts treated with TGFβ.** Lung fibroblasts were cultured with stable isotopes of glucose, lactate, and glutamine during treatment with TGFβ (*columns*).
-Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*).
-Significant differences in isotopic enrichment are indicated (\* adjusted p-value < 0.05).
-Summary data are mean ± SEM (N = 4 biological replicates).](figs/Figure 04.supplement.2.png)
-
-\newpage
-
-![**Supplemental Figure 5 - Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A-C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh.
-Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/Figure 05.supplement.png)
-
-\newpage
-
-![**Supplemental Figure 7 - Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-(**B**) MitoSOX fluorescence, a marker of mitochondrial superoxide production, was measured and normalized to MitoTracker fluorescence, a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
-(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p-value < 0.05 for the main treatment effect).
-(**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates).
-Summary data are mean ± SEM.](figs/Figure 07.supplement.png)
-
-\newpage
-
-![**Supplemental Figure 8 - Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly.
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/Figure 08.supplement.png)
-
-\newpage
-
-![**Supplemental Figure 9 - Metabolic reprogramming by lactate transport inhibition *in vivo*.** (**A-D**) Plasma (A-B) and lung (C-D) metabolomic profile and metabolite set enrichment analyses in bleomycin-treated mice.
-(**E-H**) Plasma (E-F) and lung (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment in bleomycin-treated mice.
-(**I-L**) Plasma (I-J) and lung (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment in bleomycin-treated mice.
-Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by bleomycin treatment or MCT inhibition ordered by normalized enrichment score (NES).
-Positive NES indicate enrichment in treated mice while negative NES indicate relative enrichment in control or untreated mice.](figs/Figure 09.supplement.1.png)
-
-\newpage
-
-![**Supplemental Figure 10 - VB253 does not attenuate canonical TGFβ signaling.** Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
-Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/Figure 10.supplement.png)
-
-\newpage
-
-# Fin
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_10.png)

### 2024-04-12 manuscript/supplement.qmd
revise figures for submission

@@
 ---
+
+```{r}
+#| label: setup
+#| include: false
+
+# ensure refs are up-to-date
+targets::tar_load(refs)
+targets::tar_load(pkg_citations)
+
+# ensure figures are up-to-date
+targets::tar_load(
+  c(
+    fig01s,
+    fig02s, 
+    fig03s, 
+    fig04s, 
+    fig05s, 
+    fig06s, 
+    fig07s, 
+    fig08s, 
+    fig09s, 
+    fig10s,
+    NULL
+  )
+)
+```
+
+\newpage
+
+# Materials and Methods
+
+\newpage
+
+# Figures
+
+![**Fig.
+S1.
+Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
+(**B**) Small molecule inhibitors of MCT1/2 (AR-C155858, AR) or MCT4 (VB124, VB) decrease TGFβ-stimulated α-SMA expression in normal human lung fibroblasts.
+(**C**-**D**) Cell counts following pharmacologic inhibition of lactate transporters.
+Summary data are mean ± SEM (\* p-value-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+](figs/SF_01.png)
+
+![**Fig.
+S2.
+Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A**-**C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh.
+Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/SF_02.png)
+
+\newpage
+
+![**Fig.
+S3.
+Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+(**B**) Extracellular glucose was determined following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
+(**C**) Summary data from Seahorse extracellular flux analysis.
+(**D**) Cell bioenergetic stress testing results from Seahorse analysis.
+Glycolytic capacity measures the increase in proton efflux rate following oligomycin treatment.
+Spare respiratory capacity measures the increase in oxygen consumption rate (OCR) following FCCP treatment compared to basal OCR.
+Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin).
+The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p-value < 0.05 compared to TGFβ/Veh).
+Summary data are mean ± SEM.](figs/SF_03.png)
+
+\newpage
+
+![**Fig.
+S4.
+Metabolomic profiling of lung fibroblasts treated with lactate transport inhibitors.** (**A**-**D**) Extracellular (A-B) and intracellular (C-D) metabolomic profile and metabolite set enrichment analyses of lung fibroblasts treated with TGFβ.
+(**E-H**) Extracellular (E-F) and intracellular (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment.
+(**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment.
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Positive NES indicates enrichment in treated cells while negative NES indicates enrichment in Vehicle-treated cells.
+(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+Summary data are mean ± SEM.](figs/SF_04.png)
+
+\newpage
+
+![**Fig.
+S5.
+Stable isotope tracing in lung fibroblasts treated with TGFβ.** Lung fibroblasts were cultured with stable isotopes of glucose, lactate, and glutamine during treatment with TGFβ (*columns*).
+Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*).
+Significant differences in isotopic enrichment are indicated (\* adjusted p-value < 0.05).
+Summary data are mean ± SEM (N = 4 biological replicates).](figs/SF_05.png)
+
+\newpage
+
+![**Fig.
+S6.
+Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**B**) MitoTracker fluorescence was measured as a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
+(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p-value < 0.05 for the main treatment effect).
+(**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates).
+Summary data are mean ± SEM.](figs/SF_06.png)
+
+\newpage
+
+![**Fig.
+S7.
+Mechanisms of lactate signaling.** (**A**) Exogenous lactate (10 mM) does not ameliorate TGFβ-stimulated α-SMA expression.
+(**B**) Lactate transport inhibition does not attenuate HIF-1α stabilization 6 h after TGFβ treatment.
+Summary data are mean ± SEM.
+(N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).](figs/SF_07.png)
+
+\newpage
+
+![**Fig.
+S8.
+Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly.
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/SF_08.png)
+
+\newpage
+
+![**Fig.
+S9.
+Metabolic reprogramming by lactate transport inhibition *in vivo*.** (**A**-**D**) Plasma (A-B) and lung (C-D) metabolomic profile and metabolite set enrichment analyses in bleomycin-treated mice.
+(**E**-**H**) Plasma (E-F) and lung (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment in bleomycin-treated mice.
+(**I**-**L**) Plasma (I-J) and lung (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment in bleomycin-treated mice.
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by bleomycin treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Positive NES indicate enrichment in treated mice while negative NES indicate relative enrichment in control or untreated mice.](figs/SF_09.png)
+
+\newpage
+
+![**Fig.
+S10.
+VB253 does not attenuate canonical TGFβ signaling.** Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
+Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/SF_10.png)

### 2024-04-12 R/plots.R

@@ -580,5 +580,8 @@ plot_msea_table <- function(
     gt::cols_align("center", c(.data$NES)) |>
     gtExtras::gt_theme_538() |>
     gt::opt_table_font(font = "Calibri") |>
-    gt::tab_options(table_body.border.bottom.color = "black")
+    gt::tab_options(
+      table.width = 350,
+      table_body.border.bottom.color = "black"
+    )
 }

### 2024-04-12 R/utils.R

@@ -36,14 +36,84 @@ manuscript_path <- function(nm) {
 #     )
 # }
 
-plot_image <- function(img, scale = 1, hjust = 0, vjust = 0) {
-  png::readPNG(img, native = TRUE) |>
-    patchwork::wrap_elements(plot = _) +
-    theme_plot() +
-    ggplot2::theme(
-      panel.border = ggplot2::element_blank(),
-      axis.line = ggplot2::element_blank()
-    )
+# plot_image <- function(img, scale = 1, hjust = 0, vjust = 0) {
+#   png::readPNG(img, native = TRUE) |>
+#     patchwork::wrap_elements(plot = _) +
+#     theme_plot() +
+#     ggplot2::theme(
+#       panel.border = ggplot2::element_blank(),
+#       axis.line = ggplot2::element_blank()
+#     )
+# }
+
+plot_image <- function(x) {
+
+  path <- NULL
+
+  if (is.character(x) && file.exists(x)) {
+    path <- x
+    img <- magick::image_read(x)
+  } else {
+    img <- x
+  }
+
+  x_dim <- magick::image_info(img)$width
+  y_dim <- magick::image_info(img)$height
+  max_dim <- max(x_dim, y_dim)
+  aspect_ratio <- y_dim / x_dim
+
+  fig <-
+    patchwork::wrap_elements(
+      full =
+        ggplot2::ggplot() +
+        ggplot2::annotation_custom(
+          grid::rasterGrob(
+            image = img,
+            interpolate = TRUE,
+            width = ggplot2::unit(1, "npc"),
+            height = ggplot2::unit(1, "npc")
+          )
+        ) +
+        ggplot2::theme(
+          panel.border = ggplot2::element_blank(),
+          axis.line = ggplot2::element_blank(),
+          aspect.ratio = aspect_ratio
+        )
+    ) +
+    theme_plot()
+
+  fig$fig_data <- list(
+    path = path,
+    aspect_ratio = aspect_ratio,
+    x_dim = x_dim,
+    y_dim = y_dim
+  )
+
+  fig
+}
+
+plot_scale <- function(...) {
+  figs <- list(...)
+
+  max_x <- max(purrr::map_dbl(figs, \(x) x$fig_data$x_dim))
+  max_y <- max(purrr::map_dbl(figs, \(x) x$fig_data$y_dim))
+
+  purrr::map(
+    figs,
+    \(x) {
+      img <- magick::image_read(x$fig_data$path)
[…]

### 2024-04-16 manuscript/manuscript.qmd
format for STM and co-author review

@@
 - id: utah_gen
-  name: Department of Human Genetics, University of Utah, Salt Lake City, UT
+  name: Department of Internal Medicine, University of Utah, Salt Lake City, UT
 - id: umd
@@
 - id: tj
-  name: Department of Pathology, Anatomy & Cell Biology, Thomas Jefferson University, Philadelphia, PA
+  name: Department of Pathology and Genomic Medicine, Thomas Jefferson University, Philadelphia, PA
 - id: mgh_path
@@
 bibliography: 
-- library.bib
+- ms.bib
 - pkgs.bib
-csl: jci.csl
+csl: stm.csl
 
@@
 # ensure refs are up-to-date
-targets::tar_load(refs)
+targets::tar_load(refs_ms)
 targets::tar_load(pkg_citations)
@@
 
-# Conflicts of Interest
-
-*Email Will with COI.*
-
-B.A.M. has received consulting fees from Actelion and Tenax and has performed investigator-initiated research with support from Deerfield, all outside the scope of this research.
-
 \newpage
@@
 
-Myofibroblast differentiation requires metabolic reprogramming to support the extracellular matrix synthesis that drives pulmonary fibrosis.
-While prior studies demonstrate that targeting fibroblast metabolism mitigates experimental pulmonary fibrosis, this approach has been slow to translate to clinical practice.
-Here, we examined the contribution of lactate transporters MCT1 and MCT4 to myofibroblast differentiation and experimental pulmonary fibrosis using novel drugs suitable for human clinical trials, including a newly developed small molecule inhibitor of MCT4, VB253.
-We demonstrate that inhibition of MCT1 or MCT4 decreases TGFβ-stimulated myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
-Using bioenergetic analyses, stable isotope tracing, metabolomic profiling, and imaging mass spectrometry in cells and mice, we demonstrate lactate transport inhibition promotes oxidative phosphorylation and decreases glucose metabolite incorporation into fibrotic lung regions.
-We characterize the transcriptional consequences of these inhibitors and propose that metabolic signaling from lactate by protein lactylation and by reactive oxygen species detoxification account for changes in gene expression.
-Collectively, these data establish lactate transport as a requirement for myofibroblast differentiation, establish these transporters as a pharmacologic target in experimental fibrosis, and support continued clinical study in patients with pulmonary fibrosis for whom limited therapeutic options exist.
+Myofibroblast differentiation, essential for driving extracellular matrix synthesis in pulmonary fibrosis, requires increased glycolysis.
+While glycolytic cells must export lactate, the contributions of lactate transporters to myofibroblast differentiation are unknown.
+In this study, we investigated how MCT1 and MCT4, key lactate transporters, influence myofibroblast differentiation and experimental pulmonary fibrosis.
+Our findings reveal that inhibiting MCT1 or MCT4 reduces TGFβ-stimulated myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
+Through comprehensive metabolic analyses, including bioenergetics, stable isotope tracing, metabolomics, and imaging mass spectrometry in both cells and mice, we demonstrate that inhibiting lactate transport enhances oxidative phosphorylation, reduces reactive oxygen species production, and diminishes glucose metabolite incorporation into fibrotic lung regions.
+Furthermore, we introduce VB253, a novel MCT4 inhibitor, which effectively ameliorated pulmonary fibrosis in both young and aged mice, comparable to established antifibrotic therapies.
+These results underscore the necessity of lactate transport for myofibroblast differentiation, identify MCT1 and MCT4 as promising pharmacologic targets in pulmonary fibrosis, and support further evaluation of lactate transport inhibitors for patients with pulmonary fibrosis for whom limited therapeutic options currently exist.
 
@@
 Fundamental changes in myofibroblast metabolism support myofibroblast differentiation and extracellular matrix production [@bernard2015; @xie2015; @goodwin2018; @hamanaka2019; @selvarajah2019].
-In particular, increased glycolysis and lactate production have been observed in IPF myofibroblasts *ex vivo* and following TGF-β1 (TGFβ)-induced myofibroblast differentiation *in vitro* [@kottmann2012; @xie2015; @kottmann2015; @bernard2015].
+In particular, increased glycolysis and lactate production have been observed in IPF myofibroblasts *ex vivo* and following transforming growth factor β1 (TGFβ)-induced myofibroblast differentiation *in vitro* [@kottmann2012; @xie2015; @kottmann2015; @bernard2015].
 These metabolic changes are critical for fibrogenesis, as small molecule inhibitors of glucose uptake, glycolysis, and lactate fermentation prevent myofibroblast differentiation *in vitro* and attenuate pulmonary fibrosis in animal models [@bernard2015; @kottmann2012; @xie2015; @goodwin2018; @judge2018; @cho2017].
@@
 
-Toward this end, we sought to characterize the effects of a novel metabolic approach - inhibition of lactate transport - on myofibroblast differentiation and experimental pulmonary fibrosis.
-Lactate must be secreted to sustain glycolysis in myofibroblasts.
-A family of monocarboxylate transporters (MCT1-4) maintains lactate homeostasis and small molecule inhibitors of these transporters have been under active clinical development for oncologic applications where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
-Importantly, compared to previously investigated inhibitors of glycolysis, MCT inhibitors exhibit favorable pharmacologic profiles and have successfully translated to human clinical trials for advanced solid tumors [@halford2023].
+Toward this end, we aimed to examine the impact of a novel metabolic strategy --- lactate transport inhibition --- on myofibroblast differentiation and experimental pulmonary fibrosis.
+Myofibroblasts rely on lactate secretion to sustain glycolysis, a process regulated by a family of monocarboxylate transporters (MCT1-4).
+Inhibitors targeting these transporters have been actively explored in clinical trials for oncological conditions where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
+Importantly, MCT inhibitors present favorable pharmacologic profiles compared to previously studied glycolysis inhibitors, with successful translation to human clinical trials for advanced solid tumors [@halford2023].
 Before the promise of this therapeutic approach in IPF may be realized, however, the preclinical efficacy and molecular mechanisms-of-action of lactate transport inhibitors must be demonstrated experimentally.
@@
 In this work, we evaluated the contribution of lactate transporters to experimental pulmonary fibrosis.
-We found increased expression of the lactate transporters MCT1 and MCT4 IPF patient lungs, consistent with increased glycolysis previously observed.
+We found increased expression of the lactate transporters MCT1 and MCT4 IPF patient lungs.
 Inhibition of these transporters attenuated bleomycin-induced lung fibrosis *in vivo* and TGFβ-induced myofibroblast differentiation *in vitro*, where MCT4 inhibition demonstrated increased therapeutic efficacy.
-Using metabolomics, stable isotope tracing, and high-resolution spatial metabolomic imaging, we find that lactate transport inhibition promotes glucose oxidation.
-These metabolic changes signal, in part, through increased protein lactylation and decreased intracellular reactive oxygen species (ROS) production.
-Based on these findings, we report a novel MCT4 inhibitor suitable for human clinical studies.
-Together, these data illuminate a novel approach to disrupt the metabolic program fueling the exuberant fibrosis at the core of IPF and provide proof-of-concept to target lactate transporters in the treatment of this deadly disease.
+Using metabolomics, stable isotope tracing, and high-resolution spatial metabolomic imaging, we find that lactate transport inhibition promotes glucose oxidation and decreases pro-fibrogenic reactive oxygen species (ROS) production.
+Building on these results, we introduce VB253, a novel MCT4 inhibitor suitable for human clinical studies.
+Together, our findings offer insights into disrupting the metabolic pathways driving IPF fibrosis and present a promising strategy targeting lactate transporters for the treatment of this fatal condition.
 
@@
 
-Of the four lactate transporters, MCT1 and MCT4 are the most highly expressed in the lung [@frangogiannis2020].
-Based on this observation, we characterized MCT1 and MCT4 expression in human IPF lung explants obtained at the time of transplantation.
-Consistent with a pathologic role for MCT1 and MCT4 in IPF, we found significantly increased expression of these proteins in human IPF lung compared to non-fibrotic control samples (**Figure 1A-B**).
-These findings were recapitulated in experimental pulmonary fibrosis where intratracheal bleomycin increased both MCT1 and MCT4 expression (**Figure 1C-D**).
+Among the four lactate transporters, MCT1 and MCT4 exhibit the highest expression levels in the lung [@frangogiannis2020].
+Based on this observation, we investigated the expression of MCT1 and MCT4 in lung explants from patients with IPF obtained during transplantation.
+Consistent with a pathologic role for MCT1 and MCT4 in IPF, we observed a significant upregulation of both MCT1 and MCT4 proteins in IPF lung tissues compared to non-fibrotic controls (**Fig.
+1A**).
+These findings were corroborated in an experimental model of pulmonary fibrosis, where intratracheal bleomycin administration led to increased expression of both MCT1 and MCT4 (**Fig.
+1B**).
 
-A defining feature of IPF is the activation of tissue myofibroblasts, which are characterized by the *de novo* expression of smooth muscle α-actin (α-SMA); stress fiber formation; and increased migration, contraction, and extracellular matrix production [@selvarajah2021; @xie2015; @tomasek2002].
+A hallmark of IPF is the activation of tissue myofibroblasts, distinguished by their *de novo* expression of smooth muscle α-actin (α-SMA); stress fiber formation; and increased migration, contraction, and extracellular matrix production [@selvarajah2021; @xie2015; @tomasek2002].
 TGFβ is the most potent inducer of myofibroblast activation *in vitro* and *in vivo*.
-TGFβ-dependent α-SMA expression is a well-accepted and widely utilized model of myofibroblast activation relevant to pulmonary fibrosis [@goodwin2018; @cui2019; @rangarajan2018; @locy2020].
-Consistent with our findings in human IPF lungs, we observed increased expression of MCT1 and MCT4 in normal human lung fibroblasts following TGFβ treatment (**Figure 1E-F**).
-These findings are consistent with increased expression of other glycolytic enzymes and the associated metabolic changes that have been previously observed in these cells [@xie2015].
-Together, these data demonstrate that pulmonary fibrosis is associated with increased expression of lactate transporters in the lung generally and in myofibroblasts specifically.
+The TGFβ-dependent upregulation of α-SMA expression serves as a well-established and widely utilized model for studying myofibroblast activation pertinent to pulmonary fibrosis [@goodwin2018; @cui2019; @rangarajan2018; @locy2020].
+Consistent with our findings in human IPF lungs, we observed increased expression of MCT1 and MCT4 in normal human lung fibroblasts following TGFβ treatment (**Fig.
+1C**).
+These findings align with increased expression of other glycolytic enzymes and the associated metabolic changes previously documented in these cells [@xie2015].
+Taken together, these data underscore the association between pulmonary fibrosis and lactate transporter expression in the lung generally and in myofibroblasts specifically.
 
@@
 
-Using RNA interference and pharmacologic approaches, we next tested whether MCT expression and activity were required for myofibroblast differentiation *in vitro*.
-Lung fibroblasts were transfected with siRNA targeting MCT1 and MCT4 separately and together.
+We proceeded to investigate whether MCT expression and activity were essential for myofibroblast differentiation *in vitro* using RNA interference and pharmacologic approaches.
+Lung fibroblasts were transfected with siRNA targeting MCT1 and MCT4 individually and in combination.
 After 24 h, the cells were treated with TGFβ for 48 h to induce myofibroblast differentiation.
-The siRNAs reduced lactate transporter protein levels > 80% (**Figure 2A-B**).
-Decreasing MCT1 or MCT4 expression caused a marked reduction in TGFβ-stimulated α-SMA expression.
+The siRNAs effectively reduced lactate transporter protein levels (**Fig.
+2A**).
+Reduction in either MCT1 or MCT4 expression caused a marked decrease in TGFβ-stimulated α-SMA expression.
 Notably, siMCT1 also decreased MCT4 expression.
-The mRNA sequence homology of these transporters is low and the MCT1 siRNA sequences do not target the MCT4 mRNA sequence, suggesting a post-translational mechanism accounting for decreased MCT1 expression decreasing MCT4 protein.
-No adverse effects of lactate transporter knockdown on cell viability were observed, and siMCT1 significantly increased cell count in both control and TGFβ-treated cells (**Supplemental Figure 2A**).
+No adverse effects on cell viability were noted following lactate transporter knockdown, and siMCT1 significantly increased cell count in both control and TGFβ-treated cells (**Fig.
+S1A**).
 
-Given the reduction of MCT4 expression following siMCT1 treatment, we next evaluated the effects of pharmacologic MCT inhibitors on myofibroblast differentiation.
-AZD3965 is a high-affinity (*K~i~* 1.6 nM) inhibitor of MCT1 [@curtis2017].
-VB124 is a recently developed high-affinity (IC~50~ 19 nM) inhibitor of MCT4 [@cluntun2021].
-IPF lung fibroblasts were differentiated with TGFβ in the presence of these MCT inhibitors (**Figure 2C-D**).
+After observing a reduction in MCT4 expression following siMCT1 treatment, we proceeded to evaluate the impact of pharmacological MCT inhibitors on myofibroblast differentiation to examine the independent effects of MCT1 and MCT4 inhibition.
+AZD3965, a high-affinity (*K~i~* 1.6 nM) inhibitor of MCT1 [@curtis2017], and VB124, a recently developed high-affinity (*K~i~* 11 nM) inhibitor of MCT4 [@cluntun2021], were used in this evaluation.
+IPF lung fibroblasts were differentiated with TGFβ in the presence of these MCT inhibitors (**Fig.
+2B**).
 MCT4 inhibition by VB124 alone, or in combination with AZD3965, decreased Col1a1 and α-SMA expression.
-These effects were also observed in normal lung fibroblasts where AZD3965 as well as VB124, separately or together, decreased α-SMA expression (**Figure 2E-F**).
-AR-C155858, which is a high-affinity (*K~i~* 2 nM) inhibitor of both MCT1 and MCT2 [@ovens2010], also decreased α-SMA expression alone and in combination with VB124 (**Supplemental Figure 2B-C**).
-Pharmacologic MCT inhibition did not affect cell counts over 48 h of treatment (**Supplemental Figure 2D-E**).
-As anticipated, decreased Col1a1 and α-SMA expression was associated with decreased myofibroblast contractility as assessed by gel contraction assay (**Figure 2G-H**).
+These effects were consistent with normal human lung fibroblasts where both AZD3965 and VB124, either individually or in combination, decreased α-SMA expression (**Fig.
+2C**).
+Similarly, AR-C155858, an inhibitor with high-affinity (*K~i~* 2 nM) for both MCT1 and MCT2 [@ovens2010], also reduced α-SMA expression, alone and in combination with VB124 (**Fig.
+S2B**).
+Importantly, pharmacologic MCT inhibition did not significantly impact cell count during the 48 h treatment (**Fig.
+S2C-D**).
+As expected, decreased Col1a1 and α-SMA expression correlated with reduced myofibroblast contractility, as demonstrated by gel contraction assay (**Fig.
+2D**).
 Together, these data indicate that MCT expression and activity are required for myofibroblast differentiation *in vitro*.
+
+## MCT inhibition attenuates pro-fibrotic transcriptional programs
+
+To further characterize the antifibrotic effects of lactate transport inhibition, we conducted RNA sequencing on lung fibroblasts treated with TGFβ in conjunction with AZD3965 or VB124 (**Fig.
+2E-H, S2**).
+Principal components analysis (PCA) revealed that the first principal component predominantly represented the effect of TGFβ treatment, while MCT4 inhibition, either alone or combined with MCT1 inhibition, aligned with the second principal component (**Fig.
+2E**).
+Samples treated with the MCT1 inhibitor AZD3965 were similar to vehicle-treated controls.
+
+Differential expression analysis of TGFβ-treated cells revealed the anticipated upregulation of extracellular matrix proteins (**Fig.
+S2A**) and enrichment of the epithelial-to-mesenchymal (EMT) gene set, among others (**Fig.
+2F**).
+In line with the PCA results, only GRIK4 (glutamate ionotropic receptor kainate type subunit 4) and BRI3 (brain protein I3) were differentially expressed following AZD3965 treatment (**Fig.
+S2B**).
+By contrast, VB124, either alone (**Fig.
+S2C**) or in combination with AZD3965 (**Fig.
+2G**), induced more significant alterations in fibroblast transcription, with `r deg_frac`% of `r deg_tot` genes being differentially expressed at a false discovery rate (FDR) < 0.05.
+
+Both AZD3965 and VB124 reversed TGFβ-dependent enrichment of the EMT gene set, which is the Hallmark gene set containing genes related to fibrosis (**Fig.
+2F**).
+Leading edge analysis of the EMT gene set identified seven genes shared among all three comparisons (*i.e.*, genes increased by TGFβ and decreased by both AZD3965 and VB124) (**Fig.
+2H**).
+These genes, including biglycan (BGN), COL6A3, Frizzled 8 (FZD8), matrix Gla protein (MGP), Prostate Transmembrane Protein, Androgen Induced 1 (PMEPA1), TIMP metallopeptidase inhibitor 1 (TIMP1), and tenascin-C (TNC), are known contributors to pulmonary fibrosis pathobiology or serve as biomarkers of disease or treatment response [@kwapiszewska2018; @williams2020; @baarsma2017; @todd2020; @bhattacharyya2016].
+Taken together, these findings suggest that lactate transport inhibitors attenuate the pro-fibrotic transcriptional program in TGFβ-treated lung fibroblasts.
 
@@
 
-MCTs are critical regulators of cellular lactate homeostasis.
-MCT1 is considered to be the main lactate importer in cells that use lactate to fuel oxidative phosphorylation or gluconeogenesis and is ubiquitously expressed.
-MCT1 also mediates lactate export in glycolytic cells [@halestrap2013; @brooks2018].
-MCT4 is considered to be the main lactate exporter in glycolytic cells and is up-regulated by hypoxia-inducible transcription factors [@ullah2006; @halestrap2013].
-MCT4 can also function as a lactate importer with a *K*~M~ of 1 mM [@contreras-baeza2019].
+MCTs play pivotal roles in maintaining cellular lactate homeostasis.
+MCT1 serves as the primary lactate importer in cells utilizing lactate for oxidative phosphorylation or gluconeogenesis and exhibits ubiquitous expression.
+Additionally, MCT1 facilitates lactate export in glycolytic cells [@halestrap2013; @brooks2018].
+By contrast, MCT4 functions as the main lactate exporter in glycolytic cells and is up-regulated by hypoxia-inducible transcription factors [@ullah2006; @halestrap2013].
+Importantly, MCT4 can also act as a lactate importer with a *K*~M~ of 1 mM [@contreras-baeza2019].
 
-To evaluate the metabolic consequences of lactate transporter inhibition, we measured extracellular lactate in the conditioned medium from cells treated with MCT siRNA or pharmacologic inhibitors (**Figure 3A-B**).
-Silencing MCT1 and MCT4, separately or together, decreased TGFβ-stimulated lactate efflux in lung fibroblasts (**Figure 3A**).
-Silencing both transporters was required to block TGFβ-stimulated lactate efflux completely.
-By contrast, pharmacologic inhibition of either MCT1 or MCT4 did not decrease TGFβ-stimulated lactate efflux (**Figure 3B**), and inhibition of both transporters was required to decrease extracellular lactate.
-Similarly, inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to prevent increases in lactate efflux (**Supplemental Figure 3A**).
-Compared to lactate production, the effect of MCT inhibitors on extracellular glucose was less pronounced, although we did observe a trend toward decreased glucose consumption by cells treated with both AZD3965 and VB124 (**Supplemental Figure 3B**).
-These findings are consistent with prior studies suggesting compensatory roles for MCT1 and MCT4 in lactate export [@benjamin2018].
-Moreover, these data also suggest that inhibition of glycolysis is not the mechanism by which lactate transport inhibition attenuates myofibroblast differentiation.
+To assess the metabolic consequences of lactate transporter inhibition, we quantified extracellular lactate in the conditioned medium from cells treated with MCT siRNA or pharmacologic inhibitors.
+Silencing MCT1 and MCT4, either individually or concurrently, decreased TGFβ-stimulated lactate efflux in lung fibroblasts (**Fig.
+3A**).
+By contrast, pharmacologic inhibition of either MCT1 or MCT4 alone did not decrease TGFβ-stimulated lactate efflux (**Fig.
+3B**); inhibition of both transporters was required to lower extracellular lactate levels.
+Inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to prevent increases in lactate efflux (**Fig.
+S3A**).
+Although less pronounced compared to lactate production, we observed a trend toward decreased extracellular glucose consumption in cells treated with both AZD3965 and VB124 (**Fig.
+S3B**).
+These findings align with previous studies indicating compensatory roles for MCT1 and MCT4 in lactate export [@benjamin2018].
+Furthermore, these results suggest that glycolysis inhibition is not the principal mechanism underlying the attenuation of myofibroblast differentiation by lactate transport inhibition.
 
-To further interrogate the metabolic consequences of lactate transport inhibition, we next measured proton efflux (PER) and oxygen consumption (OCR) rates of lung fibroblasts treated with TGFβ in combination with MCT inhibitors (**Figure 3C, Supplemental Figure 3C**).
-Since MCTs co-transport protons with lactate, PER correlates with lactate efflux.
-Consistent with prior reports [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER CR following TGFβ stimulation for 48 h, indicating increases in both glycolysis and oxidative phosphorylation.
-Similar to our direct measurements of extracellular lactate and glucose, simultaneous treatment with AZD3965 and VB124 was required to reduce PER.
-As anticipated, this decrease in PER was associated with a commensurate increase in OCR as myofibroblasts shifted their metabolism from glycolysis to oxidative phosphorylation.
-Unexpectedly, an increase in OCR was observed with AZD3965 and VB124 when administered individually.
-Indeed, the primary consequence of MCT1 or MCT4 inhibition alone was an increase in cellular ATP production rates driven by an increase in oxidative phosphorylation (**Figure 3D-E**).
-MCT inhibition decreased spare respiratory capacity, indicating that the basal respiratory rate of treated cells was closer to their maximal oxidative capacity (**Supplemental Figure 3D**).
-No significant differences in glycolytic capacity or electron transport chain coupling efficiency were observed (**Supplemental Figure 3D**).
-Together, these data suggest that the primary metabolic consequence of MCT inhibition is the activation of oxidative phosphorylation rather than inhibition of glycolysis.
+To further elucidate the metabolic consequences of lactate transport inhibition, we next measured proton efflux (PER) and oxygen consumption (OCR) rates in lung fibroblasts treated with TGFβ in combination with MCT inhibitors (**Fig.
+3C, S3C**).
+Given that MCTs co-transport protons with lactate, PER serves as a surrogate measure of lactate efflux.
+In line with previous findings [@schruf2019; @cho2017; @xie2015; @bernard2015; @goodwin2018], we observed increases in both PER and OCR following 48 h of TGFβ stimulation, indicative of increased glycolysis and oxidative phosphorylation.
+Consistent with our direct measures of extracellular lactate and glucose, concurrent treatment with AZD3965 and VB124 was necessary to reduce PER.
+Consequently, this reduction in PER coincided with a corresponding increase in OCR as myofibroblasts transitioned their metabolism from glycolysis to oxidative phosphorylation.
 
-To develop additional support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of extracellular and intracellular metabolites from cells treated with AZD3965 and VB124.
-As suggested by the Seahorse analysis, inhibiting a single lactate transporter had modest effects on extracellular metabolite levels (**Figure 4A-D, Supplemental Figure 4.1**).
-This analysis confirmed the results of extracellular lactate measurements by enzyme assay, which showed dual inhibition was required to decrease lactate efflux (**Figure 4B**).
-In addition to lactate, dual lactate transporter inhibition also had broader impacts on amino acid fluxes (**Figure 4C-D**).
-Several metabolites were differentially regulated by TGFβ treatment (**Supplemental Figure 4.1A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, and ketoleucine.
+Interestingly, individual administration of AZD3965 and VB124 unexpectedly increased OCR without corresponding decreases in PER.
+Indeed, the primary consequence of MCT1 or MCT4 inhibition alone was increased cellular ATP production rates driven by upregulation of oxidative phosphorylation (**Fig.
+3D-E**).
+MCT inhibition decreased spare respiratory capacity, indicating that the basal respiratory rate of treated cells approached their maximal oxidative capacity (**Fig.
+S3D**).
+No significant differences were observed in glycolytic capacity or electron transport chain coupling efficiency (**Fig.
+S3D**).
+Together, these data suggest that the principal metabolic effect of MCT inhibition is the stimulation of oxidative phosphorylation rather than inhibition of glycolysis.
 
-Similar effects on intracellular metabolites were observed.
-Principal components (PC) analysis demonstrated excellent clustering of treatment groups with the drug effect corresponding to PC1 and the TGFβ treatment effect corresponding to PC2.
-Similar to the extracellular flux results, the magnitude of drug-induced perturbations was increased from MCT1 to MCT4 to combined inhibition (**Figure 4E**), as demonstrated by intracellular lactate levels (**Figure 4F**).
-As expected from measures of extracellular lactate following MCT inhibition, intracellular lactate accumulation increases modestly with MCT4 inhibition and markedly with MCT1/4 inhibition.
-TGFβ treatment was associated with diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Supplemental Figure 4.1C-D**).
-The effect of AZD3965 alone on intracellular metabolite levels was modest (**Supplemental Figure 4.1G-H**).
-By contrast, MCT4 inhibition by VB124 alone (**Supplemental Figure 4.1K-L**) or combined with AZD3965 (**Figure 4G-H**) caused substantial perturbations to intracellular metabolism.
+To provide further support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of extracellular and intracellular metabolites from cells treated with AZD3965 and VB124.
+In accordance with the Seahorse analysis, inhibiting a single lactate transporter had modest effects on extracellular metabolite levels (**Fig.
+4A-D, S4**).
+This analysis confirmed the results of extracellular lactate measurements by enzyme assay, which showed that dual inhibition was required to decrease lactate efflux (**Fig.
+4B**).
+Besides lactate, dual inhibition of lactate transporters primarily altered transcellular fluxes of amino acids (**Fig.
+4C-D**), including several metabolites that were differentially regulated by TGFβ treatment (**Fig.
+S4A**) and dual lactate transport inhibition, including leucine, alanine, ornithine, and ketoleucine.
+
+Similar effects were observed on intracellular metabolites.
+PCA showed distinct clustering of treatment groups, with the drug effects aligning with PC1 and TGFβ treatment effects with PC2 (**Fig.
+4E**).
+Similar to the extracellular flux results, the magnitude of drug-induced perturbations was increased from MCT1 to MCT4 to combined inhibition (**Fig.
+4E**), reflected by intracellular lactate levels (**Fig.
+4F**).
+As expected from extracellular lactate measures following MCT inhibition, intracellular lactate accumulation moderately increased with MCT4 inhibition and substantially with MCT1/4 inhibition.
+TGFβ treatment resulted in diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Fig.
+S4C-D**).
+The effect of AZD3965 alone on intracellular metabolite levels was modest (**Fig.
+S4G-H**).
+Conversely, MCT4 inhibition by VB124 alone (**Fig.
+SK-L**) or in combination with AZD3965 (**Fig.
+4G-H**) caused substantial perturbations to intracellular metabolism.
 Specifically, we observed enrichment of the glycolysis and tricarboxylic acid (TCA) cycle metabolite sets with MCT4 inhibition.
-Together, these data indicate that inhibition of lactate export leads to accumulation of upstream glycolytic intermediates that are rerouted to mitochondrial oxidative metabolic pathways.
+These findings collectively suggest that inhibiting lactate export leads to the accumulation of upstream glycolytic intermediates that are rerouted to mitochondrial oxidative metabolic pathways.
 
-To test this hypothesis, we next labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM) in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Figure 4I**).
-TGFβ increases ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), consistent with increased flux from glucose into the TCA cycle.
-While AZD3965 had little impact on these labeling patterns, MCT4 inhibition significantly increased the fractions of these metabolites labeled by glucose, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the bioenergetic and steady-state metabolomics experiments described above.
+To test this hypothesis, we labeled lung fibroblasts with [U-^13^C~6~]-glucose (8 mM) in medium containing lactate (2 mM), glutamine (1 mM), and pyruvate (1 mM) during TGFβ stimulation and treatment with MCT inhibitors (**Fig.
+4J, S5**).
+TGFβ increases ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), indicated enhanced flux from glucose into the TCA cycle.
+While AZD3965 had minimal impact on these labeling patterns, MCT4 inhibition significantly elevated the fractions of these metabolites labeled by glucose, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the bioenergetic and steady-state metabolomics experiments described above.
 
 Increased metabolite labeling by glucose must by offset by decreased labeling from other substrates.
-Since recent data suggest that lactate is a major oxidative fuel source in the lung [@hui2017; @hui2020; @copeland2023], we reasoned that MCT inhibition would decrease oxidation of exogenous lactate.
-To test this, lung fibroblasts were cultured with [U-^13^C~3~]-lactate (2 mM) in medium containing naturally labeled glucose, glutamine, and pyruvate (**Figure 4J**).
-Extracellular [U-^13^C~3~]-lactate labeled ~ 50% of intracellular pyruvate and lactate at baseline with significant downstream incorporation into TCA metabolites.
-This labeling was decreased following TGFβ treatment, mirroring increased fractional labeling from glucose.
-MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Supplemental Figure 4.1M**).
-Interestingly, MCT4 inhibition alone or with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, emphasizing the importance of bi-directional lactate transport mediated by MCTs.
-Similar to our measurements of extracellular lactate, the effects of MCT inhibition were more pronounced when both inhibitors were used simultaneously, again demonstrating some functional redundancy of MCT1 and MCT4 in lung fibroblasts.
+Given recent data suggesting lactate as a major oxidative fuel source in the lung [@hui2017; @hui2020; @copeland2023], we hypothesized that MCT inhibition would decrease exogenous lactate oxidation.
+To test this, lung fibroblasts were cultured with [U-^13^C~3~]-lactate (2 mM) in medium containing naturally labeled glucose, glutamine, and pyruvate (**Fig.
+4K, S5**).
+Extracellular [U-^13^C~3~]-lactate labeled approximately 50% of intracellular pyruvate and lactate at baseline, with significant downstream incorporation into TCA metabolites.
+This labeling decreased following TGFβ treatment, mirroring increased fractional labeling from glucose (**Fig.
+4J**).
+MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Fig.
+S4M, S5**).
+Interestingly, MCT4 inhibition alone or in combination with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, underscoring the importance of bi-directional lactate transport mediated by MCTs.
+Consistent with our measurements of extracellular lactate, the effects of MCT inhibition were more pronounced when both inhibitors were used simultaneously, again highlighting functional redundancy of MCT1 and MCT4 in lung fibroblasts.
 
-## MCT inhibition attenuates pro-fibrotic transcriptional programs
-
-To identify anti-fibrotic transcriptional programs mediated by MCT inhibition, we performed RNA-seq on lung fibroblasts treated with TGFβ in combination with AZD3965 or VB124 (**Figure 5**).
-Principal components analysis (PCA) of the transcriptional changes mirrored the metabolomic findings.
-The first principal component corresponded to TGFβ treatment, accounting for 71% of the overall variance, while MCT4 inhibition, either alone or in combination with MCT1 inhibition aligned with the second principal component (**Figure 5A**).
-Samples treated with the MCT1 inhibitor AZD3965 were similar to vehicle-treated controls.
-Differential expression analysis of TGFβ-treated cells identified the expected increased expression of extracellular matrix proteins (**Supplemental Figure 5A**) and enrichment of the epithelial-to-mesenchymal (EMT) gene set, among others (**Figure 5B**).
-Consistent with the PCA results, only GRIK4 and BRI3 were differentially expressed following AZD3965 (**Supplemental Figure 5B**).
-By contrast, VB124, alone (**Supplemental Figure 5C**) or in combination with AZD3965 (**Figure 5C**), demonstrated more substantial reprogramming of myofibroblast transcription with `r deg_diff` of `r deg_tot` genes (`r deg_frac`%) differentially expressed at a FDR < 0.05.
-
-Co-administration of either AZD3965 or VB124 reversed TGFβ-dependent enrichment of the EMT gene set, which is the Hallmark gene set containing genes related to fibrosis (**Figure 5B**).
-Leading edge analysis of the EMT gene set identified seven genes shared among all three comparisons (*i.e.*, genes increased by TGFβ and decreased by both AZD3965 and VB124) (**Figure 5D**).
-These genes include: biglycan (BGN), COL6A3, Frizzled 8 (FZD8), matrix Gla protein (MGP), Prostate Transmembrane Protein, Androgen Induced 1 (PMEPA1), TIMP metallopeptidase inhibitor 1 (TIMP1), and tenascin-C (TNC).
-Many of these genes contribute to the pathobiology of pulmonary fibrosis or serve as biomarkers of disease or treatment response [@kwapiszewska2018; @williams2020; @baarsma2017; @todd2020; @bhattacharyya2016].
-Together, these data suggest lactate transport inhibitors attenuate the pro-fibrotic transcriptional program in TGFβ-treated LFs.
-
-## Lactate transport inhibition may signal through histone lactylation
-
-We next sought to identify the signaling mechanism(s) linking lactate transport inhibition to anti-fibrotic transcriptional programs.
-Since increasing intracellular lactate was associated with more potent inhibition of α-SMA expression, we first treated cells with TGFβ in combination with extracellular lactate (10 mM).
-A prior study demonstrated that extracellular lactate modestly increased α-SMA expression in the absence of TGFβ [@kottmann2012].
-We found that extracellular lactate did not impact α-SMA expression in TGFβ-treated LFs (**Figure 6A-B**), suggesting that accumulation of intracellular lactate alone does not mediate the anti-fibrotic effects of MCT inhibition.
-
-Prior work has also linked inhibition of glycolysis and lactate production to hypoxia-inducible factor 1α activation (HIF-1α), which contributes to myofibroblast differentiation [@kottmann2012; @goodwin2018].
-We found maximal HIF-1α activation 6 h following TGFβ stimulation.
-At this time point, HIF-1α protein levels were similar across vehicle-, AZD3965-, and VB124-treated cells (**Figure 6C-D**).
-In LFs treated with dual inhibitors, HIF-1α protein was increased compared to vehicle, consistent with enrichment of the "Hypoxia" gene set in our RNA-seq data (**Figure 5B**).
-These findings, in light of prior studies, indicate that lactate transport inhibition acts downstream of HIF-1α-dependent transcriptional programs to inhibit myofibroblast differentiation.
-
-Next, we examined TGFβ-dependent signaling.
-TGFβ activates both SMAD and non-SMAD signaling pathways through a cascade of protein phosphorylation events.
-MCT inhibitors did not decrease Smad3 or ERK phosphorylation after 48 h of TGFβ stimulation (**Figure 6E-H**).
-
-Post-translational protein modification of histone lysine residues by lactylation has been identified as a novel mechanism of metabolic regulation of gene expression [@zhang2019a].
-Given the increase in intracellular lactate following MCT4 inhibition, we measured H3 histone lactylation by immunoblot (**Figure 6I-J**).
-We found that inhibition of both MCT1 and MCT4 increased histone lactylation and may account for some of the transcriptional changes observed following these treatments.
-
-## Lactate transport inhibition contributes to antioxidant defense
+## Lactate transport inhibition contributes to antioxidant defense mechanisms
 
 Lactate metabolism is closely coupled to cellular redox homeostasis through its metabolism by lactate dehydrogenases, which transfer electrons from lactate to NADH.
-Since prior studies have suggested that reactive oxygen species (ROS) are important for TGFβ-mediated gene expression [@jain2013; @schworer2020], we next examined the impact of lactate transport inhibition on cellular redox homeostasis.
-Consistent with an increase in intracellular lactate with MCT4 inhibition, we observed a concomitant increase in intracellular NADH/NAD^+^ (**Figure 7A**), which was associated with a reduction in total ROS as measured by the CellROX fluorescent probe (**Figure 7B**).
-AZD3965, but not VB124, also decreased mitochondrial superoxide production (**Supplemental Figure 7B**).
-By contrast, the NADPH/NADP^+^ was unchanged by either TGFβ or MCT inhibitors (**Supplemental Figure 7A**).
-Notably, we did not observe TGFβ-dependent increases in total ROS production in our model system.
+Building on previous findings highlighting the significance of ROS in TGFβ-mediated gene expression [@jain2013; @schworer2020], we investigated the impact of lactate transport inhibition on cellular redox balance.
+Consistent with increased intracellular lactate upon MCT4 inhibition, we observed a corresponding rise in intracellular NADH/NAD^+^ (**Fig.
+4L**), coupled with a reduction in total ROS as measured using the CellROX fluorescent probe (**Fig.
+4M**).
+AZD3965, but not VB124, decreased mitochondrial superoxide production (**Fig.
+4N**).
+Neither inhibitor affected mitochondrial biomass measured by MitoTracker fluorescence (**Fig.
+S6B**).
+Less substantial changes were observed in the NADPH/NADP^+^ ratio, where VB124 decreased NADPH/NADP^+^ in TGFβ-treated cells (**Fig.
+S6A**).
+Notably, TGFβ did not induce ROS production in our experimental system.
 
-Proline biosynthesis has been previously suggested as a mechanism by which NIH-3T3 fibroblasts prevent excessive ROS accumulation following TGFβ stimulation [@schworer2020].
+Previous research has proposed proline biosynthesis as a mechanism for NIH-3T3 fibroblasts to mitigate ROS accumulation following TGFβ stimulation [@schworer2020].
 In this model, proline synthesis from glutamine consumes reducing equivalents from NADPH and NADH, thereby ameliorating reductive stress and decreasing ROS production.
-Compared to this prior study, we did not find that TGFβ increased proline in primary lung fibroblasts (**Supplemental Figure 7C**).
-Overall, the fractional labeling of proline from [U-^13^C~5~]-glutamine was modest at 10% compared to the 40% previously reported (**Supplemental Figure 7D**).
-While we observed trends toward increased proline production from glutamine with MCT4 inhibition, this mechanism seems unlikely to account for the effects of MCT inhibitors on ROS generation.
+Contrary to these findings, we did not observe TGFβ-induced proline elevations in primary lung fibroblasts (**Fig.
+S6C**).
+Overall, the fractional labeling of proline from [U-^13^C~5~]-glutamine was modest at 10% compared to the previously reported 40% (**Fig.
+S6D**).
+While we noted trends toward increased proline production from glutamine with MCT4 inhibition, this mechanism appears insufficient to explain the effects of MCT inhibitors on ROS generation.
+
+## Lactate transport inhibition does not alter classical TGFβ signaling pathways
+
+Next, we examined TGFβ-dependent signaling, which activates both SMAD and non-SMAD signaling pathways through a cascade of protein phosphorylation events.
+Following 48 h of TGFβ stimulation, MCT inhibitors did not diminish Smad3 or ERK phosphorylation (**Fig.
+5A-B**).
+
+Given that increased intracellular lactate correlated with more potent inhibition of α-SMA expression, we treated cells with TGFβ in combination with extracellular lactate (10 mM).
+A previous study suggested that extracellular lactate modestly increased α-SMA expression independently of TGFβ [@kottmann2012].
+However, we observed no impact of extracellular lactate on α-SMA expression in TGFβ-treated lung fibroblasts (**Fig.
+S7A**), indicating intracellular lactate accumulation alone does not mediate the antifibrotic effects of MCT inhibition.
+
+Prior research has linked inhibition of glycolysis and lactate production to hypoxia-inducible factor 1α activation (HIF-1α), contributing to myofibroblast differentiation [@kottmann2012; @goodwin2018].
+We observed maximal HIF-1α activation 6 h after TGFβ stimulation.
+At this time point, HIF-1α protein levels remained consistent across vehicle-, AZD3965-, and VB124-treated cells (**Fig.
+S7B**).
+However, in lung fibroblasts treated with dual inhibitors, HIF-1α protein  increased compared to vehicle, consistent with the enrichment of the "Hypoxia" gene set in our RNA-seq analysis (**Fig.
+2F**).
+These findings, alongside previous studies, suggest that lactate transport inhibition acts downstream of HIF-1α-dependent transcriptional programs to inhibit myofibroblast differentiation.
 
@@
 
-Based on our *in vitro* data suggesting an anti-fibrotic effect of lactate transporter inhibition, we next tested the efficacy of MCT inhibitors in the bleomycin-induced mouse model of pulmonary fibrosis.
+Building on our *in vitro* findings indicating an antifibrotic effect of lactate transporter inhibition, we proceeded to evaluate the efficacy of MCT inhibitors in a bleomycin-induced mouse model of pulmonary fibrosis.
 Mice received 1.2 U/kg bleomycin by intratracheal administration.
-After seven days, the animals began treatment with AZD3965 (100 mg/kg twice daily) or VB124 (30 mg/kg once daily) or vehicle by oral gavage (**Figure 8A**).
-We anticipated that dual inhibition would likely be toxic *in vivo* and did not assess the combination of inhibitors.
-Compared to vehicle, mice treated with VB124 had increased weight recovery 21 days after bleomycin administration (**Supplemental Figure 8**).
-Lung mechanics were improved by ~ 50% of baseline compared to vehicle controls following 14 days of MCT inhibitor treatment (**Figure 8B-C**).
-Consistent with improved elastance and compliance measures, histologic fibrosis severity assessed by Ashcroft scoring [@ashcroft1988] also improved.
+Seven days later, the animals began treatment with AZD3965 (100 mg/kg twice daily) or VB124 (30 mg/kg once daily) or vehicle by oral gavage (**Fig.
+6A**).
+Considering likely toxicity, we did not assess the combination of inhibitors.
+Compared to vehicle, mice treated with VB124 had increased weight recovery 21 days after bleomycin administration (**Fig.
+S8**).
+Lung mechanics improved by approximately 50% of baseline compared to vehicle controls following 14 days of MCT inhibitor treatment (**Fig.
+6B-C**).
+This improvement in measures of lung stiffness was further supported by histologic assessment of fibrosis severity using Ashcroft scoring [@ashcroft1988] (**Fig.
+6D-E**) and hydroxyproline content measurement (**Fig.
+6F**).
+Taken together, these data demonstrate substantially decreased pulmonary fibrosis severity after 14 days of treatment with lactate transport inhibitors.
 
@@
 
-To investigate metabolic perturbations following lactate transporter inhibition *in vivo*, we next performed metabolomic profiling of lung and plasma from mice treated with bleomycin and MCT inhibitors (**Supplemental Figure 9.1**).
-While bleomycin treatment substantially altered the lung metabolic profile, treatment with AZD3965 or VB124 had little additional impact on total metabolite levels.
-Unlike bleomycin alone or following AZD3965 treatment, VB124 treatment was associated with significant changes in several circulating metabolites (**Supplemental Figure 9.1I**).
-Among these, VB124 significantly increased circulating lactate, while AZD3965 was associated with decreased circulating lactate (**Figure 9A**).
-This pattern was also reflected in the lung, where only modest changes were observed following bleomycin, contrary to prior reports [@kottmann2012].
-Given the impact of our treatments on circulating lactate, we calculated the ratio of lung-to-plasma lactate.
-This analysis demonstrated the expected increase in lactate in fibrotic lungs with levels returning to baseline following the addition of VB124 (**Figure 9A**).
+To explore metabolic changes following lactate transporter inhibition *in vivo*, we conducted metabolomic profiling of lung and plasma from mice treated with bleomycin and MCT inhibitors (**Fig.
+S9**).
+While bleomycin administration substantially altered the lung metabolic profile, treatment with AZD3965 or VB124 had little additional impact on total metabolite levels.
+However, unlike bleomycin alone or following AZD3965 treatment, VB124 treatment resulted in significant alterations in several circulating metabolites (**Fig.
+S9I**).
+Among these, VB124 significantly increased circulating lactate, while AZD3965 led to decreased circulating lactate (**Fig.
+7A**).
+This trend was also reflected in the lung, where only modest changes were noted following bleomycin, contrary to prior reports [@kottmann2012].
+Considering the influence of our treatments on circulating lactate, we calculated the lung-to-plasma lactate ratio.
+This analysis revealed the anticipated increase in lactate in fibrotic lungs, with levels returning to baseline following the addition of VB124 (**Fig.
+7A**).
 
-We were surprised to observe a relative paucity of metabolic changes in whole lung lysates following MCT inhibition compared to our *in vitro* studies.
-We hypothesized that this might be a consequence of measuring steady-state metabolite levels that do not account for differences in metabolic flux.
-To examine this hypothesis, we performed  multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following administration of ^15^N-proline and ^2^H-glucose or ^13^C-glucose.
+We were surprised to observe relatively few metabolic changes in whole lung lysates following MCT inhibition compared to our *in vitro* findings.
+We speculated that this discrepancy might arise from measuring steady-state metabolite levels that do not reflect differences in metabolic flux.
+To test this hypothesis, we performed multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following administration of ^15^N-proline and either ^2^H-glucose or ^13^C-glucose.
 MIMS enables the quantification of stable isotope tracer flux into tissue biomass with subcellular spatial resolution [@steinhauser2012; @gyngard2019].
-For three days prior to tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a fibrosis tracer and 50 mg ^2^H-glucose or ^13^C-glucose as a metabolic tracer.
-Lung tissue sections were subsequently imaged by nanoscale secondary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B-C**).
-This labeling strategy was quite effective, demonstrating ~1.5-fold ^15^N, ~4.6-fold for ^2^H, and ~2.4-fold for ^13^C enrichment above natural isotope abundance.
-Pulmonary fibrosis was associated with a significant increase in ^15^N labeling from proline, consistent with increased collagen synthesis and deposition during the labeling period.
-Similar to the lung function and histological analyses, less ^15^N incorporation per tissue area was observed in drug-treated animals.
-Glucose labeling followed a similar pattern where lactate transporter inhibition caused a decrease in glucose incorporation into tissue biomass.
-Importantly, these  findings are not merely a consequence of decreased tissue fibrosis as we chose more fibrotic areas to image (**Figure 9B**), and the enrichment values are normalized to tissue area.
+For three days preceding tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a fibrosis tracer and 50 mg glucose isotope as a metabolic tracer.
+Subsequently, lung tissue sections were imaged by nanoscale secondary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Fig.
+7B-C**).
+This labeling approach proved effective, demonstrating approximately 1.5-fold enrichment ^15^N, 4.6-fold for ^2^H, and 2.4-fold for ^13^C above natural isotope abundance.
+Pulmonary fibrosis correlated with a significant increase in ^15^N labeling from proline, consistent with increased collagen synthesis and deposition during the labeling period.
+Similarly, drug-treated animals exhibited less ^15^N incorporation per tissue area, consistent with lung function and histological analyses.
+Glucose labeling displayed a similar pattern, wherein lactate transporter inhibition led to reduced glucose incorporation into tissue biomass.
+Importantly, these findings are not merely attributable to decreased tissue fibrosis, as we selectively imaged more fibrotic areas (**Fig.
+7B**) and normalized the enrichment values to tissue area.
 
-## Novel MCT4 inhibitor mitigates experimental pulmonary fibrosis
+## VB253, a novel MCT4 inhibitor, alleviates experimental pulmonary fibrosis
 
-Our preclinical data suggest that MCT4 inhibition, as a single target therapy, is more potently antifibrotic than MCT1 inhibition, both in terms of inhibiting myofibroblast differentiation *in vitro* and decreasing bleomycin-induced fibrosis.
+Our preclinical findings indicate that MCT4 inhibition, as a single therapeutic target, exhibits greater antifibrotic efficacy than MCT1 inhibition, both in suppressing myofibroblast differentiation *in vitro* and reducing bleomycin-induced fibrosis.
 VB253 is a novel inhibitor of MCT4 with similar selectivity, but ~10-fold increased potency, for MCT4 inhibition.
-Like VB124, VB253 dose-dependently decreases TGFβ-stimulated α-SMA in human IPF lung fibroblasts *ex vivo* (**Figure 10A**).
-Given the robust effect of MCT4 inhibition on myofibroblast differentiation, we compared VB253 to nintedanib, an anti-fibrotic medication clinically indicated for pulmonary fibrosis (**Figure 10A-B**).
-Nintedanib inhibits TGFβ-mediated fibroblast to myofibroblast differentiation and decreases Col1a1 expression *in vitro* and *in vivo*  [@wollin2014].
-While both drugs potently attenuated α-SMA expression, nintedanib demonstrated a moderate cytotoxic effect that was not observed with VB253 (**Figure 10B**).
-Nintedanib prevents myofibroblast differentiation, in part, through inhibition of TGFβ receptor phosphorylation and activation of Smad-dependent signaling pathways [@rangarajan2016].
-Consistent with these findings, nintedanib dose-dependently attenuates TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 have no impact on this signaling pathway (**Figure 6**, **Supplemental Figure 10**).
+Like VB124, VB253 dose-dependently decreases TGFβ-stimulated α-SMA in human IPF lung fibroblasts *ex vivo* (**Fig.
+8A**).
+To assess its efficacy compared to established therapies, we compared VB253 to nintedanib, a clinically approved antifibrotic medication for pulmonary fibrosis (**Fig.
+8A-B**).
+Nintedanib inhibits TGFβ-mediated myofibroblast differentiation and decreases Col1a1 expression *in vitro* and *in vivo*  [@wollin2014].
+While both compounds effectively attenuated α-SMA expression, nintedanib exhibited moderate cytotoxicity absent with VB253 (**Fig.
+8B**).
+Nintedanib mitigates myofibroblast differentiation partly by inhibiting TGFβ receptor phosphorylation and Smad-dependent signaling pathways [@rangarajan2016].
+Correspondingly, nintedanib dose-dependently attenuated TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 have no impact on this signaling pathway (**Fig.
+5, S10**).
 
-We next assessed the ability of VB253 to inhibit bleomycin-induced pulmonary fibrosis.
-In experiments conducted independently from those reported above, VB253 was administered to young mice (8-10 weeks) beginning on day 7 following bleomycin (**Figure 10C**).
-Mice treated with nintedanib or pirfenidone provided comparisons to currently available therapies for pulmonary fibrosis.
-Three weeks after bleomycin administration, whole-body plethysmography was performed to assess breathing patterns in unrestrained mice.
+Subsequently, we evaluated the potential of VB253 to counter bleomycin-induced pulmonary fibrosis.
+In experiments conducted independently from those reported above, VB253 was administered to young mice (8-10 weeks) beginning on day 7 following bleomycin (**Fig.
+8C**).
+Nintedanib and pirfenidone served as comparators to existing pulmonary fibrosis therapies.
+Three weeks after bleomycin, whole-body plethysmography was performed to assess breathing patterns in unrestrained mice.
 Enhanced pause (Penh) is a dimensionless index describing airflow during tidal breathing found to increase significantly in bleomycin-treated mice [@vanoirbeek2010].
-VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterns in bleomycin-treated mice (**Figure 10D**).
-This was associated with decreased pulmonary fibrosis by histologic assessment (**Figure 10E**) and α-SMA expression (**Figure 10F-G**).
-Moreover, the effects of VB253 compared favorably to the antifibrotic effects of nintedanib and pirfenidone.
+VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterns in bleomycin-treated mice (**Fig.
+8D**).
+Histologic assessment confirmed reduced pulmonary fibrosis (**Fig.
+8E**) and α-SMA expression (**Fig.
+8F**).
+Moreover, the effects of VB253 were comparable to the antifibrotic effects of nintedanib and pirfenidone with decreased cytotoxicity in *in vitro* assays.
 
-Compared to young mice, bleomycin-induced fibrosis in aged mice is more severe and does not resolve spontaneously [@redente2011].
-Given that IPF incidence increases with age, some studies have suggested such aged mouse models may be more clinically relevant.
-Thus, we examined the antifibrotic effects of VB253 in aged (60+ weeks) mice.
-Similar to young mice, VB253 decreased fibrosis severity as quantified by Ashcroft score and α-SMA expression (**Figure 10H-J**).
-The magnitude of improvement was similar to that observed with nintedanib and pirfenidone.
-As anticipated, VB253 decreased total lung lactate similar to VB124 (**Figure 9A**).
-Taken together, these data provide compelling preclinical data for the development of lactate transporter inhibition as a therapeutic strategy for fibrotic lung disease.
+Compared to young mice, aged mice exhibit more severe and persistent bleomycin-induced pulmonary fibrosis [@redente2011].
+Given the clinical relevance of age-related IPF incidence, we investigated the antifibrotic effects of VB253 in aged (60+ weeks) mice.
+Similar to young mice, VB253 decreased fibrosis severity, as quantified by Ashcroft score and α-SMA expression (**Fig.
+8G-I**).
+The magnitude of improvement paralleled that of nintedanib and pirfenidone.
+As expected, VB253 decreased total lung lactate akin to VB124 (**Fig.
+7A**).
+Collectively, these findings provide compelling preclinical evidence supporting lactate transporter inhibition as a novel therapeutic strategy for fibrotic lung disease.
 
@@
 
-Our findings demonstrate that lactate transport is an essential component of the metabolic reprogramming associated with myofibroblast differentiation *in vitro* and *in vivo*.
-We observed increased expression of the lactate transporters MCT1 and MCT4 in IPF lung explants and experimental model systems.
-Inhibition of these transporters decreased TGFβ-stimulated myofibroblast differentiation and the severity of bleomycin-induced pulmonary fibrosis.
-Metabolically, MCT antagonists promoted glucose oxidation and decreased glucose carbon incorporation into fibrotic lung regions, which were associated with decreased oxidative stress.
-Beyond metabolic reprogramming, we observed changes in gene expression profiles consistent with antifibrotic metabolic signaling downstream of lactate transport inhibition, which may be mediated by changes in ROS production, protein lactylation, or other mechanisms.
-MCT4 inhibition consistently proved more potently antifibrotic than MCT1 inhibition, and we report a novel MCT4 inhibitor, VB253, with a more favorable pharmacologic profile than VB124 that is now in Phase 1 clinical trials [**REF?**].
-Collectively, these data establish lactate transport as a metabolic opportunity for therapeutic intervention in pulmonary fibrosis.
+Our findings identify the pivotal role of lactate transport in the metabolic reprogramming associated with myofibroblast differentiation both *in vitro* and *in vivo*.
+Elevated expression of the lactate transporters MCT1 and MCT4 was observed in IPF lung explants and experimental models, underscoring their significance.
+Inhibiting these transporters mitigated TGFβ-stimulated myofibroblast differentiation and attenuated the severity of bleomycin-induced pulmonary fibrosis without a demonstrable impact on classical TGFβ receptor signaling pathways.
+Metabolically, MCT antagonists promoted glucose oxidation and while reducing glucose carbon incorporation into fibrotic lung regions, correlating with decreased oxidative stress.
+MCT4 inhibition consistently exhibited superior antifibrotic potency compared to MCT1 inhibition.
+We introduce a novel MCT4 inhibitor, VB253, with a more favorable pharmacologic profile than VB124 currently undergoing Phase 1 clinical trials [**REF?**].
+Altogether, these data establish lactate transport as a promising metabolic target for therapeutic intervention in pulmonary fibrosis.
 
-Metabolic reprogramming is a hallmark of myofibroblast differentiation.
-Previous studies have identified alterations in carbohydrate, amino acid, and lipid metabolic pathways that support fibrogenesis [@selvarajah2021].
-Of these, increased lactate production was among the earliest recognized metabolic changes associated with pulmonary fibrosis [@kottmann2012].
-Subsequent work demonstrated increased lactate production was driven by activation of glycolysis in myofibroblasts [@xie2015] and that inhibition of glycolysis prevented myofibroblast differentiation [@xie2015; @kottmann2015; @bernard2015] and, in turn, experimental pulmonary fibrosis [@judge2018; @cho2017].
-Consistent with increased glycolysis in myofibroblasts, we have now demonstrated increased expression of the lactate transporters MCT1 and MCT4 support this metabolic phenotype.
+Metabolic reprogramming characterizes myofibroblast differentiation, with previous studies revealing alterations in carbohydrate, amino acid, and lipid metabolic pathways that promote fibrogenesis [@selvarajah2021].
+Of these, increased lactate production emerged as an early metabolic hallmark associated with pulmonary fibrosis [@kottmann2012].
+Subsequent investigations identified glycolysis activation as the driving force behind increased lactate production by myofibroblasts [@xie2015].
+Furthermore, inhibition of glycolysis not only prevented myofibroblast differentiation [@xie2015; @kottmann2015; @bernard2015], but also attenuated experimental pulmonary fibrosis [@judge2018; @cho2017].
+Our findings align with this metabolic shift, as we observed upregulated expression of lactate transporters MCT1 and MCT4, supporting this glycolytic phenotype in myofibroblasts.
 
 MCTs are proton-coupled monocarboxylate symporters with varying affinities for lactate, pyruvate, and other monocarboxylates [@halestrap2013].
-MCT1 is constitutively and ubiquitously expressed and primarily considered to be a lactate importer with a *K*~M~ for lactate of 3-6 mM.
-By contrast, MCT4 expression is dynamically regulated, including by HIF-1α, and has primarily been considered to be a lactate exporter with a low lactate affinity (*K*~M~ 30-40 mM).
-Recent data suggests, however, that MCT4 has a much higher affinity for lactate import than previously appreciated (*K*~M~ 1 mM) [@contreras-baeza2019].
-These data are most consistent with our findings that inhibition of both MCT1 and MCT4 was required to inhibit lactate export (**Figure 3**) and isotope incorporation from extracellular [U-^13^C~3~]-lactate import (**Figure 4J**).
-Previous studies of cancer cells have similarly recognized that MCT4-expressing cells are resistant to the cytotoxic effects of MCT1 inhibition [@curtis2017].
-Interestingly, inhibition of either MCT1 or MCT4 reduced myofibroblast differentiation without affecting lactate export.
+MCT1 is constitutively and ubiquitously expressed and primarily considered to be a lactate importer with a lactate affinity (*K*~M~) ranging from 3-6 mM.
+By contrast, MCT4 expression is dynamically regulated, including by HIF-1α, and traditionally considered a lactate exporter with a lower lactate affinity (*K*~M~ 30-40 mM).
+However, recent evidence suggests that MCT4 has a much higher affinity for lactate import than previously appreciated (*K*~M~ 1 mM) [@contreras-baeza2019].
+These findings align with our results where inhibiting both MCT1 and MCT4 was required to block lactate export (**Fig.
+3, S3**) and isotope incorporation from extracellular [U-^13^C~3~]-lactate import (**Fig.
+4K**).
+Previous studies in cancer cells have also noted that MCT4-expressing cells are resistant to the cytotoxic effects of MCT1 inhibition [@curtis2017].
+Interestingly, inhibiting either MCT1 or MCT4 decreased myofibroblast differentiation without affecting lactate export, adding further complexity to their roles in cellular metabolism.
 
-To clarify the metabolic consequences of lactate transporter inhibition, we performed a comprehensive metabolic analysis of myofibroblasts treated with MCT inhibitors, including bioenergetic measurements, metabolomic profiling, and stable isotope tracing experiments.
-The primary metabolic consequence of MCT inhibition is activation of oxidative phosphorylation.
-MCT inhibitors increased the fraction of mitochondrial ATP production, which was associated with increased levels of tricarboxylic acid cycle intermediates, and increased isotope incorporation from glucose substrates.
-Collectively, these data indicate that MCT inhibition reroutes carbon flow from glucose away from lactate fermentation and toward glucose oxidation.
-Several mechanisms may account for this phenomenon.
-Since the magnitude of these effects generally correlates with intracellular lactate levels, mass action may be an important driver of this metabolic phenomenon.
-Lactate accumulation is also closely coupled in NADH production through lactate dehydrogenase activity.
+To clarify the metabolic consequences of lactate transporter inhibition, we performed a comprehensive metabolic analysis of myofibroblasts treated with MCT inhibitors, encompassing bioenergetic measurements, metabolomic profiling, and stable isotope tracing.
+The primary metabolic consequence of MCT inhibition is stimulation of oxidative phosphorylation.
+MCT inhibitors increased the fraction of mitochondrial ATP production, which was associated with increased levels of TCA intermediates and increased isotope incorporation from glucose.
+Collectively, these data suggest that MCT inhibition redirects glucose carbon flux away from lactate fermentation and toward glucose oxidation.
+Several mechanisms could underlie this shift.
+Given that the observed effects generally scale with intracellular lactate levels, mass action likely plays an important role in driving this metabolic phenomenon.
+Additionally, lactate accumulation is closely coupled in NADH production through lactate dehydrogenase activity.
 Cytoplasmic NADH may be transported into the mitochondria through the malate-aspartate shuttle and oxidized by the electron transport chain.
-Recently, lactate itself was shown activate the mitochondrial electron transport chain independently of its metabolism [@cai2023], although the molecular mechanism for this effect is unknown.
+Recently, lactate itself was shown activate the mitochondrial electron transport chain independent of its metabolism [@cai2023], though the molecular mechanism remains unknown.
 
-How these metabolic effects signal to antifibrotic transcriptional programs remains unclear.
-We investigated several hypotheses based on previous studies.
-First, we did not observe significant changes in canonical (*i.e.*, Smad3) or non-canonical (*i.e.*, ERK) TGFβ signaling pathways.
-Second, TGFβ-dependent activation of HIF-1α seems to be required for myofibroblast differentiation [@goodwin2018].
-MCT1 and MCT4 potentiated HIF-1α stabilization following TGFβ (**Figure 6D**), indicating that MCT inhibition acts downstream of HIF signaling pathways.
-Third, post-translational modification of histone lysines by lactate serves as an epigenetic modification that stimulates gene expression [@zhang2019a].
-We observed increased histone lactylation following dual MCT inhibition, which may account for some of the transcriptional changes we observed.
-In addition to histones, protein lactylation has been increasingly identified as a critical regulator of protein function, including cytoskeletal proteins [@yang2023].
-Fourth, increased ROS production has been observed following TGFβ treatment [@jain2013; @schworer2020].
+Our data reveal that lactate transport inhibition exploits novel antifibrotic mechanisms distinct from TGFβ signaling pathways and suggest that these metabolic effects signal to antifibrotic transcriptional programs by dampening ROS levels.
+Increased ROS production has been observed following TGFβ treatment and seems to be essential for myofibroblast differentiation [@jain2013; @schworer2020].
 While we were unable to replicate TGFβ-dependent increases in ROS in our cells and model system, we did observe decreased CellROX oxidation with MCT4 inhibition and decreased MitoSOX oxidation with AZD3965.
-Collectively, these studies identify two putative mechanisms for antifibrotic lactate signaling following MCT inhibition, protein lactylation and antioxidant activity; however, novel lactate targets are being increasingly recognized as important mediators of metabolic signaling [@liu2023a; @liu2023b].
+Additional signaling mechanisms may also contribute, such as post-translational modification of protein lysines by lactate (*i.e.*, protein lactylation).
+Histone lactylation serves as an epigenetic modification that stimulates gene expression [@zhang2019a] and this modification has been increasingly identified as a critical regulator of protein function, including cytoskeletal proteins [@yang2023].
+Furthermore, ongoing research is uncovering novel lactate targets and mechanisms as significant mediators of metabolic signaling [@liu2023a; @liu2023b].
 
-Through our detailed metabolic investigation of lactate transport inhibition, we have also generated valuable data on the metabolic and transcriptional consequences of human lung myofibroblast differentiation.
-TGFβ stimulated marked changes in amino acid and nucleic acid metabolic pathways (**Supplemental Figure 4.1A-D**).
-Previous studies have identified important roles for glutamine, proline, and taurine [@selvarajah2021; @choudhury2023; @phan1995], while the contributions of branched chain amino acid metabolism, for example, remain unexplored.
-Integrating these multi-omics data sets may identify novel molecular targets for future drug development.
+Through our comprehensive metabolic investigation of lactate transport inhibition, we have also generated valuable data on the metabolic and transcriptional consequences of human lung myofibroblast differentiation.
+TGFβ stimulation induced notable alterations in amino acid and nucleic acid metabolic pathways (**Fig.
+S4**).
+While previous studies have underscored the significance of glutamine, proline, and taurine [@selvarajah2021; @choudhury2023; @phan1995], the role of branched chain amino acid metabolism, for instance, remains unexplored.
+Integrating these multi-omics data sets could unveil novel molecular targets for future drug development.
 
-In addition to metabolomic profiling, we have performed stable isotope tracing using glucose, lactate, and glutamine substrates in TGFβ-treated primary human lung fibroblasts.
-To our knowledge, these are the first comprehensive data on intracellular substrate metabolism in this commonly utilized model of myofibroblast differentiation.
-Overall, isotope labeling patterns changed little following TGFβ treatment (**Supplemental Figure 4.2**).
-Together with the results of our extracellular flux experiments, this finding suggests that TGFβ primarily increases the flow of metabolites through metabolic pathways without substantially altering the pathways themselves.
-Consistent with increased glucose uptake and accumulation of glycolytic metabolites, a greater fraction of pyruvate, lactate, alanine, and serine are labeled by [U-^13^C~6~]-glucose, which is offset by a decrease in the fractional labeling from [U-^13^C~3~]-lactate.
-Similar to our previous work [@copeland2023], nearly 50% of tricarboxylic acid metabolites (citrate, 2-oxoglutarate, succinate, and malate) are labeled by [U-^13^C~3~]-lactate, highlighting the importance of lactate as a respiratory fuel source in these cells.
+In addition to metabolomic profiling, we conducted stable isotope tracing using glucose, lactate, and glutamine substrates in TGFβ-treated primary human lung fibroblasts.
+To our knowledge, this represents the first comprehensive dataset on intracellular substrate metabolism in this widely employed model of myofibroblast differentiation.
+Overall, isotope labeling patterns changed little following TGFβ treatment (**Fig.
+S5**).
+This observation, coupled with the results of our extracellular flux experiments, suggests that TGFβ primarily enhances metabolite flow through metabolic pathways without substantially altering the pathways themselves.
+Consistent with increased glucose uptake and accumulation of glycolytic metabolites, a greater fraction of pyruvate, lactate, alanine, and serine are labeled by [U-^13^C~6~]-glucose, countered by a reduction in the fractional labeling from [U-^13^C~3~]-lactate.
+Consistent with our prior findings [@copeland2023], nearly 50% of TCA metabolites (citrate, 2-oxoglutarate, succinate, and malate) are labeled by [U-^13^C~3~]-lactate, highlighting the importance of lactate as a respiratory fuel source in these cells.
 
-In addition to cellular bioenergetics, the contribution of fibroblast metabolic pathways to extracellular matrix production has been an intense area of investigation [@selvarajah2021].
-Serine and glycine may be synthesized from the glycolytic intermediate 3-phosphoglycerate, and proline may be synthesized from glutamine.
-These pathways have been implicated in myofibroblast differentiation and pulmonary fibrosis [@bernard2018; @hamanaka2019; @selvarajah2019; @schworer2020; @choudhury2023].
-Interestingly, our data suggest little incorporation of glucose or lactate carbon into serine and little incorporation of glutamine carbon into proline.
-These differences may be secondary to the cells studied (IMR-90 fetal lung fibroblasts or NIH-3T3 spontaneously immortalized mouse embryonic fibroblasts) or the culture medium utilized (Eagle's minimum essential medium, which lacks serine, glycine, and proline).
-In this context, our study highlights that fibroblasts will preferentially use available amino acids rather than reroute substrates into biosynthetic pathways.
-Developing metabolic flux models using human physiologic medium [@cantor2017] may provide a more accurate and complete characterization of substrate flow into energetic and biosynthetic pathways that more closely reflects fibroblast metabolism *in vivo*.
+Recently, the contributions of fibroblast metabolic pathways to extracellular matrix production has garnered significant attention [@selvarajah2021].
+Serine and glycine synthesis from the glycolytic intermediate 3-phosphoglycerate, and proline synthesis from glutamine, have been implicated in myofibroblast differentiation and pulmonary fibrosis [@bernard2018; @hamanaka2019; @selvarajah2019; @schworer2020; @choudhury2023].
+Intriguingly, our data indicate limited incorporation of glucose or lactate carbon into serine and minimal glutamine carbon incorporation into proline.
+These differences may stem from differences in the cell types (IMR-90 fetal lung fibroblasts or NIH-3T3 spontaneously immortalized mouse embryonic fibroblasts) or the culture medium composition (Eagle's minimum essential medium, which lacks serine, glycine, and proline).
+Our study suggests that fibroblasts preferentially utilize available amino acids over rerouting substrates into biosynthetic pathways.
+Developing metabolic flux models using human physiologic medium [@cantor2017] could provide a more accurate and comprehensive understanding of substrate flow into energetic and biosynthetic pathways, better mimicking fibroblast metabolism *in vivo*.
 
-Even better than the use of *in vitro* model systems are approaches that enable the study of cell metabolism *in vivo*.
-Here, we performed metabolomic profiling of lung and plasma samples from mice treated with bleomycin and MCT inhibitors (**Supplemental Figure 9**).
-These studies did not reveal consistent changes in whole lung metabolite levels and demonstrate a potential limitation of using bulk metabolomics to monitor metabolic changes in heterogeneous cell populations *in vivo*.
-We sought to overcome this limitation with spatial metabolic imaging.
-Mice received stable isotope tracers of proline and glucose prior to euthanasia.
-Since these tracers are administered to live animals, multi-isotope imaging mass spectrometry identifies lung regions that are metabolically active over the labeling period.
-Since tissues are fixed and processed *ex vivo*, the isotope signals indicate substrate incorporation into fixable biomass [@zhang2020; @wertheim2023].
-Thus, using multi-isotope imaging mass spectrometry, we are directly quantifying isotope flux from glucose into fibrotic lung regions.
-We found that ^15^N-proline enrichment levels correlate with tissue fibrosis from histologic and biochemical assessments.
-This finding emphasizes the ability of cells to incorporate circulating proline for protein synthesis, and likely de-emphasizes the importance of *de novo* proline biosynthesis *in vivo*.
-The ^2^H-glucose signal provides some direct evidence for MCT-dependent metabolic reprogramming in mice, with less carbohydrate incorporation into matrix proteins.
-Owing to the financial and time costs of isotope tracing and imaging, only a few animals per group were studied.
-Nonetheless, metabolic imaging at high spatial resolution offers tremendous promise for correlating cell identities determined from spatial transcriptomic profiles with metabolic features.
-Stable isotopes may be safely administered to human patients prior lung biopsy or explant, an approach that would considerably advance our understanding of how cell metabolism contributes to pulmonary fibrosis.
+Even more informative than *in vitro* model systems are approaches that enable the study of cell metabolism *in vivo*.
+Here, we performed metabolomic profiling of lung and plasma samples from mice treated with bleomycin and MCT inhibitors (**Fig.
+S9**).
+These analyses did not reveal changes in whole lung metabolite levels, demonstrating a potential limitation of using bulk metabolomics to monitor metabolic changes in heterogeneous cell populations *in vivo*.
+To overcome this challenge, we employed spatial metabolic imaging.
+Mice were administered stable isotope tracers of proline and glucose before euthanasia.
+Since these tracers are administered to live animals, multi-isotope imaging mass spectrometry identifies lung regions that are metabolically active during the labeling period.
+Since tissues were fixed and processed *ex vivo*, the isotope signals indicate substrate incorporation into fixable biomass [@zhang2020; @wertheim2023].
+Thus, using multi-isotope imaging mass spectrometry, we directly quantify isotope flux from glucose into fibrotic lung regions.
 
-MCT1 and MCT4 are expressed by many cells in the lung, particularly macrophages and dendritic cells [@travaglini2020].
-Although our *ex vivo* and *in vitro* experiments suggest that inhibiting myofibroblast differentiation is a primary antifibrotic mechanism, inhibition of lactate transporters expressed by other cell types may also play a role in their therapeutic effect.
-For example, MCT4 expression is increased as part of a HIF-1α gene expression signature in transitional AT2 cells that accumulate in pulmonary fibrosis and contribute to aberrant repair [@choi2020; @zhu2022].
-We are in the process of generating fibroblast-specific conditional knockout mice to explore the role of these transporters in fibroblasts specifically.
+Our findings demonstrated that ^15^N-proline enrichment levels correlated with tissue fibrosis as assessed histologic and biochemically.
+This finding emphasizes the ability of cells to utilize  circulating proline for protein synthesis, potentially diminishing the significance of *de novo* proline biosynthesis *in vivo*.
+Moreover, the ^2^H-glucose signal provided some direct evidence for MCT-dependent metabolic reprogramming in mice, showing reduced carbohydrate incorporation into matrix proteins.
+Owing to the financial and time costs of isotope tracing and imaging, we were able to study only a few animals per group.
+Nevertheless, metabolic imaging at high spatial resolution holds great promise for correlating cell identities identified from spatial transcriptomic profiles with metabolic features.
+Furthermore, stable isotopes may be safely administered to human patients prior lung biopsy or explant [@steinhauser2012; @guillermier2017], offering a strategy that could significantly advance our understanding of how cell metabolism contributes to pulmonary fibrosis.
 
+MCT1 and MCT4 are expressed by many cells in the lung, notably macrophages and dendritic cells [@travaglini2020].
+Although our *ex vivo* and *in vitro* experiments suggest that inhibiting myofibroblast differentiation is a primary antifibrotic mechanism, the inhibition of lactate transporters expressed by other cell types may also contribute to their therapeutic effect.
+For example, MCT4 expression is upregulated as part of a HIF-1α gene expression signature in transitional AT2 cells that accumulate in pulmonary fibrosis and contribute to aberrant repair processes [@choi2020; @zhu2022].
+Future work will explore the cell-type-specific effects of these transporters in conditional knockout mice.
+
+One of the barriers to translating metabolic therapies to humans has been the poor pharmacologic properties of the small molecule laboratory compounds that have been studied previously.
+Both AZD3965 and VB253 exhibit low IC~50~ values of approximately 2 nM, compared to the next most potent inhibitor studied in pulmonary fibrosis models, lonidamine, with an IC~50~ of 7,000 nM [@yin2019].
+As metabolic targets downstream of glycolysis, lactate transport inhibitors may also offer better tolerance by allowing glycolysis to continue supporting glucose oxidation, contrasting with upstream glycolysis inhibitors that can more severely impact cellular bioenergetics.
 Pharmacologic interest in lactate transporters has been driven by the recognition of increased lactate transporter expression in a variety of cancers [@payen2020].
@@
 In this Phase 1 study, AZD3965 was generally well-tolerated with 7 of 40 patients experiencing dose-limiting toxicities including asymptomatic, reversible ocular changes; acidosis; and increased troponin.
-VB124 was the first selective MCT4 inhibitor developed [@cluntun2021] and here we report a second generation MCT4 inhibitor, VB253.
-A Phase 1 clinical trial of VB253 is currently enrolling [**REF?**].
-MCT4 global knockout mice are viable and breed normally [@bisetto2018; @bisetto2019], increasing optimism that VB253 will be well tolerated in humans.
-One of the barriers to translating metabolic therapies to humans has been the poor pharmacologic properties of the small molecule laboratory compounds that have been studied previously.
-The IC~50~ values for AZD3965 and VB253 are ~ 2 nM, compared to the next most potent inhibitor, lonidamine, with an IC~50~ of 7,000 nM [@yin2019].
-As a metabolic target downstream of glycolysis, lactate transport inhibitors may also be better tolerated by allowing glycolysis to continue to support of glucose oxidation, while upstream inhibitors of glycolysis have a more severe impact on cellular bioenergetics.
+VB124 was the first selective MCT4 inhibitor developed [@cluntun2021] and we now introduce VB253 as a second generation MCT4 inhibitor.
+Our data demonstrate that VB253 has similar efficacy in experimental pulmonary fibrosis models as the established antifibrotics, nintedanib and pirfenidone, with potentially less cytotoxicity (**Fig.
+8**).
+MCT4 global knockout mice are viable and breed normally [@bisetto2018; @bisetto2019], raising optimism that VB253 will be well tolerated in humans, and a Phase 1 clinical trial of VB253 is currently underway [**REF?**].
 
-In conclusion, we have demonstrated that lactate transporters promote myofibroblast differentiation and pulmonary fibrosis.
-Using a comprehensive metabolic phenotyping strategy, we have identified antifibrotic mechanisms for MCT inhibition and demonstrated evidence for metabolic reprogramming in animal models.
-We have demonstrated the antifibrotic efficacy of currently available lactate transport inhibitors with accepted preclinical disease models.
-Collectively, our data advance lactate transport inhibition as a viable therapeutic strategy for patients with pulmonary fibrosis.
+In summary, our findings highlight the pivotal role of lactate transporter in driving myofibroblast differentiation and pulmonary fibrosis.
+Through a comprehensive metabolic phenotyping approach, we have characterized the antifibrotic mechanisms associated with MCT inhibition and provided compelling evidence of metabolic reprogramming in animal models.
+Furthermore, we have validated the antifibrotic effectiveness of existing lactate transport inhibitors using established preclinical disease models.
+Altogether, our results significantly advance lactate transport inhibition as a promising therapeutic approach for patients suffering from pulmonary fibrosis.
 
@@
 
-# Methods
+# Materials and Methods
 
-## Key Resources
+## Study Design
 
-```{r}
-#| results: asis
-#| tbl-cap: "Key Resources Table"
+This study was designed to investigate the role of lactate transporters in the myofibroblast differentiation and pulmonary fibrosis.
+The objectives of this study were (i) to determine the expression of lactate transporters in IPF lungs and model systems, (ii) to characterize the phenotypic effects of lactate transport inhibition in model systems, and (iii) to profile the metabolic consequences of lactate transport inhibitors.
 
-# flextable::flextable_to_rmd(resources_table)
-```
+Cell culture experiments were performed at least three times.
+The number of animals per experimental group was chosen based on prior publications and were repeated at least once.
+Stable isotope tracing experiments were performed on three animals per group owing to resource availability.
+Animals were randomly assigned to treatment.
+The pathologist scoring histologic fibrosis severity was blinded to treatment assignment.
+The number of unique patient samples was determined by clinical availability.
 
@@
 
-## Chemicals
-
-Recombinant human TGF-β1 was purchased commercially, dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
-AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
-VB124 (10 μM) was a generous gift from Vettore, LLC.
-The drugs were prepared in DMSO and used at the indicated doses unless otherwise indicated.
-These doses were chosen based on prior reports and their target affinities.
-
-## Human lung tissue and fibroblasts
-
-Human IPF lung samples were obtained at the time of lung transplant.
-Control samples were prepared from non-fibrotic and unaffected border regions of lung nodule biopsy specimens or from failed donor controls.
-To isolate human lung fibroblasts, lung tissue was minced, digested with liberase and DNase, and passed through a sterile 70 μm filter.
-The filtrate was centrifuged at 300 ×*g* and washed once with RPMI medium (Lonza).
-Cells were plated in DMEM (Lonza) supplemented with 10% FBS, penicillin, and streptomycin.
-After two passages, medium was changed to FGM-2 (Lonza).
-
-## Bleomycin model
-
-C57Bl6/N mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
-Control animals received an equivalent volume of PBS.
-For drug treatments, AZD3965 (100 mg/kg twice daily) and VB124 (30 mg/kg daily) were prepared in 0.5% methylcellulose and 0.1% Tween-20 with sonication and administered by oral gavage every day beginning on day 7 following bleomycin.
-Control animals received vehicle alone twice daily.
-After 21 d, animals were euthanized.
-Lung function measurements were performed.
-Plasma samples were obtained by right ventricular puncture and collected in EDTA-coated tubes.
-The lungs were dissected and perfused with PBS.
-The right lung was isolated while the left lung was inflated to 20 cmH~2~O prior to fixation.
-The right lungs were snap frozen.
-
-For lung function measurements, animals were anesthetized with pentobarbital (100 μg/g; Covetrus, Portland, ME) and tracheally cannulated for pulmonary function measurements using a flexiVent (SCIREQ).
-Animals were ventilated for 3-5 min with 150 breaths per minute at 10 mL/kg and 3 cmH~2~O positive end-expiratory pressure.
-After two total lung capacity maneuvers, respiratory elastance was measured using two Prime-8 maneuvers.
-A pressure-volume loop was then generated, and static compliance was calculated using the Salazar-Knowles equation.
-
-Histologic fibrosis severity was assessed by Ashcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
-The average of twenty 10× fields is reported for each animal.
-
-Lung hydroxyproline content was determined as described previously [@tager2008].
-The four lobes of the R lung were homogenized in PBS with Halt protease inhibitors (Thermo) and hydrolyzed in 6 N HCl at 120 °C for 12 h, with an aliquot added to 1.4% chloramine-T, 10% 1-propanol, and 0.5 M sodium acetate, pH 6.0.
-After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-value-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
-Samples were incubated at 65 °C for 15 minutes.
-Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
-
-## Myofibroblast differentiation *in vitro*
-
-Normal lung fibroblasts were purchased commercially (Lonza).
-All fibroblasts were cultured in FGM-2 medium (Lonza).
-Passages 3-8 were used for experiments.
-Myofibroblast differentiation was induced *in vitro* using recombinant human TGF-β1 (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
-
-## Isotope labeling
-
-Cells were seeded and serum starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
-The concentrations of glucose and glutamine match the concentrations of these substrates determined in standard (FGM-2) growth medium.
-After 24 h, cells were washed with PBS and the medium was changed to "heavy" labeling medium containing [1,2-^13^C~2~]-glucose, [U-^13^C~6~]-glucose, [U-^13^C~5~]-glutamine, or [U-^13^C~2~]-lactate (Cambridge Isotope Labs) along with TGFβ and MCT inhibitors.
-Cell lysates were prepared for metabolomics as described below.
-
-## Immunoblotting
-
-Human and mouse lung samples were mechanically homogenized in RIPA lysis buffer containing Halt Protease Inhibitor Cocktail (Thermo).
-Cells were washed with one volume of PBS and collected by scraping in PBS.
-Cell suspensions were centrifuged at 5,000 ×*g* for 5 min at 4 °C.
-Pellets were lysed in buffer containing Tris 10 mM, pH 7.4, NaCl 150 mM, EDTA 1 mM, EGTA 1 mM, Triton X-100 1% v/v, NP-40 0.5% v/v, and protease inhibitors.
-Protein concentrations were determined by BCA Protein Assay (Thermo).
-Lysates were normalized for protein concentration and subjected to SDS-PAGE separation on stain-free tris-glycine gels (Bio-Rad), cross-linked and imaged with the Chemidoc system (Bio-Rad), transferred to PVDF membranes with the Trans-Blot Turbo transfer system (Bio-Rad), imaged, blocked in 5% blocking buffer (Bio-Rad), blotted in primary and secondary antibodies, and developed using WesternBright ECL (Advansta).
-Band signal intensity was normalized to total protein per lane as determined from the stain-free gel or membrane images using Image Lab software (Bio-Rad).
-
-## Histone purification
-
-Histones were extracted from lung fibroblast cell lysates using the EpiQuik Total Histone Extraction Kit (Epigentek).
-
-## RNA interference
-
-Lung fibroblasts were reverse transfected with ON-TARGETplus SMARTpool siRNA (Dharmacon) targeting MCT1 or MCT4 using Lipofectamine RNAiMAX (Thermo).
-A non-targeting siRNA pool was used a transfection control (Dharmacon).
-For combined siMCT1 and siMCT4 treatment, the total siRNA dose was kept constant.
-After 24 h, cells were serum starved for an additional 24 h prior to treatment with TGFβ.
-
-## Gel contraction assay
-
-Lung fibroblasts in FBM and neutralized collagen solution (TeloCol-3, Advanced Biomatrix) were combined 1:2 to yield 100,000 cells/mL and 1.8 mg/mL collagen.
-This solution was added to 24-well plates (0.5 mL) and allowed to harden for 10 min at 37 °C prior to the addition of 1 mL FBM.
-Cells were serum starved overnight prior to releasing the gels and treating them with TGFβ and MCT inhibitors in triplicate.
-Images were acquired after 24 and 48 h using a ChemiDoc imager (Bio-Rad).
-Gel areas were measured using Image Lab software (Bio-Rad).
-Data are expressed as the ratio of TGFβ-treated gel area to control gel area.
-
-## Cell count
-
-Cell counts were estimated from total DNA quantified using the Quant-iT PicoGreen dsDNA Assay Kit (Thermo) as described previously [@copeland2023].
-
-## Lactate assay
-
-Lactate was determined using an enzymatic cycling assay resulting in the reduction of resazurin to the fluorescent chemical resorufin.
-Medium samples were diluted 1:10 in PBS.
-Samples were combined with a solution of diaphorase (125 mU), lactate dehydrogenase, NAD (250 μM), and resazurin (25 μM) in Tris buffer (100 mM, pH 8.5) then incubated for 30 min at room temperature.
-Resorufin fluorescence was measured using a SpectraMax i3x microplate reader (Molecular Devices) and lactate concentration was interpolated from a standard curve.
-Data were normalized to estimated cell count.
-
-## Glucose assay
-
-Medium samples were diluted 10-fold in PBS.
-Glucose concentration was determined using the Glucose Colorimetric Assay Kit (Cayman) according to the manufacturer's protocol.
-Standards were prepared in PBS.
-
-## Seahorse assays
-
-Lung fibroblasts were seeded in Seahorse assay plates at a density of 20,000 cells per well.
-On the following day, the cells were serum starved for 24 h prior to treatment with TGFβ and MCT inhibitors for 48 h.
-On the day of the assay, the medium was exchanged to Seahorse DMEM without phenol red supplemented with 5 mM HEPES, 10 mM glucose, 2 mM glutamine, and 1 mM pyruvate.
-OCR and extracellular acidification rates (ECAR) were measured at baseline and following the sequential injections of oligomycin (1 μM final), FCCP (2 μM final), and rotenone plus antimycin A (0.5 μM final each) using an XFe24 bioanalyzer (Agilent).
-At the end of the assay, cells were stained for 30 min with the fluorescent dye Nuclear Green LCS1 (10 μM; Abcam).
-Stained cells were quantified using a SpectraMax MiniMax 300 Imaging Cytometer (Molecular Devices) and these cell counts were used to normalize the rate data.
-Proton efflux rate (PER) was calculated from the ECAR and buffer capacity of the medium.
-PER from glycolysis was determined by subtracting the contribution of oxidative phosphorylation to total PER using the empirically measured CO~2~ correction factor of 0.41.
-ATP from glycolysis was assumed to be equal to basal glycolytic PER.
-ATP from oxidative phosphorylation was assumed to be 2 × 2.75 × oligomycin-sensitive OCR [@mookerjee2017].
-
-## Metabolomics
-
-Metabolomics data acquisition and analysis were performed as described previously [@copeland2023].
-
-### Metabolite extraction
-
-Extracellular metabolites were obtained by mixing conditioned medium 1:4 with 100% MeOH precooled to -80 °C.
-Intracellular metabolites were obtained after washing cells with 2 volumes of ice-cold PBS and floating on liquid nitrogen.
-Plates were stored at -80 °C until extraction.
-Metabolites were extracted with 1 mL 80% MeOH pre-cooled to -80 °C.
-Mouse plasma and lung homogenates were mixed 1:4 with 100% MeOH precooled to -80 °C.
-Samples were extracted at -80 °C for 4 h.
-Insoluble material from these samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
-The supernatant was evaporated to dryness at 42 °C using a SpeedVac concentrator (Thermo Savant).
-Samples were resuspended in 35 μL 20 mM ammonium phosphate in LC-MS-grade water prior to analysis.
-
-### Acquisition parameters
-
-LC-MS analysis was performed on a Vanquish ultra-high-performance liquid chromatography system coupled to a Q Exactive orbitrap mass spectrometer by a HESI-II electrospray ionization probe (Thermo).
-External mass calibration was performed weekly.
-Metabolite samples (2.5 μL) were separated using a ZIC-pHILIC stationary phase (2.1 × 150 mm, 5 μm) (Merck).
-The autosampler temperature was 4 °C and the column compartment was maintained at 25 °C.
-Mobile phase A was 20 mM ammonium carbonate, 0.1% ammonium hydroxide, and 5 μM medronic acid.
-Mobile phase B was acetonitrile.
-The flow rate was 0.1 mL/min.
-Solvent was introduced to the mass spectrometer *via* electrospray ionization with the following source parameters: sheath gas 40, auxiliary gas 15, sweep gas 1, spray voltage +3.0 kV for positive mode and -3.1 kV for negative mode, capillary temperature 275 °C, S-lens RF level 40, and probe temperature 350 °C.
-Data were acquired and peaks integrated using TraceFinder 4.1 (Thermo).
-
-### Metabolomic profiling
-
-For metabolomic profiling, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B.
-The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*.
-Resolution was set to 70,000 and the AGC target was 1×10^6^ ions.
-Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method.
-For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run.
-Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas.
-Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018].
-Missing values were imputed using random forest.
-Sample peak areas were normalized using probabilistic quotient normalization [@dieterle2006].
-Differentially regulated metabolites were identified using limma [@ritchie2015].
-Exploratory data analysis and review of PCA plots suggested orthogonal effects of TGFβ and MCT inhibitor treatment.
-For this reason, differentially regulated metabolites were determined based on the main effects of TGFβ or MCT inhibitor.
-Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021] with KEGG metabolite pathways [@kanehisa2000].
-
-### Stable isotope quantification
-
-Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B.
-The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite.
-The resolution was set at 70,000 and AGC target was 1×10^5^ ions.
-Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015].
-Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].
-## RNA-seq transcriptomics
-
-RNA was collected from LFs treated with TGFβ and MCT inhibitors as described above.
-Four biological replicates were analyzed.
-Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample.
-Sequences were mapped to the human GRCh38 primary assembly and counts summarized using salmon [@patro2017].
-Differentially expressed genes were identified using DESeq2 [@love2014].
-Gene set enrichment was performed using the fgsea R package [@korotkevich2021].
-
-## Pyridine dinucleotide measurements
-
-Cellular NAD^+^ and NADH were measured using an enzymatic fluorometric cycling assay as previously described [@oldham2015; @copeland2023].
-Samples for NADP^+^ and NADPH quantification were similarly prepared and analyzed using the luminescence-based NADP/NADPH-Glo assay (Promega).
-
-## Reactive oxygen species
-
-*Fei Li to write method for CellROX and MitoSOX/MitoTracker.*
-
-## Multi-isotope imaging mass spectrometry
-
-For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on day 21 after bleomycin.
-
-*Christelle to write.*
-
-Isotopic enrichment for each animal/image was summarized as follows.
-The ratio of "heavy" to "light" ion intensities were calculated.
-An image mask was generated by Otsu thresholding to exclude background regions.
-Pixel ratios equal to zero were excluded as these are non-biologic.
-The average of the remaining pixel ratios was then calculated to provide a mean tissue isotope enrichment value for the image.
-
-## VB253
-
-VB253 was developed by Vettore, LLC as a more potent small molecule inhibitor of MCT4.
-Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC.
-These experiments were completed independently of the other studies presented in the manuscript and offer independently validation of the antifibrotic effects of MCT4 inhibition in cells and mice.
-
-*Nathan, Mark, and Jared.
-Add the names of the CROs that did the experiments?
-I think the aSMA/Smad3/nuclei experiments will need a description.
-The others are pretty similar to what is described above.*
-
-## Statistics
+## Statistical Analysis
 
 Data analysis, statistical comparisons, and visualization were performed in R [@R-base].
-Experiments included technical and biological replicates as noted above.
+Experiments included technical and biological replicates as noted in the Materials and Methods.
 The number of biological replicates (N) is indicated in the figure legends.
@@
 Outliers were identified using twice the median absolute deviation as a cutoff threshold.
-Comparisons were performed using linear mixed-effects models with condition (±TGFβ), treatment and their interaction as fixed effects and biological replicate as a random effect.
+Comparisons were performed using linear mixed-effects models with condition (±TGFβ), treatment, and their interaction as fixed effects and biological replicate or donor as a random effect.
 Significant differences in estimated marginal means were identified by comparisons to the multivariate *t* distribution.
-Metabolomics and RNA-seq data were analyzed as described above.
+Metabolomics and RNA-seq data were analyzed as described in the Materials and Methods.
 Probability values less than 0.05 were considered statistically significant.
@@
 Human samples were obtained through the BWH Biorepository for Understanding Inflammatory Lung Disease (BUILD) or the MGH ILD Translational Research Program and their collection was approved by the Mass General Brigham Institutional Review Board (2013P002332, 2016P001890, 2019P003592, 2020P002765).
-
 All animal experiments were approved by the Brigham and Women's Hospital IACUC (2020N000199).
 
-## Data availability
+# List of Supplementary Materials
+
+Materials and Methods
+
+References and Notes
+
+Fig. S1-S10
+
+\newpage
+
+# References and Notes
+
+::: {#refs}
+:::
+
+\newpage
+
+# Acknowledgements
+
+The authors thank Louise Trakimas and the HMS Electron Microscopy Facility for MIMS sample preparation and consultation.
+The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health.
+VB124 was a gift from Vettore, LLC to the Oldham laboratory.
+
+## Funding
+
+National Institutes of Health grant R01GM987654 (WMO)
+Brigham and Women's Hospital Research Institute (WMO)
+
+## Author contributions
+
+Conceptualization: KMP, MLS, JR, WMO
+Methodology: DRZ, FL, KMP, CG, JV, BAM, MLS, RSK, JR, WMO
+Investigation: DRZ, FL, KMP, NMK, KJL, CG, JV, LPH, RSK, WMO
+Visualization: WMO
+Funding acquisition: KMP, JR, WMO
+Project administration: KMP, JR, WMO
+Resources: DRZ, KMP, NMK, RMB, NJP, EYK, RSK
+Supervision: KMP, JR, WMO
+Writing - original draft: WMO
+Writing - review and editing: DRZ, FL, KMP, NMK, CG, JV, RMB, BAM, NJP, EYK, MLS, RSK, JR, WMO
+
+## Competing interests
+
+W.M.O. has received consulting fees from Nikang Therapeutics outside the scope of this research.
+
+J.R. is a consultant and shareholder for Vettore Biosciences.
+
+R.S.K. received a Discovery ILD Award from Boehringer Ingelheim and received support through the Partners Drug Development Lab from Bayer Pharmaceuticals, all outside the scope of this research.
+
+B.A.M. has received consulting fees from Actelion and Tenax and has performed investigator-initiated research with support from Deerfield, all outside the scope of this research.
+
+The remaining authors declare that they have no competing interests.
+
+*Here, also include any awarded or filed patents pertaining to the results presented in the paper, stating patent filing number and title and which authors are co-inventors*
+
+## Data and materials availability
 
@@
 Details of the processing pipeline and summarized data are available from the Oldham Lab GitHub repository (github.com/oldhamlab/rnaseq.lf.tgfb.mcti).
-
-\newpage
-
-# Author Contributions
-
-The authors confirm contribution to the paper as follows: designed research studies - D.R.Z., F.L., K.M.P., R.M.B., B.A.M., N.J.P., E.Y.K., M.L.S., R.S.K., J.R., W.M.O.; conducted experiments - D.R.Z., F.L., K.J.L., W.M.O.; acquired data - D.R.Z., F.L., N.M.K., K.J.L., K.M.P., C.G., J.V., L.P.H., R.S.K., W.M.O.; analyzed data - D.R.Z., F.L., N.M.K., K.M.P., C.G., M.L.S., R.S.K., J.R., W.M.O.; wrote the manuscript - W.M.O.
-All authors reviewed the results, revised the manuscript, and approve the final version.
-
-# Acknowledgements
-
-The authors thank Louise Trakimas and the HMS Electron Microscopy Facility for MIMS sample preparation and consultation.
-Research reported in this publication was supported by the National Heart Lung and Blood Institute of the National Institutes of Health under award number R01GM987654 (to W.M.O.), and by the Brigham and Women's Hospital Research Institute.
-The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Institutes of Health.
-VB124 was a gift from Vettore, LLC.
-
-\newpage
-
-# References
-
-::: {#refs}
-:::
 
@@
 (**D**) Representative trichrome-stained histologic sections of mouse lungs.
+Scale bar is 200 μm.
 (**E**) Ashcroft scores from histologic sections from (D) determined by a lung pathologist blinded to treatment assignment (N = 6-7).
 (**F**) Right lung hydroxyproline content (N = 10-13).
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_06.png)
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_06.png)
 
@@
 The ^12^C~2~^2^H/^12^C~2~^1^H ratio shows enrichment above natural abundance from ^2^H-glucose (colorbar 1-10-fold of natural abundance).
-(**C**) Quantification of isotope enrichment (N = 3 biological replicates, \* adjusted p-value-value < 0.05 compared to bleomycin-treated control).](figs/F_07.png)
+(**C**) Quantification of isotope enrichment (N = 3 biological replicates, \* adjusted p-value < 0.05 compared to bleomycin-treated control).](figs/F_07.png)
 
@@
 (**D**) Enhanced pause (Penh) is a non-invasive index of mouse tidal breathing that increases with bleomycin treatment.
-(**E**-**G**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (E) and decreased α-SMA protein expression (F-G).
-(**H**-**J**) Similar histologic improved was observed in aged mice with decreased histologic fibrosis (H) and decreased α-SMA protein expression (I-J).
-(**K**) VB253 decreased lung lactate, similar to established antifibrotics.
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_10.png)
+(**E**-**G**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (E) and decreased α-SMA protein expression (F).
+(**G**-**I**) Similar histologic improved was observed in aged mice with decreased histologic fibrosis (G) and decreased α-SMA protein expression (H).
+(**I**) VB253 decreased lung lactate, similar to established antifibrotics.
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_08.png)

### 2024-04-16 _targets.R

@@ -2063,8 +2063,8 @@ list(
   # figure 8 ----------------------------------------------------------------
 
   tar_target(
-    fig10,
-    make_fig10(
+    fig08,
+    make_fig08(
       vb_invitro_plot_sma_ipf,
       vb_invitro_plot_nuclei_ipf,
       fig_img_vb_timeline,
@@ -2089,7 +2089,7 @@ list(
         ),
       vb_mice_plot_old_lactate
     ) |>
-      write_figures("F_10"),
+      write_figures("F_08"),
     format = "file"
   ),
 
@@ -2272,11 +2272,20 @@ list(
     format = "file"
   ),
   tar_target(
-    refs,
+    refs_ms,
     rbbt::bbt_update_bib(
       path = "manuscript/manuscript.qmd",
       ignore = c("R-base"),
-      path_bib = "manuscript/library.bib"
+      path_bib = "manuscript/ms.bib"
+    ),
+    cue = tar_cue("always")
+  ),
+  tar_target(
+    refs_supp,
+    rbbt::bbt_update_bib(
+      path = "manuscript/supplement.qmd",
+      ignore = c("R-base"),
+      path_bib = "manuscript/supp.bib"
     ),
     cue = tar_cue("always")
   ),

### 2024-04-16 manuscript/supplement.qmd
format for STM and co-author review

@@
 - id: utah_gen
-  name: Department of Human Genetics, University of Utah, Salt Lake City, UT
+  name: Department of Internal Medicine, University of Utah, Salt Lake City, UT
 - id: vettore
@@
 bibliography: 
-- library.bib
+- supp.bib
 - pkgs.bib
-csl: jci.csl
+csl: stm.csl
 
@@
 # ensure refs are up-to-date
-targets::tar_load(refs)
+targets::tar_load(refs_supp)
 targets::tar_load(pkg_citations)
@@
   c(
+    resources_table, 
     fig01s,
@@
 # Materials and Methods
+
+## Key Resources
+
+```{r}
+#| results: asis
+#| tbl-cap: "Key Resources Table"
+
+flextable::flextable_to_rmd(resources_table)
+```
+
+\newpage
+
+## Chemicals
+
+Recombinant human TGF-β1 was purchased commercially , dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
+AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
+VB124 (10 μM) was a gift from Vettore, LLC.
+The drugs were prepared in DMSO and used at the indicated doses unless otherwise indicated.
+These doses were chosen based on prior reports and their target affinities.
+
+## Human lung tissue and fibroblasts
+
+Human IPF lung samples were obtained at the time of lung transplant.
+Control samples were prepared from non-fibrotic and unaffected border regions of lung nodule biopsy specimens or from failed donor controls.
+To isolate human lung fibroblasts, lung tissue was minced, digested with liberase and DNase, and passed through a sterile 70 μm filter.
+The filtrate was centrifuged at 300 ×*g* and washed once with RPMI medium (Lonza).
+Cells were plated in DMEM (Lonza) supplemented with 10% FBS, penicillin, and streptomycin.
+After two passages, medium was changed to FGM-2 (Lonza).
+
+## Bleomycin model
+
+C57Bl6/N mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
+Control animals received an equivalent volume of PBS.
+For drug treatments, AZD3965 (100 mg/kg twice daily) and VB124 (30 mg/kg daily) were prepared in 0.5% methylcellulose and 0.1% Tween-20 with sonication and administered by oral gavage every day beginning on day 7 following bleomycin.
+Control animals received vehicle alone twice daily.
+After 21 d, lung function measurements were performed.
+Animals were euthanized and plasma samples were obtained by right ventricular puncture and collected in EDTA-coated tubes.
+The lungs were dissected and perfused with PBS.
+The right lung was isolated while the left lung was inflated to 20 cmH~2~O prior to fixation.
+The right lungs were snap frozen.
+
+Lung physiology measurements were performed as previously described [@lee2021].
+Briefly, animals were anesthetized with pentobarbital (100 μg/g; Covetrus, Portland, ME) and tracheally cannulated for pulmonary function measurements using a flexiVent (SCIREQ).
+Animals were ventilated for 3-5 min with 150 breaths per minute at 10 mL/kg and 3 cmH~2~O positive end-expiratory pressure.
+After two total lung capacity maneuvers to prevent atelectasis and standardize volume history, respiratory elastance was measured using two Prime-8 maneuvers; values reported are the average of these two maneuvers.
+A pressure-volume loop was then generated, and static compliance was calculated using the Salazar-Knowles equation.
+
+Histologic fibrosis severity was assessed by Ashcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
+The average of twenty 10× fields is reported for each animal.
+
+Lung hydroxyproline content was determined as described previously [@tager2008].
+The four lobes of the R lung were homogenized in PBS with Halt protease inhibitors (Thermo) and hydrolyzed in 6 N HCl at 120 °C for 12 h, with an aliquot added to 1.4% chloramine-T, 10% 1-propanol, and 0.5 M sodium acetate, pH 6.0.
+After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-value-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
+Samples were incubated at 65 °C for 15 minutes.
+Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
+
+## Myofibroblast differentiation *in vitro*
+
+Normal lung fibroblasts were purchased commercially (Lonza).
+All fibroblasts were cultured in FGM-2 medium (Lonza).
+Passages 3-8 were used for experiments.
+Myofibroblast differentiation was induced *in vitro* using recombinant human TGF-β1 (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
+
+## Isotope labeling
+
+Cells were seeded and serum starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
+The concentrations of glucose and glutamine match the concentrations of these substrates determined in standard (FGM-2) growth medium.
+After 24 h, cells were washed with PBS and the medium was changed to "heavy" labeling medium containing [1,2-^13^C~2~]-glucose, [U-^13^C~6~]-glucose, [U-^13^C~5~]-glutamine, or [U-^13^C~2~]-lactate (Cambridge Isotope Labs) along with TGFβ and MCT inhibitors for 48 h.
+Cell lysates were prepared for metabolomics as described below.
+
+## Immunoblotting
+
+Human and mouse lung samples were mechanically homogenized in RIPA lysis buffer containing Halt Protease Inhibitor Cocktail (Thermo).
+Cells were washed with one volume of PBS and collected by scraping in PBS.
+Cell suspensions were centrifuged at 5,000 ×*g* for 5 min at 4 °C.
+Pellets were lysed in buffer containing Tris 10 mM, pH 7.4, NaCl 150 mM, EDTA 1 mM, EGTA 1 mM, Triton X-100 1% v/v, NP-40 0.5% v/v, and protease inhibitors.
+Protein concentrations were determined by BCA Protein Assay (Thermo).
+Lysates were normalized for protein concentration and subjected to SDS-PAGE separation on stain-free tris-glycine gels (Bio-Rad), cross-linked and imaged with the Chemidoc system (Bio-Rad), transferred to PVDF membranes with the Trans-Blot Turbo transfer system (Bio-Rad), imaged, blocked in 5% blocking buffer (Bio-Rad), blotted in primary and secondary antibodies, and developed using WesternBright ECL (Advansta).
+Band signal intensity was normalized to total protein per lane as determined from the stain-free gel or membrane images using Image Lab software (Bio-Rad).
+
+## RNA interference
+
+Lung fibroblasts were reverse transfected with ON-TARGETplus SMARTpool siRNA (Dharmacon) targeting MCT1 or MCT4 using Lipofectamine RNAiMAX (Thermo).
+A non-targeting siRNA pool was used a transfection control (Dharmacon).
+For combined siMCT1 and siMCT4 treatment, the total siRNA dose was kept constant.
+After 24 h, cells were serum starved for an additional 24 h prior to treatment with TGFβ.
+
+## Gel contraction assay
+
+Lung fibroblasts in FBM and neutralized collagen solution (TeloCol-3, Advanced Biomatrix) were combined 1:2 to yield 100,000 cells/mL and 1.8 mg/mL collagen.
+This solution was added to 24-well plates (0.5 mL) and allowed to harden for 10 min at 37 °C prior to the addition of 1 mL FBM.
+Cells were serum starved overnight prior to releasing the gels and treating them with TGFβ and MCT inhibitors in triplicate.
+Images were acquired after 24 and 48 h using a ChemiDoc imager (Bio-Rad).
+Gel areas were measured using Image Lab software (Bio-Rad).
+Data are expressed as the ratio of TGFβ-treated gel area to control gel area.
+
+## Cell count
+
+Cell counts were estimated from total DNA quantified using the Quant-iT PicoGreen dsDNA Assay Kit (Thermo) as described previously [@copeland2023].
+
+## Lactate assay
+
+Lactate was determined using an enzymatic cycling assay resulting in the reduction of resazurin to the fluorescent chemical resorufin.
+Medium samples were diluted 1:10 in PBS.
+Samples were combined with a solution of diaphorase (125 mU), lactate dehydrogenase, NAD (250 μM), and resazurin (25 μM) in Tris buffer (100 mM, pH 8.5) then incubated for 30 min at room temperature.
+Resorufin fluorescence was measured using a SpectraMax i3x microplate reader (Molecular Devices) and lactate concentration was interpolated from a standard curve.
+Data were normalized to estimated cell count.
+
+## Glucose assay
+
+Medium samples were diluted 10-fold in PBS.
+Glucose concentration was determined using the Glucose Colorimetric Assay Kit (Cayman) according to the manufacturer's protocol.
+Standards were prepared in PBS.
+
+## Seahorse assays
+
+Lung fibroblasts were seeded in Seahorse assay plates at a density of 20,000 cells per well.
+On the following day, the cells were serum starved for 24 h prior to treatment with TGFβ and MCT inhibitors for 48 h.
+On the day of the assay, the medium was exchanged to Seahorse DMEM without phenol red supplemented with 5 mM HEPES, 10 mM glucose, 2 mM glutamine, and 1 mM pyruvate.
+OCR and extracellular acidification rates (ECAR) were measured at baseline and following the sequential injections of oligomycin (1 μM final), FCCP (2 μM final), and rotenone plus antimycin A (0.5 μM final each) using an XFe24 bioanalyzer (Agilent).
+At the end of the assay, cells were stained for 30 min with the fluorescent dye Nuclear Green LCS1 (10 μM; Abcam).
+Stained cells were quantified using a SpectraMax MiniMax 300 Imaging Cytometer (Molecular Devices) and these cell counts were used to normalize the rate data.
+Proton efflux rate (PER) was calculated from the ECAR and buffer capacity of the medium.
+PER from glycolysis was determined by subtracting the contribution of oxidative phosphorylation to total PER using the empirically measured CO~2~ correction factor of 0.41.
+ATP from glycolysis was assumed to be equal to basal glycolytic PER.
+ATP from oxidative phosphorylation was assumed to be 2 × 2.75 × oligomycin-sensitive OCR [@mookerjee2017].
+
+## Metabolomics
+
+Metabolomics data acquisition and analysis were performed as described previously [@copeland2023].
+
+### Metabolite extraction
+
+Extracellular metabolites were obtained by mixing conditioned medium 1:4 with 100% MeOH precooled to -80 °C.
+Intracellular metabolites were obtained after washing cells with 2 volumes of ice-cold PBS and floating on liquid nitrogen.
+Plates were stored at -80 °C until extraction.
+Metabolites were extracted with 1 mL 80% MeOH pre-cooled to -80 °C.
+Mouse plasma and lung homogenates were mixed 1:4 with 100% MeOH precooled to -80 °C.
+Samples were extracted at -80 °C for 4 h.
+Insoluble material from these samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
+The supernatant was evaporated to dryness at 42 °C using a SpeedVac concentrator (Thermo Savant).
+Samples were resuspended in 35 μL 20 mM ammonium phosphate in LC-MS-grade water prior to analysis.
+
+### Acquisition parameters
+
+LC-MS analysis was performed on a Vanquish ultra-high-performance liquid chromatography system coupled to a Q Exactive orbitrap mass spectrometer by a HESI-II electrospray ionization probe (Thermo).
+External mass calibration was performed weekly.
+Metabolite samples (2.5 μL) were separated using a ZIC-pHILIC stationary phase (2.1 × 150 mm, 5 μm) (Merck).
+The autosampler temperature was 4 °C and the column compartment was maintained at 25 °C.
+Mobile phase A was 20 mM ammonium carbonate, 0.1% ammonium hydroxide, and 5 μM medronic acid.
+Mobile phase B was acetonitrile.
+The flow rate was 0.1 mL/min.
+Solvent was introduced to the mass spectrometer *via* electrospray ionization with the following source parameters: sheath gas 40, auxiliary gas 15, sweep gas 1, spray voltage +3.0 kV for positive mode and -3.1 kV for negative mode, capillary temperature 275 °C, S-lens RF level 40, and probe temperature 350 °C.
+Data were acquired and peaks integrated using TraceFinder 4.1 (Thermo).
+
+### Metabolomic profiling
+
+For metabolomic profiling, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B.
+The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*.
+Resolution was set to 70,000 and the AGC target was 1×10^6^ ions.
+Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method.
+For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run.
+Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas.
+Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018].
+Missing values were imputed using random forest.
+Sample peak areas were normalized using probabilistic quotient normalization [@dieterle2006].
+Differentially regulated metabolites were identified using limma [@ritchie2015].
+Exploratory data analysis and review of PCA plots suggested orthogonal effects of TGFβ and MCT inhibitor treatment.
+For this reason, differentially regulated metabolites were determined based on the main effects of TGFβ or MCT inhibitor.
+Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021] with KEGG metabolite pathways [@kanehisa2000].
+
+### Stable isotope quantification
+
+Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B.
+The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite.
+The resolution was set at 70,000 and AGC target was 1×10^5^ ions.
+Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015].
+Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].
+
+## RNA-seq transcriptomics
+
+RNA was collected from LFs treated with TGFβ and MCT inhibitors as described above.
+Four biological replicates were analyzed.
+Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample.
+Sequences were mapped to the human GRCh38 primary assembly and counts summarized using salmon [@patro2017].
+Differentially expressed genes were identified using DESeq2 [@love2014].
+Gene set enrichment was performed using the fgsea R package [@korotkevich2021].
+
+## Pyridine dinucleotide measurements
+
+Cellular NAD^+^ and NADH were measured using an enzymatic fluorometric cycling assay as previously described [@oldham2015; @copeland2023].
+Samples for NADP^+^ and NADPH quantification were similarly prepared and analyzed using the luminescence-based NADP/NADPH-Glo assay (Promega).
+
+## Reactive oxygen species
+
+Intracellular and mitochondrial ROS were measured using the fluorescent probes CellROX Deep Red (1 μM) and MitoSOX Green (2 μM), respectively.
+Mitochondria were also labeled with MitoTracker (200 nM).
+Briefly, lung fibroblasts were seeded in a 24-well plate and serum starved for 24 h prior to treatment with TGFβ, with or without MCT inhibitors, for 3 hours.
+This was followed by staining with the specified probes for 30 min at 37 °C.
+The cells were then harvested and analyzed using the CytoFLEX flow cytometry analyzer (Beckman Coulter).
+The mean fluorescence intensity of DAPI-negative cells was analyzed using FlowJo V10.8.1.
+
+## Multi-isotope imaging mass spectrometry
+
+For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on day 21 after bleomycin.
+Perfused and inflated lungs were fixed with formaldehyde/glutaraldehyde 2.5% in sodium cacodylate buffer, pH 7.4, embedded in EPON, section to 0.5 μm, mounted on silicon wafers, and gold coated.
+Sections were analyzed with the NanoSIMS 50L (CAMECA) housed in the Brigham and Women’s Hospital Center for NanoImaging, using previously developed analytical protocols [@wertheim2023; @guillermier2017a].
+The instrument was tuned to simultaneously acquire images of ^12^C~2~^1^H, ^12^C^14^N, ^12^C~2~^2^H, and ^12^C^15^N secondary ions.
+Spatial distribution visualization and quantification of ^15^N-proline and ^2^H-glucose labeling was derived from ^12^C^15^N/^12^C^14^N and ^12^C~2~^2^H/^12^C~2~^1^H ratios, respectively, using an open source plugin to ImageJ, OpenMIMS 3.0 (github.com/BWHCNI/OpenMIMS).
+Ratio images are presented visually as hue saturation intensity images, with the blue lower bound of the scale set to the natural background ratio and the red upper bound of the scale set such that within tissue labeling differences are visually apparent.
+Changes to the scaling modify the visual appearance but not affect the quantitative data.
+Quantification of isotopic enrichment was based on a pixel ratio analysis of several MIMS images acquired per animal.
+Specifically, image masks were generated by Otsu thresholding the ^12^C^14^N to exclude background regions.
+The ratio of "heavy" to "light" ion intensities in tissue regions was calculated.
+Pixel ratios equal to zero were excluded.
+The average of the remaining pixel ratios was then averaged to provide a mean tissue isotope enrichment value for the image.
+
+## VB253
+
+VB253 was developed by Vettore, LLC as a more potent small molecule inhibitor of MCT4.
+Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC.
+These experiments were completed independently of the other studies presented in the manuscript and offer independent validation of the antifibrotic effects of MCT4 inhibition in cells and mice.
+
+*Nathan, Mark, and Jared.
+Add the names of the CROs that did the experiments?
+I think the aSMA/Smad3/nuclei experiments will need a description.
+The others are pretty similar to what is described above.*
+
+\newpage
+
+# References and Notes
+
+::: {#refs}
+:::
 
@@
 S3.
-Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
-(**B**) Extracellular glucose was determined following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
+Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined 48 h following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+(**B**) Extracellular glucose was determined 48 h following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
 (**C**) Summary data from Seahorse extracellular flux analysis.

### 2024-04-16 R/figures.R

@@ -229,6 +229,76 @@ make_fig07 <- function(p1, p2, p3) {
     )
 }
 
+make_fig08 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
+  # design <- "abcc \n defg \n hijk"
+  # p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + p10 + p11 +
+  #   theme_patchwork(
+  #     design = design,
+  #     widths = ggplot2::unit(1.5, "in"),
+  #     heights = ggplot2::unit(1.5, "in")
+  #   )
+
+  top <-
+    ((p1 + p2) +
+       patchwork::plot_layout(guides = "collect") &
+       ggplot2::theme(
+         legend.position = "bottom",
+         legend.box.margin = ggplot2::margin(t = -10)
+       )) -
+    p3 +
+    theme_patchwork(
+      # design = "abc",
+      widths = c(2, 2),
+      tags = list(c("A", "B", "C")),
+      guides = "collect"
+    ) &
+    ggplot2::theme(legend.position = "bottom")
+
+  mid <-
+    p4 + p5 + p6 + p7 +
+    patchwork::plot_annotation(
+      title = "Young mice",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
+    ) +
+    theme_patchwork(
+      design = "abcd",
+      tags = list(c("D", "E", "F", ""))
+    )
+
+  bot <-
+    p8 + p9 + p10 + p11 +
+    patchwork::plot_annotation(
+      title = "Aged mice",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          size = ggplot2::rel(0.8),
+          face = "bold"
+        )
+      )
+    ) +
+    theme_patchwork(
+      design = "abcd",
+      tags = list(c("G", "H", "", "I"))
+    )
+
+  patchwork::wrap_elements(full = top) +
+    patchwork::wrap_elements(full = mid) +
+    patchwork::wrap_elements(full = bot) +
+    theme_patchwork(
+      design = "a \n b \n c",
+      widths = ggplot2::unit(7.5, "in"),
+      heights = ggplot2::unit(2, "in"),
+      tags = NULL
+    )
+}
[…]

### 2024-04-16 R/plots.R

@@ -534,7 +534,7 @@ plot_msea_table <- function(
     dplyr::select("pathway", "NES") |>
     dplyr::mutate(
       pathway = stringr::str_replace(.data$pathway, " - Homo.*$", ""),
-      pathway = stringr::str_replace(.data$pathway, "_", "")
+      pathway = stringr::str_replace_all(.data$pathway, "_", " ")
     )
 
   y <-

### 2024-04-25 manuscript/supplement.qmd
updates in response to co-author feedback

@@
   affiliations: 
+  - ref: bwh
   - ref: mgh
+  - ref: hms
+- name: Fei Li
+  affiliations:
   - ref: bwh
   - ref: hms
+- name: K. Mark Parnell
+  affiliations:
+  - ref: vettore
 - name: Nathan M. Krah
@@
   - ref: utah_gen
-- name: Kevin Leahy 
+  - ref: utah_biochem
+- name: Kevin J. Leahy 
   affiliations:
   - ref: bwh
-- name: K. Mark Parnell
+- name: Christelle Guillermier
   affiliations:
-  - ref: vettore
+  - ref: bwh
+  - ref: hms
 - name: Jack Varon
@@
   - ref: hms
+- name: Bradley A. Maron
+  affiliations:
+  - ref: umd
+  - ref: ihc
 - name: Nancy J. Philp
@@
 affiliations:
+- id: bwh
+  name: Department of Medicine, Brigham and Women's Hospital, Boston, MA
 - id: mgh
   name: Department of Medicine, Massachusetts General Hospital, Boston, MA
-- id: bwh
-  name: Department of Medicine, Brigham and Women's Hospital and Harvard Medical School, Boston, MA
 - id: hms
   name: Department of Medicine, Harvard Medical School, Boston, MA
+- id: vettore
+  name: Vettore Biosciences, San Francisco, CA
 - id: utah_gen
   name: Department of Internal Medicine, University of Utah, Salt Lake City, UT
-- id: vettore
-  name: Vettore Biosciences, San Francisco, CA
+- id: utah_biochem
+  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
+- id: umd
+  name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MD
+- id: ihc
+  name: University of Maryland Institute for Health Computing, Bethesda, MD
 - id: tj
@@
   name: Department of Pathology, Massachusetts General Hospital, Boston, MA
-- id: utah_biochem
-  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
+- id: pitt
+  name: Aging Institute, University of Pittsburgh, Pittsburgh, PA
+- id: upmc
+  name: UPMC Heart and Vascular Institute, UPMC Presbyterian, Pittsburgh, PA
 - id: utah_hhmi
@@
 
-Recombinant human TGF-β1 was purchased commercially , dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
+Recombinant human TGFβ was purchased commercially , dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
 AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
@@
 Passages 3-8 were used for experiments.
-Myofibroblast differentiation was induced *in vitro* using recombinant human TGF-β1 (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
+Myofibroblast differentiation was induced *in vitro* using recombinant human TGFβ (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
 
@@
 VB253 was developed by Vettore, LLC as a more potent small molecule inhibitor of MCT4.
-Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC.
+Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC, as described in detail below.
 These experiments were completed independently of the other studies presented in the manuscript and offer independent validation of the antifibrotic effects of MCT4 inhibition in cells and mice.
 
-*Nathan, Mark, and Jared.
-Add the names of the CROs that did the experiments?
-I think the aSMA/Smad3/nuclei experiments will need a description.
-The others are pretty similar to what is described above.*
+All cell culture experiments were conducted by Charles River Laboratories (Wilmington, MA, USA).
+Lung fibroblasts from three independent IPF donors were seeded at a density of 750 cells/well in 384-well plates.
+Fibroblasts were subsequently treated with medium with 0.1% DMSO (vehicle control), 1.25 ng/ml TGFβ alone, 1.25 ng/ml TGFβ in the presence of 1 μM SB525334 (positive control), or 1.25 ng/ml TGFβ with 8-point concentration response curves (CRC) in biologic duplicate with semi-log dilutions of nintedanib or VB253.
+The top concentration tested was 10 μM.
+Cells were fixed with 4% paraformaldehyde 72 h post treatment with TGFβ.
+The principal readout was immunofluorescence staining to for α-SMA and DAPI to quantify the proportion of α-SMA-positive cells.
+
+SMAD3 nuclear translocation was evaluated in lung fibroblasts from one IPF donor.
+Cells were plated on Purecol-coated 96-well plates.
+Unstimulated (medium only) and stimulated cells (0.5 ng/ml TGFβ) were treated with 0.1% DMSO (vehicle control), 1 μM SB525334 (positive control), nintedanib in 8-point CRC, ALK5 inhibitor (SB525334) in 8-point CRC, and VB253 in a 16-point CRC, with a maximum concentration of 30 μM.
+All cells were fixed 2 h after treatment.
+The readout was immunofluorescence staining for pSMAD3 and DAPI.
+
+The *in vivo* therapeutic evaluation of VB253 in bleomycin-induced pulmonary fibrosis in both young and aged mice was conducted by Aragen Biosciences (Morgan Hill, CA, USA).
+In two separate experiments, C57Bl/6 mice, aged either 6-7 weeks or 68-70 weeks, were obtained from Jackson Labs (Bar Harbor, ME, USA).
+Under anesthesia, mice were administered either PBS vehicle (N = 10 per experiment) or 1.5 U/kg bleomycin in PBS intratracheally.
+After 7 days, mice treated with bleomycin were randomized to receive either vehicle (0.5% methylcellulose in saline) (N = 10), VB253 dissolved in 0.5% methylcellulose (3 mg/kg BID), Pirfenidone (Glentham Life Sciences Lid.
+Cat #GP4948, 100 mg/kg BID), or Nintedinib (Axon MedChem Cat#2648, 50 mg/kg, daily) (N = 10 per group).
+Mice injected with saline, rather than bleomycin, were administered vehicle.
+Treatments were administered daily until the experiment was terminated 21 days after bleomycin instillation, approximately 2-4 h after the final treatment dose.
+Whole-body plethysmography was performed on day 20 using Buxco WBP instrument system measuring Enhanced Pause (Penh) and respiratory rate on unrestrained animals.
+Upon experiment termination, lungs were inflated with 10% neutral buffered formalin (NBF) and fixed in 10% NBF for histology.
+
+Histology, imaging, and pathologic analyses were performed by HistoTox Labs (Boulder, CO, USA).
+Multiple slides per block were sectioned at 5 μm and stained with hematoxylin and eosin (H&E) or by routine immunohistochemistry (IHC) methods for α-SMA.
+H&E-stained glass slides were evaluated using light microscopy by an ACVP board-certified veterinary pathologist.
+Lung sections were scored according to the modified Ashcroft scale.
+Briefly, scores for five representative 200× microscopic fields per sample were averaged to obtain a mean score for each animal.
+α-SMA-stained glass slides were scanned using an Aperio AT2 whole slide scanner.
+Whole slide images were annotated to delineate regions of interest (ROI).
+Exclusions were applied to remove tissue artifacts (folds, tears, non-specific staining), large airways, and blood vessels.
+Visiopharm (VIS) image analysis software, using imaging filters to separate positive staining from counterstaining and background were applied to quantify α-SMA signal within each sample.
+Differences in expression levels between samples and treatments were evaluated by calculating the percent positive detection of cells or area within the viable lung tissue.
 

### 2024-04-25 manuscript/manuscript.qmd
updates in response to co-author feedback

@@
   - ref: utah_gen
+  - ref: utah_biochem
 - name: Kevin J. Leahy 
@@
   name: Department of Internal Medicine, University of Utah, Salt Lake City, UT
+- id: utah_biochem
+  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
 - id: umd
-  name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MA
+  name: Department of Medicine, University of Maryland School of Medicine, Baltimore, MD
 - id: ihc
@@
   name: UPMC Heart and Vascular Institute, UPMC Presbyterian, Pittsburgh, PA
-- id: utah_biochem
-  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
 - id: utah_hhmi
@@
 
-Small molecule inhibitors of lactate transporters, including the novel MCT4 inhibitor VB253, reprogram fibroblast metabolism to prevent myofibroblast differentiation and to decrease bleomycin-induced pulmonary fibrosis.
+Small molecule inhibitors of lactate transporters, including the novel MCT4 inhibitor VB253, reprogram fibroblast metabolism to prevent myofibroblast differentiation and decrease bleomycin-induced pulmonary fibrosis.
 
@@
 In this study, we investigated how MCT1 and MCT4, key lactate transporters, influence myofibroblast differentiation and experimental pulmonary fibrosis.
-Our findings reveal that inhibiting MCT1 or MCT4 reduces TGFβ-stimulated myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
+Our findings reveal that inhibiting MCT1 or MCT4 reduces TGFβ-stimulated pulmonary myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
 Through comprehensive metabolic analyses, including bioenergetics, stable isotope tracing, metabolomics, and imaging mass spectrometry in both cells and mice, we demonstrate that inhibiting lactate transport enhances oxidative phosphorylation, reduces reactive oxygen species production, and diminishes glucose metabolite incorporation into fibrotic lung regions.
-Furthermore, we introduce VB253, a novel MCT4 inhibitor, which effectively ameliorated pulmonary fibrosis in both young and aged mice, comparable to established antifibrotic therapies.
-These results underscore the necessity of lactate transport for myofibroblast differentiation, identify MCT1 and MCT4 as promising pharmacologic targets in pulmonary fibrosis, and support further evaluation of lactate transport inhibitors for patients with pulmonary fibrosis for whom limited therapeutic options currently exist.
+Furthermore, we introduce VB253, a novel MCT4 inhibitor, which ameliorates pulmonary fibrosis in both young and aged mice, with comparable efficacy to established antifibrotic therapies.
+These results underscore the necessity of lactate transport for myofibroblast differentiation, identify MCT1 and MCT4 as promising pharmacologic targets in pulmonary fibrosis, and support further evaluation of lactate transport inhibitors for patients for whom limited therapeutic options currently exist.
 
@@
 
-Accumulating evidence suggests that metabolic reprogramming may be one such therapeutic vulnerability in IPF [@bueno2020; @selvarajah2021].
+Accumulating evidence suggests that metabolic reprogramming may be one such therapeutic strategy in IPF [@bueno2020; @selvarajah2021].
 Lung fibrosis is driven by the excessive deposition of extracellular matrix by myofibroblasts [@martinez2017].
@@
 Toward this end, we aimed to examine the impact of a novel metabolic strategy --- lactate transport inhibition --- on myofibroblast differentiation and experimental pulmonary fibrosis.
-Myofibroblasts rely on lactate secretion to sustain glycolysis, a process regulated by a family of monocarboxylate transporters (MCT1-4).
+Sustained glycolysis in myofibroblasts relies on lactate secretion, which is conducted by a family of monocarboxylate transporters (MCT1-4).
 Inhibitors targeting these transporters have been actively explored in clinical trials for oncological conditions where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
@@
 Using metabolomics, stable isotope tracing, and high-resolution spatial metabolomic imaging, we find that lactate transport inhibition promotes glucose oxidation and decreases pro-fibrogenic reactive oxygen species (ROS) production.
-Building on these results, we introduce VB253, a novel MCT4 inhibitor suitable for human clinical studies.
+Building on these results, we introduce VB253, a novel MCT4 inhibitor suitable for human clinical studies that performs similarly to current standard-of-care antifibrotic therapies in bleomycin-induced pulmonary fibrosis.
 Together, our findings offer insights into disrupting the metabolic pathways driving IPF fibrosis and present a promising strategy targeting lactate transporters for the treatment of this fatal condition.
@@
 After 24 h, the cells were treated with TGFβ for 48 h to induce myofibroblast differentiation.
-The siRNAs effectively reduced lactate transporter protein levels (**Fig.
+The siRNAs reduced lactate transporter protein levels (**Fig.
 2A**).
@@
 
-MCTs play pivotal roles in maintaining cellular lactate homeostasis.
-MCT1 serves as the primary lactate importer in cells utilizing lactate for oxidative phosphorylation or gluconeogenesis and exhibits ubiquitous expression.
-Additionally, MCT1 facilitates lactate export in glycolytic cells [@halestrap2013; @brooks2018].
-By contrast, MCT4 functions as the main lactate exporter in glycolytic cells and is up-regulated by hypoxia-inducible transcription factors [@ullah2006; @halestrap2013].
+MCTs play pivotal roles in maintaining cellular lactate and redox homeostasis.
+MCT1 predominantly imports lactate in cells utilizing lactate for oxidative phosphorylation or gluconeogenesis and is ubiquitously expressed.
+MCT1 also facilitates lactate export in some glycolytic cells [@halestrap2013; @brooks2018].
+By contrast, MCT4 functions as the main lactate exporter in glycolytic cells and is up-regulated when the glycolytic transcriptional program is activated by, for example, the c-Myc or hypoxia-inducible transcription factors [@ullah2006; @halestrap2013].
 Importantly, MCT4 can also act as a lactate importer with a *K*~M~ of 1 mM [@contreras-baeza2019].
@@
 To assess the metabolic consequences of lactate transporter inhibition, we quantified extracellular lactate in the conditioned medium from cells treated with MCT siRNA or pharmacologic inhibitors.
-Silencing MCT1 and MCT4, either individually or concurrently, decreased TGFβ-stimulated lactate efflux in lung fibroblasts (**Fig.
+Silencing MCT1 and MCT4, either individually or concurrently, decreased net TGFβ-stimulated lactate efflux in lung fibroblasts (**Fig.
 3A**).
@@
 Consistent with our direct measures of extracellular lactate and glucose, concurrent treatment with AZD3965 and VB124 was necessary to reduce PER.
-Consequently, this reduction in PER coincided with a corresponding increase in OCR as myofibroblasts transitioned their metabolism from glycolysis to oxidative phosphorylation.
+Consequently, this reduction in PER coincided with an increase in OCR as myofibroblasts transitioned their metabolism from glycolysis to oxidative phosphorylation.
 
@@
 
-To provide further support for this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of extracellular and intracellular metabolites from cells treated with AZD3965 and VB124.
+To further test this hypothesis, we performed liquid chromatography-mass spectrometry-based profiling of extracellular and intracellular metabolites from cells treated with AZD3965 and VB124.
 In accordance with the Seahorse analysis, inhibiting a single lactate transporter had modest effects on extracellular metabolite levels (**Fig.
@@
 4E**).
-Similar to the extracellular flux results, the magnitude of drug-induced perturbations was increased from MCT1 to MCT4 to combined inhibition (**Fig.
+Similar to the extracellular flux results, the magnitude of drug-induced perturbations increased from MCT1 to MCT4 to combined inhibition (**Fig.
 4E**), reflected by intracellular lactate levels (**Fig.
 4F**).
-As expected from extracellular lactate measures following MCT inhibition, intracellular lactate accumulation moderately increased with MCT4 inhibition and substantially with MCT1/4 inhibition.
+As expected from extracellular lactate measures following MCT inhibition, intracellular lactate accumulated moderately with MCT4 inhibition and substantially with MCT1/4 inhibition.
 TGFβ treatment resulted in diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Fig.
@@
 4J, S5**).
-TGFβ increases ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), indicated enhanced flux from glucose into the TCA cycle.
-While AZD3965 had minimal impact on these labeling patterns, MCT4 inhibition significantly elevated the fractions of these metabolites labeled by glucose, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the bioenergetic and steady-state metabolomics experiments described above.
+TGFβ increased ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), indicating enhanced flux from glucose into the TCA cycle.
+While AZD3965 had minimal impact on these labeling patterns, MCT4 inhibition significantly elevated the fractions of these metabolites labeled by ^13^C, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the results of our bioenergetic and steady-state metabolomics experiments described above.
 
-Increased metabolite labeling by glucose must by offset by decreased labeling from other substrates.
+Increased metabolite labeling from [U-^13^C~6~]-glucose must be offset by decreased labeling from other substrates.
 Given recent data suggesting lactate as a major oxidative fuel source in the lung [@hui2017; @hui2020; @copeland2023], we hypothesized that MCT inhibition would decrease exogenous lactate oxidation.
@@
 4J**).
-MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C]-glutamine (**Fig.
+MCT inhibition had no impact on fractional labeling of TCA intermediates by [U-^13^C~5~]-glutamine (**Fig.
 S4M, S5**).
-Interestingly, MCT4 inhibition alone or in combination with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate, underscoring the importance of bi-directional lactate transport mediated by MCTs.
-Consistent with our measurements of extracellular lactate, the effects of MCT inhibition were more pronounced when both inhibitors were used simultaneously, again highlighting functional redundancy of MCT1 and MCT4 in lung fibroblasts.
+MCT4 inhibition alone or in combination with AZD3965 decreased ^13^C labeling of intracellular metabolites by lactate.
+These findings demonstrate the importance of MCT4 for lactate import at physiologic lactate concentrations, in contrast to the prevailing sentiment that MCT4 is primarily a lactate exporter.
+Moreover, these data further underscore the relatively greater importance of MCT4 activity in fibroblast metabolism compared to MCT1.
+Consistent with our measurements of extracellular lactate, the effects of MCT inhibition were more pronounced when both inhibitors were used simultaneously, again highlighting some functional redundancy of MCT1 and MCT4 in these cells.
 
@@
 Given that increased intracellular lactate correlated with more potent inhibition of α-SMA expression, we treated cells with TGFβ in combination with extracellular lactate (10 mM).
-A previous study suggested that extracellular lactate modestly increased α-SMA expression independently of TGFβ [@kottmann2012].
+A previous study suggested that extracellular lactate modestly increases α-SMA expression independently of TGFβ [@kottmann2012].
 However, we observed no impact of extracellular lactate on α-SMA expression in TGFβ-treated lung fibroblasts (**Fig.
@@
 Our preclinical findings indicate that MCT4 inhibition, as a single therapeutic target, exhibits greater antifibrotic efficacy than MCT1 inhibition, both in suppressing myofibroblast differentiation *in vitro* and reducing bleomycin-induced fibrosis.
-VB253 is a novel inhibitor of MCT4 with similar selectivity, but ~10-fold increased potency, for MCT4 inhibition.
-Like VB124, VB253 dose-dependently decreases TGFβ-stimulated α-SMA in human IPF lung fibroblasts *ex vivo* (**Fig.
+VB253 is a novel inhibitor of MCT4 with ~10-fold increased selectivity for MCT4 *v*.
+MCT1 and ~10-fold increased potency for MCT4 inhibition.
+Like VB124, VB253 dose-dependently decreased TGFβ-stimulated α-SMA in human IPF lung fibroblasts *ex vivo* (**Fig.
 8A**).
@@
 Nintedanib mitigates myofibroblast differentiation partly by inhibiting TGFβ receptor phosphorylation and Smad-dependent signaling pathways [@rangarajan2016].
-Correspondingly, nintedanib dose-dependently attenuated TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 have no impact on this signaling pathway (**Fig.
+Correspondingly, nintedanib dose-dependently attenuated TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 had no impact on this signaling pathway (**Fig.
 5, S10**).
@@
 8C**).
-Nintedanib and pirfenidone served as comparators to existing pulmonary fibrosis therapies.
+The existing pulmonary fibrosis therapies, nintedanib and pirfenidone, served as comparators.
 Three weeks after bleomycin, whole-body plethysmography was performed to assess breathing patterns in unrestrained mice.
@@
 Inhibiting these transporters mitigated TGFβ-stimulated myofibroblast differentiation and attenuated the severity of bleomycin-induced pulmonary fibrosis without a demonstrable impact on classical TGFβ receptor signaling pathways.
-Metabolically, MCT antagonists promoted glucose oxidation and while reducing glucose carbon incorporation into fibrotic lung regions, correlating with decreased oxidative stress.
+Metabolically, MCT antagonists promoted glucose oxidation while reducing glucose carbon incorporation into fibrotic lung regions, correlating with decreased oxidative stress.
 MCT4 inhibition consistently exhibited superior antifibrotic potency compared to MCT1 inhibition.
-We introduce a novel MCT4 inhibitor, VB253, with a more favorable pharmacologic profile than VB124 currently undergoing Phase 1 clinical trials [**REF?**].
+We introduce a novel MCT4 inhibitor, VB253, which has a more favorable pharmacologic profile than VB124, and is currently undergoing Phase 1 clinical trials.
 Altogether, these data establish lactate transport as a promising metabolic target for therapeutic intervention in pulmonary fibrosis.
@@
 
-Our data reveal that lactate transport inhibition exploits novel antifibrotic mechanisms distinct from TGFβ signaling pathways and suggest that these metabolic effects signal to antifibrotic transcriptional programs by dampening ROS levels.
+Our data reveal that lactate transport inhibition exploits novel antifibrotic mechanisms distinct from TGFβ signaling pathways and suggest that these metabolic effects may signal to antifibrotic transcriptional programs by dampening ROS levels.
 Increased ROS production has been observed following TGFβ treatment and seems to be essential for myofibroblast differentiation [@jain2013; @schworer2020].
-While we were unable to replicate TGFβ-dependent increases in ROS in our cells and model system, we did observe decreased CellROX oxidation with MCT4 inhibition and decreased MitoSOX oxidation with AZD3965.
+While our cells and model system did not replicate TGFβ-dependent increases in ROS, we observed decreased CellROX oxidation with MCT4 inhibition and decreased MitoSOX oxidation with AZD3965.
 Additional signaling mechanisms may also contribute, such as post-translational modification of protein lysines by lactate (*i.e.*, protein lactylation).
@@
 
-Recently, the contributions of fibroblast metabolic pathways to extracellular matrix production has garnered significant attention [@selvarajah2021].
+Recently, the contributions of fibroblast metabolic pathways to extracellular matrix production have garnered significant attention [@selvarajah2021].
 Serine and glycine synthesis from the glycolytic intermediate 3-phosphoglycerate, and proline synthesis from glutamine, have been implicated in myofibroblast differentiation and pulmonary fibrosis [@bernard2018; @hamanaka2019; @selvarajah2019; @schworer2020; @choudhury2023].
@@
 This finding emphasizes the ability of cells to utilize  circulating proline for protein synthesis, potentially diminishing the significance of *de novo* proline biosynthesis *in vivo*.
-Moreover, the ^2^H-glucose signal provided some direct evidence for MCT-dependent metabolic reprogramming in mice, showing reduced carbohydrate incorporation into matrix proteins.
+Moreover, the ^2^H-glucose signal provided direct evidence for MCT-dependent metabolic reprogramming in mice, showing reduced carbohydrate incorporation into matrix proteins.
 Owing to the financial and time costs of isotope tracing and imaging, we were able to study only a few animals per group.
-Nevertheless, metabolic imaging at high spatial resolution holds great promise for correlating cell identities identified from spatial transcriptomic profiles with metabolic features.
+Nevertheless, metabolic imaging at high spatial resolution holds great promise for correlating cell identities from spatial transcriptomic profiles with metabolic features.
 Furthermore, stable isotopes may be safely administered to human patients prior lung biopsy or explant [@steinhauser2012; @guillermier2017], offering a strategy that could significantly advance our understanding of how cell metabolism contributes to pulmonary fibrosis.
@@
 
-One of the barriers to translating metabolic therapies to humans has been the poor pharmacologic properties of the small molecule laboratory compounds that have been studied previously.
+The poor pharmacologic properties of small molecules studied previously have prevented translation into metabolic therapies for human pulmonary fibrosis.
 Both AZD3965 and VB253 exhibit low IC~50~ values of approximately 2 nM, compared to the next most potent inhibitor studied in pulmonary fibrosis models, lonidamine, with an IC~50~ of 7,000 nM [@yin2019].
-As metabolic targets downstream of glycolysis, lactate transport inhibitors may also offer better tolerance by allowing glycolysis to continue supporting glucose oxidation, contrasting with upstream glycolysis inhibitors that can more severely impact cellular bioenergetics.
+As metabolic targets downstream of glycolysis, lactate transport inhibitors also offer better tolerance by allowing glycolysis to continue supporting glucose oxidation, contrasting with upstream glycolysis inhibitors that more severely impact cellular bioenergetics.
 Pharmacologic interest in lactate transporters has been driven by the recognition of increased lactate transporter expression in a variety of cancers [@payen2020].
@@
 8**).
-MCT4 global knockout mice are viable and breed normally [@bisetto2018; @bisetto2019], raising optimism that VB253 will be well tolerated in humans, and a Phase 1 clinical trial of VB253 is currently underway [**REF?**].
+MCT4 global knockout mice are viable and breed normally [@bisetto2018; @bisetto2019], raising optimism that VB253 will be well tolerated in humans, and a Phase 1 clinical trial of VB253 is currently underway.
 
@@
 
-This study was designed to investigate the role of lactate transporters in the myofibroblast differentiation and pulmonary fibrosis.
+This study was designed to investigate the role of lactate transporters in myofibroblast differentiation and pulmonary fibrosis.
 The objectives of this study were (i) to determine the expression of lactate transporters in IPF lungs and model systems, (ii) to characterize the phenotypic effects of lactate transport inhibition in model systems, and (iii) to profile the metabolic consequences of lactate transport inhibitors.
@@
 Cell culture experiments were performed at least three times.
-The number of animals per experimental group was chosen based on prior publications and were repeated at least once.
+The number of animals per experimental group was chosen based on prior publications and experiments were repeated at least once.
 Stable isotope tracing experiments were performed on three animals per group owing to resource availability.
@@
 The number of unique patient samples was determined by clinical availability.
-
-\newpage
 
@@
 All animal experiments were approved by the Brigham and Women's Hospital IACUC (2020N000199).
+
+\newpage
 
@@
 
-National Institutes of Health grant R01GM987654 (WMO)
-Brigham and Women's Hospital Research Institute (WMO)
+* National Institutes of Health grant R01GM987654 (WMO)
+* Brigham and Women's Hospital Research Institute (WMO)
+* American Lung Association grant DA-827785 (EYK)
+* BWH Bell Family Award (EYK)
+* National Institutes of Health grant R01HL152075 (LH)
+* National Institutes of Health grant T32HL007633 (JV)
 
@@
 
-Conceptualization: KMP, MLS, JR, WMO
-Methodology: DRZ, FL, KMP, CG, JV, BAM, MLS, RSK, JR, WMO
-Investigation: DRZ, FL, KMP, NMK, KJL, CG, JV, LPH, RSK, WMO
-Visualization: WMO
-Funding acquisition: KMP, JR, WMO
-Project administration: KMP, JR, WMO
-Resources: DRZ, KMP, NMK, RMB, NJP, EYK, RSK
-Supervision: KMP, JR, WMO
-Writing - original draft: WMO
-Writing - review and editing: DRZ, FL, KMP, NMK, CG, JV, RMB, BAM, NJP, EYK, MLS, RSK, JR, WMO
+* Conceptualization: KMP, MLS, JR, WMO
+* Methodology: DRZ, FL, KMP, CG, JV, BAM, MLS, RSK, JR, WMO
+* Investigation: DRZ, FL, KMP, NMK, KJL, CG, JV, LPH, RSK, WMO
+* Visualization: WMO
+* Funding acquisition: KMP, JR, WMO
+* Project administration: KMP, JR, WMO
+* Resources: DRZ, KMP, NMK, RMB, NJP, EYK, RSK
+* Supervision: KMP, JR, WMO
+* Writing - original draft: WMO
+* Writing - review and editing: DRZ, FL, KMP, NMK, CG, JV, RMB, BAM, NJP, EYK, MLS, RSK, JR, WMO
 
@@
 
-W.M.O. has received consulting fees from Nikang Therapeutics outside the scope of this research.
-
-J.R. is a consultant and shareholder for Vettore Biosciences.
-
-R.S.K. received a Discovery ILD Award from Boehringer Ingelheim and received support through the Partners Drug Development Lab from Bayer Pharmaceuticals, all outside the scope of this research.
-
-B.A.M. has received consulting fees from Actelion and Tenax and has performed investigator-initiated research with support from Deerfield, all outside the scope of this research.
-
-The remaining authors declare that they have no competing interests.
-
-*Here, also include any awarded or filed patents pertaining to the results presented in the paper, stating patent filing number and title and which authors are co-inventors*
+* W.M.O. has received consulting fees from Nikang Therapeutics outside the scope of this research. 
+* J.R. is a consultant and shareholder for Vettore Biosciences. 
+* R.S.K. received a Discovery ILD Award from Boehringer Ingelheim and received support through the Partners Drug Development Lab from Bayer Pharmaceuticals, all outside the scope of this research.
+* E.Y.K. received unrelated research funding from Bayer AG, Roche Pharma Research and Early Development, and 10X Genomics. E.Y.K. has a financial interest in Novartis AG unrelated to this work.
+* B.A.M. has received consulting fees from Actelion and Tenax and has performed investigator-initiated research with support from Deerfield, all outside the scope of this research. 
+* L.P.H. reports grants from Boehringer Ingelheim Pharmaceuticals, Inc. (BIPI) and has received personal consulting fees from BIPI, Pliant Therapeutics, Clario, and Abbvie Pharmaceuticals.
+* The remaining authors declare that they have no competing interests.
 

### 2024-04-25 manuscript/supplement.qmd
small edits to Supplement

@@
 
-For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on day 21 after bleomycin.
+For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on either day 21 or day 14 after bleomycin.
 Perfused and inflated lungs were fixed with formaldehyde/glutaraldehyde 2.5% in sodium cacodylate buffer, pH 7.4, embedded in EPON, section to 0.5 μm, mounted on silicon wafers, and gold coated.
@@
 All cells were fixed 2 h after treatment.
-The readout was immunofluorescence staining for pSMAD3 and DAPI.
+The readout was immunofluorescence staining for pSmad3 and DAPI.
 

### 2024-04-25 README.Rmd
add draft README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Ziehr.2023.ipf.mcti

<!-- badges: start -->
<!-- badges: end -->

### Author

William Oldham (woldham@bwh.harvard.edu)

### Contents

This research compendium contains all data, code, and text associated with our publication:

> **Lactate transport inhibition therapeutically reprograms fibroblast metabolism in experimental pulmonary fibrosis.** David R. Ziehr, Fei Li, K. Mark Parnell, Nathan M. Krah, Kevin J. Leahy, Christelle Guillermier, Jack Varon, Rebecca M. Baron, Bradley A. Maron, Nancy J. Philp, Lida P. Hariri, Edy Y. Kim, Matthew L. Steinhauser, Rachel S. Knipe, Jared Rutter, and William M. Oldham. *bioRxiv*

### Licenses

| **Text / Figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
| **Code :** See the [DESCRIPTION](DESCRIPTION) file.
| **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/). Attribution requested in reuse.

### 2024-04-25 manuscript/manuscript.qmd
fix grant support and add email addresses

@@
   - ref: hms
+  email: dziehr@mgb.org
 - name: Fei Li
@@
   - ref: hms
+  email: feli@bwh.harvard.edu
 - name: K. Mark Parnell
@@
   - ref: vettore
+  email: markparnell1@gmail.com
 - name: Nathan M. Krah
@@
   - ref: utah_biochem
+  email: Nathan.Krah@hci.utah.edu
 - name: Kevin J. Leahy 
@@
   - ref: bwh
+  email: kjleahy@bwh.harvard.edu
 - name: Christelle Guillermier
@@
   - ref: hms
+  email: cguillermier@bwh.harvard.edu
 - name: Jack Varon
@@
   - ref: hms
+  email: varon.jack@gmail.com
 - name: Rebecca M. Baron
@@
   - ref: hms
+  email: rbaron@bwh.harvard.edu
 - name: Bradley A. Maron
@@
   - ref: ihc
+  email: BMaron@som.umaryland.edu
 - name: Nancy J. Philp
@@
   - ref: tj
+  email: Nancy.Philp@jefferson.edu
 - name: Lida P. Hariri
@@
   - ref: mgh_path
+  email: lhariri@mgh.harvard.edu
 - name: Edy Y. Kim
@@
   - ref: hms
+  email: ekim11@bwh.harvard.edu
 - name: Matthew L. Steinhauser
@@
   - ref: upmc
+  email: msteinhauser@pitt.edu
 - name: Rachel S. Knipe
@@
   - ref: hms
+  email: rknipe@mgh.harvard.edu
 - name: Jared Rutter
@@
   - ref: utah_hhmi
+  email: rutter@biochem.utah.edu
 - name: William M. Oldham
@@
 
-* National Institutes of Health grant R01GM987654 (WMO)
+* National Institutes of Health grant R01HL167718 (WMO)
 * Brigham and Women's Hospital Research Institute (WMO)

### 2024-05-01 analysis/bcecf.qmd
add BCECF data and analysis

---
title: "BCECF"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
    toc: true
    toc-depth: 3
    toc-location: left
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/blots.qmd")

tar_config_set(store = here("_targets"))
```

## Experiment

Monocarboxylate transporters co-transport lactate and H^+^. We measured the effects of lactate transport inhibition on intracellular pH by loading cells with the pH-sensitive dye BCECF-AM. For these experiments, 8,000 LFs were seeded in 96-well plates. They were serum starved for 24 h prior to treatment with TGFβ and lactate transport inhibitors for 48 h. After 48 h, cells were loaded with 400 μM BCECF-AM in HBSS for 30 min in the presence of lactate transport inhibitors. Medium was then replaced with basal medium containing MCT inhibitors. Fluorescence was measured using excitation 440 and 490 nm with emission wavelength 535 nm.

## Results

```{r}
#| label: plot

tar_read(bcecf_plot)
```

### 2024-05-01 R/bcecf.R

@@ -1,0 +1,138 @@
+# bcecf.R
+
+format_bcecf <- function(x) {
+  x |>
+    tidyr::pivot_longer(
+      `1`:tidyselect::last_col(),
+      names_to = "col",
+      values_to = "value",
+      names_transform = list(col = as.numeric)
+    ) |>
+    dplyr::mutate(
+      type = dplyr::case_when(
+        col %in% 1:2 ~ "blank",
+        col %in% 3:10 ~ "sample",
+        col %in% 11:12 ~ "standard"
+      ),
+      condition = dplyr::case_when(
+        col %in% 3:6 ~ "Ctl",
+        col %in% 7:10 ~ "TGFβ"
+      ),
+      condition = factor(condition, levels = c("Ctl", "TGFβ")),
+      treatment = dplyr::case_when(
+        col %in% c(3, 7) ~ "Veh",
+        col %in% c(4, 8) ~ "AZD",
+        col %in% c(5, 9) ~ "VB",
+        col %in% c(6, 10) ~ "AZD/VB"
+      ),
+      treatment = factor(treatment, levels = c("Veh", "AZD", "VB", "AZD/VB")),
+      group = dplyr::case_when(
+        col == 1 ~ "blank",
+        col == 2 ~ "cells",
+        col %in% 3:10 ~ stringr::str_c(condition, ".", treatment),
+        col == 11 & row %in% 1:4 ~ "4.5",
+        col == 11 & row %in% 5:8 ~ "5.5",
+        col == 12 & row %in% 1:4 ~ "6.5",
+        col == 12 & row %in% 5:8 ~ "7.5"
+      ),
+      .before = value
+    )
+}
+
+clean_bcecf <- function(x) {
+  x |>
+    dplyr::filter(col %in% 2:10) |>
+    dplyr::group_by(date, excitation) |>
+    dplyr::mutate(value = value - mean(value[group == "cells"])) |>
+    dplyr::filter(type != "blank") |>
+    dplyr::group_by(date, row, col, condition, treatment, group) |>
+    dplyr::summarize(ratio = value[excitation == 490] / value[excitation == 440]) |>
+    dplyr::group_by(date, condition, treatment, group) |>
+    wmo::remove_nested_outliers(ratio, remove = TRUE) |>
+    dplyr::summarise(ratio = mean(ratio)) |>
+    dplyr::select(group = date, condition, treatment, ratio) |>
+    dplyr::ungroup()
+}
+
[…]

### 2024-05-01 _targets.R

@@ -634,6 +634,34 @@ list(
     NULL
   ),
 
+  # bcecf -------------------------------------------------------------------
+
+  tar_target(
+    bcecf_file,
+    raw_data_path("bcecf.csv"),
+    format = "file_fast"
+  ),
+  tar_target(
+    bcecf_data,
+    readr::read_csv(bcecf_file, show_col_types = FALSE)
+  ),
+  tar_target(
+    bcecf_raw,
+    format_bcecf(bcecf_data)
+  ),
+  tar_target(
+    bcecf_clean,
+    clean_bcecf(bcecf_raw)
+  ),
+  tar_target(
+    bcecf_stats,
+    twofactor(bcecf_clean, "ratio", comps = comps_azd_2)
+  ),
+  tar_target(
+    bcecf_plot,
+    plot_bcecf(bcecf_clean, bcecf_stats)
+  ),
+
   # seahorse ----------------------------------------------------------------
 
   tar_target(
@@ -1893,6 +1921,10 @@ list(
   tar_quarto(
     mims_analysis,
     path = "analysis/mims.qmd"
+  ),
+  tar_quarto(
+    bcecf_analysis,
+    path = "analysis/bcecf.qmd"
   ),
 
   # figure images -----------------------------------------------------------

### 2024-07-24 analysis/bcecf.qmd
fix a lot of problems and get running

@@
     embed-resources: true
-    echo: false
-    warning: false
+    html-math-method: katex
+    theme: lumen
+    df-print: kable
     toc: true
@@
     toc-location: left
+execute:
+  echo: false
+  warning: false
+  error: false
 ---

### 2024-07-24 analysis/mims.qmd
fix a lot of problems and get running

@@
 date-modified: today
-format: html
-excecute:
-  message: false
+format: 
+  html:
+    embed-resources: true
+    html-math-method: katex
+    theme: lumen
+    df-print: kable
+    toc: true
+    toc-depth: 3
+    toc-location: left
+execute:
+  echo: false
+  warning: false
+  error: false
 ---

### 2024-07-24 manuscript/manuscript.qmd
fix a lot of problems and get running

@@
 Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-(**D**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
+(**D**) KEGG pathways significantly enriched (p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by normalized enrichment score (NES).
 Positive NES indicates enrichment in AZD/VB-treated cells while negative NES indicates enrichment in Vehicle-treated cells.
@@
 Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-(**H**) KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by NES.
+(**H**) KEGG pathways significantly enriched (p-value < 0.1) with metabolites differentially regulated by combined MCT1 and MCT4 inhibition ordered by NES.
 Positive NES indicates enrichment in AZD/VB-treated cells while negative NES indicates enrichment in Vehicle-treated cells.

### 2024-07-24 manuscript/supplement.qmd
fix a lot of problems and get running

@@
 Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
+KEGG pathways significantly enriched (p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
 Positive NES indicates enrichment in treated cells while negative NES indicates enrichment in Vehicle-treated cells.

### 2024-07-24 R/bcecf.R

@@ -66,6 +66,15 @@ plot_bcecf <- function(df, stats) {
       color = condition,
       fill = condition
     ) +
+    ggbeeswarm::geom_quasirandom(
+      pch = 1,
+      color = "black",
+      width = 0.05,
+      size = 0.5,
+      stroke = 0.25,
+      dodge.width = 0.8,
+      show.legend = FALSE
+    ) +
     ggplot2::geom_errorbar(
       ggplot2::aes(color = .data$condition),
       stat = "summary",
@@ -83,15 +92,6 @@ plot_bcecf <- function(df, stats) {
       size = 1.5,
       color = "white",
       position = ggplot2::position_dodge(width = 0.8),
-      show.legend = FALSE
-    ) +
-    ggbeeswarm::geom_quasirandom(
-      pch = 1,
-      color = "black",
-      width = 0.05,
-      size = 0.5,
-      stroke = 0.25,
-      dodge.width = 0.8,
       show.legend = FALSE
     ) +
     ggplot2::geom_text(

### 2024-07-24 R/blots.R

@@ -97,7 +97,7 @@ analyze_blot <- function(df, paired, comp = NULL) {
   }
 
   if (length(unique(df$treatment)) == 1) {
-    out <- ratio_ttest(df, "norm", paired)
+    out <- ratio_ttest(df, "norm")
   } else {
     out <- twofactor(df, "norm", mixed = paired, comp)
   }

### 2024-07-24 R/stats.R

@@ -19,10 +19,10 @@ annot_stats <- function(stats) {
 }
 
 
-ratio_ttest <- function(df, y, paired){
+ratio_ttest <- function(df, y){
   fo <- stats::formula(paste("log(", y, ") ~ condition"))
 
-  stats::t.test(fo, data = df, paired = paired) |>
+  stats::t.test(fo, data = df) |>
     broom::tidy() |>
     dplyr::select("p.value") |>
     dplyr::mutate(
@@ -32,13 +32,14 @@ ratio_ttest <- function(df, y, paired){
     annot_stats()
 }
 
+
 group_ratio_ttest <- function(df, y, ...) {
   cols <- c("protein", "name", "rate", "stage", "stats")
 
   df |>
     tidyr::nest() |>
     dplyr::mutate(
-      stats = purrr::map(.data$data, ratio_ttest, y = y, ...)
+      stats = purrr::map(.data$data, ratio_ttest, y = y)
     ) |>
     dplyr::select(tidyselect::any_of(cols)) |>
     tidyr::unnest(c("stats"))

### 2024-07-24 _targets.R

@@ -182,9 +182,9 @@ list(
         NULL
       ),
       comp = list(
-        FALSE,
-        FALSE,
-        FALSE,
+        NULL,
+        NULL,
+        NULL,
         rlang::sym("comps_sirna"),
         rlang::sym("comps_azd_2"),
         rlang::sym("comps_ar_2"),

### 2024-07-24 analysis/km.R

@@ -1,0 +1,56 @@
+# survival analysis
+
+
+# setup -------------------------------------------------------------------
+
+library(tidyverse)
+library(lubridate)
+library(survival)
+library(survminer)
+conflict_prefer("filter", "dplyr")
+
+
+# data format -------------------------------------------------------------
+
+df <-
+  read_csv("~/Dropbox (Partners HealthCare)/lactate_detritus/km.csv") %>% 
+  mutate(
+    time = as.numeric(date - as_date("2021-07-29")), 
+    event = if_else(event == "died", 1, 0)
+  )
+
+
+# early deaths ------------------------------------------------------------
+
+df <- 
+  df %>% 
+  mutate(
+    treatment = replace(treatment, time < 8 & disease == "Bleo", "Control"), 
+    treatment = factor(treatment, levels = c("Control", "AZD3965", "VB124")), 
+    disease = factor(disease, levels = c("Control", "Bleo")), 
+    group = interaction(disease, treatment),
+  ) %>% 
+  filter(disease == "Bleo")
+
+
+# survival analysis -------------------------------------------------------
+
+f1 <- survfit(Surv(time, event) ~ 1, data = df)
+ggsurvplot(
+  fit = f1, 
+  xlab = "Days", 
+  ylab = "Survival Probability"
+)
+
+f2 <- survfit(Surv(time, event) ~ (disease + treatment), data = df)
+ggsurvplot(
+  fit = f2, 
+  xlab = "Days", 
+  ylab = "Survival Probability"
+)
+
+
+# statistics --------------------------------------------------------------
+
+pairwise_survdiff(Surv(time, event) ~ treatment, data = df)
+coxph(Surv(time, event) ~ treatment, data = df)

### 2024-07-31 manuscript/manuscript.qmd
add pcls data

@@
 Taken together, these data demonstrate substantially decreased pulmonary fibrosis severity after 14 days of treatment with lactate transport inhibitors.
+Similarly, dual MCT inhibition decreased TGFβ-stimulated extracellular matrix protein production in human precision-cut lung slices (**Fig.
+6G**).
 
@@
 6.
-Lactate transporter inhibition attenuates bleomycin-induced pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
+Lactate transporter inhibition attenuates experimental pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
 Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21.
@@
 (**F**) Right lung hydroxyproline content (N = 10-13).
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_06.png)
+(**G**) Human precision-cut lung slices were stimulated with TGFβ (5 ng/mL) and treated with AZD3965 and VB124.
+After 6 d of treatment, protein lysates were analyzed for extracellular matrix proteins by immunoblot.
+Dual MCT inhibition decreased fibronectin (FN1) and Col1a2 production.
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) or TGFβ-treated vehicle controls.](figs/F_06.png)
 

### 2024-07-31 manuscript/supplement.qmd
add pcls data

@@
 Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
+
+## Precision-cut lung slices
+
+Human precision-cut lung slices (PCLS) were purchased from Anabios (240326LuHA, 26 y.o.
+female without lung disease).
+The slices were incubated for 48 h in DMEM/F12 with 10% serum and then treated with TGFβ (5 ng/mL) and AZD3965 (100 nM) plus VB124 (10 μM) for 6 d, exchanging the medium every 48 h with fresh TGFβ and inhibitors.
+PCLS were lysed in RIPA buffer containing protease inhibitors, homogenized by bead-beating, and analyzed for extracellular matrix production by immunoblot.
 

### 2024-07-31 R/blots.R

@@ -143,3 +143,26 @@ blot_ratios <- function(df) {
     dplyr::mutate(ratio = .data$density / .data$background) |>
     dplyr::filter(!is.na(.data$ratio))
 }
+
+
+combine_blots <- function(nms, df, stats) {
+  x <-
+    dplyr::bind_rows(df, .id = "protein") |>
+    dplyr::mutate(protein = factor(protein, levels = nms))
+  y <-
+    dplyr::bind_rows(stats, .id = "protein") |>
+    dplyr::mutate(protein = factor(protein, levels = nms))
+  plot_one_factor(
+    df = x,
+    stats = y,
+    x = "grp",
+    y = "norm",
+    ytitle = "Fold Change"
+  ) +
+    ggplot2::facet_wrap(
+      ggplot2::vars(.data[["protein"]]),
+      scales = "free_y",
+      nrow = 1,
+      strip.position = "top"
+    )
+}

### 2024-07-31 R/figures.R

@@ -209,13 +209,33 @@ make_fig05 <- function(p1, p2, p3, p4) {
     ggplot2::theme(legend.position = "bottom")
 }
 
-make_fig06 <- function(p1, p2, p3, p4, p5, p6) {
-  design <- "abc \n def"
-  p1 + p2 + p3 + p4 + p5 + p6 +
+make_fig06 <- function(p1, p2, p3, p4, p5, p6, p7, p8) {
+  design <- "abc \n def \n gg#"
+
+  g <-
+    p7 + p8 +
+    patchwork::plot_annotation(
+      title = "Human PCLS",
+      theme = ggplot2::theme(
+        plot.title = ggplot2::element_text(
+          hjust = 0.5,
+          face = "bold",
+          size = ggplot2::rel(0.8)
+        )
+      )
+    ) +
+    theme_patchwork(
+      tags = list(c("G", "")),
+      widths = c(1, 2)
+    )
+
+
+  p1 + p2 + p3 + p4 + p5 + p6 + patchwork::wrap_elements(full = g) +
     theme_patchwork(
       design = design,
       widths = ggplot2::unit(c(2, 1.5, 1.5), "in"),
-      heights = ggplot2::unit(1.5, "in")
+      heights = ggplot2::unit(c(1.5, 1.5, 2), "in"),
+      tags = list(c(LETTERS[1:6], ""))
     )
 }
 

### 2024-07-31 _targets.R

@@ -173,6 +173,15 @@ list(
     ),
     NULL
   ),
+  tar_target(
+    blot_fig_pcls,
+    combine_blots(
+      nms = c("FN1", "Col1a2"),
+      df = list(FN1 = blot_norm_pcls_fn1, Col1a2 = blot_norm_pcls_col1a2),
+      stats = list(FN1 = blot_stats_pcls_fn1, Col1a2 = blot_stats_pcls_col1a2)
+    ),
+    format = "rds"
+  ),
   tar_map(
     values = list(
       experiment = list(
@@ -1973,27 +1982,28 @@ list(
 
   tar_map(
     values = tibble::tribble(
-      ~path,                        ~scale, ~hjust, ~vjust, ~names,
-      "ipf-mct-blots.png",          1.1,    0,      -0.05,  "mct_ipf",
-      "bleo-mct-blots.png",         1.1,    0,      -0.05,  "mct_bleo",
-      "tgfb-mct-blots.png",         1.1,    0,      -0.05,  "mct_tgfb",
-      "sirna-mct-blots.png",        1,      0,      0,      "sma_sirna",
-      "dual-azd-ipf-lf-blot-2.png", 0.9,    0,      0,      "ipf_lf_azd",
-      "dual-azd-sma-blot.png",      1,      0,      -0.05,  "sma_azd",
-      "dual-ar-sma-blot.png",       1.2,    0.1,    -0.05,  "sma_ar",
-      "dual-azd-contract.png",      1,      0,      -0.05,  "contraction",
-      "bleo-trichrome.png",         1.1,    0,      0,      "trichrome",
-      "mct-bleo-timeline.png",          1,      0,      0,  "timeline",
-      "dual-azd-smad3-blot.png",    1,      0,      -0.05,  "smad3_azd",
-      "dual-azd-erk-blot.png",      1,      0,      -0.05,  "erk_azd",
-      "dual-azd-hif1a-blot.png",    1,      0,      -0.05,  "hif_azd",
-      "lactate-sma-blot.png",       1.2,    0,      -0.05,  "sma_lac",
-      "dual-azd-kla-h3.png",        1,      0,      0,      "azd_kla",
-      "mims-panel.png",             1.05,   0.05,   0,      "mims",
-      "vb-bleo-timeline.png",       1,      0,      0,      "vb_timeline",
-      "bleo-vb-young-sma.png",      1,      0,      0,      "vb_young_sma",
[…]

### 2024-08-05 manuscript/manuscript.qmd
add VB253 toxicology

@@
 Our preclinical findings indicate that MCT4 inhibition, as a single therapeutic target, exhibits greater antifibrotic efficacy than MCT1 inhibition, both in suppressing myofibroblast differentiation *in vitro* and reducing bleomycin-induced fibrosis.
-VB253 is a novel inhibitor of MCT4 with ~10-fold increased selectivity for MCT4 *v*.
-MCT1 and ~10-fold increased potency for MCT4 inhibition.
+VB253 is a novel inhibitor of MCT4 with a 30,000-fold increased selectivity for MCT4 *v*.
+MCT1 (**Fig.
+8A-B, S10**) and 10-fold increased potency for MCT4 inhibition compared to VB124 [@cluntun2021].
 Like VB124, VB253 dose-dependently decreased TGFβ-stimulated α-SMA in human IPF lung fibroblasts *ex vivo* (**Fig.
-8A**).
+8C**).
 To assess its efficacy compared to established therapies, we compared VB253 to nintedanib, a clinically approved antifibrotic medication for pulmonary fibrosis (**Fig.
-8A-B**).
+8C-D**).
 Nintedanib inhibits TGFβ-mediated myofibroblast differentiation and decreases Col1a1 expression *in vitro* and *in vivo*  [@wollin2014].
 While both compounds effectively attenuated α-SMA expression, nintedanib exhibited moderate cytotoxicity absent with VB253 (**Fig.
-8B**).
+8D**).
 Nintedanib mitigates myofibroblast differentiation partly by inhibiting TGFβ receptor phosphorylation and Smad-dependent signaling pathways [@rangarajan2016].
@@
 In experiments conducted independently from those reported above, VB253 was administered to young mice (8-10 weeks) beginning on day 7 following bleomycin (**Fig.
-8C**).
+8E**).
 The existing pulmonary fibrosis therapies, nintedanib and pirfenidone, served as comparators.
@@
 VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterns in bleomycin-treated mice (**Fig.
-8D**).
+8F**).
 Histologic assessment confirmed reduced pulmonary fibrosis (**Fig.
-8E**) and α-SMA expression (**Fig.
-8F**).
+8G**) and α-SMA expression (**Fig.
+8H**).
 Moreover, the effects of VB253 were comparable to the antifibrotic effects of nintedanib and pirfenidone with decreased cytotoxicity in *in vitro* assays.
@@
 Similar to young mice, VB253 decreased fibrosis severity, as quantified by Ashcroft score and α-SMA expression (**Fig.
-8G-I**).
+8I-K**).
 The magnitude of improvement paralleled that of nintedanib and pirfenidone.
@@
 7A**).
+Preclinical evaluation of VB253 in pharmacokinetic and toxicology studies demonstrated a reassuring risk profile at the estimated efficacious dose in humans, and a Phase 1 clinical trial is currently ongoing (**Supplementary Material**).
 Collectively, these findings provide compelling preclinical evidence supporting lactate transporter inhibition as a novel therapeutic strategy for fibrotic lung disease.
@@
 In this Phase 1 study, AZD3965 was generally well-tolerated with 7 of 40 patients experiencing dose-limiting toxicities including asymptomatic, reversible ocular changes; acidosis; and increased troponin.
-VB124 was the first selective MCT4 inhibitor developed [@cluntun2021] and we now introduce VB253 as a second generation MCT4 inhibitor.
+VB124 was the first selective MCT4 inhibitor developed [@cluntun2021], and we now introduce VB253 as a second generation MCT4 inhibitor.
 Our data demonstrate that VB253 has similar efficacy in experimental pulmonary fibrosis models as the established antifibrotics, nintedanib and pirfenidone, with potentially less cytotoxicity (**Fig.
@@
 # List of Supplementary Materials
+
+VB253 Pharmacokinetic and Safety Data
 
@@
 8.
-VB253, a novel MCT4 inhibitor, is potently antifibrotic *in vitro* and *in vivo*.** (**A**) Dose-response of α-SMA protein expression in IPF lung fibroblasts treated with VB253 or nintedanib (NIN).
-(**B**) Number of nuclei, a marker of cell viability, following treatment with VB253 or NIN.
+VB253, a novel MCT4 inhibitor, is potently antifibrotic *in vitro* and *in vivo*.** (**A**) Structure of VB253.
+(**B**) VB253 potently and selectively inhibits lactate uptake rate in a reporter cell assay.
+IC~50~ values are noted.
+(**C**) Dose-response of α-SMA protein expression in IPF lung fibroblasts treated with VB253 or nintedanib (NIN).
+(**D**) Number of nuclei, a marker of cell viability, following treatment with VB253 or NIN.
 Summary data are mean ± SEM of cell responses from N = 3 different donors.
-(**C**) Young (8-10 w) or aged (60+ w) mice were treated with bleomycin (1 U/kg) on day 0.
+(**E**) Young (8-10 w) or aged (60+ w) mice were treated with bleomycin (1 U/kg) on day 0.
 Beginning on day 7, animals were treated with VB253 or the FDA-approved antifibrotics nintedanib (NIN) and pirfenidone (PFD) for 14 days prior to euthanasia on day 21.
-(**D**) Enhanced pause (Penh) is a non-invasive index of mouse tidal breathing that increases with bleomycin treatment.
-(**E**-**G**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (E) and decreased α-SMA protein expression (F).
-(**G**-**I**) Similar histologic improved was observed in aged mice with decreased histologic fibrosis (G) and decreased α-SMA protein expression (H).
-(**I**) VB253 decreased lung lactate, similar to established antifibrotics.
+(**F**) Enhanced pause (Penh) is a non-invasive index of mouse tidal breathing that increases with bleomycin treatment.
+(**G**-**H**) Histologic assessment of VB253-treated mouse lungs demonstrates decreased fibrosis by Ashcroft score (G) and decreased α-SMA protein expression (H).
+(**I**-**J**) Similar histologic improvement was observed in aged mice with decreased histologic fibrosis (I) and decreased α-SMA protein expression (J).
+(**K**) VB253 decreased lung lactate, similar to established antifibrotics.
 Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_08.png)

### 2024-08-05 manuscript/supplement.qmd
add VB253 toxicology

@@
 ```
+
+\newpage
+
+# VB253 Pharmacokinetic and Safety Data
+
+Safety pharmacology of VB253 was evaluated in an *in vitro* hERG assay, a behavioral study in Sprague-Dawley (SD) rats, and a respiratory and cardiovascular system assessment using ECG telemetered beagle dogs.
+VB253 had an IC*50* value > 300 µM against the hERG channel.
+There were no VB253-related behavioral effects following an oral dose of up to 400 mg/kg in male rats (the highest dose tested).
+There were no significant respiratory effects following an oral dose of up to 150 mg/kg in dogs (the highest dose tested).
+Cardiovascular safety pharmacology assessments noted decreases in PR interval duration that were detected up to 9 h post administration of VB253 and were usually dose related and associated with potential changes in heart rate.
+These changes were considered biologically insignificant due to their small magnitude (up to 10 msec compared to vehicle control), association with potential increases in heart rate, and the absence of correlated findings.
+
+The ADME properties of VB253 have been characterized in both *in vitro* and *in vivo* studies.
+Sensitive and selective bioanalytical methods have been developed and validated for rat and dog plasma.
+Following IV administration, the systemic clearance of VB253 in all species evaluated was considered low.
+Following oral administration, bioavailability was generally greater than 30% in the species evaluated.
+Initial studies utilized VB253 free form; however, due to higher bioavailability, a tris salt form of VB253 was chosen for further development.
+All GLP studies were performed using the VB253 tris salt form.
+Dose levels indicate VB253 free form equivalents unless otherwise noted.
+
+In Caco-2 cells, VB253 demonstrated high permeability with no evidence of being a substrate for an efflux transporter.
+Plasma protein binding was 98.7% in human plasma, 99.9% in rat plasma, and 98.5% in dog plasma.
+The metabolic profile of VB253 was evaluated in rat, dog, and human liver microsomes.
+Rat and dog appeared to cover the metabolites produced in human microsomes and were considered the suitable species for toxicology studies.
+Study of the oxidative metabolism of VB253 by recombinant CYP enzymes indicated that CYP2C8, CYP3A4, and CYP3A5 are involved in VB253 metabolism *in vitro*.
+VB253 demonstrated negligible inhibition of any of the CYP isozymes, with IC*50* values being greater than 30 µM.
+VB253 was a weak activator of human PXR *in vitro* and is therefore unlikely to cause induction of PXR target genes, including CYP3A, *in vivo*.
+VB253 was a weak inhibitor of BCRP, OAT1, OAT3, OAT1B1, OAT1B3, and OCT1 transporters, with IC*50* values ranging from 1.81-6.53 µM.
+
+Toxicokinetic studies generally demonstrated less-than-dose-proportional increases in exposure in the rat and dog.
+In the rat, sex differences were noted with female rats demonstrating higher exposure than male rats, while in the dog no differences in exposure between the sexes was noted.
+Generally, exposure was slightly higher following multiple doses relative to single dose administration.
+
+The toxicology of VB253 was evaluated in pivotal GLP studies in SD rats and beagle dogs, as well as mutagenicity studies.
+VB253 was not mutagenic in the bacterial reverse mutation assay or in the *in vitro* human lymphocyte micronucleus assay.
+
+VB253 was administered to SD rats by oral gavage at doses of 25, 75, and 250 mg/kg/day (females) and 40, 120, and 400 mg/kg/day (males) for 28 days.
+VB253-related mortality occurred in 8 of the 15 toxicity males administered 400 mg/kg/day beginning on Day 8 through 24.
+VB253-related effects in males administered 400 mg/kg/day and females administered 250 mg/kg/day were decreased body weight, decreased food consumption, clinical pathology changes in hematology and serum chemistry parameters, and histopathological changes.
+Target tissues included bone marrow, spleen, thymus, liver, lymph nodes, epididymides, seminal vescicles, prostate, testes, uterus, and ovaries.
+Administration of 40 or 120 mg/kg/day VB253 to male and 25 or 75 mg/kg/day in female rats *via* oral gavage once daily for 4 weeks with a 4-week recovery phase was well-tolerated.
+Based on these adverse findings the no-observed-adverse-effect level (NOAEL) was considered to be 120 mg/kg/day in male rats and 75 mg/kg/day in female rats with C*max* of 180,000 ng/mL and 198,000 ng/mL, respectively, and an AUC0-24 of 2,780,000 and 3,230,000 ng•h/mL, respectively.
+
+VB253 was administered to beagle dogs by oral gavage in a range finding study at doses of 30, 79, 240 or 790 mg/kg.
+Watery feces and emesis were occasionally observed in dogs at all dose levels.
+Male dogs administered 790 mg/kg VB253 had increased potassium and decreased sodium and chloride, indicating a possible VB253-related effect on electrolyte homeostasis.
+Decreased thymus weights were observed at ≥ 240 mg/kg and correlated with microscopic findings of decreased thymic lymphoid cellularity.
+VB253 was administered at doses of 10, 50, and 150 mg/kg/day in the pivotal repeat dose toxicity studies for 28 days.
+Based on a lack of adverse findings, the NOAEL was considered to be 150 mg/kg/day in dogs with an associated sex-combined C*max* of 80,400 ng/mL and sex-combined AUC0-24 of 474,000 ng•h/mL.
 
@@
 
+MCT4 and MCT1 inhibition by VB253 were determined using a cell-based lactate transport assay monitoring the change in intracellular pH with BCECF following the addition of extracellular lactate.
+MDA-MB-231 and BT20 breast cancer cell lines were used to examine MCT4 and MCT1 inhibition, respectively.
+
 All cell culture experiments were conducted by Charles River Laboratories (Wilmington, MA, USA).
@@
 S10.
-VB253 does not attenuate canonical TGFβ signaling.** Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
+VB253 is a novel, potent, and selective inhibitor of MCT4.** (**A**) Dose-dependent change in intracellular pH over time in BCECF-loaded MDA-MB-231 (MCT4) or BT20 (MCT1) breast cancer cell lines.
+(**B**) VB253 does not attenuate canonical TGFβ signaling.
+Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
 Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/SF_10.png)

### 2024-08-05 R/figures.R

@@ -249,7 +249,7 @@ make_fig07 <- function(p1, p2, p3) {
     )
 }
 
-make_fig08 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
+make_fig08 <- function(p12, p13, p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
   # design <- "abcc \n defg \n hijk"
   # p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8 + p9 + p10 + p11 +
   #   theme_patchwork(
@@ -257,6 +257,14 @@ make_fig08 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
   #     widths = ggplot2::unit(1.5, "in"),
   #     heights = ggplot2::unit(1.5, "in")
   #   )
+
+  very_top <-
+    p12 + p13 + patchwork::plot_spacer() +
+    theme_patchwork(
+      design = "abc",
+      widths = c(1, 1, 2),
+      tags = list(c("A", "B", ""))
+    )
 
   top <-
     ((p1 + p2) +
@@ -269,7 +277,7 @@ make_fig08 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
     theme_patchwork(
       # design = "abc",
       widths = c(2, 2),
-      tags = list(c("A", "B", "C")),
+      tags = list(c("C", "D", "E")),
       guides = "collect"
     ) &
     ggplot2::theme(legend.position = "bottom")
@@ -288,7 +296,7 @@ make_fig08 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
     ) +
     theme_patchwork(
       design = "abcd",
-      tags = list(c("D", "E", "F", ""))
+      tags = list(c("F", "G", "H", ""))
     )
 
   bot <-
@@ -305,14 +313,15 @@ make_fig08 <- function(p1, p2, p3, p4, p5, p6, p7, p8, p9, p10, p11) {
     ) +
     theme_patchwork(
       design = "abcd",
-      tags = list(c("G", "H", "", "I"))
+      tags = list(c("I", "J", "", "K"))
     )
 
+  patchwork::wrap_elements(full = very_top) +
   patchwork::wrap_elements(full = top) +
     patchwork::wrap_elements(full = mid) +
     patchwork::wrap_elements(full = bot) +
     theme_patchwork(
[…]

### 2024-08-05 R/plots.R

@@ -223,6 +223,8 @@ clrs <- c(
   "VB253"         = "#E15759",
   "NIN"           = "#499894",
   "PFD"           = "#B07AA1",
+  "MCT1"          = "#59A14F",
+  "MCT4"          = "#B6992D",
   NULL
 )
 

### 2024-08-05 R/vb.R

@@ -1,5 +1,157 @@
 # vb.R
 
+read_vb_pharm <- function(x) {
+  nms <-
+    stringr::str_extract(x, "mct\\d{1}") |>
+    toupper()
+  purrr::map(
+    x,
+    \(x) readr::read_csv(
+      file = x,
+      show_col_types = FALSE
+    ) |>
+      tidyr::pivot_longer(
+        cols = -time,
+        names_to = "conc",
+        values_to = "value",
+        names_transform = as.numeric
+      )
+  ) |>
+    rlang::set_names(nms) |>
+    dplyr::bind_rows(.id = "target") |>
+    dplyr::mutate(
+      target = factor(target, levels = c("MCT4", "MCT1")),
+      time = time - 8,
+      smooth = slider::slide_dbl(value, mean, .before = 15, .after = 5),
+      .by = c(target, conc)
+    )
+}
+
+
+plot_vb_rates <- function(z) {
+  ggplot2::ggplot(z) +
+    ggplot2::facet_wrap(
+      ggplot2::vars(target),
+      nrow = 1
+    ) +
+    ggplot2::aes(
+      x = time,
+      y = value,
+      color = conc,
+      group = conc
+    ) +
+    ggplot2::geom_point(
+      size = 0.05
+    ) +
+    ggplot2::scale_color_viridis_c(
+      option = "H",
+      transform = scales::transform_log(),
+      breaks = scales::breaks_log(),
+      labels = scales::label_log(),
+      oob = scales::oob_squish_infinite
+    ) +
+    ggplot2::labs(
+      x = "Time (min)",
+      y = "Fluorescence ratio",
+      color = "VB253 (nM)"
+    ) +
+    ggplot2::coord_cartesian(clip = "off") +
+    theme_plot() +
+    ggplot2::theme(
+      legend.position = "bottom"
+    ) +
+    NULL
+}
+
+
+read_vb_drc <- function(x) {
+  nms <-
+    stringr::str_extract(x, "mct\\d{1}") |>
+    toupper()
+  purrr::map(
+    x,
+    \(x) readr::read_csv(
+      file = x,
+      show_col_types = FALSE
+    )
+  ) |>
+    rlang::set_names(nms) |>
+    dplyr::bind_rows(.id = "target") |>
+    dplyr::group_by(target) |>
+    dplyr::mutate(
+      target = factor(target, levels = c("MCT4", "MCT1")),
+      conc = 10 ^ conc,
[…]

### 2024-08-05 _targets.R

@@ -1864,6 +1864,32 @@ list(
   # VB253 -------------------------------------------------------------------
 
   tar_target(
+    vb_rates_files,
+    raw_data_path("vb_mct\\d{1}.csv"),
+    format = "file"
+  ),
+  tar_target(
+    vb_rates_data,
+    read_vb_pharm(vb_rates_files)
+  ),
+  tar_target(
+    vb_rates_plot,
+    plot_vb_rates(vb_rates_data)
+  ),
+  tar_target(
+    vb_drc_files,
+    raw_data_path("vb_mct\\d{1}_drc.csv"),
+    format = "file"
+  ),
+  tar_target(
+    vb_drc_data,
+    read_vb_drc(vb_drc_files)
+  ),
+  tar_target(
+    vb_drc_plot,
+    plot_vb_drc(vb_drc_data)
+  ),
+  tar_target(
     vb_invitro_sma_file,
     raw_data_path("invitro_sma.csv"),
     format = "file"
@@ -2003,7 +2029,8 @@ list(
       "bleo-vb-young-sma.png",      "vb_young_sma",
       "bleo-vb-aged-sma.png",       "vb_aged_sma",
       "glucose-tracer.png",         "tracer",
-      "pcls-blot.png",              "pcls"
+      "pcls-blot.png",              "pcls",
+      "VB253.png",                  "vb253"
     ),
     names = names,
     tar_target(
@@ -2151,6 +2178,8 @@ list(
   tar_target(
     fig08,
     make_fig08(
+      fig_img_vb253,
+      vb_drc_plot,
       vb_invitro_plot_sma_ipf,
       vb_invitro_plot_nuclei_ipf,
       fig_img_vb_timeline,
@@ -2322,6 +2351,7 @@ list(
   tar_target(
     fig10s,
     make_fig10s(
+      vb_rates_plot,
       vb_invitro_smad_plot
     ) |>
       write_figures("SF_10"),

### 2024-08-08 manuscript/manuscript.qmd
revised submission for STM

@@
 targets::tar_load(refs_ms)
+targets::tar_load(refs_supp)
 targets::tar_load(pkg_citations)
@@
     fig08, 
+    resources_table, 
+    fig01s,
+    fig02s, 
+    fig03s, 
+    fig04s, 
+    fig05s, 
+    fig06s, 
+    fig07s, 
+    fig08s, 
+    fig09s, 
+    fig10s,
     NULL
@@
 While glycolytic cells must export lactate, the contributions of lactate transporters to myofibroblast differentiation are unknown.
-In this study, we investigated how MCT1 and MCT4, key lactate transporters, influence myofibroblast differentiation and experimental pulmonary fibrosis.
+In this study, we investigated how MCT1 and MCT4, key pulmonary lactate transporters, influence myofibroblast differentiation and experimental pulmonary fibrosis.
 Our findings reveal that inhibiting MCT1 or MCT4 reduces TGFβ-stimulated pulmonary myofibroblast differentiation *in vitro* and decreases bleomycin-induced pulmonary fibrosis *in vivo*.
@@
 Furthermore, we introduce VB253, a novel MCT4 inhibitor, which ameliorates pulmonary fibrosis in both young and aged mice, with comparable efficacy to established antifibrotic therapies.
-These results underscore the necessity of lactate transport for myofibroblast differentiation, identify MCT1 and MCT4 as promising pharmacologic targets in pulmonary fibrosis, and support further evaluation of lactate transport inhibitors for patients for whom limited therapeutic options currently exist.
+These results underscore the necessity of lactate transport for myofibroblast differentiation, identify MCT1 and MCT4 as promising pharmacologic targets in pulmonary fibrosis, and support further evaluation of lactate transport inhibitors for patients for whom limited therapeutic options exist currently.
 
@@
 Toward this end, we aimed to examine the impact of a novel metabolic strategy --- lactate transport inhibition --- on myofibroblast differentiation and experimental pulmonary fibrosis.
-Sustained glycolysis in myofibroblasts relies on lactate secretion, which is conducted by a family of monocarboxylate transporters (MCT1-4).
-Inhibitors targeting these transporters have been actively explored in clinical trials for oncological conditions where glycolytic reprogramming also features prominently in disease pathobiology [@puri2020; @cluntun2021].
-Importantly, MCT inhibitors present favorable pharmacologic profiles compared to previously studied glycolysis inhibitors, with successful translation to human clinical trials for advanced solid tumors [@halford2023].
+Sustained glycolysis in myofibroblasts relies on lactate secretion, which is mediated by a family of monocarboxylate transporters (MCT1-4).
+Lactate transport inhibitors are under active development for diseases where glycolytic reprogramming also features prominently, such as cancer and heart failure [@puri2020; @cluntun2021].
+Importantly, MCT inhibitors have favorable pharmacologic profiles compared to previously studied glycolysis inhibitors, with successful translation to human clinical trials for advanced solid tumors [@halford2023].
 Before the promise of this therapeutic approach in IPF may be realized, however, the preclinical efficacy and molecular mechanisms-of-action of lactate transport inhibitors must be demonstrated experimentally.
@@
 We found increased expression of the lactate transporters MCT1 and MCT4 IPF patient lungs.
-Inhibition of these transporters attenuated bleomycin-induced lung fibrosis *in vivo* and TGFβ-induced myofibroblast differentiation *in vitro*, where MCT4 inhibition demonstrated increased therapeutic efficacy.
+Inhibition of these transporters decreased bleomycin-induced lung fibrosis in mice, TGFβ-induced fibrosis in human precision-cut lung slices (PCLS), and TGFβ-induced myofibroblast differentiation *in vitro*.
 Using metabolomics, stable isotope tracing, and high-resolution spatial metabolomic imaging, we find that lactate transport inhibition promotes glucose oxidation and decreases pro-fibrogenic reactive oxygen species (ROS) production.
-Building on these results, we introduce VB253, a novel MCT4 inhibitor suitable for human clinical studies that performs similarly to current standard-of-care antifibrotic therapies in bleomycin-induced pulmonary fibrosis.
-Together, our findings offer insights into disrupting the metabolic pathways driving IPF fibrosis and present a promising strategy targeting lactate transporters for the treatment of this fatal condition.
+Across all studies, inhibition of MCT4 was more potently anti-fibrotic than MCT1 inhibition.
+Building on these results, we introduce VB253, a novel MCT4 inhibitor suitable for human clinical studies that performs similarly to current standard-of-care antifibrotic therapies in bleomycin-induced pulmonary fibrosis models.
+Together, our findings offer further insight into the role of metabolic reprogramming in pulmonary fibrosis and and present a promising strategy targeting lactate transporters for the treatment of this fatal condition.
 
@@
 Among the four lactate transporters, MCT1 and MCT4 exhibit the highest expression levels in the lung [@frangogiannis2020].
-Based on this observation, we investigated the expression of MCT1 and MCT4 in lung explants from patients with IPF obtained during transplantation.
+Thus, we investigated the expression of MCT1 and MCT4 in lung explants from patients with IPF obtained at the time of transplantation.
 Consistent with a pathologic role for MCT1 and MCT4 in IPF, we observed a significant upregulation of both MCT1 and MCT4 proteins in IPF lung tissues compared to non-fibrotic controls (**Fig.
 1A**).
-These findings were corroborated in an experimental model of pulmonary fibrosis, where intratracheal bleomycin administration led to increased expression of both MCT1 and MCT4 (**Fig.
+These findings were corroborated in an experimental model of pulmonary fibrosis, where intratracheal bleomycin administration caused increased expression of both MCT1 and MCT4 (**Fig.
 1B**).
@@
 
-After observing a reduction in MCT4 expression following siMCT1 treatment, we proceeded to evaluate the impact of pharmacological MCT inhibitors on myofibroblast differentiation to examine the independent effects of MCT1 and MCT4 inhibition.
-AZD3965, a high-affinity (*K~i~* 1.6 nM) inhibitor of MCT1 [@curtis2017], and VB124, a recently developed high-affinity (*K~i~* 11 nM) inhibitor of MCT4 [@cluntun2021], were used in this evaluation.
-IPF lung fibroblasts were differentiated with TGFβ in the presence of these MCT inhibitors (**Fig.
+After observing a reduction in MCT4 expression following siMCT1 treatment, we next evaluated the impact of pharmacological MCT inhibitors on myofibroblast differentiation to examine the independent effects of MCT1 and MCT4 inhibition.
+AZD3965 is a high-affinity (*K~i~* 1.6 nM) inhibitor of MCT1 [@curtis2017], and VB124 is a recently developed high-affinity (*K~i~* 11 nM) inhibitor of MCT4 [@cluntun2021].
+IPF lung fibroblasts were treated with TGFβ in the presence of these MCT inhibitors (**Fig.
 2B**).
@@
 
-To further characterize the antifibrotic effects of lactate transport inhibition, we conducted RNA sequencing on lung fibroblasts treated with TGFβ in conjunction with AZD3965 or VB124 (**Fig.
+To further characterize the antifibrotic effects of lactate transport inhibition, we conducted RNA sequencing on lung fibroblasts treated with TGFβ and AZD3965 or VB124 (**Fig.
 2E-H, S2**).
-Principal components analysis (PCA) revealed that the first principal component predominantly represented the effect of TGFβ treatment, while MCT4 inhibition, either alone or combined with MCT1 inhibition, aligned with the second principal component (**Fig.
+Principal components analysis (PCA) revealed that the first principal component predominantly represented the effect of TGFβ, while MCT4 inhibition, either alone or combined with MCT1 inhibition, aligned with the second principal component (**Fig.
 2E**).
@@
 
-Differential expression analysis of TGFβ-treated cells revealed the anticipated upregulation of extracellular matrix proteins (**Fig.
+Differential expression analysis of TGFβ-treated cells demonstrated the anticipated upregulation of extracellular matrix proteins (**Fig.
 S2A**) and enrichment of the epithelial-to-mesenchymal (EMT) gene set, among others (**Fig.
@@
 3A**).
-By contrast, pharmacologic inhibition of either MCT1 or MCT4 alone did not decrease TGFβ-stimulated lactate efflux (**Fig.
-3B**); inhibition of both transporters was required to lower extracellular lactate levels.
-Inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to prevent increases in lactate efflux (**Fig.
+By contrast, pharmacologic inhibition of either MCT1 or MCT4 alone had no effect (**Fig.
+3B**); inhibition of both transporters was required to decrease extracellular lactate levels.
+Inhibition of MCT1, MCT2, and MCT4 with the combination of AR-C155858 and VB124 was required to abrogate increases in lactate efflux (**Fig.
 S3A**).
-Although less pronounced compared to lactate production, we observed a trend toward decreased extracellular glucose consumption in cells treated with both AZD3965 and VB124 (**Fig.
+We observed a trend toward decreased extracellular glucose consumption in cells treated with both AZD3965 and VB124 (**Fig.
 S3B**).
 These findings align with previous studies indicating compensatory roles for MCT1 and MCT4 in lactate export [@benjamin2018].
-Furthermore, these results suggest that glycolysis inhibition is not the principal mechanism underlying the attenuation of myofibroblast differentiation by lactate transport inhibition.
+Furthermore, these results suggest that glycolysis inhibition is not the principal mechanism preventing myofibroblast differentiation by lactate transport inhibitors.
 
@@
 Consistent with our direct measures of extracellular lactate and glucose, concurrent treatment with AZD3965 and VB124 was necessary to reduce PER.
-Consequently, this reduction in PER coincided with an increase in OCR as myofibroblasts transitioned their metabolism from glycolysis to oxidative phosphorylation.
+Consequently, this reduction in PER coincided with an increase in OCR as myofibroblasts transitioned from glycolysis to oxidative phosphorylation.
 
 Interestingly, individual administration of AZD3965 and VB124 unexpectedly increased OCR without corresponding decreases in PER.
-Indeed, the primary consequence of MCT1 or MCT4 inhibition alone was increased cellular ATP production rates driven by upregulation of oxidative phosphorylation (**Fig.
+Indeed, the primary consequence of MCT1 or MCT4 inhibition alone was increased cellular ATP production rates driven by increased oxidative phosphorylation (**Fig.
 3D-E**).
-MCT inhibition decreased spare respiratory capacity, indicating that the basal respiratory rate of treated cells approached their maximal oxidative capacity (**Fig.
+MCT inhibition decreased spare respiratory capacity, indicating that the basal respiratory rate of treated cells approached the maximal oxidative capacity of these cells (**Fig.
 S3D**).
@@
 4F**).
-As expected from extracellular lactate measures following MCT inhibition, intracellular lactate accumulated moderately with MCT4 inhibition and substantially with MCT1/4 inhibition.
-TGFβ treatment resulted in diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Fig.
+As expected from extracellular lactate measures following MCT inhibition, intracellular lactate accumulated moderately with MCT4 inhibition and substantially with dual inhibition.
+TGFβ treatment caused diverse changes in the intracellular metabolomic profile of treated fibroblasts (**Fig.
 S4C-D**).
@@
 TGFβ increased ^13^C incorporation from glucose into pyruvate, lactate, citrate, succinate (SUC), and malate (MAL), indicating enhanced flux from glucose into the TCA cycle.
-While AZD3965 had minimal impact on these labeling patterns, MCT4 inhibition significantly elevated the fractions of these metabolites labeled by ^13^C, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the results of our bioenergetic and steady-state metabolomics experiments described above.
+While AZD3965 had minimal impact on these labeling patterns, MCT4 inhibition significantly elevated the fractions of these metabolites labeled by ^13^C, providing direct evidence for a proportional increase in glucose oxidation following MCT4 inhibition, consistent with the results of our bioenergetic and steady-state metabolomics experiments.
 
@@
 
-Building on our *in vitro* findings indicating an antifibrotic effect of lactate transporter inhibition, we proceeded to evaluate the efficacy of MCT inhibitors in a bleomycin-induced mouse model of pulmonary fibrosis.
+Building on our *in vitro* findings indicating an antifibrotic effect of lactate transporter inhibition, we evaluated the efficacy of MCT inhibitors in a bleomycin-induced mouse model of pulmonary fibrosis.
 Mice received 1.2 U/kg bleomycin by intratracheal administration.
@@
 S8**).
-Lung mechanics improved by approximately 50% of baseline compared to vehicle controls following 14 days of MCT inhibitor treatment (**Fig.
+Lung mechanics improved by approximately 50% compared to vehicle controls following 14 days of MCT inhibitor treatment (**Fig.
 6B-C**).
-This improvement in measures of lung stiffness was further supported by histologic assessment of fibrosis severity using Ashcroft scoring [@ashcroft1988] (**Fig.
+This improvement in pulmonary function was corroborated by histologic decrease in fibrosis severity using Ashcroft scoring [@ashcroft1988] (**Fig.
 6D-E**) and hydroxyproline content measurement (**Fig.
@@
 While bleomycin administration substantially altered the lung metabolic profile, treatment with AZD3965 or VB124 had little additional impact on total metabolite levels.
-However, unlike bleomycin alone or following AZD3965 treatment, VB124 treatment resulted in significant alterations in several circulating metabolites (**Fig.
+However, unlike bleomycin alone or following AZD3965 treatment, VB124 treatment altered several circulating metabolite levels (**Fig.
 S9I**).
-Among these, VB124 significantly increased circulating lactate, while AZD3965 led to decreased circulating lactate (**Fig.
+Among these, VB124 increased circulating lactate, while AZD3965 decreased circulating lactate (**Fig.
 7A**).
 This trend was also reflected in the lung, where only modest changes were noted following bleomycin, contrary to prior reports [@kottmann2012].
-Considering the influence of our treatments on circulating lactate, we calculated the lung-to-plasma lactate ratio.
+Considering the influence of the treatments on circulating lactate, we calculated the lung-to-plasma lactate ratio.
 This analysis revealed the anticipated increase in lactate in fibrotic lungs, with levels returning to baseline following the addition of VB124 (**Fig.
@@
 7B-C**).
-This labeling approach proved effective, demonstrating approximately 1.5-fold enrichment ^15^N, 4.6-fold for ^2^H, and 2.4-fold for ^13^C above natural isotope abundance.
+This labeling approach proved effective, demonstrating approximately 1.5-fold enrichment of ^15^N, 4.6-fold for ^2^H, and 2.4-fold for ^13^C above natural isotope abundance.
 Pulmonary fibrosis correlated with a significant increase in ^15^N labeling from proline, consistent with increased collagen synthesis and deposition during the labeling period.
-Similarly, drug-treated animals exhibited less ^15^N incorporation per tissue area, consistent with lung function and histological analyses.
-Glucose labeling displayed a similar pattern, wherein lactate transporter inhibition led to reduced glucose incorporation into tissue biomass.
-Importantly, these findings are not merely attributable to decreased tissue fibrosis, as we selectively imaged more fibrotic areas (**Fig.
+Similarly, drug-treated animals exhibited less ^15^N incorporation per tissue area, consistent with lung function and histologic analyses.
+Glucose labeling displayed a similar pattern, where lactate transporter inhibition reduced glucose incorporation into tissue biomass.
+Importantly, these findings were not merely attributable to decreased tissue fibrosis, as we selectively imaged more fibrotic areas (**Fig.
 7B**) and normalized the enrichment values to tissue area.
@@
 Our preclinical findings indicate that MCT4 inhibition, as a single therapeutic target, exhibits greater antifibrotic efficacy than MCT1 inhibition, both in suppressing myofibroblast differentiation *in vitro* and reducing bleomycin-induced fibrosis.
-VB253 is a novel inhibitor of MCT4 with a 30,000-fold increased selectivity for MCT4 *v*.
+VB253 is a novel inhibitor of MCT4 with a 30,000-fold selectivity for MCT4 *v*.
 MCT1 (**Fig.
@@
 Nintedanib inhibits TGFβ-mediated myofibroblast differentiation and decreases Col1a1 expression *in vitro* and *in vivo*  [@wollin2014].
-While both compounds effectively attenuated α-SMA expression, nintedanib exhibited moderate cytotoxicity absent with VB253 (**Fig.
+While both compounds effectively decreased α-SMA expression, nintedanib exhibited moderate cytotoxicity absent with VB253 (**Fig.
 8D**).
 Nintedanib mitigates myofibroblast differentiation partly by inhibiting TGFβ receptor phosphorylation and Smad-dependent signaling pathways [@rangarajan2016].
-Correspondingly, nintedanib dose-dependently attenuated TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 had no impact on this signaling pathway (**Fig.
+Correspondingly, nintedanib dose-dependently decreased TGFβ-mediated Smad3 phosphorylation, while VB253 and VB124 had no impact on this signaling pathway (**Fig.
 5, S10**).
@@
 Three weeks after bleomycin, whole-body plethysmography was performed to assess breathing patterns in unrestrained mice.
-Enhanced pause (Penh) is a dimensionless index describing airflow during tidal breathing found to increase significantly in bleomycin-treated mice [@vanoirbeek2010].
+Enhanced pause (Penh) is a dimensionless index describing airflow during tidal breathing found to increase in bleomycin-treated mice [@vanoirbeek2010].
 VB253 restored Penh to baseline levels, suggesting normalization of respiratory patterns in bleomycin-treated mice (**Fig.
@@
 Furthermore, inhibition of glycolysis not only prevented myofibroblast differentiation [@xie2015; @kottmann2015; @bernard2015], but also attenuated experimental pulmonary fibrosis [@judge2018; @cho2017].
-Our findings align with this metabolic shift, as we observed upregulated expression of lactate transporters MCT1 and MCT4, supporting this glycolytic phenotype in myofibroblasts.
+Our findings align with this metabolic shift, as we observed upregulated expression of lactate transporters MCT1 and MCT4, supporting a glycolytic phenotype in myofibroblasts.
 
@@
 Nevertheless, metabolic imaging at high spatial resolution holds great promise for correlating cell identities from spatial transcriptomic profiles with metabolic features.
-Furthermore, stable isotopes may be safely administered to human patients prior lung biopsy or explant [@steinhauser2012; @guillermier2017], offering a strategy that could significantly advance our understanding of how cell metabolism contributes to pulmonary fibrosis.
+Furthermore, stable isotopes may be safely administered to human patients prior to lung biopsy or explant [@steinhauser2012; @guillermier2017], offering a strategy that could significantly advance our understanding of how cell metabolism contributes to pulmonary fibrosis.
 
@@
 This study was designed to investigate the role of lactate transporters in myofibroblast differentiation and pulmonary fibrosis.
-The objectives of this study were (i) to determine the expression of lactate transporters in IPF lungs and model systems, (ii) to characterize the phenotypic effects of lactate transport inhibition in model systems, and (iii) to profile the metabolic consequences of lactate transport inhibitors.
+The objectives of this study were (i) to determine the expression of lactate transporters in IPF lungs and model systems, (ii) to characterize the phenotypic effects of lactate transport inhibition in model systems, and (iii) to profile the metabolic consequences of lactate transport inhibition.
 
@@
 Significantly enriched pathways with adjusted p-value < 0.05 are included and points are colored by normalized enrichment score (NES).
-Positive NES indicated relative enrichment following TGFβ compared to control (TGFβ) or with MCT inhibitor compared to vehicle in TGFβ-treated cells.
+Positive NES indicates relative enrichment following TGFβ compared to control (TGFβ) or with MCT inhibitor compared to vehicle in TGFβ-treated cells.
 (**G**) Volcano plot of significantly differentially expressed genes in TGFβ-stimulated cells treated with AZD3965 and VB124 compared to vehicle control.
 Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.
-(**H**) Transcript counts from the leading edge of enrichment for the epithelial mesenchymal transition Hallmark gene set demonstrating down-regulation of pro-fibrotic genes with MCT inhibition.
+(**H**) Transcript counts from the leading edge of enrichment for the epithelial mesenchymal transition Hallmark gene set demonstrating downregulation of pro-fibrotic genes with MCT inhibition.
 Individual data points are biological replicates.
@@
 6.
-Lactate transporter inhibition attenuates experimental pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
+Lactate transporter inhibition decreases experimental pulmonary fibrosis.** (**A**) C57Bl/6N mice were administered bleomycin (1.2 U/kg) on day 0.
 Beginning on day 7, animals were treated with AZD3965 (AZD) or VB124 (VB) for 14 days prior to euthanasia on day 21.
@@
 Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value < 0.05 compared to bleomycin-treated (Bleo) vehicle control.](figs/F_08.png)
+
+\newpage
+
+# VB253 Pharmacokinetic and Safety Data
+
+Safety pharmacology of VB253 was evaluated in an *in vitro* hERG assay, a behavioral study in Sprague-Dawley (SD) rats, and a respiratory and cardiovascular system assessment using ECG telemetered beagle dogs.
+VB253 had an IC~50~ value > 300 µM against the hERG channel.
+There were no VB253-related behavioral effects following an oral dose of up to 400 mg/kg in male rats (the highest dose tested).
+There were no significant respiratory effects following an oral dose of up to 150 mg/kg in dogs (the highest dose tested).
+Cardiovascular safety pharmacology assessments noted decreases in PR interval duration that were detected up to 9 h post administration of VB253 and were usually dose related and associated with potential changes in heart rate.
+These changes were considered biologically insignificant due to their small magnitude (up to 10 msec compared to vehicle control), association with potential increases in heart rate, and the absence of correlated findings.
+
+The ADME properties of VB253 have been characterized in both *in vitro* and *in vivo* studies.
+Sensitive and selective bioanalytical methods have been developed and validated for rat and dog plasma.
+Following IV administration, the systemic clearance of VB253 in all species evaluated was considered low.
+Following oral administration, bioavailability was generally greater than 30% in the species evaluated.
+Initial studies utilized VB253 free form; however, due to higher bioavailability, a tris salt form of VB253 was chosen for further development.
+All GLP studies were performed using the VB253 tris salt form.
+Dose levels indicate VB253 free form equivalents unless otherwise noted.
+
+In Caco-2 cells, VB253 demonstrated high permeability with no evidence of being a substrate for an efflux transporter.
+Plasma protein binding was 98.7% in human plasma, 99.9% in rat plasma, and 98.5% in dog plasma.
+The metabolic profile of VB253 was evaluated in rat, dog, and human liver microsomes.
+Rat and dog appeared to cover the metabolites produced in human microsomes and were considered the suitable species for toxicology studies.
+Study of the oxidative metabolism of VB253 by recombinant CYP enzymes indicated that CYP2C8, CYP3A4, and CYP3A5 are involved in VB253 metabolism *in vitro*.
+VB253 demonstrated negligible inhibition of any of the CYP isozymes, with IC~50~ values being greater than 30 µM.
+VB253 was a weak activator of human PXR *in vitro* and is therefore unlikely to cause induction of PXR target genes, including CYP3A, *in vivo*.
+VB253 was a weak inhibitor of BCRP, OAT1, OAT3, OAT1B1, OAT1B3, and OCT1 transporters, with IC~50~ values ranging from 1.81-6.53 µM.
+
+Toxicokinetic studies generally demonstrated less-than-dose-proportional increases in exposure in the rat and dog.
+In the rat, sex differences were noted with female rats demonstrating higher exposure than male rats, while in the dog no differences in exposure between the sexes was noted.
+Generally, exposure was slightly higher following multiple doses relative to single dose administration.
+
+The toxicology of VB253 was evaluated in pivotal GLP studies in SD rats and beagle dogs, as well as mutagenicity studies.
+VB253 was not mutagenic in the bacterial reverse mutation assay or in the *in vitro* human lymphocyte micronucleus assay.
+
+VB253 was administered to SD rats by oral gavage at doses of 25, 75, and 250 mg/kg/day (females) and 40, 120, and 400 mg/kg/day (males) for 28 days.
+VB253-related mortality occurred in 8 of the 15 toxicity males administered 400 mg/kg/day beginning on Day 8 through 24.
+VB253-related effects in males administered 400 mg/kg/day and females administered 250 mg/kg/day were decreased body weight, decreased food consumption, clinical pathology changes in hematology and serum chemistry parameters, and histopathological changes.
+Target tissues included bone marrow, spleen, thymus, liver, lymph nodes, epididymides, seminal vescicles, prostate, testes, uterus, and ovaries.
+Administration of 40 or 120 mg/kg/day VB253 to male and 25 or 75 mg/kg/day in female rats *via* oral gavage once daily for 4 weeks with a 4-week recovery phase was well-tolerated.
+Based on these adverse findings the no-observed-adverse-effect level (NOAEL) was considered to be 120 mg/kg/day in male rats and 75 mg/kg/day in female rats with C~max~ of 180,000 ng/mL and 198,000 ng/mL, respectively, and an AUC0-24 of 2,780,000 and 3,230,000 ng·h/mL, respectively.
+
+VB253 was administered to beagle dogs by oral gavage in a range finding study at doses of 30, 79, 240 or 790 mg/kg.
+Watery feces and emesis were occasionally observed in dogs at all dose levels.
+Male dogs administered 790 mg/kg VB253 had increased potassium and decreased sodium and chloride, indicating a possible VB253-related effect on electrolyte homeostasis.
+Decreased thymus weights were observed at ≥ 240 mg/kg and correlated with microscopic findings of decreased thymic lymphoid cellularity.
+VB253 was administered at doses of 10, 50, and 150 mg/kg/day in the pivotal repeat dose toxicity studies for 28 days.
+Based on a lack of adverse findings, the NOAEL was considered to be 150 mg/kg/day in dogs with an associated sex-combined C~max~ of 80,400 ng/mL and sex-combined AUC0-24 of 474,000 ng•h/mL.
+
+\newpage
+
+# Materials and Methods
+
+## Key Resources
+
+```{r}
+#| results: asis
+#| echo: false
+#| tbl-cap: "Key Resources Table"
+
+flextable::flextable_to_rmd(resources_table)
+```
+
+\newpage
+
+## Chemicals
+
+Recombinant human TGFβ was purchased commercially , dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
+AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
+VB124 (10 μM) was a gift from Vettore, LLC.
+The drugs were prepared in DMSO and used at the indicated doses unless otherwise indicated.
+These doses were chosen based on prior reports and their target affinities.
+
+## Human lung tissue and fibroblasts
+
+Human IPF lung samples were obtained at the time of lung transplant.
+Control samples were prepared from non-fibrotic and unaffected border regions of lung nodule biopsy specimens or from failed donor controls.
+To isolate human lung fibroblasts, lung tissue was minced, digested with liberase and DNase, and passed through a sterile 70 μm filter.
+The filtrate was centrifuged at 300 ×*g* and washed once with RPMI medium (Lonza).
+Cells were plated in DMEM (Lonza) supplemented with 10% FBS, penicillin, and streptomycin.
+After two passages, medium was changed to FGM-2 (Lonza).
+
+## Bleomycin model
+
+C57Bl6/N mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
+Control animals received an equivalent volume of PBS.
+For drug treatments, AZD3965 (100 mg/kg twice daily) and VB124 (30 mg/kg daily) were prepared in 0.5% methylcellulose and 0.1% Tween-20 with sonication and administered by oral gavage every day beginning on day 7 following bleomycin.
+Control animals received vehicle alone twice daily.
+After 21 d, lung function measurements were performed.
+Animals were euthanized and plasma samples were obtained by right ventricular puncture and collected in EDTA-coated tubes.
+The lungs were dissected and perfused with PBS.
+The right lung was isolated while the left lung was inflated to 20 cmH~2~O prior to fixation.
+The right lungs were snap frozen.
+
+Lung physiology measurements were performed as previously described [@lee2021].
+Briefly, animals were anesthetized with pentobarbital (100 μg/g; Covetrus, Portland, ME) and tracheally cannulated for pulmonary function measurements using a flexiVent (SCIREQ).
+Animals were ventilated for 3-5 min with 150 breaths per minute at 10 mL/kg and 3 cmH~2~O positive end-expiratory pressure.
+After two total lung capacity maneuvers to prevent atelectasis and standardize volume history, respiratory elastance was measured using two Prime-8 maneuvers; values reported are the average of these two maneuvers.
+A pressure-volume loop was then generated, and static compliance was calculated using the Salazar-Knowles equation.
+
+Histologic fibrosis severity was assessed by Ashcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
+The average of twenty 10× fields is reported for each animal.
+
+Lung hydroxyproline content was determined as described previously [@tager2008].
+The four lobes of the R lung were homogenized in PBS with Halt protease inhibitors (Thermo) and hydrolyzed in 6 N HCl at 120 °C for 12 h, with an aliquot added to 1.4% chloramine-T, 10% 1-propanol, and 0.5 M sodium acetate, pH 6.0.
+After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-value-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
+Samples were incubated at 65 °C for 15 minutes.
+Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
+
+## Precision-cut lung slices
+
+Human precision-cut lung slices (PCLS) were purchased from Anabios (240326LuHA, 26 y.o.
+female without lung disease).
+The slices were incubated for 48 h in DMEM/F12 with 10% serum and then treated with TGFβ (5 ng/mL) and AZD3965 (100 nM) plus VB124 (10 μM) for 6 d, exchanging the medium every 48 h with fresh TGFβ and inhibitors.
+PCLS were lysed in RIPA buffer containing protease inhibitors, homogenized by bead-beating, and analyzed for extracellular matrix production by immunoblot.
+
+## Myofibroblast differentiation *in vitro*
+
+Normal lung fibroblasts were purchased commercially (Lonza).
+All fibroblasts were cultured in FGM-2 medium (Lonza).
+Passages 3-8 were used for experiments.
+Myofibroblast differentiation was induced *in vitro* using recombinant human TGFβ (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
+
+## Isotope labeling
+
+Cells were seeded and serum starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
+The concentrations of glucose and glutamine match the concentrations of these substrates determined in standard (FGM-2) growth medium.
+After 24 h, cells were washed with PBS and the medium was changed to "heavy" labeling medium containing [1,2-^13^C~2~]-glucose, [U-^13^C~6~]-glucose, [U-^13^C~5~]-glutamine, or [U-^13^C~2~]-lactate (Cambridge Isotope Labs) along with TGFβ and MCT inhibitors for 48 h.
+Cell lysates were prepared for metabolomics as described below.
+
+## Immunoblotting
+
+Human and mouse lung samples were mechanically homogenized in RIPA lysis buffer containing Halt Protease Inhibitor Cocktail (Thermo).
+Cells were washed with one volume of PBS and collected by scraping in PBS.
+Cell suspensions were centrifuged at 5,000 ×*g* for 5 min at 4 °C.
+Pellets were lysed in buffer containing Tris 10 mM, pH 7.4, NaCl 150 mM, EDTA 1 mM, EGTA 1 mM, Triton X-100 1% v/v, NP-40 0.5% v/v, and protease inhibitors.
+Protein concentrations were determined by BCA Protein Assay (Thermo).
+Lysates were normalized for protein concentration and subjected to SDS-PAGE separation on stain-free tris-glycine gels (Bio-Rad), cross-linked and imaged with the Chemidoc system (Bio-Rad), transferred to PVDF membranes with the Trans-Blot Turbo transfer system (Bio-Rad), imaged, blocked in 5% blocking buffer (Bio-Rad), blotted in primary and secondary antibodies, and developed using WesternBright ECL (Advansta).
+Band signal intensity was normalized to total protein per lane as determined from the stain-free gel or membrane images using Image Lab software (Bio-Rad).
+
+## RNA interference
+
+Lung fibroblasts were reverse transfected with ON-TARGETplus SMARTpool siRNA (Dharmacon) targeting MCT1 or MCT4 using Lipofectamine RNAiMAX (Thermo).
+A non-targeting siRNA pool was used a transfection control (Dharmacon).
+For combined siMCT1 and siMCT4 treatment, the total siRNA dose was kept constant.
+After 24 h, cells were serum starved for an additional 24 h prior to treatment with TGFβ.
+
+## Gel contraction assay
+
+Lung fibroblasts in FBM and neutralized collagen solution (TeloCol-3, Advanced Biomatrix) were combined 1:2 to yield 100,000 cells/mL and 1.8 mg/mL collagen.
+This solution was added to 24-well plates (0.5 mL) and allowed to harden for 10 min at 37 °C prior to the addition of 1 mL FBM.
+Cells were serum starved overnight prior to releasing the gels and treating them with TGFβ and MCT inhibitors in triplicate.
+Images were acquired after 24 and 48 h using a ChemiDoc imager (Bio-Rad).
+Gel areas were measured using Image Lab software (Bio-Rad).
+Data are expressed as the ratio of TGFβ-treated gel area to control gel area.
+
+## Cell count
+
+Cell counts were estimated from total DNA quantified using the Quant-iT PicoGreen dsDNA Assay Kit (Thermo) as described previously [@copeland2023].
+
+## Lactate assay
+
+Lactate was determined using an enzymatic cycling assay resulting in the reduction of resazurin to the fluorescent chemical resorufin.
+Medium samples were diluted 1:10 in PBS.
+Samples were combined with a solution of diaphorase (125 mU), lactate dehydrogenase, NAD (250 μM), and resazurin (25 μM) in Tris buffer (100 mM, pH 8.5) then incubated for 30 min at room temperature.
+Resorufin fluorescence was measured using a SpectraMax i3x microplate reader (Molecular Devices) and lactate concentration was interpolated from a standard curve.
+Data were normalized to estimated cell count.
+
+## Glucose assay
+
+Medium samples were diluted 10-fold in PBS.
+Glucose concentration was determined using the Glucose Colorimetric Assay Kit (Cayman) according to the manufacturer's protocol.
+Standards were prepared in PBS.
+
+## Seahorse assays
+
+Lung fibroblasts were seeded in Seahorse assay plates at a density of 20,000 cells per well.
+On the following day, the cells were serum starved for 24 h prior to treatment with TGFβ and MCT inhibitors for 48 h.
+On the day of the assay, the medium was exchanged to Seahorse DMEM without phenol red supplemented with 5 mM HEPES, 10 mM glucose, 2 mM glutamine, and 1 mM pyruvate.
+OCR and extracellular acidification rates (ECAR) were measured at baseline and following the sequential injections of oligomycin (1 μM final), FCCP (2 μM final), and rotenone plus antimycin A (0.5 μM final each) using an XFe24 bioanalyzer (Agilent).
+At the end of the assay, cells were stained for 30 min with the fluorescent dye Nuclear Green LCS1 (10 μM; Abcam).
+Stained cells were quantified using a SpectraMax MiniMax 300 Imaging Cytometer (Molecular Devices) and these cell counts were used to normalize the rate data.
+Proton efflux rate (PER) was calculated from the ECAR and buffer capacity of the medium.
+PER from glycolysis was determined by subtracting the contribution of oxidative phosphorylation to total PER using the empirically measured CO~2~ correction factor of 0.41.
+ATP from glycolysis was assumed to be equal to basal glycolytic PER.
+ATP from oxidative phosphorylation was assumed to be 2 × 2.75 × oligomycin-sensitive OCR [@mookerjee2017].
+
+## Metabolomics
+
+Metabolomics data acquisition and analysis were performed as described previously [@copeland2023].
+
+### Metabolite extraction
+
+Extracellular metabolites were obtained by mixing conditioned medium 1:4 with 100% MeOH precooled to -80 °C.
+Intracellular metabolites were obtained after washing cells with 2 volumes of ice-cold PBS and floating on liquid nitrogen.
+Plates were stored at -80 °C until extraction.
+Metabolites were extracted with 1 mL 80% MeOH pre-cooled to -80 °C.
+Mouse plasma and lung homogenates were mixed 1:4 with 100% MeOH precooled to -80 °C.
+Samples were extracted at -80 °C for 4 h.
+Insoluble material from these samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
+The supernatant was evaporated to dryness at 42 °C using a SpeedVac concentrator (Thermo Savant).
+Samples were resuspended in 35 μL 20 mM ammonium phosphate in LC-MS-grade water prior to analysis.
+
+### Acquisition parameters
+
+LC-MS analysis was performed on a Vanquish ultra-high-performance liquid chromatography system coupled to a Q Exactive orbitrap mass spectrometer by a HESI-II electrospray ionization probe (Thermo).
+External mass calibration was performed weekly.
+Metabolite samples (2.5 μL) were separated using a ZIC-pHILIC stationary phase (2.1 × 150 mm, 5 μm) (Merck).
+The autosampler temperature was 4 °C and the column compartment was maintained at 25 °C.
+Mobile phase A was 20 mM ammonium carbonate, 0.1% ammonium hydroxide, and 5 μM medronic acid.
+Mobile phase B was acetonitrile.
+The flow rate was 0.1 mL/min.
+Solvent was introduced to the mass spectrometer *via* electrospray ionization with the following source parameters: sheath gas 40, auxiliary gas 15, sweep gas 1, spray voltage +3.0 kV for positive mode and -3.1 kV for negative mode, capillary temperature 275 °C, S-lens RF level 40, and probe temperature 350 °C.
+Data were acquired and peaks integrated using TraceFinder 4.1 (Thermo).
+
+### Metabolomic profiling
+
+For metabolomic profiling, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B.
+The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*.
+Resolution was set to 70,000 and the AGC target was 1×10^6^ ions.
+Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method.
+For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run.
+Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas.
+Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018].
+Missing values were imputed using random forest.
+Sample peak areas were normalized using probabilistic quotient normalization [@dieterle2006].
+Differentially regulated metabolites were identified using limma [@ritchie2015].
+Exploratory data analysis and review of PCA plots suggested orthogonal effects of TGFβ and MCT inhibitor treatment.
+For this reason, differentially regulated metabolites were determined based on the main effects of TGFβ or MCT inhibitor.
+Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021] with KEGG metabolite pathways [@kanehisa2000].
+
+### Stable isotope quantification
+
+Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B.
+The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite.
+The resolution was set at 70,000 and AGC target was 1×10^5^ ions.
+Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015].
+Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].
+
+## RNA-seq transcriptomics
+
+RNA was collected from LFs treated with TGFβ and MCT inhibitors as described above.
+Four biological replicates were analyzed.
+Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample.
+Sequences were mapped to the human GRCh38 primary assembly and counts summarized using salmon [@patro2017].
+Differentially expressed genes were identified using DESeq2 [@love2014].
+Gene set enrichment was performed using the fgsea R package [@korotkevich2021].
+
+## Pyridine dinucleotide measurements
+
+Cellular NAD^+^ and NADH were measured using an enzymatic fluorometric cycling assay as previously described [@oldham2015; @copeland2023].
+Samples for NADP^+^ and NADPH quantification were similarly prepared and analyzed using the luminescence-based NADP/NADPH-Glo assay (Promega).
+
+## Reactive oxygen species
+
+Intracellular and mitochondrial ROS were measured using the fluorescent probes CellROX Deep Red (1 μM) and MitoSOX Green (2 μM), respectively.
+Mitochondria were also labeled with MitoTracker (200 nM).
+Briefly, lung fibroblasts were seeded in a 24-well plate and serum starved for 24 h prior to treatment with TGFβ, with or without MCT inhibitors, for 3 hours.
+This was followed by staining with the specified probes for 30 min at 37 °C.
+The cells were then harvested and analyzed using the CytoFLEX flow cytometry analyzer (Beckman Coulter).
+The mean fluorescence intensity of DAPI-negative cells was analyzed using FlowJo V10.8.1.
+
+## Multi-isotope imaging mass spectrometry
+
+For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on either day 21 or day 14 after bleomycin.
+Perfused and inflated lungs were fixed with formaldehyde/glutaraldehyde 2.5% in sodium cacodylate buffer, pH 7.4, embedded in EPON, section to 0.5 μm, mounted on silicon wafers, and gold coated.
+Sections were analyzed with the NanoSIMS 50L (CAMECA) housed in the Brigham and Women’s Hospital Center for NanoImaging, using previously developed analytical protocols [@wertheim2023; @guillermier2017a].
+The instrument was tuned to simultaneously acquire images of ^12^C~2~^1^H, ^12^C^14^N, ^12^C~2~^2^H, and ^12^C^15^N secondary ions.
+Spatial distribution visualization and quantification of ^15^N-proline and ^2^H-glucose labeling was derived from ^12^C^15^N/^12^C^14^N and ^12^C~2~^2^H/^12^C~2~^1^H ratios, respectively, using an open source plugin to ImageJ, OpenMIMS 3.0 (github.com/BWHCNI/OpenMIMS).
+Ratio images are presented visually as hue saturation intensity images, with the blue lower bound of the scale set to the natural background ratio and the red upper bound of the scale set such that within tissue labeling differences are visually apparent.
+Changes to the scaling modify the visual appearance but not affect the quantitative data.
+Quantification of isotopic enrichment was based on a pixel ratio analysis of several MIMS images acquired per animal.
+Specifically, image masks were generated by Otsu thresholding the ^12^C^14^N to exclude background regions.
+The ratio of "heavy" to "light" ion intensities in tissue regions was calculated.
+Pixel ratios equal to zero were excluded.
+The average of the remaining pixel ratios was then averaged to provide a mean tissue isotope enrichment value for the image.
+
+## VB253
+
+VB253 was developed by Vettore, LLC as a more potent small molecule inhibitor of MCT4.
+Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC, as described in detail below.
+These experiments were completed independently of the other studies presented in the manuscript and offer independent validation of the antifibrotic effects of MCT4 inhibition in cells and mice.
+
+MCT4 and MCT1 inhibition by VB253 were determined using a cell-based lactate transport assay monitoring the change in intracellular pH with BCECF following the addition of extracellular lactate.
+MDA-MB-231 and BT20 breast cancer cell lines were used to examine MCT4 and MCT1 inhibition, respectively.
+
+All cell culture experiments were conducted by Charles River Laboratories (Wilmington, MA, USA).
+Lung fibroblasts from three independent IPF donors were seeded at a density of 750 cells/well in 384-well plates.
+Fibroblasts were subsequently treated with medium with 0.1% DMSO (vehicle control), 1.25 ng/ml TGFβ alone, 1.25 ng/ml TGFβ in the presence of 1 μM SB525334 (positive control), or 1.25 ng/ml TGFβ with 8-point concentration response curves (CRC) in biologic duplicate with semi-log dilutions of nintedanib or VB253.
+The top concentration tested was 10 μM.
+Cells were fixed with 4% paraformaldehyde 72 h post treatment with TGFβ.
+The principal readout was immunofluorescence staining to for α-SMA and DAPI to quantify the proportion of α-SMA-positive cells.
+
+SMAD3 nuclear translocation was evaluated in lung fibroblasts from one IPF donor.
+Cells were plated on Purecol-coated 96-well plates.
+Unstimulated (medium only) and stimulated cells (0.5 ng/ml TGFβ) were treated with 0.1% DMSO (vehicle control), 1 μM SB525334 (positive control), nintedanib in 8-point CRC, ALK5 inhibitor (SB525334) in 8-point CRC, and VB253 in a 16-point CRC, with a maximum concentration of 30 μM.
+All cells were fixed 2 h after treatment.
+The readout was immunofluorescence staining for pSmad3 and DAPI.
+
+The *in vivo* therapeutic evaluation of VB253 in bleomycin-induced pulmonary fibrosis in both young and aged mice was conducted by Aragen Biosciences (Morgan Hill, CA, USA).
+In two separate experiments, C57Bl/6 mice, aged either 6-7 weeks or 68-70 weeks, were obtained from Jackson Labs (Bar Harbor, ME, USA).
+Under anesthesia, mice were administered either PBS vehicle (N = 10 per experiment) or 1.5 U/kg bleomycin in PBS intratracheally.
+After 7 days, mice treated with bleomycin were randomized to receive either vehicle (0.5% methylcellulose in saline) (N = 10), VB253 dissolved in 0.5% methylcellulose (3 mg/kg BID), Pirfenidone (Glentham Life Sciences Lid.
+Cat #GP4948, 100 mg/kg BID), or Nintedinib (Axon MedChem Cat#2648, 50 mg/kg, daily) (N = 10 per group).
+Mice injected with saline, rather than bleomycin, were administered vehicle.
+Treatments were administered daily until the experiment was terminated 21 days after bleomycin instillation, approximately 2-4 h after the final treatment dose.
+Whole-body plethysmography was performed on day 20 using Buxco WBP instrument system measuring Enhanced Pause (Penh) and respiratory rate on unrestrained animals.
+Upon experiment termination, lungs were inflated with 10% neutral buffered formalin (NBF) and fixed in 10% NBF for histology.
+
+Histology, imaging, and pathologic analyses were performed by HistoTox Labs (Boulder, CO, USA).
+Multiple slides per block were sectioned at 5 μm and stained with hematoxylin and eosin (H&E) or by routine immunohistochemistry (IHC) methods for α-SMA.
+H&E-stained glass slides were evaluated using light microscopy by an ACVP board-certified veterinary pathologist.
+Lung sections were scored according to the modified Ashcroft scale.
+Briefly, scores for five representative 200× microscopic fields per sample were averaged to obtain a mean score for each animal.
+α-SMA-stained glass slides were scanned using an Aperio AT2 whole slide scanner.
+Whole slide images were annotated to delineate regions of interest (ROI).
+Exclusions were applied to remove tissue artifacts (folds, tears, non-specific staining), large airways, and blood vessels.
+Visiopharm (VIS) image analysis software, using imaging filters to separate positive staining from counterstaining and background were applied to quantify α-SMA signal within each sample.
+Differences in expression levels between samples and treatments were evaluated by calculating the percent positive detection of cells or area within the viable lung tissue.
+
+\newpage
+
+# Figures
+
+![**Fig.
+S1.
+Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
+(**B**) Small molecule inhibitors of MCT1/2 (AR-C155858, AR) or MCT4 (VB124, VB) decrease TGFβ-stimulated α-SMA expression in normal human lung fibroblasts.
+(**C**-**D**) Cell counts following pharmacologic inhibition of lactate transporters.
+Summary data are mean ± SEM (\* p-value-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+](figs/SF_01.png)
+
+![**Fig.
+S2.
+Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A**-**C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh.
+Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/SF_02.png)
+
+\newpage
+
+![**Fig.
+S3.
+Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined 48 h following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
+(**B**) Extracellular glucose was determined 48 h following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
+(**C**) Summary data from Seahorse extracellular flux analysis.
+(**D**) Cell bioenergetic stress testing results from Seahorse analysis.
+Glycolytic capacity measures the increase in proton efflux rate following oligomycin treatment.
+Spare respiratory capacity measures the increase in oxygen consumption rate (OCR) following FCCP treatment compared to basal OCR.
+Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin).
+The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p-value < 0.05 compared to TGFβ/Veh).
+Summary data are mean ± SEM.](figs/SF_03.png)
+
+\newpage
+
+![**Fig.
+S4.
+Metabolomic profiling of lung fibroblasts treated with lactate transport inhibitors.** (**A**-**D**) Extracellular (A-B) and intracellular (C-D) metabolomic profile and metabolite set enrichment analyses of lung fibroblasts treated with TGFβ.
+(**E-H**) Extracellular (E-F) and intracellular (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment.
+(**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment.
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Positive NES indicates enrichment in treated cells while negative NES indicates enrichment in Vehicle-treated cells.
+(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
+Summary data are mean ± SEM.](figs/SF_04.png)
+
+\newpage
+
+![**Fig.
+S5.
+Stable isotope tracing in lung fibroblasts treated with TGFβ.** Lung fibroblasts were cultured with stable isotopes of glucose, lactate, and glutamine during treatment with TGFβ (*columns*).
+Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*).
+Significant differences in isotopic enrichment are indicated (\* adjusted p-value < 0.05).
+Summary data are mean ± SEM (N = 4 biological replicates).](figs/SF_05.png)
+
+\newpage
+
+![**Fig.
+S6.
+Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
+(**B**) MitoTracker fluorescence was measured as a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
+(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p-value < 0.05 for the main treatment effect).
+(**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates).
+Summary data are mean ± SEM.](figs/SF_06.png)
+
+\newpage
+
+![**Fig.
+S7.
+Mechanisms of lactate signaling.** (**A**) Exogenous lactate (10 mM) does not ameliorate TGFβ-stimulated α-SMA expression.
+(**B**) Lactate transport inhibition does not attenuate HIF-1α stabilization 6 h after TGFβ treatment.
+Summary data are mean ± SEM.
+(N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).](figs/SF_07.png)
+
+\newpage
+
+![**Fig.
+S8.
+Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly.
+Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/SF_08.png)
+
+\newpage
+
+![**Fig.
+S9.
+Metabolic reprogramming by lactate transport inhibition *in vivo*.** (**A**-**D**) Plasma (A-B) and lung (C-D) metabolomic profile and metabolite set enrichment analyses in bleomycin-treated mice.
+(**E**-**H**) Plasma (E-F) and lung (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment in bleomycin-treated mice.
+(**I**-**L**) Plasma (I-J) and lung (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment in bleomycin-treated mice.
+Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
+KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by bleomycin treatment or MCT inhibition ordered by normalized enrichment score (NES).
+Positive NES indicates enrichment in treated mice while negative NES indicates relative enrichment in control or untreated mice.](figs/SF_09.png)
+
+\newpage
+
+![**Fig.
+S10.
+VB253 is a novel, potent, and selective inhibitor of MCT4.** (**A**) Dose-dependent change in intracellular pH over time in BCECF-loaded MDA-MB-231 (MCT4) or BT20 (MCT1) breast cancer cell lines.
+(**B**) VB253 does not attenuate canonical TGFβ signaling.
+Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
+Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/SF_10.png)

### 2024-08-08 manuscript/supplement.qmd
revised submission for STM

--- 2024-07-24 manuscript/supplement.qmd
@@
 
-```{r}
-#| label: setup
-#| include: false
-
-# ensure refs are up-to-date
-targets::tar_load(refs_supp)
-targets::tar_load(pkg_citations)
-
-# ensure figures are up-to-date
-targets::tar_load(
-  c(
-    resources_table, 
-    fig01s,
-    fig02s, 
-    fig03s, 
-    fig04s, 
-    fig05s, 
-    fig06s, 
-    fig07s, 
-    fig08s, 
-    fig09s, 
-    fig10s,
-    NULL
-  )
-)
-```
-
 \newpage
-
-# Materials and Methods
-
-## Key Resources
-
-```{r}
-#| results: asis
-#| tbl-cap: "Key Resources Table"
-
-flextable::flextable_to_rmd(resources_table)
-```
-
-\newpage
-
-## Chemicals
-
-Recombinant human TGFβ was purchased commercially , dissolved in 10 mM citric acid, pH 3.0, filtered, and diluted to 10 μg/mL in PBS with 0.1% BSA prior to aliquoting and storing at -80 °C.
-AZD3965 (100 nM; Medkoo) and AR-C155858 (200 nM; Tocris) were purchased from commercial vendors.
-VB124 (10 μM) was a gift from Vettore, LLC.
-The drugs were prepared in DMSO and used at the indicated doses unless otherwise indicated.
-These doses were chosen based on prior reports and their target affinities.
-
-## Human lung tissue and fibroblasts
-
-Human IPF lung samples were obtained at the time of lung transplant.
-Control samples were prepared from non-fibrotic and unaffected border regions of lung nodule biopsy specimens or from failed donor controls.
-To isolate human lung fibroblasts, lung tissue was minced, digested with liberase and DNase, and passed through a sterile 70 μm filter.
-The filtrate was centrifuged at 300 ×*g* and washed once with RPMI medium (Lonza).
-Cells were plated in DMEM (Lonza) supplemented with 10% FBS, penicillin, and streptomycin.
-After two passages, medium was changed to FGM-2 (Lonza).
-
-## Bleomycin model
-
-C57Bl6/N mice aged 10-12 weeks were anesthetized with ketamine (90 mg/kg) and xylazine (5 mg/kg) and administered 1.2 U/kg bleomycin in PBS (Fresenius Kabi) intratracheally.
-Control animals received an equivalent volume of PBS.
-For drug treatments, AZD3965 (100 mg/kg twice daily) and VB124 (30 mg/kg daily) were prepared in 0.5% methylcellulose and 0.1% Tween-20 with sonication and administered by oral gavage every day beginning on day 7 following bleomycin.
-Control animals received vehicle alone twice daily.
-After 21 d, lung function measurements were performed.
-Animals were euthanized and plasma samples were obtained by right ventricular puncture and collected in EDTA-coated tubes.
-The lungs were dissected and perfused with PBS.
-The right lung was isolated while the left lung was inflated to 20 cmH~2~O prior to fixation.
-The right lungs were snap frozen.
-
-Lung physiology measurements were performed as previously described [@lee2021].
-Briefly, animals were anesthetized with pentobarbital (100 μg/g; Covetrus, Portland, ME) and tracheally cannulated for pulmonary function measurements using a flexiVent (SCIREQ).
-Animals were ventilated for 3-5 min with 150 breaths per minute at 10 mL/kg and 3 cmH~2~O positive end-expiratory pressure.
-After two total lung capacity maneuvers to prevent atelectasis and standardize volume history, respiratory elastance was measured using two Prime-8 maneuvers; values reported are the average of these two maneuvers.
-A pressure-volume loop was then generated, and static compliance was calculated using the Salazar-Knowles equation.
-
-Histologic fibrosis severity was assessed by Ashcroft scoring of trichrome-stained lung sections in a blinded fashion by an experienced lung pathologist [@hubner2008].
-The average of twenty 10× fields is reported for each animal.
-
-Lung hydroxyproline content was determined as described previously [@tager2008].
-The four lobes of the R lung were homogenized in PBS with Halt protease inhibitors (Thermo) and hydrolyzed in 6 N HCl at 120 °C for 12 h, with an aliquot added to 1.4% chloramine-T, 10% 1-propanol, and 0.5 M sodium acetate, pH 6.0.
-After incubating 20 min at room temperature, 1 ml of Ehrlich's solution (1 M p-value-dimethylaminobenzaldehyde in 70% 1-propanol and 20% perchloric acid) was added.
-Samples were incubated at 65 °C for 15 minutes.
-Absorbance was measured at 550 nm and hydroxyproline content was interpolated from a standard curve.
-
-## Myofibroblast differentiation *in vitro*
-
-Normal lung fibroblasts were purchased commercially (Lonza).
-All fibroblasts were cultured in FGM-2 medium (Lonza).
-Passages 3-8 were used for experiments.
-Myofibroblast differentiation was induced *in vitro* using recombinant human TGFβ (2 ng/mL) (Peprotech) for 48 h following 24 h serum starvation unless otherwise indicated.
-
-## Isotope labeling
-
-Cells were seeded and serum starved in MCDB131 medium lacking glucose, glutamine, and phenol red (genDEPOT) which was supplemented with naturally labeled glucose (8 mM), glutamine (1 mM), and lactate (2 mM) ("light" labeling medium).
-The concentrations of glucose and glutamine match the concentrations of these substrates determined in standard (FGM-2) growth medium.
-After 24 h, cells were washed with PBS and the medium was changed to "heavy" labeling medium containing [1,2-^13^C~2~]-glucose, [U-^13^C~6~]-glucose, [U-^13^C~5~]-glutamine, or [U-^13^C~2~]-lactate (Cambridge Isotope Labs) along with TGFβ and MCT inhibitors for 48 h.
-Cell lysates were prepared for metabolomics as described below.
-
-## Immunoblotting
-
-Human and mouse lung samples were mechanically homogenized in RIPA lysis buffer containing Halt Protease Inhibitor Cocktail (Thermo).
-Cells were washed with one volume of PBS and collected by scraping in PBS.
-Cell suspensions were centrifuged at 5,000 ×*g* for 5 min at 4 °C.
-Pellets were lysed in buffer containing Tris 10 mM, pH 7.4, NaCl 150 mM, EDTA 1 mM, EGTA 1 mM, Triton X-100 1% v/v, NP-40 0.5% v/v, and protease inhibitors.
-Protein concentrations were determined by BCA Protein Assay (Thermo).
-Lysates were normalized for protein concentration and subjected to SDS-PAGE separation on stain-free tris-glycine gels (Bio-Rad), cross-linked and imaged with the Chemidoc system (Bio-Rad), transferred to PVDF membranes with the Trans-Blot Turbo transfer system (Bio-Rad), imaged, blocked in 5% blocking buffer (Bio-Rad), blotted in primary and secondary antibodies, and developed using WesternBright ECL (Advansta).
-Band signal intensity was normalized to total protein per lane as determined from the stain-free gel or membrane images using Image Lab software (Bio-Rad).
-
-## RNA interference
-
-Lung fibroblasts were reverse transfected with ON-TARGETplus SMARTpool siRNA (Dharmacon) targeting MCT1 or MCT4 using Lipofectamine RNAiMAX (Thermo).
-A non-targeting siRNA pool was used a transfection control (Dharmacon).
-For combined siMCT1 and siMCT4 treatment, the total siRNA dose was kept constant.
-After 24 h, cells were serum starved for an additional 24 h prior to treatment with TGFβ.
-
-## Gel contraction assay
-
-Lung fibroblasts in FBM and neutralized collagen solution (TeloCol-3, Advanced Biomatrix) were combined 1:2 to yield 100,000 cells/mL and 1.8 mg/mL collagen.
-This solution was added to 24-well plates (0.5 mL) and allowed to harden for 10 min at 37 °C prior to the addition of 1 mL FBM.
-Cells were serum starved overnight prior to releasing the gels and treating them with TGFβ and MCT inhibitors in triplicate.
-Images were acquired after 24 and 48 h using a ChemiDoc imager (Bio-Rad).
-Gel areas were measured using Image Lab software (Bio-Rad).
-Data are expressed as the ratio of TGFβ-treated gel area to control gel area.
-
-## Cell count
-
-Cell counts were estimated from total DNA quantified using the Quant-iT PicoGreen dsDNA Assay Kit (Thermo) as described previously [@copeland2023].
-
-## Lactate assay
-
-Lactate was determined using an enzymatic cycling assay resulting in the reduction of resazurin to the fluorescent chemical resorufin.
-Medium samples were diluted 1:10 in PBS.
-Samples were combined with a solution of diaphorase (125 mU), lactate dehydrogenase, NAD (250 μM), and resazurin (25 μM) in Tris buffer (100 mM, pH 8.5) then incubated for 30 min at room temperature.
-Resorufin fluorescence was measured using a SpectraMax i3x microplate reader (Molecular Devices) and lactate concentration was interpolated from a standard curve.
-Data were normalized to estimated cell count.
-
-## Glucose assay
-
-Medium samples were diluted 10-fold in PBS.
-Glucose concentration was determined using the Glucose Colorimetric Assay Kit (Cayman) according to the manufacturer's protocol.
-Standards were prepared in PBS.
-
-## Seahorse assays
-
-Lung fibroblasts were seeded in Seahorse assay plates at a density of 20,000 cells per well.
-On the following day, the cells were serum starved for 24 h prior to treatment with TGFβ and MCT inhibitors for 48 h.
-On the day of the assay, the medium was exchanged to Seahorse DMEM without phenol red supplemented with 5 mM HEPES, 10 mM glucose, 2 mM glutamine, and 1 mM pyruvate.
-OCR and extracellular acidification rates (ECAR) were measured at baseline and following the sequential injections of oligomycin (1 μM final), FCCP (2 μM final), and rotenone plus antimycin A (0.5 μM final each) using an XFe24 bioanalyzer (Agilent).
-At the end of the assay, cells were stained for 30 min with the fluorescent dye Nuclear Green LCS1 (10 μM; Abcam).
-Stained cells were quantified using a SpectraMax MiniMax 300 Imaging Cytometer (Molecular Devices) and these cell counts were used to normalize the rate data.
-Proton efflux rate (PER) was calculated from the ECAR and buffer capacity of the medium.
-PER from glycolysis was determined by subtracting the contribution of oxidative phosphorylation to total PER using the empirically measured CO~2~ correction factor of 0.41.
-ATP from glycolysis was assumed to be equal to basal glycolytic PER.
-ATP from oxidative phosphorylation was assumed to be 2 × 2.75 × oligomycin-sensitive OCR [@mookerjee2017].
-
-## Metabolomics
-
-Metabolomics data acquisition and analysis were performed as described previously [@copeland2023].
-
-### Metabolite extraction
-
-Extracellular metabolites were obtained by mixing conditioned medium 1:4 with 100% MeOH precooled to -80 °C.
-Intracellular metabolites were obtained after washing cells with 2 volumes of ice-cold PBS and floating on liquid nitrogen.
-Plates were stored at -80 °C until extraction.
-Metabolites were extracted with 1 mL 80% MeOH pre-cooled to -80 °C.
-Mouse plasma and lung homogenates were mixed 1:4 with 100% MeOH precooled to -80 °C.
-Samples were extracted at -80 °C for 4 h.
-Insoluble material from these samples was removed by centrifugation at 21,000 ×*g* for 15 min at 4 °C.
-The supernatant was evaporated to dryness at 42 °C using a SpeedVac concentrator (Thermo Savant).
-Samples were resuspended in 35 μL 20 mM ammonium phosphate in LC-MS-grade water prior to analysis.
-
-### Acquisition parameters
-
-LC-MS analysis was performed on a Vanquish ultra-high-performance liquid chromatography system coupled to a Q Exactive orbitrap mass spectrometer by a HESI-II electrospray ionization probe (Thermo).
-External mass calibration was performed weekly.
-Metabolite samples (2.5 μL) were separated using a ZIC-pHILIC stationary phase (2.1 × 150 mm, 5 μm) (Merck).
-The autosampler temperature was 4 °C and the column compartment was maintained at 25 °C.
-Mobile phase A was 20 mM ammonium carbonate, 0.1% ammonium hydroxide, and 5 μM medronic acid.
-Mobile phase B was acetonitrile.
-The flow rate was 0.1 mL/min.
-Solvent was introduced to the mass spectrometer *via* electrospray ionization with the following source parameters: sheath gas 40, auxiliary gas 15, sweep gas 1, spray voltage +3.0 kV for positive mode and -3.1 kV for negative mode, capillary temperature 275 °C, S-lens RF level 40, and probe temperature 350 °C.
-Data were acquired and peaks integrated using TraceFinder 4.1 (Thermo).
-
-### Metabolomic profiling
-
-For metabolomic profiling, the following mobile phase gradient was used: 0 min, 80% B; 20 min, 20% B; 20.5 min, 80% B; 28 min, 80% B; 42 min, 80% B.
-The mass spectrometer was operated in polarity switching full scan mode from 70-1000 m/*z*.
-Resolution was set to 70,000 and the AGC target was 1×10^6^ ions.
-Peak identifications were based on an in-house library of authentic metabolite standards previously analyzed utilizing this method.
-For metabolomics studies, pooled quality control (QC) samples were injected at the beginning, end, and between every four samples of the run.
-Raw peak areas for each metabolite were corrected for instrument drift using a cubic spline model of QC peak areas.
-Low quality features were removed on the basis of a relative standard deviation greater than 0.2 in the QC samples and a dispersion ratio greater than 0.4 [@broadhurst2018].
-Missing values were imputed using random forest.
-Sample peak areas were normalized using probabilistic quotient normalization [@dieterle2006].
-Differentially regulated metabolites were identified using limma [@ritchie2015].
-Exploratory data analysis and review of PCA plots suggested orthogonal effects of TGFβ and MCT inhibitor treatment.
-For this reason, differentially regulated metabolites were determined based on the main effects of TGFβ or MCT inhibitor.
-Metabolite set enrichment analysis was performed using the fgsea package [@korotkevich2021] with KEGG metabolite pathways [@kanehisa2000].
-
-### Stable isotope quantification
-
-Metabolites were measured using the following mobile phase gradient: 0 min, 80% B; 5 min, 80% B; 30 min, 20% B; 31 min, 80% B; 42 min, 80% B.
-The mass spectrometer was operated in selected ion monitoring mode with an m/*z* window width of 9.0 centered at 1.003355-times half the number of carbon atoms in the target metabolite.
-The resolution was set at 70,000 and AGC target was 1×10^5^ ions.
-Peak areas were corrected for quadrupole bias as in Kim *et al.* [-@kim2015].
-Raw mass isotopomer distributions were corrected for natural isotope abundance using a custom R package (mzrtools, github.com/oldhamlab/mzrtools) employing the method of Fernandez, *et al.* [-@fernandez1996].
-
-## RNA-seq transcriptomics
-
-RNA was collected from LFs treated with TGFβ and MCT inhibitors as described above.
-Four biological replicates were analyzed.
-Library construction and sequencing was performed by BGI Genomics using 100 bp paired end analysis and a read depth of 50 M reads per sample.
-Sequences were mapped to the human GRCh38 primary assembly and counts summarized using salmon [@patro2017].
-Differentially expressed genes were identified using DESeq2 [@love2014].
-Gene set enrichment was performed using the fgsea R package [@korotkevich2021].
-
-## Pyridine dinucleotide measurements
-
-Cellular NAD^+^ and NADH were measured using an enzymatic fluorometric cycling assay as previously described [@oldham2015; @copeland2023].
-Samples for NADP^+^ and NADPH quantification were similarly prepared and analyzed using the luminescence-based NADP/NADPH-Glo assay (Promega).
-
-## Reactive oxygen species
-
-Intracellular and mitochondrial ROS were measured using the fluorescent probes CellROX Deep Red (1 μM) and MitoSOX Green (2 μM), respectively.
-Mitochondria were also labeled with MitoTracker (200 nM).
-Briefly, lung fibroblasts were seeded in a 24-well plate and serum starved for 24 h prior to treatment with TGFβ, with or without MCT inhibitors, for 3 hours.
-This was followed by staining with the specified probes for 30 min at 37 °C.
-The cells were then harvested and analyzed using the CytoFLEX flow cytometry analyzer (Beckman Coulter).
-The mean fluorescence intensity of DAPI-negative cells was analyzed using FlowJo V10.8.1.
-
-## Multi-isotope imaging mass spectrometry
-
-For multi-isotope imaging mass spectrometry, mice received five intraperitoneal doses of 250 μL of [^2^H~7~]-glucose or [U-^13^C~6~]-glucose (200 mg/mL) and [^15^N]-proline (20 mg/mL) every 12 h prior to euthanasia on either day 21 or day 14 after bleomycin.
-Perfused and inflated lungs were fixed with formaldehyde/glutaraldehyde 2.5% in sodium cacodylate buffer, pH 7.4, embedded in EPON, section to 0.5 μm, mounted on silicon wafers, and gold coated.
-Sections were analyzed with the NanoSIMS 50L (CAMECA) housed in the Brigham and Women’s Hospital Center for NanoImaging, using previously developed analytical protocols [@wertheim2023; @guillermier2017a].
-The instrument was tuned to simultaneously acquire images of ^12^C~2~^1^H, ^12^C^14^N, ^12^C~2~^2^H, and ^12^C^15^N secondary ions.
-Spatial distribution visualization and quantification of ^15^N-proline and ^2^H-glucose labeling was derived from ^12^C^15^N/^12^C^14^N and ^12^C~2~^2^H/^12^C~2~^1^H ratios, respectively, using an open source plugin to ImageJ, OpenMIMS 3.0 (github.com/BWHCNI/OpenMIMS).
-Ratio images are presented visually as hue saturation intensity images, with the blue lower bound of the scale set to the natural background ratio and the red upper bound of the scale set such that within tissue labeling differences are visually apparent.
-Changes to the scaling modify the visual appearance but not affect the quantitative data.
-Quantification of isotopic enrichment was based on a pixel ratio analysis of several MIMS images acquired per animal.
-Specifically, image masks were generated by Otsu thresholding the ^12^C^14^N to exclude background regions.
-The ratio of "heavy" to "light" ion intensities in tissue regions was calculated.
-Pixel ratios equal to zero were excluded.
-The average of the remaining pixel ratios was then averaged to provide a mean tissue isotope enrichment value for the image.
-
-## VB253
-
-VB253 was developed by Vettore, LLC as a more potent small molecule inhibitor of MCT4.
-Experiments involving VB253 were conducted by several contract research organizations under the supervision of Vettore, LLC, as described in detail below.
-These experiments were completed independently of the other studies presented in the manuscript and offer independent validation of the antifibrotic effects of MCT4 inhibition in cells and mice.
-
-All cell culture experiments were conducted by Charles River Laboratories (Wilmington, MA, USA).
-Lung fibroblasts from three independent IPF donors were seeded at a density of 750 cells/well in 384-well plates.
-Fibroblasts were subsequently treated with medium with 0.1% DMSO (vehicle control), 1.25 ng/ml TGFβ alone, 1.25 ng/ml TGFβ in the presence of 1 μM SB525334 (positive control), or 1.25 ng/ml TGFβ with 8-point concentration response curves (CRC) in biologic duplicate with semi-log dilutions of nintedanib or VB253.
-The top concentration tested was 10 μM.
-Cells were fixed with 4% paraformaldehyde 72 h post treatment with TGFβ.
-The principal readout was immunofluorescence staining to for α-SMA and DAPI to quantify the proportion of α-SMA-positive cells.
-
-SMAD3 nuclear translocation was evaluated in lung fibroblasts from one IPF donor.
-Cells were plated on Purecol-coated 96-well plates.
-Unstimulated (medium only) and stimulated cells (0.5 ng/ml TGFβ) were treated with 0.1% DMSO (vehicle control), 1 μM SB525334 (positive control), nintedanib in 8-point CRC, ALK5 inhibitor (SB525334) in 8-point CRC, and VB253 in a 16-point CRC, with a maximum concentration of 30 μM.
-All cells were fixed 2 h after treatment.
-The readout was immunofluorescence staining for pSmad3 and DAPI.
-
-The *in vivo* therapeutic evaluation of VB253 in bleomycin-induced pulmonary fibrosis in both young and aged mice was conducted by Aragen Biosciences (Morgan Hill, CA, USA).
-In two separate experiments, C57Bl/6 mice, aged either 6-7 weeks or 68-70 weeks, were obtained from Jackson Labs (Bar Harbor, ME, USA).
-Under anesthesia, mice were administered either PBS vehicle (N = 10 per experiment) or 1.5 U/kg bleomycin in PBS intratracheally.
-After 7 days, mice treated with bleomycin were randomized to receive either vehicle (0.5% methylcellulose in saline) (N = 10), VB253 dissolved in 0.5% methylcellulose (3 mg/kg BID), Pirfenidone (Glentham Life Sciences Lid.
-Cat #GP4948, 100 mg/kg BID), or Nintedinib (Axon MedChem Cat#2648, 50 mg/kg, daily) (N = 10 per group).
-Mice injected with saline, rather than bleomycin, were administered vehicle.
-Treatments were administered daily until the experiment was terminated 21 days after bleomycin instillation, approximately 2-4 h after the final treatment dose.
-Whole-body plethysmography was performed on day 20 using Buxco WBP instrument system measuring Enhanced Pause (Penh) and respiratory rate on unrestrained animals.
-Upon experiment termination, lungs were inflated with 10% neutral buffered formalin (NBF) and fixed in 10% NBF for histology.
-
-Histology, imaging, and pathologic analyses were performed by HistoTox Labs (Boulder, CO, USA).
-Multiple slides per block were sectioned at 5 μm and stained with hematoxylin and eosin (H&E) or by routine immunohistochemistry (IHC) methods for α-SMA.
-H&E-stained glass slides were evaluated using light microscopy by an ACVP board-certified veterinary pathologist.
-Lung sections were scored according to the modified Ashcroft scale.
-Briefly, scores for five representative 200× microscopic fields per sample were averaged to obtain a mean score for each animal.
-α-SMA-stained glass slides were scanned using an Aperio AT2 whole slide scanner.
-Whole slide images were annotated to delineate regions of interest (ROI).
-Exclusions were applied to remove tissue artifacts (folds, tears, non-specific staining), large airways, and blood vessels.
-Visiopharm (VIS) image analysis software, using imaging filters to separate positive staining from counterstaining and background were applied to quantify α-SMA signal within each sample.
-Differences in expression levels between samples and treatments were evaluated by calculating the percent positive detection of cells or area within the viable lung tissue.
-
-\newpage
-
-# References and Notes
-
-::: {#refs}
-:::
-
-\newpage
-
-# Figures
-
-![**Fig.
-S1.
-Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown.
-(**B**) Small molecule inhibitors of MCT1/2 (AR-C155858, AR) or MCT4 (VB124, VB) decrease TGFβ-stimulated α-SMA expression in normal human lung fibroblasts.
-(**C**-**D**) Cell counts following pharmacologic inhibition of lactate transporters.
-Summary data are mean ± SEM (\* p-value-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-](figs/SF_01.png)
-
-![**Fig.
-S2.
-Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A**-**C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh.
-Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/SF_02.png)
-
-\newpage
-
-![**Fig.
-S3.
-Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined 48 h following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh).
-(**B**) Extracellular glucose was determined 48 h following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates).
-(**C**) Summary data from Seahorse extracellular flux analysis.
-(**D**) Cell bioenergetic stress testing results from Seahorse analysis.
-Glycolytic capacity measures the increase in proton efflux rate following oligomycin treatment.
-Spare respiratory capacity measures the increase in oxygen consumption rate (OCR) following FCCP treatment compared to basal OCR.
-Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin).
-The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p-value < 0.05 compared to TGFβ/Veh).
-Summary data are mean ± SEM.](figs/SF_03.png)
-
-\newpage
-
-![**Fig.
-S4.
-Metabolomic profiling of lung fibroblasts treated with lactate transport inhibitors.** (**A**-**D**) Extracellular (A-B) and intracellular (C-D) metabolomic profile and metabolite set enrichment analyses of lung fibroblasts treated with TGFβ.
-(**E-H**) Extracellular (E-F) and intracellular (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment.
-(**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment.
-Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-KEGG pathways significantly enriched (p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES).
-Positive NES indicates enrichment in treated cells while negative NES indicates enrichment in Vehicle-treated cells.
-(**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).
-Summary data are mean ± SEM.](figs/SF_04.png)
-
-\newpage
-
-![**Fig.
-S5.
-Stable isotope tracing in lung fibroblasts treated with TGFβ.** Lung fibroblasts were cultured with stable isotopes of glucose, lactate, and glutamine during treatment with TGFβ (*columns*).
-Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*).
-Significant differences in isotopic enrichment are indicated (\* adjusted p-value < 0.05).
-Summary data are mean ± SEM (N = 4 biological replicates).](figs/SF_05.png)
-
-\newpage
-
-![**Fig.
-S6.
-Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition).
-(**B**) MitoTracker fluorescence was measured as a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells).
-(**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p-value < 0.05 for the main treatment effect).
-(**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates).
-Summary data are mean ± SEM.](figs/SF_06.png)
-
-\newpage
-
-![**Fig.
-S7.
-Mechanisms of lactate signaling.** (**A**) Exogenous lactate (10 mM) does not ameliorate TGFβ-stimulated α-SMA expression.
-(**B**) Lactate transport inhibition does not attenuate HIF-1α stabilization 6 h after TGFβ treatment.
-Summary data are mean ± SEM.
-(N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).](figs/SF_07.png)
-
-\newpage
-
-![**Fig.
-S8.
-Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly.
-Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/SF_08.png)
-
-\newpage
-
-![**Fig.
-S9.
-Metabolic reprogramming by lactate transport inhibition *in vivo*.** (**A**-**D**) Plasma (A-B) and lung (C-D) metabolomic profile and metabolite set enrichment analyses in bleomycin-treated mice.
-(**E**-**H**) Plasma (E-F) and lung (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment in bleomycin-treated mice.
-(**I**-**L**) Plasma (I-J) and lung (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment in bleomycin-treated mice.
-Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled.
-KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by bleomycin treatment or MCT inhibition ordered by normalized enrichment score (NES).
-Positive NES indicate enrichment in treated mice while negative NES indicate relative enrichment in control or untreated mice.](figs/SF_09.png)
-
-\newpage
-
-![**Fig.
-S10.
-VB253 does not attenuate canonical TGFβ signaling.** Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus.
-Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/SF_10.png)

### 2024-08-08 R/vb.R

@@ -51,14 +51,15 @@ plot_vb_rates <- function(z) {
       oob = scales::oob_squish_infinite
     ) +
     ggplot2::labs(
-      x = "Time (min)",
+      x = "Time (s)",
       y = "Fluorescence ratio",
       color = "VB253 (nM)"
     ) +
     ggplot2::coord_cartesian(clip = "off") +
     theme_plot() +
     ggplot2::theme(
-      legend.position = "bottom"
+      legend.position = "bottom",
+      legend.key.width = ggplot2::unit(0.25, "in"),
     ) +
     NULL
 }
